### 2019-12-17 latex_manuscript/MDPI/1_energies/spamda.tex
Initial commit.

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

\usepackage{textcomp, gensymb}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{epstopdf}
\usepackage{subfig}
\usepackage[T1]{fontenc}

\epstopdfsetup{outdir=./figures/}
\definecolor{gray090}{gray}{0.90}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0001-8849-6036} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0001-9773-6783} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorC}{0000-0002-2657-776X} % Add \orcidC{} behind the author's name
\newcommand{\orcidauthorD}{0000-0003-4564-1816} % Add \orcidD{} behind the author's name

% Authors, for the paper (add full first names)
%\Author{Antonio Manuel G\'omez-Orellana $^{1,\ddagger}$*, Juan Carlos Fern\'andez $^{1,\ddagger}$*\orcidA{}, Manuel Dorado-Moreno $^{1,\ddagger}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1,\ddagger}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1,\ddagger}$*\orcidD{}}

\Author{Antonio Manuel G\'omez-Orellana $^{1}$*, Juan Carlos Fern\'andez $^{1}$*\orcidA{}, Manuel Dorado-Moreno $^{1}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1}$*\orcidD{}}

% Authors, for metadata in PDF
\AuthorNames{Antonio Manuel G\'omez-Orellana, Juan Carlos Fern\'andez, Manuel Dorado-Moreno, Pedro Antonio Guti\'errez and C\'esar Herv\'as-Mart\'inez}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}

% Contact information of the corresponding author
\corres{Correspondence: am.gomez@uco.es (A.M.G.-O); jfcaballero@uco.es (J.C.F.); manuel.dorado@uco.es (M.D.-M.); pagutierrez@uco.es (P.A.G.); chervas@uco.es (C.H.-M.) }

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning. Machine Learning (ML) techniques have become a valuable support in many research areas, improving the performance of traditional statistical procedures. However, such techniques require datasets containing information related to the topic under study, which are not always available in an appropriate format. Preparing these datasets implies a lot of time and effort by the researchers. This paper presents a software tool for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} (NDBC) and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project}. The datasets created by the software are ready to be used as inputs for ML techniques in prediction tasks (classification or regression). As the designed software is able to simplify the creation of the datasets and reduce the time needed for this task, it prevents researchers from performing repetitive technical work, allowing them to concentrate on the study of the meteorological aspects of the data. Therefore, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.}

% Keywords
\keyword{Marine energy; Environmental modelling; Sustainable energy systems; Meteorological data; Reanalysis data; Pre-processing; Prediction tasks}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\setcounter{section}{-1} %% Remove this when starting to work on the template.
	\section{Introduction}

		A better understanding of the environment is of vital importance for science, contributing not only to more efficient exploitation of natural resources but also to the development of new strategies aimed at its protection. In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g. climate change, agriculture, energy, etc.). Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind energy assessment \cite{DVORAK20101244}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}. All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.

		On the other hand, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment. In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented. This tool can also be used to analyse the performance of the solar cells. A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices. In \cite{GUTIERREZ201369}, a tool based on \textit{Fatigue, Aerodynamics, Structures, and Turbulence} (FAST) code for performing design optimisation of offshore wind turbines is presented, which can analyse a massive group of cases by means of its parametric design capability. A design tool for architects based on \textit{Lighting and Thermal} (LT) method is used in \cite{BAKER2013156}, with the purpose of comparing and optimising design solutions in terms of energy and comfort performance. Raabe et al. \cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and an interactive web application, called the \textit{Stream Hydrology And Rainfall Knowledge System} (SHARKS), for analysis of hydrologic and meteorological time series data that can be used for providing valuable insights to urban hydrologic processes is presented in \cite{BRENDEL2019}.

		Marine energy prediction is currently a hot topic where meteorological data is used in. Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others. Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{BHATTACHARYA2017157, en12091657}. Wave energy exhibits a more stable power supply than wind energy and even solar energy. In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed to transform this wave energy into electricity \cite{en11092289}. WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them. Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}. As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \cite{en80910370}. For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$). Additionally, wave predictions \cite{en11010011} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.

		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks \cite{srivastava2017large}, obtaining significant improvements in the performance of the results. ML methodologies can be used not only by experienced data and computer scientists but also by other researchers. For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms. In this way, ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction. The problem is that, in order to apply these methods, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge. Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.

		The information to create these datasets can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements. Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others. Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{FERNANDEZ201544, Adams2010}, then a merging or matching process have to be carried out by the researchers to manually create the datasets with the needed information. Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards. In summary, many important details and different intermediate steps have to be considered when creating suitable datasets for ML techniques, resulting in an extremely tedious task.

		The main purpose of this paper is to present a software tool able to prevent researchers from performing this repetitive work and greatly simplify all the steps involved in the creation of datasets. The meteorological data considered for the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}. The software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets). As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows the researchers focus on the study of the meteorological aspects of the observations. The datasets obtained are ready to be used as input for ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes. These datasets contain one or more meteorological variables as inputs and one variable as target (variable to be predicted). The format of the generated datasets will be \textit{Attribute-Relation File Format} (ARFF) \cite{WEKA_ARFF}, which is the one used by WEKA. Besides, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling the researchers to use others tools.

		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, combining meteorological data from NDBC and NNRP. The advantages that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
			\item The generation of datasets becomes a very easy and customizable task, by means of the selection of different input parameters.
			\item It makes the researcher focus on oceanic and atmospheric studies, without having to worry about mechanical tasks.
			\item It provides information about the quality and quantity of the data.
			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets.
			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
			\item It facilitates data management and well-organised storage of the datasets.
			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy research.
			\item It includes an user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
		\end{itemize}

		This paper is organised as follows: Section \ref{sec:DataSources} describes the sources of information used by SPAMDA for creating datasets. Section \ref{sec:SPAMDA} describes in detail the features of the software tool. Section \ref{sec:CaseStudy} shows a case study describing the use of SPAMDA in a practical approach. Section \ref{sec:Conclusions} provides the final conclusions and future work.

	\section{Meteorological data sources}\label{sec:DataSources}

		The data provided by the above-mentioned sources of information of SPAMDA is described below:

		\begin{itemize}[leftmargin=*,labelsep=5.8mm]

			\item NDBC is a part of the \textit{National Weather Service} (NWS). NDBC designs, develops, operates, and maintains a network of data collecting buoys (stations). The mission of the network is to collect real-time marine meteorological and oceanographic observations, such as $H_s$, dominant wave period, or wind speed and direction, among others.

			The buoys maintained by NDBC are deployed in the coastal and offshore waters around oceans and seas, and they are equipped with assorted sensors which allow them to perform different measurements. The information collected by the buoys is available on the NDBC website \cite{NOAA_1}, and it is divided into different groups. One of them corresponds to standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite{NOAA_2}. These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC (Universal Time Coordinated) and from $23$:$50$ 31th December of the previous desired year to $22$:$50$ 31th December of the desired year. In Table \ref{tab:measurementsDescription}, a comprehensive measurement description and the corresponding units are provided as a summary for the reader. A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Figure \ref{fig:fragmentAnnualTexFile}. Each column corresponds to a meteorological variable or attribute, and each row or instance corresponds to the values of the measurements collected by the buoy for each attribute at a specific date and time.

			\begin{figure}[H]
				\centering
				\includegraphics[scale=0.50]{figures/FigureFragmentAnnualTextFile.png}
				\caption{A fragment of an annual text file of the \textit{Station 46001}.}
				\label{fig:fragmentAnnualTexFile}
			\end{figure}

			\begin{table}[H]

				\caption{Measurements descriptions and units of each meteorological variable or attribute collected by the buoys.}
				\label{tab:measurementsDescription}
				\footnotesize
				\centering

				\begin{tabular}{ccm{8.60cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}

					\cline{1-4}

					\textbf{Attribute}&\textbf{Units}&\textbf{Description}&\\[0.30cm]

					\cline{1-4}

					WDIR & degT & Wind direction (the direction the wind is coming from in degrees clockwise from true North) during the same period used for WSPD. \\

					\cellcolor{gray090}WSPD & \cellcolor{gray090} m/s & \cellcolor{gray090} Wind speed (m/s) averaged over an eight-minute period for buoys and a two-minute period for land stations. Reported Hourly. \\

					GST &  m/s & Peak 5 or 8 second gust speed (m/s) measured during the eight-minute or two-minute period. \\

					\cellcolor{gray090} WVHT & \cellcolor{gray090} m & \cellcolor{gray090} Significant wave height (meters) is calculated as the average of the highest one-third of all of the wave heights during the 20-minute sampling period. \\

					DPD & sec & Dominant wave period (seconds) is the period with the maximum wave energy. \\

					\cellcolor{gray090}APD & \cellcolor{gray090} sec & \cellcolor{gray090} Average wave period (seconds) of all waves during the 20-minute period. \\

					MWD & degT & The direction from which the waves at the dominant period (DPD) are coming. The units are degrees from true North, increasing clockwise, with North as 0 (zero) degrees and East as 90 degrees. \\

					\cellcolor{gray090}PRES & \cellcolor{gray090} hPa & \cellcolor{gray090} Sea level pressure (hPa). For C-MAN sites and Great Lakes buoys, the recorded pressure is reduced to sea level using the method described in NWS Technical Procedures Bulletin 291 (11/14/80). \\

					ATMP & degC & Air temperature (Celsius degrees). &\\[0.10cm]

					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature (Celsius degrees). For buoys the depth is referenced to the hull's waterline. For fixed platforms it varies with tide, but is referenced to, or near Mean Lower Low Water (MLLW).\\

					DEWP & degC & Dewpoint temperature taken at the same height as the air temperature measurement. \\

					\cellcolor{gray090}VIS & \cellcolor{gray090} nmi & \cellcolor{gray090} Station visibility (nautical miles). Note that buoy stations are limited to reports from 0 to 1.6 nmi. \\

					TIDE & ft & The water level in feet above or below MLLW. \\

					\cline{1-4}

				\end{tabular}

			\end{table}

			Note that the data collected by the network of buoys may be incomplete due to diverse circumstances such as the weather conditions in which the buoys have to operate, failures or malfunctioning elements of the buoys, among others. Accordingly, it may be the situation that some of the measurements are completely missing (missing date or instance) or partially missing (some measurements not recorded), by a buoy or by a set of buoys, once in a while or over a period of time. It may be also possible that the measurements have been recorded at a time different from the expected one. This aspect have to be taken into account when creating the datasets. This casuistry is explained in detail in Appendix \ref{app:AppendixA}.

			\item NNRP provides three-dimensional global reanalysis of numerous meteorological variables (e.g. air temperature, components South-North and West-East of wind speed, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$ Z (Zulu time), $06$ Z, $12$ Z and $18$ Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid. Weather observations are from different sources, such as ships, satellites and radar, among others. Reanalysis data is created assimilating such observations using the same climate model throughout the entire reanalysis period in order to reduce the effects of modelling changes on climate statistics. Such information has become a substantial support of the needs of the research community, even more in locations where instrumental (real time) data is not available.

			The reanalysis data is available in the NNRP website \cite{NNRP}, which it is accessible through different sections. Such data can be fully (a global $2.5\degree$ x $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data, which provides a description of the file contents and also includes the spatial and temporal properties of the data. Each reanalysis file contains the values of a meteorological variable estimated by a mathematical model for each reanalysis node. For a better understanding, in Figure \ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical location of a buoy (obtained from NDBC) is shown.

			\begin{figure}[H]
				\centering
				\includegraphics[scale=0.52]{figures/FigureSubGrid.jpg}
				\caption{Example of a six sub-grid reanalysis nodes around the \textit{Station 46001}.}
				\label{fig:subGrid}
			\end{figure}

		\end{itemize}

		Therefore, with both sources of information, which complement each other, and carrying out a matching process, SPAMDA will create datasets for prediction tasks. In this way, the dataset input variables will be one or more reanalysis variables from NNRP and one or more measurements from NDBC. The dataset output variable will always be one measurement from NDBC.

	\section{SPAMDA}\label{sec:SPAMDA}

		SPAMDA combines  meteorological information from NDBC and NNRP to obtain new datasets for oceanic and atmospheric studies. In order to do so, SPAMDA manages three different types of datasets, which will be described in detail in the following sections, but are briefly introduced bellow for giving the reader a better general understanding:
			\begin{itemize}[leftmargin=*,labelsep=5.8mm]
				\item \textit{Intermediate datasets}: They contain the meteorological observations from NDBC.
				\item \textit{Pre-processed datasets}: They are obtained as a result of pre-processing tasks performed on the intermediate datasets.
				\item \textit{Final datasets}: Created by merging an intermediate or pre-processed dataset (which contain the information from NDBC) with the reanalysis data from NNRP. This procedure is referenced in SPAMDA as matching process and will be carried out according to the study to be performed (classification or regression).
			\end{itemize}

		\begin{figure}[H]
			\centering
			\includegraphics[scale=0.32]{figures/FigureSPAMDA.eps}
			\caption{Brief outline of the functionality provided by SPAMDA.}
			\label{fig:SPAMDA}
		\end{figure}
		SPAMDA consists of three main functional modules, whose main features, represented in Figure \ref{fig:SPAMDA}, are the following:
		\begin{itemize}[leftmargin=*,labelsep=5.8mm]

			\item \textit{Manage buoys data}: The aim of this module is to provide features for the management and analysis of the information related to the buoys from NDBC. This includes:
			\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
				\item Entering and updating the information of each buoy.
				\item Creation of intermediate datasets with the collected measurements.
				\item Pre-processing tasks for obtaining the pre-processed datasets.
				\item Matching process to merge the information from NDBC and NNRP
				\item Creation of the final datasets accordingly to the ML technique to use (classification or regression).
			\end{enumerate}

			\item \textit{Manage reanalysis data}: This module is used for the management of the reanalysis data provided by the NNRP. In this way, the researchers can keep the reanalysis data files updated for their studies. Such files will be used, depending on the researchers needs, in the matching process when obtaining the final datasets.

			\item \textit{Tools}: This module includes features for converting intermediate or pre-processed datasets to ARFF or CSV format and for opening ARFF files with WEKA software.

		\end{itemize}

		In the following subsections each integrated functional module is described in detail.

			\subsection{Buoys}\label{sec:Buoys}

				When a new buoy is included in SPAMDA the following information, which can be obtained from NDBC, is requested:
				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
					\item \textit{Station ID}: An alphanumeric identifier that allows easy identification of the buoy.
					\item \textit{Description}: A short description of the buoy.
					\item \textit{Latitude}: North or South geographical localisation (degrees) of the buoy.
					\item \textit{Longitude}: West or East geographical localisation (degrees) of the buoy.
					\item \textit{Measurements files}: The above-mentioned annual text files of the standard meteorological information collected by the buoy and downloaded from the NDBC website. This will be used for the creation of the intermediate datasets. One file per year is expected.
				\end{itemize}

				For clarification, an example is presented in Figure \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.

				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.42]{figures/FigureBuoys.eps}
					\caption{Example of entering two buoys with its annual text files.}
					\label{fig:buoys}
				\end{figure}

			\subsection{Datasets}\label{sec:Datasets}

				Once a buoy has been included as described in Section \ref{sec:Buoys}, it is possible to create datasets with one or more annual text files, which are referenced in SPAMDA as intermediate datasets. In this module, the researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary ones.

				When an intermediate dataset is created, it is associated with its corresponding buoy. Besides, a summary of its content is also created, providing relevant information such as the number of instances, the dates of the first and last measurements, the annual text files included, and the missing and duplicated dates.

				An example where three intermediate datasets have been created is presented in Figure \ref{fig:datasets}. The two intermediate datasets of the buoy ID1 contain meteorological data of different years, and the intermediate dataset of the buoy ID2 contains meteorological data of two years. For each buoy, as many intermediate datasets as needed can be created.

				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.40]{figures/FigureDatasets.eps}
					\caption{Example of the creation of the intermediate datasets.}
					\label{fig:datasets}
				\end{figure}

			\subsection{Pre-process} \label{sec:Preprocess}

				Data pre-processing prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms. In this way, the quality of dat can be improved prior to the learning phase, by applying pre-processing tasks (filters). The result will be referenced as pre-processed datasets.

				SPAMDA provides several filters grouped in three categories, \textit{Attribute}, \textit{Instance} and \textit{Recover missing data}, including the configuration of their parameters and a short description of them:

				\begin{itemize}[leftmargin=*,labelsep=5.8mm]

				 \item \textit{Attribute}: All theses filters can be applied to the attributes (variables of the buoy from NDBC) of the intermediate dataset.

					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
						\item \textit{Normalize} \cite{WEKA_Filter_Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in the interval [0,1].
						\item \textit{Remove} \cite{WEKA_Filter_Remove}: It removes an attribute or a range of them.
						\item \textit{RemoveByName} \cite{WEKA_Filter_RemoveByName}: It removes attributes based on a regular expression matched against their names.
						\item \textit{ReplaceMissingValues} \cite{WEKA_Filter_ReplaceMissingValues}: For each attribute, all the missing values will be replaced by the average value of the attribute.
						\item \textit{ReplaceMissingWithUserConstant} \cite{WEKA_Filter_ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
					\end{itemize}

				 \item \textit{Instance}: All theses filters can be applied to the instances (hourly measurements of the buoy from NDBC) of the intermediate dataset.
					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
						\item \textit{RemoveDuplicates} \cite{WEKA_Filter_RemoveDuplicates}: With this filter, all duplicated instances are removed.
						\item \textit{RemoveWithValues} \cite{WEKA_Filter_RemoveWithValues}: This filter removes all the instances that match the attribute and the value supplied by the user.
						\item \textit{SubsetByExpression} \cite{WEKA_Filter_SubsetByExpression}: It removes all the instances which do not match a user-specified expression.
					\end{itemize}

				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the intermediate dataset.
					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
						\item \textit{Replace missing values with next nearest hour}: The missing values of each attribute are replaced with the next nearest non missing value.
						\item \textit{Replace missing values with previous nearest hour}: This filter replaces the missing values of each attribute with the previous nearest non missing value.
						\item \textit{Replace missing values with next $n$ hours mean}: The missing values of each attribute are replaced with the next $n$ nearest non missing values mean, where $n$ can be configured by the user.
						\item \textit{Replace missing values with previous $n$ hours mean}: This filter replaces the missing values of each attribute in the intermediate dataset with the previous $n$ nearest non missing values mean.
						\item \textit{Replace missing values with symmetric $n$ hours mean}: The missing values of each attribute in the intermediate dataset are replaced with the $n$ previous and $n$ next non missing values mean.
					\end{itemize}

				\end{itemize}

				SPAMDA allows the researchers to undo the last filter applied or to restore the initial content of the intermediate dataset. Besides, the content and relevant statistical information (number of instances with missing values, minimum and maximum values, mean and standard deviation) of the intermediate and the pre-processed datasets can be visualised in this module.

				Figure \ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one. The intermediate dataset 1 of the buoy ID2 has been also pre-processed. \textit{Pre-processed dataset n} represents that the researchers can create as many pre-processed datasets as they consider opportune.

				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.36]{figures/FigurePreprocess.eps}
					\caption{Example of the creation of pre-processed datasets.}
					\label{fig:preprocess}
				\end{figure}

			\subsection{Matching configuration}
			\label{sec:matching_conf}

				In order to merge and format the data provided by the two sources of information described in Section \ref{sec:DataSources}, it is necessary to carry out a matching process. The matching procedure is performed using an intermediate or pre-processed dataset, which includes the measurements collected by a buoy from NDBC, and the needed reanalysis data files from NNRP. Note that SPAMDA is able to manage the NetCDF binary format for handling the information stored in the reanalysis files.

				Such process merges the information of both sources that match on time, but, given that the measurements of the buoys are hourly collected from $00$:$50$ to $23$:$50$ UTC, and the reanalysis data is available every $6$ hours at $00$ Z, $06$ Z, $12$ Z and $18$ Z, the matching can only be carried every $6$ hours (discarding the rest of measurements from the buoy data). Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest buoy measurement (before or after) within a maximum of 60 minutes of difference. Finally, the matched instances of both sources will form the final datasets.

				Figure \ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year. In this way, only the instances from both sources that are linked with arrows (highlighted in green colour) will be used in the creation of the final datasets. Although the reanalysis dates have been presented in a human readable format, note that reanalysis dates are stored in hours from $01$-$01$-$1800$, and they have to be transformed for comparison taking into account the time zone. Such transformation is automatically done by SPAMDA when matching the instances.

				The reader can check in Appendix \ref{app:AppendixA} an example with a more complex case of the procedure.
				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.38]{figures/FigureMatchingProcess.png}
					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
					\label{fig:matchingProcess}
				\end{figure}

				SPAMDA allows the researchers to perform a customisable matching process, for obtaining as many different versions of the same meteorological data as needed.
				Prediction tasks are based on the estimation of the output attribute using the information provided by the input attributes. Depending on the task, the datasets must be prepared and configured differently:				
				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
					\item \textit{Classification}: The final datasets will be ready to use as input for ML classifiers, requiring a nominal output attribute, whose specific preparation is detailed in Section \ref{sec:FinalDatasets}.
					\item \textit{Regression}: The final datasets will be ready to use as input for regression methods, requiring a real output attribute, whose preparation is also explained in Section \ref{sec:FinalDatasets}.
					\item \textit{Direct matching}: In this case the inputs attributes have a direct correspondence with the output attribute, and it is not necessary to perform any additional preparation. Both input and target attributes are synchronised in time, in such a way that the final dataset is not intended for prediction purposes. For example, the final datasets may be used in lost data recovering tasks, in correlation studies, in descriptive analyses, etc.
				\end{itemize}

				The following parameters can be specified for the matching process:
				\begin{itemize}[leftmargin=*,labelsep=5.8mm]

					\item \textit{Flux of energy} \cite{FERNANDEZ201544}: When the $F_e$ is selected, it will be used as output. This attribute is not collected by the buoys, but it can be calculated from two wave parameters: $H_s$ and $T_e$, which are collected as WVHT and APD attributes, respectively, and were described in Table \ref{tab:measurementsDescription}. In this way, SPAMDA obtains the $F_e$ of each instance using the following equation:
					\begin{equation}
							F_e = 0.49 \cdot H^2_s \cdot T_e \;,
							\label{eq:fluxOfEnergy}
					\end{equation}

					where $F_e$ is measured in kilowatts per meter, $H_s$ is measured in meters and $T_e$ is measured in seconds. Note also that $F_e$ is defined in Equation \ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.

					\item \textit{Attribute to predict}: Instead of using $F_e$, researchers can select any of the attributes collected by the buoys as output (e.g. significant wave height, WVHT, wind direction, WDIR, sea level pressure, PRES, etc.). Therefore, they can conduct different studies by selecting one attribute or other.

					\item \textit{Reanalysis data files}: In order to have a possible better description of the problem under study, more than one reanalysis variable can be considered as input. Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Figure \ref{fig:subGrid}) for each variable of reanalysis.

					In that sense, the reanalysis data files must have the same spatial and temporal properties but related to different variables. SPAMDA simplifies this task by showing the reanalysis data files that are compatibles each other, and checking that the selection made by the researches meets that condition.

					\item \textit{Buoys attributes}: In addition to the reanalysis variables, the final datasets will also include the selected attributes as inputs (of the intermediate or pre-processed dataset used), providing a possible better characterisation of the problem under study, although it will depend on how correlated the attributes are.

					\item \textit{Include missing dates}: As above-mentioned, the information collected by a buoy may be incomplete due to measurements not recorded by it. As a consequence, the matching of instances between both sources of information may not be possible (missing dates). In that situation, the researchers can consider two options: 1) discard the instances affected or 2) include them. In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft\textit{?}\guillemotright.

					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Figure \ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes. In this way, the researcher can:

						\begin{itemize}[leftmargin=*,labelsep=5.8mm]

							\item Consider all the reanalysis nodes contained in each file: in this case, the information provided by each reanalysis node contained in each selected reanalysis data file will be used.

							\item Consider only some of the reanalysis nodes contained in each file: in this case, only the information of the $N$ closest reanalysis nodes to the buoy will be used  ($N$ given by the user). To do that, SPAMDA uses the \textit{Haversine} equation \cite{Haversine_2009} to calculate the distance from each reanalysis node to the location of the buoy and obtain the closest ones. Haversine equation is also known as the great circle distance and performs calculation from main point to destination point with a trigonometric function using latitude and longitude:
								\begin{eqnarray}
									\label{eq:Haversine}
									d(p_0,p_j) & = & \arccos(\sin(lat_0)\cdot \sin(lat_j) \nonumber \\
									& & \cdot \cos(lon_0-lon_j) + \cos(lat_0) \\
									& & \cdot \cos(lat_j))\;, \nonumber 
								\end{eqnarray}								
							where $p_0$ is the buoy geographical location, $p_j$ stands for the location of each reanalysis node, and $lat$ and $lon$ are the latitude and longitude of the points, respectively.
						\end{itemize}

					\item \textit{Number of final datasets}: Depending on the number of nearest reanalysis nodes to consider, the number of final datasets to create and the content of them can be configured according to the following options:
						\begin{itemize}[leftmargin=*,labelsep=5.8mm]

							\item \textit{One (using weighted mean of the $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of the reanalysis data used (one per selected reanalysis data file). This weighted mean is obtained by SPAMDA and uses Equation \ref{eq:Haversine} to obtain the distance from each reanalysis node to the location of the buoy. Once the distances have been calculated they are inverted and normalised as follows:							
								\begin{linenomath*}
									\begin{equation}
										w_i=\frac{\sum_{j=1}^{N} d(p_0,p_j)}{d(p_0,p_i)}, ~~i=1, \ldots, N.
										\label{eq:weightedMean}
									\end{equation}
								\end{linenomath*}

							With these weights, a weighted mean of each variable of reanalysis is obtained for each of the $N$ nodes. Therefore, the closest reanalysis nodes to the localisation of the buoy will provide more information.

							Considering as example the two nearest reanalysis nodes represented in Figure \ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.

							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by the researcher will be created. Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.

							In this case, and considering as example the four closest reanalysis nodes (see Figure \ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.

						\end{itemize}

				\end{itemize}

				Once the matching parameters have been described, for a better understanding of them, Figure \ref{fig:directMatching} presents an example of the matched information considering the data shown in Figure \ref{fig:matchingProcess} and using the following matching configuration\footnote{Note that the date is shown just for better understanding, but it will not be included in the final dataset.}:
					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
						\item Variable WVHT as attribute to predict.
						\item Variable Pres as reanalysis input attribute.
						\item Variable WSDP as buoy input attribute.
						\item Not including missing dates.
						\item Considering the closest reanalysis node.
						\item Task to be used: \textit{Direct matching}.
					\end{itemize}

				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.49]{figures/FigureDirectMatching.png}
					\caption{An example of the matched information for \textit{Direct matching}.}
					\label{fig:directMatching}
				\end{figure}

			\subsection{Final datasets} \label{sec:FinalDatasets}

				Once the matching process has been performed with the desired configuration, it is necessary to prepare the matched information for the desired prediction task (\textit{Regression} or \textit{Classification}), obtaining as a result the final datasets. Remember that \textit{Direct matching}, as it was described in Section \ref{sec:matching_conf}, performs a direct correspondence between the attributes used as inputs and the output one, and it is not necessary to carry out any preparation.

				SPAMDA allows the researchers to make such preparation by means of the following options:

					\begin{itemize}[leftmargin=*,labelsep=5.8mm]

						\item \textit{Prediction horizon} (Classification and Regression): This option indicates the time gap for moving backward the attribute to predict (output attribute). In this way, the input attributes (variables of the buoy and reanalysis data) will be used to predict the output attribute in a specific future time (e.g. +6h, +12h, +18h, +1 day, etc.).

						The minimum interval for increasing and decreasing the prediction horizon is $6$h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out. Therefore, for each increment of the prediction horizon, an instance is lost of the dataset is lost (as this future information is not available). As the minimum prediction horizon is $6$h, at least one instance will be lost. The relation between the inputs and the attribute to predict will be defined as follows:
						\begin{linenomath*}
							\begin{equation}
								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t})\;,
								\label{eq:noSynchronisingRD}
							\end{equation}
						\end{linenomath*}
						where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to be predicted, $\mathbf{b}_t$ is the vector containing the selected NDBC variables and $\mathbf{r}_t$ is the vector containing the selected reanalysis variables. In this way and considering the matched information shown in Figure \ref{fig:directMatching}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.

						Optionally, the reanalysis variables can be synchronised with the attribute to predict. Given that these variables are estimated by a mathematical model, we can obtain very good future estimations, which can improve the performance of the results. In this case, the relation between the inputs and the attribute to predict would be:						
						\begin{linenomath*}
							\begin{equation}
								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t+\Delta t}).
								\label{eq:synchronisingRD}
							\end{equation}
						\end{linenomath*}

						Note that the selected NDBC variables as input cannot be synchronised with the attribute to predict.

 						For better understanding, considering the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of one \textit{Regression} dataset is shown in Figure \ref{fig:regressionNoSync}. As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one). The options considered for the preparation of each final dataset are the following:
 							\begin{itemize}[leftmargin=*,labelsep=5.8mm]
 								\item Do not synchronise the reanalysis data (see Equation \ref{eq:noSynchronisingRD} for the relation between the inputs and the output).
 								\item A prediction horizon of $6$h.
 							\end{itemize}

							\begin{figure}[H]
								\centering
								\includegraphics[scale=0.49]{figures/FigureRegressionNoSync.png}
								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and without synchronisation.}
								\label{fig:regressionNoSync}
							\end{figure}

 						Note that, due to prediction horizon is $6$h, the values of WVHT attribute are moved backward one instance (up). As a consequence, the last instance ($2017$/$12$/$31$ $18$:$00$) is lost and is not included in the final dataset. Besides, and because the reanalysis data has not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation \ref{eq:noSynchronisingRD}).

 						Moreover, considering again the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of the same dataset but applying synchronisation (see Equation \ref{eq:synchronisingRD}) is shown in Figure \ref{fig:regressionSync}.

							\begin{figure}[H]
								\centering
								\includegraphics[scale=0.49]{figures/FigureRegressionSync.png}
								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and with synchronisation.}
								\label{fig:regressionSync}
							\end{figure}

						Again, and due to the prediction horizon selected ($6$h), the values of the WVHT attribute are moved backward one instance (up) and the last instance ($2017$/$12$/$31$ $18$:$00$) is not included in the final dataset. But now, the values of the Pres variable are also moved backward one instance (due to the synchronisation). Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Equation \ref{eq:synchronisingRD}).

						\item \textit{Thresholds of the output attribute} (Classification): Since the values of the variables collected by the buoys are real numbers, it is necessary to discretise them (convert them from real to nominal values) for the attribute selected as output (attribute to be predicted). SPAMDA allows the researchers to perform this process by defining the necessary classes with their thresholds, which will be used to carry out such discretisation.

						Considering again the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of a \textit{Classification} dataset is shown in Figure \ref{fig:prediction}. The options considered for the preparation of the final dataset are the following:
							\begin{itemize}[leftmargin=*,labelsep=5.8mm]
								\item Do not synchronise the reanalysis data.
								\item A prediction horizon of $6$h.
								\item The thresholds shown in Table \ref{tab:thresholds}.
							\end{itemize}

						\begin{table}[!h]

							\caption{Thresholds for the classification example represented in Figure \ref{fig:prediction}}
							\label{tab:thresholds}
							\footnotesize
							\centering

							\begin{tabular}{cm{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}

								\cline{1-5}

								\textbf{Class}&\textbf{Description}&\textbf{Inferior [}&\textbf{Superior )}&\\[0.20cm]

								\cline{1-5}

								Low & Low wave height & $0.36$ & $1.5$&\\[0.15cm]

								\cellcolor{gray090}Average & \cellcolor{gray090}Average wave height & \cellcolor{gray090}$1.5$ & \cellcolor{gray090}$2.5$&\\[0.15cm]

								Big & Big wave height & $2.5$ & $4.0$&\\[0.15cm]

								\cellcolor{gray090}Huge & \cellcolor{gray090}Huge wave height & \cellcolor{gray090}$4.0$ & \cellcolor{gray090}$9.9$&\\[0.15cm]

								\cline{1-5}

							\end{tabular}

						\end{table}

						\begin{figure}[H]
							\centering
							\includegraphics[scale=0.49]{figures/FigureClassification.png}
							\caption{An example of the creation a \textit{Classification} dataset, with a prediction horizon of $6$h and without synchronisation.}
							\label{fig:prediction}
						\end{figure}

						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable, because its values have been discretised according to the thresholds. Besides, and due to the $6$h prediction horizon, the last instance is lost ($2017$/$12$/$31$ $18$:$00$) and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up). As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation \ref{eq:noSynchronisingRD}).

					\end{itemize}

				The content of the final datasets, obtained as the result of the preparation of the matched data, can be visualised to check everything before saving them on disk. Such preparation can be performed as many times as required and considering the different options in each moment. Although the date will not be included in the final datasets, it can be shown to properly check the matching.

				Finally, it is necessary to define the output configuration to create the final datasets:			
				\begin{itemize}[leftmargin=*,labelsep=5.8mm]

					\item \textit{Output path file}: Name of the final datasets and folder to save them on disk.

					\item \textit{Final datasets format}:

						\begin{itemize}[leftmargin=*,labelsep=5.8mm]

							\item \textit{ARFF}: \textit{Attribute-Relation File Format} \cite{WEKA_ARFF}, which is used by WEKA. SPAMDA allows the researchers to directly open the final datasets in the Explorer environment of WEKA (in the same context of work), enabling them to choose the most appropriate ML method to tackle the problem under study.

							\item \textit{CSV}: \textit{Comma-Separated Values}. This format is included in order to consider other different tasks of software tools.

						\end{itemize}

				\end{itemize}

				A text file that summarises the configuration used in matching process and in the preparation of the matched data is also generated. It can be saved and loaded, enabling the researchers to resume their studies at any other time.

			\subsection{Manage reanalysis data}

				As mentioned in Section \ref{sec:DataSources}, the reanalysis data files provided by NNRP contain the estimated values by a mathematical model of one meteorological variable.

				In this module (see Figure \ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones. Besides, useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final time, latitude and longitude. All these fields summarise the temporal and spatial properties of the data. Thus, the researcher can quickly and easily identify each reanalysis file entered in SPAMDA.

				An example where two reanalysis data files have been entered in SPAMDA is shown in Figure \ref{fig:manageReanalisys}.

				\begin{figure}[H]
					\centering
					\includegraphics[scale=0.45]{figures/FigureManageReanalisys.eps}
					\caption{Example of entering two reanalysis data files.}
					\label{fig:manageReanalisys}
				\end{figure}

			\subsection{Tools}

				SPAMDA also contains another module that provides two utilities: one of them is \textit{Dataset converter} used for converting the desired intermediate or pre-processed datasets to ARFF or CSV formats; the other utility can be used for opening ARFF files with WEKA Explorer environment, which is useful for easily checking the results of different configurations of the pre-processing.

	\section{Case study}\label{sec:CaseStudy}

		Although the software includes a user manual, this section will describe how the application works in a practical approach. To do so, an example showing how to create a fully processed dataset (final dataset), starting from the raw data, will be described. The objective of this final dataset is to be used with ML algorithms to classify waves in the Gulf of Alaska depending on their height. The data collected to perform this example is:
		\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
		\item The measurements obtained from 2013 to 2017 by the buoy with ID 46001, placed in the Gulf of Alaska, which are provided by NDBC as annual text files. This data is publicly available at the NDBC website. 
		\item Complementary information collected from reanalysis data containing air temperature, pressure and two components of wind speed measurements (South-North and West-East). This information will be collected from the four closest reanalysis nodes surrounding the geographical location of the buoy. This data is publicly available at the NNRP website and can be downloaded in NetCDF format.
		\end{enumerate}

		After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.

		In Figure \ref{fig:main_view}, the main window is shown. In order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher has to select the option \textcolor{blue}{\textit{Manage reanalysis data}}.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.60\textwidth]{figures/FigureMain_view.png}
			\caption{SPAMDA main window.}\label{fig:main_view}
		\end{figure}

		Then, the window of Figure \ref{fig:reanalysis} is shown. Here, using the buttons located at the bottom, it is possible to add, delete or consult any data from the different reanalysis files. After the information has been introduced in the application, this window can be closed and the user can go back to the main window to continue entering the information related to the buoy under study.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.89\textwidth]{figures/FigureManage_reanalysis_data.png}
			\caption{\textit{Manage reanalysis data} window.}\label{fig:reanalysis}
		\end{figure}

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.89\textwidth]{figures/FigureManage_buoys.png}
			\caption{\textit{Buoys} tab.}\label{fig:manage_buoys}
		\end{figure}

		The researcher has to select \textcolor{blue}{\textit{Manage buoys data}} to open the window shown in Figure \ref{fig:manage_buoys}. In this tab, the researcher can consult, modify, add or delete different data related to the buoy. In order to enter such data, click on the \textcolor{blue}{\textit{New}} button, and then the window shown in Figure \ref{fig:new_buoy} pops up.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.70\textwidth]{figures/FigureNew_buoy.png}
			\caption{\em{Entering a new buoy} window.}\label{fig:new_buoy}
		\end{figure}

		Here the information about the buoy has to be included: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files. In this case, the files containing the data from year $2013$ to $2017$ are inserted by clicking on the \textcolor{blue}{\textit{Add file}} button. Once the data has been introduced, it is necessary to click on the \textcolor{blue}{\textit{Save}} button to insert the buoy in SPAMDA database. After that, the window can be closed.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.90\textwidth]{figures/FigureDatasets.png}
			\caption{\textit{Datasets} tab.}\label{fig:show_datasets}
		\end{figure}

		To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the corresponding view (see Figure \ref{fig:show_datasets}). In this view, the researcher can delete or consult a summary of each intermediate or pre-processed dataset by selecting it from the corresponding list. It can also create new ones. To proceed with the creation of the intermediate dataset, the user clicks on the \textcolor{blue}{\textit{New}} button, and the view shown in Figure \ref{fig:intermediate} appears.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.70\textwidth]{figures/FigureNew_intermediate_dataset.png}
			\caption{\textit{New intermediate dataset} view.}\label{fig:intermediate}
		\end{figure}

		Here the researcher can select the annual text files to be included in the intermediate dataset. In this case, all the files introduced before, which correspond to the buoy under study, are selected. When the file selection is finished, \textcolor{blue}{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textcolor{blue}{\textit{Save}} button, the creation process starts, showing the status of the process during it.

		After that, in order to prepare the intermediate dataset, the dataset is selected, and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}). In this tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.90\textwidth]{figures/FigurePreprocess.png}
			\caption{\textit{Pre-process} tab.}\label{fig:preprocess_data}
		\end{figure}

		Here the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter. As mentioned earlier, this case study is focused on classifying waves considering their height, so any missing data from wave height ($376$ values) and the remaining attributes are recovered, using the filter \textit{Replace missing values with symmetric $3$ hours mean}. Furthermore, the attributes MWD, DEWP, VIS and TIDE are removed from the dataset by applying the filter \textit{RemoveByName}, since the first two had more than $92$\% of missing data and the last two $100$\%. After finishing the pre-processing of the dataset, the researcher can click on the \textcolor{blue}{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.

		At this point, the researcher has registered the buoy in SPAMDA, then entered its raw data and selected the required data for the problem (intermediate dataset). Finally, the data has been pre-processed in order to be ready for its future use in ML algorithms. In order to achieve a more accurate description of the problem under study, a matching process can be carried out to merge the processed data from NDBC with the reanalysis data (also entered previously) from NNRP. The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.90\textwidth]{figures/FigureMatching_configuration.png}
			\caption{\textit{Matching configuration} tab.}\label{fig:matching_conf}
		\end{figure}

		In this view, the researcher can customise (or load) the parameters of the matching process according to their needs and select the prediction task  (described in Section \ref{sec:matching_conf}) that the final dataset will be used for. In this example, the following parameters were selected:
		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
			\item Attribute to predict: WVHT.
			\item Reanalysis data: Air, pressure, u-wind and v-wind.
			\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP.
			\item Reanalysis nodes to consider: $1$.
			\item Number of final datasets: In this example that option is disabled, because only one reanalysis node is considered.
			\item Prediction task: Classification.
		\end{itemize}

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.90\textwidth]{figures/FigureFinal_datasets.png}
			\caption{\textit{Final datasets} tab.}\label{fig:final_dataset}
		\end{figure}

		After configuring the matching process, the researcher can click on the \textcolor{blue}{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task. Given that, in the previous window, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable. After this, the next step is to set the time horizon desired and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output, as explained in Section {\ref{sec:FinalDatasets}}. Then the researcher can click on the \textcolor{blue}{\textit{Update final dataset}} button to see the content shown in the bottom left corner. Finally, after checking that everything is correct, the last step would be to select the name (and path) of the dataset file and its output format and click on the \textcolor{blue}{\textit{Create final datasets}} button. For this case study, the following configuration was applied:		
		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
			\item Thresholds: see Table \ref{tab:thresholds}.
			\item Prediction horizon: 6 hours
			\item Synchronisation: Disabled
		\end{itemize}

		At this point the final dataset would be created and stored in the computer of the researcher. Also, there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Figure \ref{fig:openigFinalDatasetWeka}.

		\begin{figure}[H]
			\centering
			\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka.png}
			\caption{Final dataset opened with the environment Explorer of WEKA.}
			\label{fig:openigFinalDatasetWeka}
		\end{figure}

	\section{Conclusions}\label{sec:Conclusions}

		A software tool for creating datasets using meteorological data from NDBC and NNRP has been presented. These datasets will be ready to use as input for ML techniques in prediction tasks (classification or regression). As a result, the researchers will benefit from a great support when carrying out their oceanic and atmospheric studies, related to energy and environmental modelling. Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed. In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.

		In order to improve SPAMDA, some future work could be focused on new functional modules for managing meteorological data of different formats \cite{NOAA_3}, so that the developed tool can be extended to any other research, new pre-processing functionalities such as filters to analyse the correlation between attributes or new functional modules for recovering missing values using nearby buoys data \cite{DuranRosal2016}. Furthermore, the developed software could manage other sources of reanalysis data (with different spatial and temporal resolution), and new output formats for the datasets which could be used as input by other tools for ML such as KEEL (\textit{Knowledge Extraction based on Evolutionary Learning}) \cite{AlcalFdez2009KEELAS}. However, such new functionalities can be developed with a reasonable effort to be able to manage each particular casuistry. For example, when dealing with incomplete data, interpreting different data and files structures or carrying out the matching process of two environmental data sources.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \authorcontributions{All authors contributed to this work. Design and implementation of the software, A.M.G.-O., J.F.C. and M.D.-M.; Test of the software, P.A.G. and C.H.-M.; Conceptualization, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Methodology, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Investigation, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Writing-original draft preparation, A.M.G.-O., J.F.C. and M.D.-M.; Funding acquisition, P.A.G. and C.H.-M.; Resources, P.A.G. and C.H.-M.; Supervision, P.A.G. and C.H.-M.}

\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We thank to NVIDIA Corporation for the transfer of computational resources for research works. The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest. The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
$F_e$ & Flux of energy\\
$H_s$ & Significant wave height\\
$T_e$ & Wave energy period\\
$p_0$ & Geographical location of the buoy\\
$p_j$ & Geographical location of each reanalysis node\\
$lat$ & Latitude of the point\\
$lon$ & Longitude of the point\\
$o_{t}$ & The attribute to be predicted at the time instant to study\\
$\Delta t$ & The prediction horizon\\
$\mathbf{b}_t$ & The vector containing the selected NDBC variables\\
$\mathbf{r}_{t}$ & The vector containing the selected reanalysis variables\\
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendix
\section{Managing incomplete data}\label{app:AppendixA}
%\unskip
%\subsection{}

		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.

		The measurements collected by the buoys may be incomplete or recorded at a different time than the expected one, due to the weather conditions in which the buoys have to operate. To illustrate this casuistry, the following examples are shown in Figure \ref{fig:measurements}:
		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
			\item In the instance marked with a), the measurement of $17$:$50$ was collected at $17$:$45$, $5$ minutes earlier.
			\item In the instance marked with b), the measurement of $23$:$50$ was collected at $23$:$30$, $20$ minutes earlier.
			\item In the instance marked with c), the measurement of $05$:$50$ is duplicated.
			\item In the instance marked with d), the measurement of $11$:$50$ is missing (missing date or instance).
			\item In the instance marked with e), the measurement of $17$:$50$ and $18$:$50$ are missing (missing dates or instances).
			\item Missing values highlighted in red colour.
		\end{itemize}

		\begin{figure}[H]
			\centering
			\includegraphics[scale=0.45]{figures/FigureMeasurements.png}
			\caption{A fragment of an annual text file with different missing value examples.}
			\label{fig:measurements}
		\end{figure}

		SPAMDA has been designed to tackle these situations, and it informs the researchers of any incidence found while reading the annual text files for creating the intermediate datasets. For the case of measurements that were recorded at a different time than expected, it has been established a time gap of 6 minutes ($10\%$ of an hour). Therefore, if the time difference exceeds such value the date will be considered as an unexpected.

		Figure \ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Figure \ref{fig:measurements}. Note that the instance marked with a) has not been informed by SPAMDA as an unexpected date because its time difference is less than $6$ minutes. Depending on the affected attribute, NDBC uses a specific value \cite{NOAA_3} to indicate the presence of lost data (e.g. $99$ for VIS and TIDE attributes, $999$ for DEWP, MWD and WDIR, etc.). SPAMDA interprets these specific values and, after creating the intermediate dataset, the researchers can check if it contains missing values by visualising its statistical information or content. Remember that SPAMDA provides several filters for recovering missing data, which were described in Section \ref{sec:Datasets}.

		\begin{figure}[H]
			\centering
			\includegraphics[scale=0.40]{figures/FigureCreatingDataset.png}
			\caption{Status of the creation of the intermediate dataset for the example of Fig \ref{fig:measurements}.}
			\label{fig:creatingDataset}
		\end{figure}

		SPAMDA takes into account this casuistry when carrying out the matching process. An example is given in Figure \ref{fig:matchingMeasurements}. As above-mentioned, the matching process is performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference. However, in the instance marked with $e)$, given that the measurements dates $01$/$05$/$2017$ $17$:$50$ and $01$/$05$/$2017$ $18$:$50$ are missing, the reanalysis date $01$/$05$/$2017$ $18$:$00$ cannot be matched with buoy data (this date is highlighted in mauve colour in the Figure). Depending on the selection made by the researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.

		\begin{figure}[H]
			\centering
			\includegraphics[scale=0.36]{figures/FigureMatchingMeasurements.png}
			\caption{Matching the measurements (left) and the reanalysis data (right).}
			\label{fig:matchingMeasurements}
		\end{figure}

%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{bibfile}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{Samples of the compounds ...... are available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2019-12-17 latex_manuscript/elsevier/2_Environmental_Modelling_and_Software/spamda.tex
Initial commit.

--- 2019-12-17 latex_manuscript/MDPI/1_energies/spamda.tex
@@
-%  LaTeX support: latex@mdpi.com 
-%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
-
-%=================================================================
-\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
-
-% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
-The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
-This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
-For preprints.org, the editorial staff will make this change immediately prior to posting.
-
-%--------------------
-% Class Options:
-%--------------------
-%----------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by: 
-% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
-% supfile = supplementary materials
-
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
-This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
-Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
-
-%=================================================================
-\firstpage{1} 
-\makeatletter 
-\setcounter{page}{\@firstpage} 
-\makeatother
-\pubvolume{xx}
-\issuenum{1}
-\articlenumber{5}
-\pubyear{2019}
-\copyrightyear{2019}
-%\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
-%\updates{yes} % If there is an update available, un-comment this line
-
-%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
-%\continuouspages{yes}
-
-%------------------------------------------------------------------
-% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
-%\pdfoutput=1
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
+% !TeX spellcheck = en_GB
+\documentclass[review]{elsarticle}
 
@@
 \usepackage[T1]{fontenc}
+\usepackage{lineno}
+\usepackage{amsmath}
+\usepackage{hyperref}
 
 \epstopdfsetup{outdir=./figures/}
+
 \definecolor{gray090}{gray}{0.90}
 
-%=================================================================
-%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
-%=================================================================
-% Full title of the paper (Capitalized)
-\Title{SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
-
-% Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0001-8849-6036} % Add \orcidA{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0001-9773-6783} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0002-2657-776X} % Add \orcidC{} behind the author's name
-\newcommand{\orcidauthorD}{0000-0003-4564-1816} % Add \orcidD{} behind the author's name
-
-% Authors, for the paper (add full first names)
-%\Author{Antonio Manuel G\'omez-Orellana $^{1,\ddagger}$*, Juan Carlos Fern\'andez $^{1,\ddagger}$*\orcidA{}, Manuel Dorado-Moreno $^{1,\ddagger}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1,\ddagger}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1,\ddagger}$*\orcidD{}}
-
-\Author{Antonio Manuel G\'omez-Orellana $^{1}$*, Juan Carlos Fern\'andez $^{1}$*\orcidA{}, Manuel Dorado-Moreno $^{1}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1}$*\orcidD{}}
-
-% Authors, for metadata in PDF
-\AuthorNames{Antonio Manuel G\'omez-Orellana, Juan Carlos Fern\'andez, Manuel Dorado-Moreno, Pedro Antonio Guti\'errez and C\'esar Herv\'as-Mart\'inez}
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-$^{1}$ \quad Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}
-
-% Contact information of the corresponding author
-\corres{Correspondence: am.gomez@uco.es (A.M.G.-O); jfcaballero@uco.es (J.C.F.); manuel.dorado@uco.es (M.D.-M.); pagutierrez@uco.es (P.A.G.); chervas@uco.es (C.H.-M.) }
-
-% Current address and/or shared authorship
-%\firstnote{Current address: Affiliation 3} 
-%\secondnote{These authors contributed equally to this work.}
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-%\simplesumm{} % Simple summary
-
-%\conference{} % An extended version of a conference paper
-
-% Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
+\modulolinenumbers[5]
+
+\journal{Environmental Modelling \& Software}
+
+\bibliographystyle{elsarticle-num}
+
+\begin{document}
+
+	\begin{frontmatter}
+
+		\title{SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
+
+		\author[UCO]{Antonio M. G\'omez-Orellana}
+		\ead{am.gomez@uco.es}
+
+		\author[UCO]{Juan C. Fern\'andez\corref{JCFC}}
+		\cortext[JCFC]{Corresponding author.}
+		\ead{jfcaballero@uco.es}
+
+		\author[UCO]{Manuel Dorado-Moreno}
+		\ead{manuel.dorado@uco.es}
+
+		\author[UCO]{Pedro A. Guti\'errez}
+		\ead{pagutierrez@uco.es}
+
+		\author[UCO]{C\'esar Herv\'as-Mart\'inez}
+		\ead{chervas@uco.es}
+
+		\address[UCO]{Department of Computer Science and Numerical Analysis, University of Cordoba, C\'ordoba, Spain.}
+
+		\begin{abstract}
+			%Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
 Machine Learning (ML) techniques have become a valuable support in many research areas, improving the performance of traditional statistical procedures.
@@
 As the designed software is able to simplify the creation of the datasets and reduce the time needed for this task, it prevents researchers from performing repetitive technical work, allowing them to concentrate on the study of the meteorological aspects of the data.
-Therefore, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.}
-
-% Keywords
-\keyword{Marine energy; Environmental modelling; Sustainable energy systems; Meteorological data; Reanalysis data; Pre-processing; Prediction tasks}
-
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%\setcounter{secnumdepth}{4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\setcounter{section}{-1} %% Remove this when starting to work on the template.
+Therefore, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
+			Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
+Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation implies a lot of time and effort by the researchers.
+This paper presents a software for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
+As the designed software is able to simplify the creation of the datasets (ready to be used as inputs for ML techniques in prediction tasks) and reduce the time needed for this task, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
+		\end{abstract}
+
+		\begin{keyword}
+			Meteorological data\sep Reanalysis data\sep  Pre-processing\sep Creating datasets\sep  Prediction tasks\sep Marine energy
+		\end{keyword}
+
+	\end{frontmatter}
+
+	\linenumbers
+
 	\section{Introduction}
@@
 		Marine energy prediction is currently a hot topic where meteorological data is used in.
-Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{BHATTACHARYA2017157, en12091657}.
+Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
+Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{BHATTACHARYA2017157, BREDE201344}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed to transform this wave energy into electricity \cite{en11092289}.
+In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed to transform this wave energy into electricity.
 WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them.
 Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \cite{en80910370}.
+As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction.
 For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-Additionally, wave predictions \cite{en11010011} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
+Additionally, wave predictions are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
 
@@
 Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
-Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{FERNANDEZ201544, Adams2010}, then a merging or matching process have to be carried out by the researchers to manually create the datasets with the needed information.
+Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a merging or matching process have to be carried out by the researchers to manually create the datasets with the needed information.
 Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
@@
 The advantages that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+		\begin{itemize}
 			\item The generation of datasets becomes a very easy and customizable task, by means of the selection of different input parameters.
@@
 
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+		\begin{itemize}
 
@@
 The information collected by the buoys is available on the NDBC website \cite{NOAA_1}, and it is divided into different groups.
-One of them corresponds to standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite{NOAA_2}.
+One of them corresponds to standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite {NOAA_2}.
 These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC (Universal Time Coordinated) and from $23$:$50$ 31th December of the previous desired year to $22$:$50$ 31th December of the desired year.
 In Table \ref{tab:measurementsDescription}, a comprehensive measurement description and the corresponding units are provided as a summary for the reader.
-A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Figure \ref{fig:fragmentAnnualTexFile}.
+A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Fig.
+\ref{fig:fragmentAnnualTexFile}.
 Each column corresponds to a meteorological variable or attribute, and each row or instance corresponds to the values of the measurements collected by the buoy for each attribute at a specific date and time.
 
-			\begin{figure}[H]
+			\begin{figure}[ht!]
 				\centering
@@
 
-			\begin{table}[H]
+			\begin{table}[!ht]
 
@@
 This aspect have to be taken into account when creating the datasets.
-This casuistry is explained in detail in Appendix \ref{app:AppendixA}.
+This casuistry is explained in detail in Appendix A.
 
@@
 Each reanalysis file contains the values of a meteorological variable estimated by a mathematical model for each reanalysis node.
-For a better understanding, in Figure \ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical location of a buoy (obtained from NDBC) is shown.
-
-			\begin{figure}[H]
+For a better understanding, in Fig.
+\ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical location of a buoy (obtained from NDBC) is shown.
+
+			\begin{figure}[ht!]
 				\centering
@@
 In order to do so, SPAMDA manages three different types of datasets, which will be described in detail in the following sections, but are briefly introduced bellow for giving the reader a better general understanding:
-			\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+			\begin{itemize}
 				\item \textit{Intermediate datasets}: They contain the meteorological observations from NDBC.
@@
 
-		\begin{figure}[H]
+		\begin{figure}[ht!]
 			\centering
@@
 		\end{figure}
-		SPAMDA consists of three main functional modules, whose main features, represented in Figure \ref{fig:SPAMDA}, are the following:
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+		SPAMDA consists of three main functional modules, whose main features, represented in Fig. \ref{fig:SPAMDA}, are the following:
+		\begin{itemize}
 
 			\item \textit{Manage buoys data}: The aim of this module is to provide features for the management and analysis of the information related to the buoys from NDBC. This includes:
-			\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
+			\begin{enumerate}
 				\item Entering and updating the information of each buoy.
@@
 				When a new buoy is included in SPAMDA the following information, which can be obtained from NDBC, is requested:
-				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+				\begin{itemize}
 					\item \textit{Station ID}: An alphanumeric identifier that allows easy identification of the buoy.
@@
 
-				For clarification, an example is presented in Figure \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.
-
-				\begin{figure}[H]
+				For clarification, an example is presented in Fig. \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.
+
+				\begin{figure}[ht!]
 					\centering
@@
 
-				An example where three intermediate datasets have been created is presented in Figure \ref{fig:datasets}.
+				An example where three intermediate datasets have been created is presented in Fig.
+\ref{fig:datasets}.
 The two intermediate datasets of the buoy ID1 contain meteorological data of different years, and the intermediate dataset of the buoy ID2 contains meteorological data of two years.
@@
 
-				\begin{figure}[H]
+				\begin{figure}[ht!]
 					\centering
@@
 
-				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+				\begin{itemize}
 
@@
 
-					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+					\begin{itemize}
 						\item \textit{Normalize} \cite{WEKA_Filter_Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in the interval [0,1].
@@
 				 \item \textit{Instance}: All theses filters can be applied to the instances (hourly measurements of the buoy from NDBC) of the intermediate dataset.
-					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+					\begin{itemize}
 						\item \textit{RemoveDuplicates} \cite{WEKA_Filter_RemoveDuplicates}: With this filter, all duplicated instances are removed.
@@
 				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the intermediate dataset.
-					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+					\begin{itemize}
 						\item \textit{Replace missing values with next nearest hour}: The missing values of each attribute are replaced with the next nearest non missing value.
@@
 
-				Figure \ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one.
+				Fig.
+\ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one.
 The intermediate dataset 1 of the buoy ID2 has been also pre-processed.
@@
 
-				Figure \ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year.
+				Fig.
+\ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year.
 In this way, only the instances from both sources that are linked with arrows (highlighted in green colour) will be used in the creation of the final datasets.
@@
 
-				The reader can check in Appendix \ref{app:AppendixA} an example with a more complex case of the procedure.
-				\begin{figure}[H]
+				The reader can check in Appendix A \ref{app:AppendixA} an example with a more complex case of the procedure.
+				\begin{figure}[ht!]
 					\centering
@@
 Depending on the task, the datasets must be prepared and configured differently:
-				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+				\begin{itemize}
 					\item \textit{Classification}: The final datasets will be ready to use as input for ML classifiers, requiring a nominal output attribute, whose specific preparation is detailed in Section \ref{sec:FinalDatasets}.
@@
 				The following parameters can be specified for the matching process:
-				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+				\begin{itemize}
 
@@
 					\begin{equation}
-							F_e = 0.49 \cdot H^2_s \cdot T_e \;,
+							F_e = 0.49 \cdot H^2_s \cdot T_e,
 							\label{eq:fluxOfEnergy}
@@
 					where $F_e$ is measured in kilowatts per meter, $H_s$ is measured in meters and $T_e$ is measured in seconds.
-Note also that $F_e$ is defined in Equation \ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.
+Note also that $F_e$ is defined in Eq.
+\ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.
 
@@
 					\item \textit{Reanalysis data files}: In order to have a possible better description of the problem under study, more than one reanalysis variable can be considered as input.
-Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Figure \ref{fig:subGrid}) for each variable of reanalysis.
+Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Fig.
+\ref{fig:subGrid}) for each variable of reanalysis.
 
@@
 
-					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Figure \ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
+					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Fig.
+\ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
 In this way, the researcher can:
 
-						\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+						\begin{itemize}
 
@@
 									& & \cdot \cos(lon_0-lon_j) + \cos(lat_0) \\
-									& & \cdot \cos(lat_j))\;, \nonumber 
+									& & \cdot \cos(lat_j)), \nonumber 
 								\end{eqnarray}								
@@
 					\item \textit{Number of final datasets}: Depending on the number of nearest reanalysis nodes to consider, the number of final datasets to create and the content of them can be configured according to the following options:
-						\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+						\begin{itemize}
 
 							\item \textit{One (using weighted mean of the $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of the reanalysis data used (one per selected reanalysis data file).
-This weighted mean is obtained by SPAMDA and uses Equation \ref{eq:Haversine} to obtain the distance from each reanalysis node to the location of the buoy.
+This weighted mean is obtained by SPAMDA and uses Eq.
+\ref{eq:Haversine} to obtain the distance from each reanalysis node to the location of the buoy.
 Once the distances have been calculated they are inverted and normalised as follows:
@@
 
-							Considering as example the two nearest reanalysis nodes represented in Figure \ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.
+							Considering as example the two nearest reanalysis nodes represented in Fig.
+\ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.
 
@@
 
-							In this case, and considering as example the four closest reanalysis nodes (see Figure \ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.
-
-						\end{itemize}
-
-				\end{itemize}
-
-				Once the matching parameters have been described, for a better understanding of them, Figure \ref{fig:directMatching} presents an example of the matched information considering the data shown in Figure \ref{fig:matchingProcess} and using the following matching configuration\footnote{Note that the date is shown just for better understanding, but it will not be included in the final dataset.}:
-					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+							In this case, and considering as example the four closest reanalysis nodes (see Fig.
+\ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.
+
+						\end{itemize}
+
+				\end{itemize}
+
+				Once the matching parameters have been described, for a better understanding of them,  Fig.
+\ref{fig:directMatching} presents an example of the matched information considering the data shown in Fig.
+\ref{fig:matchingProcess} and using the following matching configuration\footnote{Note that the date is shown just for better understanding, but it will not be included in the final dataset.}:
+					\begin{itemize}
 						\item Variable WVHT as attribute to predict.
@@
 
-					\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+					\begin{itemize}
 
@@
 							\begin{equation}
-								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t})\;,
+								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t}),
 								\label{eq:noSynchronisingRD}
@@
 						where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to be predicted, $\mathbf{b}_t$ is the vector containing the selected NDBC variables and $\mathbf{r}_t$ is the vector containing the selected reanalysis variables.
-In this way and considering the matched information shown in Figure \ref{fig:directMatching}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
+In this way and considering the matched information shown in Fig.
+\ref{fig:directMatching}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
 
@@
 
- 						For better understanding, considering the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of one \textit{Regression} dataset is shown in Figure \ref{fig:regressionNoSync}.
+ 						For better understanding, considering the matched information shown in Fig.
+\ref{fig:directMatching}, an example of the creation of one \textit{Regression} dataset is shown in Fig.
+\ref{fig:regressionNoSync}.
 As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one).
 The options considered for the preparation of each final dataset are the following:
- 							\begin{itemize}[leftmargin=*,labelsep=5.8mm]
- 								\item Do not synchronise the reanalysis data (see Equation \ref{eq:noSynchronisingRD} for the relation between the inputs and the output).
+ 							\begin{itemize}
+ 								\item Do not synchronise the reanalysis data (see Eq. \ref{eq:noSynchronisingRD} for the relation between the inputs and the output).
  								\item A prediction horizon of $6$h.
@@
 
-							\begin{figure}[H]
+							\begin{figure}[ht!]
 								\centering
@@
 As a consequence, the last instance ($2017$/$12$/$31$ $18$:$00$) is lost and is not included in the final dataset.
-Besides, and because the reanalysis data has not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation \ref{eq:noSynchronisingRD}).
-
- 						Moreover, considering again the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of the same dataset but applying synchronisation (see Equation \ref{eq:synchronisingRD}) is shown in Figure \ref{fig:regressionSync}.
-
-							\begin{figure}[H]
+Besides, and because the reanalysis data has not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
+\ref{eq:noSynchronisingRD}).
+
+ 						Moreover, considering again the matched information shown in Fig.
+\ref{fig:directMatching}, an example of the creation of the same dataset but applying synchronisation (see Eq.
+\ref{eq:synchronisingRD}) is shown in Fig.
+\ref{fig:regressionSync}.
+
+							\begin{figure}[ht!]
 								\centering
@@
 But now, the values of the Pres variable are also moved backward one instance (due to the synchronisation).
-Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Equation \ref{eq:synchronisingRD}).
+Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Eq.
+\ref{eq:synchronisingRD}).
 
@@
 
-						Considering again the matched information shown in Figure \ref{fig:directMatching}, an example of the creation of a \textit{Classification} dataset is shown in Figure \ref{fig:prediction}.
+						Considering again the matched information shown in Fig.
+\ref{fig:directMatching}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
+\ref{fig:prediction}.
 The options considered for the preparation of the final dataset are the following:
-							\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+							\begin{itemize}
 								\item Do not synchronise the reanalysis data.
@@
 
-							\caption{Thresholds for the classification example represented in Figure \ref{fig:prediction}}
+							\caption{Thresholds for the classification example represented in Fig. \ref{fig:prediction}}
 							\label{tab:thresholds}
@@
 
-						\begin{figure}[H]
+						\begin{figure}[ht!]
 							\centering
@@
 Besides, and due to the $6$h prediction horizon, the last instance is lost ($2017$/$12$/$31$ $18$:$00$) and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up).
-As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation \ref{eq:noSynchronisingRD}).
+As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
+\ref{eq:noSynchronisingRD}).
 
@@
 				Finally, it is necessary to define the output configuration to create the final datasets:			
-				\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+				\begin{itemize}
 
@@
 
-						\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+						\begin{itemize}
 
@@
 
-				In this module (see Figure \ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones.
+				In this module (see Fig.
+\ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones.
 Besides, useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final time, latitude and longitude.
@@
 
-				An example where two reanalysis data files have been entered in SPAMDA is shown in Figure \ref{fig:manageReanalisys}.
-
-				\begin{figure}[H]
+				An example where two reanalysis data files have been entered in SPAMDA is shown in Fig. \ref{fig:manageReanalisys}.
+
+				\begin{figure}[ht!]
 					\centering
@@
 The data collected to perform this example is:
-		\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
+		\begin{enumerate}
 		\item The measurements obtained from 2013 to 2017 by the buoy with ID 46001, placed in the Gulf of Alaska, which are provided by NDBC as annual text files.
@@
 
-		\begin{figure}[H]
+		\begin{figure}[ht]
 			\centering
@@
 In this example, the following parameters were selected:
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+		\begin{itemize}
 			\item Attribute to predict: WVHT.
@@
 For this case study, the following configuration was applied:
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+		\begin{itemize}
 			\item Thresholds: see Table \ref{tab:thresholds}.
@@
 		At this point the final dataset would be created and stored in the computer of the researcher.
-Also, there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Figure \ref{fig:openigFinalDatasetWeka}.
-
-		\begin{figure}[H]
+Also, there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Fig.
+\ref{fig:openigFinalDatasetWeka}.
+
+		\begin{figure}[ht!]
 			\centering
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \authorcontributions{All authors contributed to this work.
-Design and implementation of the software, A.M.G.-O., J.F.C.
-and M.D.-M.; Test of the software, P.A.G.
-and C.H.-M.; Conceptualization, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Methodology, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Investigation, A.M.G.-O., J.F.C., M.D.-M., P.A.G., and C.H.-M.; Writing-original draft preparation, A.M.G.-O., J.F.C.
-and M.D.-M.; Funding acquisition, P.A.G.
-and C.H.-M.; Resources, P.A.G.
-and C.H.-M.; Supervision, P.A.G.
-and C.H.-M.}
-
-\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{We thank to NVIDIA Corporation for the transfer of computational resources for research works.
-The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{The authors declare no conflict of interest.
-The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
-
-\noindent 
-\begin{tabular}{@{}ll}
-$F_e$ & Flux of energy\\
-$H_s$ & Significant wave height\\
-$T_e$ & Wave energy period\\
-$p_0$ & Geographical location of the buoy\\
-$p_j$ & Geographical location of each reanalysis node\\
-$lat$ & Latitude of the point\\
-$lon$ & Longitude of the point\\
-$o_{t}$ & The attribute to be predicted at the time instant to study\\
-$\Delta t$ & The prediction horizon\\
-$\mathbf{b}_t$ & The vector containing the selected NDBC variables\\
-$\mathbf{r}_{t}$ & The vector containing the selected reanalysis variables\\
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{Managing incomplete data}\label{app:AppendixA}
-%\unskip
-%\subsection{}
+	\section*{Acknowledgments}
+
+		This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union.
+We also thank to NVIDIA Corporation for the transfer of computational resources for research works.
+
+		The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.
+
+	\section*{References}
+
+		\bibliography{bibfile}
+
+	\section*{Appendix A. Managing incomplete data}\label{app:AppendixA}
 
@@
 		The measurements collected by the buoys may be incomplete or recorded at a different time than the expected one, due to the weather conditions in which the buoys have to operate.
-To illustrate this casuistry, the following examples are shown in Figure \ref{fig:measurements}:
-		\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+To illustrate this casuistry, the following examples are shown in Fig.
+\ref{fig:measurements}:
+		\begin{itemize}
 			\item In the instance marked with a), the measurement of $17$:$50$ was collected at $17$:$45$, $5$ minutes earlier.
@@
 
-		Figure \ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Figure \ref{fig:measurements}.
+		Fig.
+\ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Fig \ref{fig:measurements}.
 Note that the instance marked with a) has not been informed by SPAMDA as an unexpected date because its time difference is less than $6$ minutes.
@@
 SPAMDA interprets these specific values and, after creating the intermediate dataset, the researchers can check if it contains missing values by visualising its statistical information or content.
-Remember that SPAMDA provides several filters for recovering missing data, which were described in Section \ref{sec:Datasets}.
-
-		\begin{figure}[H]
+Remember that SPAMDA provides several filters for recovering missing data, which were described in subsection \ref{sec:Datasets}.
+
+		\begin{figure}[ht!]
 			\centering
@@
 		SPAMDA takes into account this casuistry when carrying out the matching process.
-An example is given in Figure \ref{fig:matchingMeasurements}.
+An example is given in Fig.
+\ref{fig:matchingMeasurements}.
 As above-mentioned, the matching process is performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference.
@@
 
-%\section{}
-%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: internal bibliography
-%=====================================
-\reftitle{References}
-%\begin{thebibliography}{999}
-% Reference 1
-%\bibitem[Author1(year)]{ref-journal}
-%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-%\bibitem[Author2(year)]{ref-book}
-%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-%\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%=====================================
-% References, variant B: external bibliography
-%=====================================
-\externalbibliography{yes}
-\bibliography{bibfile}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2019-12-17 latex_manuscript/elsevier/3_Journal_of_Environmental_Management/spamda.tex
Initial commit.

--- 2019-12-17 latex_manuscript/elsevier/2_Environmental_Modelling_and_Software/spamda.tex
@@
 
-\journal{Environmental Modelling \& Software}
+\journal{Journal of Environmental Management}
 
@@
 
-		\title{SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
+		\title{SPAMDA: A software tool to manage meteorological data for environmental modelling}
 
-		\author[UCO]{Antonio M. G\'omez-Orellana}
+		\author[UCO]{Antonio M. G\'omez-Orellana\corref{AMGO}}
+		\cortext[AMGO]{Corresponding author.}
 		\ead{am.gomez@uco.es}
@@
 
-		\address[UCO]{Department of Computer Science and Numerical Analysis, University of Cordoba, C\'ordoba, Spain.}
+		\address[UCO]{Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}
 
 		\begin{abstract}
-			%Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
-Machine Learning (ML) techniques have become a valuable support in many research areas, improving the performance of traditional statistical procedures.
-However, such techniques require datasets containing information related to the topic under study, which are not always available in an appropriate format.
-Preparing these datasets implies a lot of time and effort by the researchers.
-This paper presents a software tool for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} (NDBC) and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project}.
-The datasets created by the software are ready to be used as inputs for ML techniques in prediction tasks (classification or regression).
-As the designed software is able to simplify the creation of the datasets and reduce the time needed for this task, it prevents researchers from performing repetitive technical work, allowing them to concentrate on the study of the meteorological aspects of the data.
-Therefore, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
 			Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
@@
 This paper presents a software for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-As the designed software is able to simplify the creation of the datasets (ready to be used as inputs for ML techniques in prediction tasks) and reduce the time needed for this task, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
+As the designed software is able to simplify the creation of the datasets (ready to be used as inputs for ML techniques in prediction tasks) and reduce the time needed for this task, researchers can benefit from it in order to achieve a more efficient exploitation and protection of environmental resources.
 		\end{abstract}
@@
 		\begin{keyword}
-			Meteorological data\sep Reanalysis data\sep  Pre-processing\sep Creating datasets\sep  Prediction tasks\sep Marine energy
+			Meteorological data\sep Reanalysis data\sep Pre-processing\sep Creating datasets\sep Environmental prediction\sep Marine energy
 		\end{keyword}
@@
 climate change, agriculture, energy, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind energy assessment \cite{DVORAK20101244}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: patterns in transmission of air pollution \cite{LU2019109377}, directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind energy assessment \cite{DVORAK20101244}, analysis of ecological typologies of large marine areas \cite{DELAHOZ201859}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{GUTIERREZ201369}, a tool based on \textit{Fatigue, Aerodynamics, Structures, and Turbulence} (FAST) code for performing design optimisation of offshore wind turbines is presented, which can analyse a massive group of cases by means of its parametric design capability.
-A design tool for architects based on \textit{Lighting and Thermal} (LT) method is used in \cite{BAKER2013156}, with the purpose of comparing and optimising design solutions in terms of energy and comfort performance.
+In \cite{PERERA2005291}, a modelling tool for water allocation and management is presented.
+An interactive system for multidisciplinary wind turbine site selection is developed in \cite{RAFIEE2018107}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and an interactive web application, called the \textit{Stream Hydrology And Rainfall Knowledge System} (SHARKS), for analysis of hydrologic and meteorological time series data that can be used for providing valuable insights to urban hydrologic processes is presented in \cite{BRENDEL2019}.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and an intuitive tool for stormwater runoff evaluation and management is developed in \cite{CHEN201645}.
 
@@
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed to transform this wave energy into electricity.
+In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
 WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them.
@@
 
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks \cite{srivastava2017large}, obtaining significant improvements in the performance of the results.
-ML methodologies can be used not only by experienced data and computer scientists but also by other researchers.
+		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{MARTINSSILVA2019109368, TORREGROSSA20181061}.
+ML methodologies can be used not only by experienced computer scientists but also by other researchers.
 For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms.
@@
 Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
-Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a merging or matching process have to be carried out by the researchers to manually create the datasets with the needed information.
+Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by the researchers to manually create the datasets with the needed information.
 Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
@@
 			\item The generation of datasets becomes a very easy and customizable task, by means of the selection of different input parameters.
-			\item It makes the researcher focus on oceanic and atmospheric studies, without having to worry about mechanical tasks.
+			\item The created datasets can be easily used by ML tools.
+			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks.
 			\item It provides information about the quality and quantity of the data.
-			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets.
+			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets, such as time and geographical coordinates conversion or missing values handling, among others.
 			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
 			\item It facilitates data management and well-organised storage of the datasets.
-			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy research.
+			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for environmental research.
 			\item It includes an user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
@@
 It may be also possible that the measurements have been recorded at a time different from the expected one.
-This aspect have to be taken into account when creating the datasets.
-This casuistry is explained in detail in Appendix A.
+These aspects have to be taken into account when creating the datasets.
+This casuistry is explained in detail in Appendix A \ref{app:AppendixA}.
 
@@
 					\begin{itemize}
-						\item \textit{Normalize} \cite{WEKA_Filter_Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in the interval [0,1].
-						\item \textit{Remove} \cite{WEKA_Filter_Remove}: It removes an attribute or a range of them.
-						\item \textit{RemoveByName} \cite{WEKA_Filter_RemoveByName}: It removes attributes based on a regular expression matched against their names.
-						\item \textit{ReplaceMissingValues} \cite{WEKA_Filter_ReplaceMissingValues}: For each attribute, all the missing values will be replaced by the average value of the attribute.
-						\item \textit{ReplaceMissingWithUserConstant} \cite{WEKA_Filter_ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
+						\item \textit{Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in the interval [0,1].
+						\item \textit{Remove}: It removes an attribute or a range of them.
+						\item \textit{RemoveByName}: It removes attributes based on a regular expression matched against their names.
+						\item \textit{ReplaceMissingValues}: For each attribute, all the missing values will be replaced by the average value of the attribute.
+						\item \textit{ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
 					\end{itemize}
@@
 					\begin{itemize}
-						\item \textit{RemoveDuplicates} \cite{WEKA_Filter_RemoveDuplicates}: With this filter, all duplicated instances are removed.
-						\item \textit{RemoveWithValues} \cite{WEKA_Filter_RemoveWithValues}: This filter removes all the instances that match the attribute and the value supplied by the user.
-						\item \textit{SubsetByExpression} \cite{WEKA_Filter_SubsetByExpression}: It removes all the instances which do not match a user-specified expression.
+						\item \textit{RemoveDuplicates}: With this filter, all duplicated instances are removed.
+						\item \textit{RemoveWithValues}: This filter removes all the instances that match the attribute and the value supplied by the user.
+						\item \textit{SubsetByExpression}: It removes all the instances which do not match a user-specified expression.
 					\end{itemize}
@@
 
-				In order to merge and format the data provided by the two sources of information described in Section \ref{sec:DataSources}, it is necessary to carry out a matching process.
+				In order to use the data provided by the two sources of information described in Section \ref{sec:DataSources}, it is necessary to carry out a data integration process, denominated as the matching process in this document, to merge and format such data.
 The matching procedure is performed using an intermediate or pre-processed dataset, which includes the measurements collected by a buoy from NDBC, and the needed reanalysis data files from NNRP.
@@
 
-		Although the software includes a user manual, this section will describe how the application works in a practical approach.
+		This section will describe how the application works in a practical approach.
 To do so, an example showing how to create a fully processed dataset (final dataset), starting from the raw data, will be described.
-The objective of this final dataset is to be used with ML algorithms to classify waves in the Gulf of Alaska depending on their height.
+The objective of this final dataset is to be used with ML algorithms for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
+In this case, the problem will be tackled as a multi-class classification problem, given that any continuous variable can be discretised in different classes.
+Such waves modelling can be applied with different purposes, energy flux range forecasting, missing buoy data reconstruction, or extreme significant wave heights detection, among others.
+
+		\subsection{Obtaining the final dataset}\label{sec:ObtainingFinalDataset}
+
 			The data collected to perform this example is:
@@
 This data is publicly available at the NDBC website.
-		\item Complementary information collected from reanalysis data containing air temperature, pressure and two components of wind speed measurements (South-North and West-East).
+			\item Complementary information collected from reanalysis data containing air temperature (air), pressure (pres) and two components of wind speed measurements: South-North (vwind) and West-East (uwind).
 This information will be collected from the four closest reanalysis nodes surrounding the geographical location of the buoy.
 This data is publicly available at the NNRP website and can be downloaded in NetCDF format.
+Concretely, the closest reanalysis nodes downloaded are $57.5$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150$ W, $55$ N $\times$ $147.5$ W and $55$ N $\times$ $150$ W.
+However, as will be seen later, only the information from the nearest node will be used in the data integration process.
 			\end{enumerate}
@@
 			After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.
-
 In Figure \ref{fig:main_view}, the main window is shown.
@@
 Here, using the buttons located at the bottom, it is possible to add, delete or consult any data from the different reanalysis files.
-After the information has been introduced in the application, this window can be closed and the user can go back to the main window to continue entering the information related to the buoy under study.
+Once the information has been introduced in the application, this window can be closed and the user can go back to the main window to continue entering the information related to the buoy under study.
 
@@
 				\includegraphics[width=0.89\textwidth]{figures/FigureManage_reanalysis_data.png}
-			\caption{\textit{Manage reanalysis data} window.}\label{fig:reanalysis}
+				\caption{\textit{Manage reanalysis data} window: downloaded files containing the four closest reanalysis nodes.}\label{fig:reanalysis}
 			\end{figure}
@@
 				\includegraphics[width=0.89\textwidth]{figures/FigureManage_buoys.png}
-			\caption{\textit{Buoys} tab.}\label{fig:manage_buoys}
+				\caption{\textit{Buoys} tab: buoy ID 46001.}\label{fig:manage_buoys}
 			\end{figure}
 
-		The researcher has to select \textcolor{blue}{\textit{Manage buoys data}} to open the window shown in Figure \ref{fig:manage_buoys}.
-In this tab, the researcher can consult, modify, add or delete different data related to the buoy.
+			After that, the researcher has to select \textcolor{blue}{\textit{Manage buoys data}} to open the window shown in Figure \ref{fig:manage_buoys}, where several tabs are available.
+In \textcolor{blue}{\textit{Buoys}} tab, the researcher can consult, modify, add or delete different data related to the buoy.
+
 			In order to enter such data, click on the \textcolor{blue}{\textit{New}} button, and then the window shown in Figure \ref{fig:new_buoy} pops up.
@@
 				\includegraphics[width=0.70\textwidth]{figures/FigureNew_buoy.png}
-			\caption{\em{Entering a new buoy} window.}\label{fig:new_buoy}
+				\caption{\textit{New buoy} window: information of the buoy ID 46001.}\label{fig:new_buoy}
 			\end{figure}
@@
 				\includegraphics[width=0.90\textwidth]{figures/FigureDatasets.png}
-			\caption{\textit{Datasets} tab.}\label{fig:show_datasets}
+				\caption{\textit{Datasets} tab: intermediate datasets of the buoy ID 46001.}\label{fig:show_datasets}
 			\end{figure}
 
-		To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the corresponding view (see Figure \ref{fig:show_datasets}).
+			To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab (see Figure \ref{fig:manage_buoys}) to switch to the corresponding view (see Figure \ref{fig:show_datasets}).
 In this view, the researcher can delete or consult a summary of each intermediate or pre-processed dataset by selecting it from the corresponding list.
@@
 				\includegraphics[width=0.70\textwidth]{figures/FigureNew_intermediate_dataset.png}
-			\caption{\textit{New intermediate dataset} view.}\label{fig:intermediate}
+				\caption{\textit{New intermediate dataset} view: creating the intermediate dataset with $5$ annual text files.}\label{fig:intermediate}
 			\end{figure}
 
-		Here the researcher can select the annual text files to be included in the intermediate dataset.
+			Here the researcher can select the annual text files to be included in the intermediate dataset, by clicking on the \textcolor{blue}{\textit{->}} and \textcolor{blue}{\textit{<-}} buttons.
 In this case, all the files introduced before, which correspond to the buoy under study, are selected.
 When the file selection is finished, \textcolor{blue}{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textcolor{blue}{\textit{Save}} button, the creation process starts, showing the status of the process during it.
-
-		After that, in order to prepare the intermediate dataset, the dataset is selected, and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}).
-In this tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
+After that, in order to prepare the intermediate dataset, the dataset is selected (see Figure \ref{fig:show_datasets}), and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}).
 
@@
 				\includegraphics[width=0.90\textwidth]{figures/FigurePreprocess.png}
-			\caption{\textit{Pre-process} tab.}\label{fig:preprocess_data}
+				\caption{\textit{Pre-process} tab: preprocessing the created intermediate dataset.}\label{fig:preprocess_data}
 			\end{figure}
 
+			In \textcolor{blue}{\textit{Pre-process}} tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
 Here the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter.
@@
 In order to achieve a more accurate description of the problem under study, a matching process can be carried out to merge the processed data from NDBC with the reanalysis data (also entered previously) from NNRP.
-The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.
 
@@
 				\includegraphics[width=0.90\textwidth]{figures/FigureMatching_configuration.png}
-			\caption{\textit{Matching configuration} tab.}\label{fig:matching_conf}
+				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files.}\label{fig:matching_conf}
 			\end{figure}
 
+			The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.
 In this view, the researcher can customise (or load) the parameters of the matching process according to their needs and select the prediction task  (described in Section \ref{sec:matching_conf}) that the final dataset will be used for.
@@
 				\item Reanalysis data: Air, pressure, u-wind and v-wind.
-			\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP.
-			\item Reanalysis nodes to consider: $1$.
+				\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription}).
+				\item Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).
 				\item Number of final datasets: In this example that option is disabled, because only one reanalysis node is considered.
@@
 				\includegraphics[width=0.90\textwidth]{figures/FigureFinal_datasets.png}
-			\caption{\textit{Final datasets} tab.}\label{fig:final_dataset}
+				\caption{\textit{Final datasets} tab: content of the final dataset created after data integration and discretisation of the output variable in $4$ classes.}\label{fig:final_dataset}
 			\end{figure}
@@
 Given that, in the previous window, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable.
-After this, the next step is to set the time horizon desired and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output, as explained in Section {\ref{sec:FinalDatasets}}.
+After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output, as explained in Section {\ref{sec:FinalDatasets}}.
 Then the researcher can click on the \textcolor{blue}{\textit{Update final dataset}} button to see the content shown in the bottom left corner.
-Finally, after checking that everything is correct, the last step would be to select the name (and path) of the dataset file and its output format and click on the \textcolor{blue}{\textit{Create final datasets}} button.
+Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textcolor{blue}{\textit{Create final datasets}} button.
 For this case study, the following configuration was applied:
@@
 
-		At this point the final dataset would be created and stored in the computer of the researcher.
-Also, there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Fig.
+			At this point, the final dataset would be created according to the tailored configuration and stored in the computer of the researcher, which already can apply the ML techniques to address the problem of wave classification.
+Concretely, the final dataset consists of $7302$ instances and whose distribution is represented in Table \ref{tab:datasetDistribution}.
+
+				\begin{table}[!h]
+
+					\caption{Distribution of instances of the final dataset}
+					\label{tab:datasetDistribution}
+					\centering
+
+					\begin{tabular}{cc}
+
+						\cline{1-2}
+
+						\textbf{Year}&\textbf{Number of instances}\\
+
+						\cline{1-2}
+
+						$2013$&$1460$\\
+						$2014$&$1460$\\
+						$2015$&$1460$\\
+						$2016$&$1464$\\
+						$2017$&$1458$\\
+
+						\cline{1-2}
+
+						&$7302$\\
+
+						\cline{1-2}
+
+					\end{tabular}
+
+				\end{table}
+
+		\subsection{Obtaining models with ML algorithms}
+
+			Now, it will be described how to obtain wave classification models using the final dataset previously created with SPAMDA.
+The modelling will be performed using WEKA as ML tool, which can be opened through SPAMDA, as shown in Fig.
 \ref{fig:openigFinalDatasetWeka}.
+Nevertheless, as mentioned above, the researcher can create the final dataset in CSV format in order to use other ML tools, such as KEEL, Python or R, among others.
 
@@
 
+			Since the final dataset is a time series of meteorological data (collected from $2013$ to $2017$), a hold-out scheme (60\% train / 40\% test) will be used.
+In this way, years from $2013$ to $2015$ will be used for the training phase ($4380$ instances) whereas $2016$ and $2017$ years will be used for the test phase ($2922$ instances).
+Previously to the learning phase, the attributes are normalised to avoid that some attributes could be more important than others because their values could be on a larger scale.
+
+			The classification algorithms that will be considered for the wave modelling are Logistic Regression \cite{hosmer2013applied}, C4.5 \cite{quinlan2014c4}, Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}, which will be applied with the default values of the parameters configured in WEKA.
+Given that Logistic Regression and C4.5 algorithms are deterministic, it is only necessary to launch one execution for each one.
+However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so in this case 30 executions for each one will be carried out.
+The results obtained are shown in Table \ref{tab:results}.
+
+				\begin{table}[!h]
+
+					\caption{Results (mean$\pm$SD) obtained by the algorithms}
+					\label{tab:results}
+					\footnotesize
+					\centering
+
+					\begin{tabular}{m{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
+
+						\cline{1-4}
+
+						\textbf{Algorithm}&\textbf{Accuracy (CCR)}&\textbf{Kappa}&\\[0.20cm]
+
+						\cline{1-4}
+
+						Logistic Regression & $59.0691$ & $0.44447$&\\[0.15cm]
+
+						\cellcolor{gray090}C4.5 & \cellcolor{gray090}$61.7385$ & \cellcolor{gray090}$0.47852$&\\[0.15cm]
+
+						Random Forest & $\textit{68.6516} \pm \textit{0.3083}$ & $\textit{0.57040} \pm \textit{0.0042}$&\\[0.15cm]
+
+						\cellcolor{gray090}Support Vector Machine &\cellcolor{gray090} $61.0016 \pm 0.0522$ &\cellcolor{gray090} $0.46770 \pm 0.0007$&\\[0.15cm]
+
+						Multilayer Perceptron & $\mathbf{69.7045} \pm \mathbf{1.3033}$ & $\mathbf{0.58576} \pm \mathbf{0.0178}$&\\[0.15cm]
+
+						\cline{1-4}
+
+					\end{tabular}
+
+				\end{table}
+
+			As can be seen, Random Forest and Multilayer Perceptron algorithms have achieved similar accuracy, but the performance of the latter is slightly better.
+
+			Although it has been performed a first classification approach, to illustrate the use of datasets created with SPAMDA for environmental modelling employing ML algorithms, both models have obtained a good performance despite the problem tackled is to predict wave height six hours in advance.
+
+		\subsection{Important remarks}
+
+			In this section, it has been described how to use SPAMDA to create a final dataset with the aim of classifying waves.
+However, using the same data described in Section \ref{sec:ObtainingFinalDataset} the researcher can quickly address other objectives or different studies by merely tailoring the matching configuration of the data integration process.
+For example, longer-term wave prediction can be addressed by changing the time horizon, wave modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
+
+			Furthermore, environmental modelling in other geographical location can be carried out by merely using other collected data.
+
+			As SPAMDA performs all data processing and management to create the datasets, it not only prevents the researchers from performing repetitive tasks but also prevents them from making possible errors.
+In this way, the researchers can focus on the studies they are carrying out.
+
 	\section{Conclusions}\label{sec:Conclusions}
@@
 These datasets will be ready to use as input for ML techniques in prediction tasks (classification or regression).
-As a result, the researchers will benefit from a great support when carrying out their oceanic and atmospheric studies, related to energy and environmental modelling.
+As a result, the researchers will benefit from significant support when carrying out environmental modelling related to atmospheric, oceanic or energy studies, among others.
 Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.
@@
 		The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.
+
+	\section*{Additional material}
+
+		The source code and the software tool are available at \url{https://github.com/ayrna}.
 

### 2019-12-17 latex_user_manual/appendix/GFDL_license.tex
Initial commit.

% This is set up to run with pdflatex.
%---------The file header---------------------------------------------
%\documentclass[a4paper,12pt]{book}

%\usepackage[english]{babel} %language selection
%\selectlanguage{english}

%\pagenumbering{arabic}

%\usepackage{hyperref}
%\hypersetup{colorlinks, 
%           citecolor=black,
%           filecolor=black,
%           linkcolor=black,
%           urlcolor=black,
%           bookmarksopen=true,
%           pdftex}

%\hfuzz = .6pt % avoid black boxes

%\begin{document}
%---------------------------------------------------------------------
\fancyhead[LO]{GNU Free Documentation License}
\fancyhead[RE]{GNU Free Documentation License}
%\chapter*{\rlap{GNU Free Documentation License}} \label{GFDL_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{GNU Free Documentation License}
\chapter{GNU Free Documentation License} \label{GFDL_LICENSE}
%\fancyhead[L]{GNU Free Documentation License}
%\label{label_fdl}

 \begin{center}

       Version 1.3, 3 November 2008

 Copyright \copyright{} 2000, 2001, 2002, 2007, 2008  Free Software Foundation, Inc.

 \bigskip

     \texttt{<https://fsf.org/>}

 \bigskip

 Everyone is permitted to copy and distribute verbatim copies
 of this license document, but changing it is not allowed.
\end{center}

\begin{center}
{\bf\large Preamble}
\end{center}

The purpose of this License is to make a manual, textbook, or other
functional and useful document ``free'' in the sense of freedom: to
assure everyone the effective freedom to copy and redistribute it,
with or without modifying it, either commercially or noncommercially.
Secondarily, this License preserves for the author and publisher a way
to get credit for their work, while not being considered responsible
for modifications made by others.

This License is a kind of ``copyleft'', which means that derivative
works of the document must themselves be free in the same sense.  It
complements the GNU General Public License, which is a copyleft
license designed for free software.

We have designed this License in order to use it for manuals for free
software, because free software needs free documentation: a free
program should come with manuals providing the same freedoms that the
software does.  But this License is not limited to software manuals;
it can be used for any textual work, regardless of subject matter or
whether it is published as a printed book.  We recommend this License
principally for works whose purpose is instruction or reference.

\begin{center}
{\Large\bf 1. APPLICABILITY AND DEFINITIONS\par}
\phantomsection
\addcontentsline{toc}{section}{1. APPLICABILITY AND DEFINITIONS}
\end{center}

This License applies to any manual or other work, in any medium, that
contains a notice placed by the copyright holder saying it can be
distributed under the terms of this License.  Such a notice grants a
world-wide, royalty-free license, unlimited in duration, to use that
work under the conditions stated herein.  The ``\textbf{Document}'', below,
refers to any such manual or work.  Any member of the public is a
licensee, and is addressed as ``\textbf{you}''.  You accept the license if you
copy, modify or distribute the work in a way requiring permission
under copyright law.

A ``\textbf{Modified Version}'' of the Document means any work containing the
Document or a portion of it, either copied verbatim, or with
modifications and/or translated into another language.

A ``\textbf{Secondary Section}'' is a named appendix or a front-matter section of
the Document that deals exclusively with the relationship of the
publishers or authors of the Document to the Document's overall subject
(or to related matters) and contains nothing that could fall directly
within that overall subject.  (Thus, if the Document is in part a
textbook of mathematics, a Secondary Section may not explain any
mathematics.)  The relationship could be a matter of historical
connection with the subject or with related matters, or of legal,
commercial, philosophical, ethical or political position regarding
them.

The ``\textbf{Invariant Sections}'' are certain Secondary Sections whose titles
are designated, as being those of Invariant Sections, in the notice
that says that the Document is released under this License.  If a
section does not fit the above definition of Secondary then it is not
allowed to be designated as Invariant.  The Document may contain zero
Invariant Sections.  If the Document does not identify any Invariant
Sections then there are none.

The ``\textbf{Cover Texts}'' are certain short passages of text that are listed,
as Front-Cover Texts or Back-Cover Texts, in the notice that says that
the Document is released under this License.  A Front-Cover Text may
be at most 5 words, and a Back-Cover Text may be at most 25 words.

A ``\textbf{Transparent}'' copy of the Document means a machine-readable copy,
represented in a format whose specification is available to the
general public, that is suitable for revising the document
straightforwardly with generic text editors or (for images composed of
pixels) generic paint programs or (for drawings) some widely available
drawing editor, and that is suitable for input to text formatters or
for automatic translation to a variety of formats suitable for input
to text formatters.  A copy made in an otherwise Transparent file
format whose markup, or absence of markup, has been arranged to thwart
or discourage subsequent modification by readers is not Transparent.
An image format is not Transparent if used for any substantial amount
of text.  A copy that is not ``Transparent'' is called ``\textbf{Opaque}''.

Examples of suitable formats for Transparent copies include plain
ASCII without markup, Texinfo input format, LaTeX input format, SGML
or XML using a publicly available DTD, and standard-conforming simple
HTML, PostScript or PDF designed for human modification.  Examples of
transparent image formats include PNG, XCF and JPG.  Opaque formats
include proprietary formats that can be read and edited only by
proprietary word processors, SGML or XML for which the DTD and/or
processing tools are not generally available, and the
machine-generated HTML, PostScript or PDF produced by some word
processors for output purposes only.

The ``\textbf{Title Page}'' means, for a printed book, the title page itself,
plus such following pages as are needed to hold, legibly, the material
this License requires to appear in the title page.  For works in
formats which do not have any title page as such, ``Title Page'' means
the text near the most prominent appearance of the work's title,
preceding the beginning of the body of the text.

The ``\textbf{publisher}'' means any person or entity that distributes
copies of the Document to the public.

A section ``\textbf{Entitled XYZ}'' means a named subunit of the Document whose
title either is precisely XYZ or contains XYZ in parentheses following
text that translates XYZ in another language.  (Here XYZ stands for a
specific section name mentioned below, such as ``\textbf{Acknowledgements}'',
``\textbf{Dedications}'', ``\textbf{Endorsements}'', or ``\textbf{History}''.)  
To ``\textbf{Preserve the Title}''
of such a section when you modify the Document means that it remains a
section ``Entitled XYZ'' according to this definition.

The Document may include Warranty Disclaimers next to the notice which
states that this License applies to the Document.  These Warranty
Disclaimers are considered to be included by reference in this
License, but only as regards disclaiming warranties: any other
implication that these Warranty Disclaimers may have is void and has
no effect on the meaning of this License.

\begin{center}
{\Large\bf 2. VERBATIM COPYING\par}
\phantomsection
\addcontentsline{toc}{section}{2. VERBATIM COPYING}
\end{center}

You may copy and distribute the Document in any medium, either
commercially or noncommercially, provided that this License, the
copyright notices, and the license notice saying this License applies
to the Document are reproduced in all copies, and that you add no other
conditions whatsoever to those of this License.  You may not use
technical measures to obstruct or control the reading or further
copying of the copies you make or distribute.  However, you may accept
compensation in exchange for copies.  If you distribute a large enough
number of copies you must also follow the conditions in section~3.

You may also lend copies, under the same conditions stated above, and
you may publicly display copies.

\begin{center}
{\Large\bf 3. COPYING IN QUANTITY\par}
\phantomsection
\addcontentsline{toc}{section}{3. COPYING IN QUANTITY}
\end{center}

If you publish printed copies (or copies in media that commonly have
printed covers) of the Document, numbering more than 100, and the
Document's license notice requires Cover Texts, you must enclose the
copies in covers that carry, clearly and legibly, all these Cover
Texts: Front-Cover Texts on the front cover, and Back-Cover Texts on
the back cover.  Both covers must also clearly and legibly identify
you as the publisher of these copies.  The front cover must present
the full title with all words of the title equally prominent and
visible.  You may add other material on the covers in addition.
Copying with changes limited to the covers, as long as they preserve
the title of the Document and satisfy these conditions, can be treated
as verbatim copying in other respects.

If the required texts for either cover are too voluminous to fit
legibly, you should put the first ones listed (as many as fit
reasonably) on the actual cover, and continue the rest onto adjacent
pages.

If you publish or distribute Opaque copies of the Document numbering
more than 100, you must either include a machine-readable Transparent
copy along with each Opaque copy, or state in or with each Opaque copy
a computer-network location from which the general network-using
public has access to download using public-standard network protocols
a complete Transparent copy of the Document, free of added material.
If you use the latter option, you must take reasonably prudent steps,
when you begin distribution of Opaque copies in quantity, to ensure
that this Transparent copy will remain thus accessible at the stated
location until at least one year after the last time you distribute an
Opaque copy (directly or through your agents or retailers) of that
edition to the public.

It is requested, but not required, that you contact the authors of the
Document well before redistributing any large number of copies, to give
them a chance to provide you with an updated version of the Document.

\begin{center}
{\Large\bf 4. MODIFICATIONS\par}
\phantomsection
\addcontentsline{toc}{section}{4. MODIFICATIONS}
\end{center}

You may copy and distribute a Modified Version of the Document under
the conditions of sections 2 and 3 above, provided that you release
the Modified Version under precisely this License, with the Modified
Version filling the role of the Document, thus licensing distribution
and modification of the Modified Version to whoever possesses a copy
of it.  In addition, you must do these things in the Modified Version:

\begin{itemize}
\item[A.] 
   Use in the Title Page (and on the covers, if any) a title distinct
   from that of the Document, and from those of previous versions
   (which should, if there were any, be listed in the History section
   of the Document).  You may use the same title as a previous version
   if the original publisher of that version gives permission.

\item[B.]
   List on the Title Page, as authors, one or more persons or entities
   responsible for authorship of the modifications in the Modified
   Version, together with at least five of the principal authors of the
   Document (all of its principal authors, if it has fewer than five),
   unless they release you from this requirement.

\item[C.]
   State on the Title page the name of the publisher of the
   Modified Version, as the publisher.

\item[D.]
   Preserve all the copyright notices of the Document.

\item[E.]
   Add an appropriate copyright notice for your modifications
   adjacent to the other copyright notices.

\item[F.]
   Include, immediately after the copyright notices, a license notice
   giving the public permission to use the Modified Version under the
   terms of this License, in the form shown in the Addendum below.

\item[G.]
   Preserve in that license notice the full lists of Invariant Sections
   and required Cover Texts given in the Document's license notice.

\item[H.]
   Include an unaltered copy of this License.

\item[I.]
   Preserve the section Entitled ``History'', Preserve its Title, and add
   to it an item stating at least the title, year, new authors, and
   publisher of the Modified Version as given on the Title Page.  If
   there is no section Entitled ``History'' in the Document, create one
   stating the title, year, authors, and publisher of the Document as
   given on its Title Page, then add an item describing the Modified
   Version as stated in the previous sentence.

\item[J.]
   Preserve the network location, if any, given in the Document for
   public access to a Transparent copy of the Document, and likewise
   the network locations given in the Document for previous versions
   it was based on.  These may be placed in the ``History'' section.
   You may omit a network location for a work that was published at
   least four years before the Document itself, or if the original
   publisher of the version it refers to gives permission.

\item[K.]
   For any section Entitled ``Acknowledgements'' or ``Dedications'',
   Preserve the Title of the section, and preserve in the section all
   the substance and tone of each of the contributor acknowledgements
   and/or dedications given therein.

\item[L.]
   Preserve all the Invariant Sections of the Document,
   unaltered in their text and in their titles.  Section numbers
   or the equivalent are not considered part of the section titles.

\item[M.]
   Delete any section Entitled ``Endorsements''.  Such a section
   may not be included in the Modified Version.

\item[N.]
   Do not retitle any existing section to be Entitled ``Endorsements''
   or to conflict in title with any Invariant Section.

\item[O.]
   Preserve any Warranty Disclaimers.
\end{itemize}

If the Modified Version includes new front-matter sections or
appendices that qualify as Secondary Sections and contain no material
copied from the Document, you may at your option designate some or all
of these sections as invariant.  To do this, add their titles to the
list of Invariant Sections in the Modified Version's license notice.
These titles must be distinct from any other section titles.

You may add a section Entitled ``Endorsements'', provided it contains
nothing but endorsements of your Modified Version by various
parties---for example, statements of peer review or that the text has
been approved by an organization as the authoritative definition of a
standard.

You may add a passage of up to five words as a Front-Cover Text, and a
passage of up to 25 words as a Back-Cover Text, to the end of the list
of Cover Texts in the Modified Version.  Only one passage of
Front-Cover Text and one of Back-Cover Text may be added by (or
through arrangements made by) any one entity.  If the Document already
includes a cover text for the same cover, previously added by you or
by arrangement made by the same entity you are acting on behalf of,
you may not add another; but you may replace the old one, on explicit
permission from the previous publisher that added the old one.

The author(s) and publisher(s) of the Document do not by this License
give permission to use their names for publicity for or to assert or
imply endorsement of any Modified Version.

\begin{center}
{\Large\bf 5. COMBINING DOCUMENTS\par}
\phantomsection
\addcontentsline{toc}{section}{5. COMBINING DOCUMENTS}
\end{center}

You may combine the Document with other documents released under this
License, under the terms defined in section~4 above for modified
versions, provided that you include in the combination all of the
Invariant Sections of all of the original documents, unmodified, and
list them all as Invariant Sections of your combined work in its
license notice, and that you preserve all their Warranty Disclaimers.

The combined work need only contain one copy of this License, and
multiple identical Invariant Sections may be replaced with a single
copy.  If there are multiple Invariant Sections with the same name but
different contents, make the title of each such section unique by
adding at the end of it, in parentheses, the name of the original
author or publisher of that section if known, or else a unique number.
Make the same adjustment to the section titles in the list of
Invariant Sections in the license notice of the combined work.

In the combination, you must combine any sections Entitled ``History''
in the various original documents, forming one section Entitled
``History''; likewise combine any sections Entitled ``Acknowledgements'',
and any sections Entitled ``Dedications''.  You must delete all sections
Entitled ``Endorsements''.

\begin{center}
{\Large\bf 6. COLLECTIONS OF DOCUMENTS\par}
\phantomsection
\addcontentsline{toc}{section}{6. COLLECTIONS OF DOCUMENTS}
\end{center}

You may make a collection consisting of the Document and other documents
released under this License, and replace the individual copies of this
License in the various documents with a single copy that is included in
the collection, provided that you follow the rules of this License for
verbatim copying of each of the documents in all other respects.

You may extract a single document from such a collection, and distribute
it individually under this License, provided you insert a copy of this
License into the extracted document, and follow this License in all
other respects regarding verbatim copying of that document.

\begin{center}
{\Large\bf 7. AGGREGATION WITH INDEPENDENT WORKS\par}
\phantomsection
\addcontentsline{toc}{section}{7. AGGREGATION WITH INDEPENDENT WORKS}
\end{center}

A compilation of the Document or its derivatives with other separate
and independent documents or works, in or on a volume of a storage or
distribution medium, is called an ``aggregate'' if the copyright
resulting from the compilation is not used to limit the legal rights
of the compilation's users beyond what the individual works permit.
When the Document is included in an aggregate, this License does not
apply to the other works in the aggregate which are not themselves
derivative works of the Document.

If the Cover Text requirement of section~3 is applicable to these
copies of the Document, then if the Document is less than one half of
the entire aggregate, the Document's Cover Texts may be placed on
covers that bracket the Document within the aggregate, or the
electronic equivalent of covers if the Document is in electronic form.
Otherwise they must appear on printed covers that bracket the whole
aggregate.

\begin{center}
{\Large\bf 8. TRANSLATION\par}
\phantomsection
\addcontentsline{toc}{section}{8. TRANSLATION}
\end{center}

Translation is considered a kind of modification, so you may
distribute translations of the Document under the terms of section~4.
Replacing Invariant Sections with translations requires special
permission from their copyright holders, but you may include
translations of some or all Invariant Sections in addition to the
original versions of these Invariant Sections.  You may include a
translation of this License, and all the license notices in the
Document, and any Warranty Disclaimers, provided that you also include
the original English version of this License and the original versions
of those notices and disclaimers.  In case of a disagreement between
the translation and the original version of this License or a notice
or disclaimer, the original version will prevail.

If a section in the Document is Entitled ``Acknowledgements'',
``Dedications'', or ``History'', the requirement (section~4) to Preserve
its Title (section~1) will typically require changing the actual
title.

\begin{center}
{\Large\bf 9. TERMINATION\par}
\phantomsection
\addcontentsline{toc}{section}{9. TERMINATION}
\end{center}

You may not copy, modify, sublicense, or distribute the Document
except as expressly provided under this License.  Any attempt
otherwise to copy, modify, sublicense, or distribute it is void, and
will automatically terminate your rights under this License.

However, if you cease all violation of this License, then your license
from a particular copyright holder is reinstated (a) provisionally,
unless and until the copyright holder explicitly and finally
terminates your license, and (b) permanently, if the copyright holder
fails to notify you of the violation by some reasonable means prior to
60 days after the cessation.

Moreover, your license from a particular copyright holder is
reinstated permanently if the copyright holder notifies you of the
violation by some reasonable means, this is the first time you have
received notice of violation of this License (for any work) from that
copyright holder, and you cure the violation prior to 30 days after
your receipt of the notice.

Termination of your rights under this section does not terminate the
licenses of parties who have received copies or rights from you under
this License.  If your rights have been terminated and not permanently
reinstated, receipt of a copy of some or all of the same material does
not give you any rights to use it.

\begin{center}
{\Large\bf 10. FUTURE REVISIONS OF THIS LICENSE\par}
\phantomsection
\addcontentsline{toc}{section}{10. FUTURE REVISIONS OF THIS LICENSE}
\end{center}

The Free Software Foundation may publish new, revised versions
of the GNU Free Documentation License from time to time.  Such new
versions will be similar in spirit to the present version, but may
differ in detail to address new problems or concerns.  See
\texttt{https://www.gnu.org/licenses/}.

Each version of the License is given a distinguishing version number.
If the Document specifies that a particular numbered version of this
License ``or any later version'' applies to it, you have the option of
following the terms and conditions either of that specified version or
of any later version that has been published (not as a draft) by the
Free Software Foundation.  If the Document does not specify a version
number of this License, you may choose any version ever published (not
as a draft) by the Free Software Foundation.  If the Document
specifies that a proxy can decide which future versions of this
License can be used, that proxy's public statement of acceptance of a
version permanently authorizes you to choose that version for the
Document.

\begin{center}
{\Large\bf 11. RELICENSING\par}
\phantomsection
\addcontentsline{toc}{section}{11. RELICENSING}
\end{center}

``Massive Multiauthor Collaboration Site'' (or ``MMC Site'') means any
World Wide Web server that publishes copyrightable works and also
provides prominent facilities for anybody to edit those works.  A
public wiki that anybody can edit is an example of such a server.  A
``Massive Multiauthor Collaboration'' (or ``MMC'') contained in the
site means any set of copyrightable works thus published on the MMC
site.

``CC-BY-SA'' means the Creative Commons Attribution-Share Alike 3.0
license published by Creative Commons Corporation, a not-for-profit
corporation with a principal place of business in San Francisco,
California, as well as future copyleft versions of that license
published by that same organization.

``Incorporate'' means to publish or republish a Document, in whole or
in part, as part of another Document.

An MMC is ``eligible for relicensing'' if it is licensed under this
License, and if all works that were first published under this License
somewhere other than this MMC, and subsequently incorporated in whole
or in part into the MMC, (1) had no cover texts or invariant sections,
and (2) were thus incorporated prior to November 1, 2008.

The operator of an MMC Site may republish an MMC contained in the site
under CC-BY-SA on the same site at any time before August 1, 2009,
provided the MMC is eligible for relicensing.

\begin{center}
{\Large\bf ADDENDUM: How to use this License for your documents\par}
\phantomsection
\addcontentsline{toc}{section}{ADDENDUM: How to use this License for your documents}
\end{center}

To use this License in a document you have written, include a copy of
the License in the document and put the following copyright and
license notices just after the title page:

\bigskip
\begin{quote}
    Copyright \copyright{}  YEAR  YOUR NAME.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.3
    or any later version published by the Free Software Foundation;
    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled ``GNU
    Free Documentation License''.
\end{quote}
\bigskip

If you have Invariant Sections, Front-Cover Texts and Back-Cover Texts,
replace the ``with \dots\ Texts.''\ line with this:

\bigskip
\begin{quote}
    with the Invariant Sections being LIST THEIR TITLES, with the
    Front-Cover Texts being LIST, and with the Back-Cover Texts being LIST.
\end{quote}
\bigskip

If you have Invariant Sections without Cover Texts, or some other
combination of the three, merge those two alternatives to suit the
situation.

If your document contains nontrivial examples of program code, we
recommend releasing these examples in parallel under your choice of
free software license, such as the GNU General Public License,
to permit their use in free software.

%---------------------------------------------------------------------
%\end{document}

### 2019-12-17 latex_user_manual/appendix/GPL3_SPAMDA_license.tex
Initial commit.

\fancyhead[LO]{License of SPAMDA}
\fancyhead[RE]{License of SPAMDA}
%\chapter{\rlap{License of SPAMDA}} \label{licenseGPLSPAMDA}
\chapter{License of SPAMDA} \label{licenseGPLSPAMDA}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{License of SPAMDA}
%\fancyhead[L]{License of SPAMDA}

\parindent=0pt

\small

SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets

Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
	\begin{quote}
	Authors: 
		\begin{quote}
			Antonio Manuel Gómez Orellana, Juan Carlos Fernández Caballero,
			Manuel Dorado Moreno, Pedro Antonio Gutiérrez Peña and 
			César Hervás Martínez.
		\end{quote}
	\end{quote}

This program is free software: you can redistribute it and/or modify it under the
terms of the GNU General Public License as published by the Free Software Foundation,
either version 3 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
See the GNU General Public License for more details.

A copy of the license is included in the appendix entitled "GNU GENERAL PUBLIC LICENSE" \ref{GPL_LICENSE}.
If not, see <http://www.gnu.org/licenses/>.

Additional permissions under GNU GPL version 3 section 7:

1. Redistributions of source code, with or without modification, must retain
the above full copyright notice as author attributions.

2. Redistributions in binary form and/or the use of the documentation,
with or without modification, must reproduce the above full copyright notice
as author attributions in the documentation and/or materials provided with
the distribution.

3. Modified versions of source code and/or documentation, as well as binary
distributions, must be marked in reasonable ways as different from the original version.

4. Neither name of copyright holders nor the names of its contributors may be used
to endorse or promote products derived from this software for publicity purposes
without specific prior written permission.

5. Redistribution and/or use of source code, binary format and documentation,
with or without modification, could require indemnification of licensors
and authors by anyone who conveys the material (or modified versions of it)
with contractual assumptions of liability to the recipient, for any liability
that these contractual assumptions directly impose on those licensors and authors.

SPAMDA uses some external libraries. You can see their respective notices about license,
copyright and disclaimer in the following appendices. For a more complete information about
such licenses, see the distributions provided by their authors:

-Library NetCDF Java, version 4.6.10

\hspace{0.75cm}Notice of license in the appendix entitled ``NetCDF-LICENSE'' \ref{NetCDF_LICENSE}.

-Library SLF4J, version 1.7.25

\hspace{0.75cm}Notice of license in the appendix entitled ``SLF4J-LICENSE`` \ref{SLF4J_LICENSE}.

-Library WEKA, version 3.8.1

\hspace{0.75cm}Notice of license in the appendix entitled ''WEKA-LICENSE'' \ref{WEKA_LICENSE}.

Contact information:

Juan Carlos Fernández Caballero, PhD.

email: jfcaballero@uco.es

Address: University of Cordoba, Department of Computer Science
and Numerical Analysis, AYRNA Research Group, Rabanales Campus,
Einstein Building, 3rd floor. Road Madrid-Cadiz, Km 396-A.
14071 - Cordoba (Spain).

### 2019-12-17 latex_user_manual/appendix/GPL3_license.tex
Initial commit.

%\documentclass[11pt]{article}

\fancyhead[LO]{GNU GENERAL PUBLIC LICENSE}
\fancyhead[RE]{GNU GENERAL PUBLIC LICENSE}
%\chapter*{\rlap{GNU GENERAL PUBLIC LICENSE}} \label{GPL_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{GNU GENERAL PUBLIC LICENSE}
\chapter{GNU GENERAL PUBLIC LICENSE} \label{GPL_LICENSE}

%\title{GNU GENERAL PUBLIC LICENSE}
%\date{Version 3, 29 June 2007}

%\begin{document}
%\maketitle

\begin{center}
{\parindent 0in

       Version 3, 29 June 2007

Copyright \copyright\  2007 Free Software Foundation, Inc. \texttt{https://fsf.org/}

\bigskip
Everyone is permitted to copy and distribute verbatim copies of this

license document, but changing it is not allowed.}

\end{center}

%\renewcommand{\abstractname}{Preamble}
\begin{center}
{\bf\large Preamble}
\end{center}

%\begin{abstract}
The GNU General Public License is a free, copyleft license for
software and other kinds of works.

The licenses for most software and other practical works are designed
to take away your freedom to share and change the works.  By contrast,
the GNU General Public License is intended to guarantee your freedom to
share and change all versions of a program--to make sure it remains free
software for all its users.  We, the Free Software Foundation, use the
GNU General Public License for most of our software; it applies also to
any other work released this way by its authors.  You can apply it to
your programs, too.

When we speak of free software, we are referring to freedom, not
price.  Our General Public Licenses are designed to make sure that you
have the freedom to distribute copies of free software (and charge for
them if you wish), that you receive source code or can get it if you
want it, that you can change the software or use pieces of it in new
free programs, and that you know you can do these things.

To protect your rights, we need to prevent others from denying you
these rights or asking you to surrender the rights.  Therefore, you have
certain responsibilities if you distribute copies of the software, or if
you modify it: responsibilities to respect the freedom of others.

For example, if you distribute copies of such a program, whether
gratis or for a fee, you must pass on to the recipients the same
freedoms that you received.  You must make sure that they, too, receive
or can get the source code.  And you must show them these terms so they
know their rights.

Developers that use the GNU GPL protect your rights with two steps:
(1) assert copyright on the software, and (2) offer you this License
giving you legal permission to copy, distribute and/or modify it.

For the developers' and authors' protection, the GPL clearly explains
that there is no warranty for this free software.  For both users' and
authors' sake, the GPL requires that modified versions be marked as
changed, so that their problems will not be attributed erroneously to
authors of previous versions.

Some devices are designed to deny users access to install or run
modified versions of the software inside them, although the manufacturer
can do so.  This is fundamentally incompatible with the aim of
protecting users' freedom to change the software.  The systematic
pattern of such abuse occurs in the area of products for individuals to
use, which is precisely where it is most unacceptable.  Therefore, we
have designed this version of the GPL to prohibit the practice for those
products.  If such problems arise substantially in other domains, we
stand ready to extend this provision to those domains in future versions
of the GPL, as needed to protect the freedom of users.

Finally, every program is threatened constantly by software patents.
States should not allow patents to restrict development and use of
software on general-purpose computers, but in those that do, we wish to
avoid the special danger that patents applied to a free program could
make it effectively proprietary.  To prevent this, the GPL assures that
patents cannot be used to render the program non-free.

The precise terms and conditions for copying, distribution and
modification follow.
%\end{abstract}

\begin{center}
{\Large \sc Terms and Conditions}
\end{center}

\begin{enumerate}

\addtocounter{enumi}{-1}

\item Definitions.

``This License'' refers to version 3 of the GNU General Public License.

``Copyright'' also means copyright-like laws that apply to other kinds of
works, such as semiconductor masks.

``The Program'' refers to any copyrightable work licensed under this
License.  Each licensee is addressed as ``you''.  ``Licensees'' and
``recipients'' may be individuals or organizations.

To ``modify'' a work means to copy from or adapt all or part of the work
in a fashion requiring copyright permission, other than the making of an
exact copy.  The resulting work is called a ``modified version'' of the
earlier work or a work ``based on'' the earlier work.

A ``covered work'' means either the unmodified Program or a work based
on the Program.

To ``propagate'' a work means to do anything with it that, without
permission, would make you directly or secondarily liable for
infringement under applicable copyright law, except executing it on a
computer or modifying a private copy.  Propagation includes copying,
distribution (with or without modification), making available to the
public, and in some countries other activities as well.

To ``convey'' a work means any kind of propagation that enables other
parties to make or receive copies.  Mere interaction with a user through
a computer network, with no transfer of a copy, is not conveying.

An interactive user interface displays ``Appropriate Legal Notices''
to the extent that it includes a convenient and prominently visible
feature that (1) displays an appropriate copyright notice, and (2)
tells the user that there is no warranty for the work (except to the
extent that warranties are provided), that licensees may convey the
work under this License, and how to view a copy of this License.  If
the interface presents a list of user commands or options, such as a
menu, a prominent item in the list meets this criterion.

\item Source Code.

The ``source code'' for a work means the preferred form of the work
for making modifications to it.  ``Object code'' means any non-source
form of a work.

A ``Standard Interface'' means an interface that either is an official
standard defined by a recognized standards body, or, in the case of
interfaces specified for a particular programming language, one that
is widely used among developers working in that language.

The ``System Libraries'' of an executable work include anything, other
than the work as a whole, that (a) is included in the normal form of
packaging a Major Component, but which is not part of that Major
Component, and (b) serves only to enable use of the work with that
Major Component, or to implement a Standard Interface for which an
implementation is available to the public in source code form.  A
``Major Component'', in this context, means a major essential component
(kernel, window system, and so on) of the specific operating system
(if any) on which the executable work runs, or a compiler used to
produce the work, or an object code interpreter used to run it.

The ``Corresponding Source'' for a work in object code form means all
the source code needed to generate, install, and (for an executable
work) run the object code and to modify the work, including scripts to
control those activities.  However, it does not include the work's
System Libraries, or general-purpose tools or generally available free
programs which are used unmodified in performing those activities but
which are not part of the work.  For example, Corresponding Source
includes interface definition files associated with source files for
the work, and the source code for shared libraries and dynamically
linked subprograms that the work is specifically designed to require,
such as by intimate data communication or control flow between those
subprograms and other parts of the work.

The Corresponding Source need not include anything that users
can regenerate automatically from other parts of the Corresponding
Source.

The Corresponding Source for a work in source code form is that
same work.

\item Basic Permissions.

All rights granted under this License are granted for the term of
copyright on the Program, and are irrevocable provided the stated
conditions are met.  This License explicitly affirms your unlimited
permission to run the unmodified Program.  The output from running a
covered work is covered by this License only if the output, given its
content, constitutes a covered work.  This License acknowledges your
rights of fair use or other equivalent, as provided by copyright law.

You may make, run and propagate covered works that you do not
convey, without conditions so long as your license otherwise remains
in force.  You may convey covered works to others for the sole purpose
of having them make modifications exclusively for you, or provide you
with facilities for running those works, provided that you comply with
the terms of this License in conveying all material for which you do
not control copyright.  Those thus making or running the covered works
for you must do so exclusively on your behalf, under your direction
and control, on terms that prohibit them from making any copies of
your copyrighted material outside their relationship with you.

Conveying under any other circumstances is permitted solely under
the conditions stated below.  Sublicensing is not allowed; section 10
makes it unnecessary.

\item Protecting Users' Legal Rights From Anti-Circumvention Law.

No covered work shall be deemed part of an effective technological
measure under any applicable law fulfilling obligations under article
11 of the WIPO copyright treaty adopted on 20 December 1996, or
similar laws prohibiting or restricting circumvention of such
measures.

When you convey a covered work, you waive any legal power to forbid
circumvention of technological measures to the extent such circumvention
is effected by exercising rights under this License with respect to
the covered work, and you disclaim any intention to limit operation or
modification of the work as a means of enforcing, against the work's
users, your or third parties' legal rights to forbid circumvention of
technological measures.

\item Conveying Verbatim Copies.

You may convey verbatim copies of the Program's source code as you
receive it, in any medium, provided that you conspicuously and
appropriately publish on each copy an appropriate copyright notice;
keep intact all notices stating that this License and any
non-permissive terms added in accord with section 7 apply to the code;
keep intact all notices of the absence of any warranty; and give all
recipients a copy of this License along with the Program.

You may charge any price or no price for each copy that you convey,
and you may offer support or warranty protection for a fee.

\item Conveying Modified Source Versions.

You may convey a work based on the Program, or the modifications to
produce it from the Program, in the form of source code under the
terms of section 4, provided that you also meet all of these conditions:
  \begin{enumerate}
  \item The work must carry prominent notices stating that you modified
  it, and giving a relevant date.

  \item The work must carry prominent notices stating that it is
  released under this License and any conditions added under section
  7.  This requirement modifies the requirement in section 4 to
  ``keep intact all notices''.

  \item You must license the entire work, as a whole, under this
  License to anyone who comes into possession of a copy.  This
  License will therefore apply, along with any applicable section 7
  additional terms, to the whole of the work, and all its parts,
  regardless of how they are packaged.  This License gives no
  permission to license the work in any other way, but it does not
  invalidate such permission if you have separately received it.

  \item If the work has interactive user interfaces, each must display
  Appropriate Legal Notices; however, if the Program has interactive
  interfaces that do not display Appropriate Legal Notices, your
  work need not make them do so.
\end{enumerate}
A compilation of a covered work with other separate and independent
works, which are not by their nature extensions of the covered work,
and which are not combined with it such as to form a larger program,
in or on a volume of a storage or distribution medium, is called an
``aggregate'' if the compilation and its resulting copyright are not
used to limit the access or legal rights of the compilation's users
beyond what the individual works permit.  Inclusion of a covered work
in an aggregate does not cause this License to apply to the other
parts of the aggregate.

\item Conveying Non-Source Forms.

You may convey a covered work in object code form under the terms
of sections 4 and 5, provided that you also convey the
machine-readable Corresponding Source under the terms of this License,
in one of these ways:
  \begin{enumerate}
  \item Convey the object code in, or embodied in, a physical product
  (including a physical distribution medium), accompanied by the
  Corresponding Source fixed on a durable physical medium
  customarily used for software interchange.

  \item Convey the object code in, or embodied in, a physical product
  (including a physical distribution medium), accompanied by a
  written offer, valid for at least three years and valid for as
  long as you offer spare parts or customer support for that product
  model, to give anyone who possesses the object code either (1) a
  copy of the Corresponding Source for all the software in the
  product that is covered by this License, on a durable physical
  medium customarily used for software interchange, for a price no
  more than your reasonable cost of physically performing this
  conveying of source, or (2) access to copy the
  Corresponding Source from a network server at no charge.

  \item Convey individual copies of the object code with a copy of the
  written offer to provide the Corresponding Source.  This
  alternative is allowed only occasionally and noncommercially, and
  only if you received the object code with such an offer, in accord
  with subsection 6b.

  \item Convey the object code by offering access from a designated
  place (gratis or for a charge), and offer equivalent access to the
  Corresponding Source in the same way through the same place at no
  further charge.  You need not require recipients to copy the
  Corresponding Source along with the object code.  If the place to
  copy the object code is a network server, the Corresponding Source
  may be on a different server (operated by you or a third party)
  that supports equivalent copying facilities, provided you maintain
  clear directions next to the object code saying where to find the
  Corresponding Source.  Regardless of what server hosts the
  Corresponding Source, you remain obligated to ensure that it is
  available for as long as needed to satisfy these requirements.

  \item Convey the object code using peer-to-peer transmission, provided
  you inform other peers where the object code and Corresponding
  Source of the work are being offered to the general public at no
  charge under subsection 6d.
  \end{enumerate}

A separable portion of the object code, whose source code is excluded
from the Corresponding Source as a System Library, need not be
included in conveying the object code work.

A ``User Product'' is either (1) a ``consumer product'', which means any
tangible personal property which is normally used for personal, family,
or household purposes, or (2) anything designed or sold for incorporation
into a dwelling.  In determining whether a product is a consumer product,
doubtful cases shall be resolved in favor of coverage.  For a particular
product received by a particular user, ``normally used'' refers to a
typical or common use of that class of product, regardless of the status
of the particular user or of the way in which the particular user
actually uses, or expects or is expected to use, the product.  A product
is a consumer product regardless of whether the product has substantial
commercial, industrial or non-consumer uses, unless such uses represent
the only significant mode of use of the product.

``Installation Information'' for a User Product means any methods,
procedures, authorization keys, or other information required to install
and execute modified versions of a covered work in that User Product from
a modified version of its Corresponding Source.  The information must
suffice to ensure that the continued functioning of the modified object
code is in no case prevented or interfered with solely because
modification has been made.

If you convey an object code work under this section in, or with, or
specifically for use in, a User Product, and the conveying occurs as
part of a transaction in which the right of possession and use of the
User Product is transferred to the recipient in perpetuity or for a
fixed term (regardless of how the transaction is characterized), the
Corresponding Source conveyed under this section must be accompanied
by the Installation Information.  But this requirement does not apply
if neither you nor any third party retains the ability to install
modified object code on the User Product (for example, the work has
been installed in ROM).

The requirement to provide Installation Information does not include a
requirement to continue to provide support service, warranty, or updates
for a work that has been modified or installed by the recipient, or for
the User Product in which it has been modified or installed.  Access to a
network may be denied when the modification itself materially and
adversely affects the operation of the network or violates the rules and
protocols for communication across the network.

Corresponding Source conveyed, and Installation Information provided,
in accord with this section must be in a format that is publicly
documented (and with an implementation available to the public in
source code form), and must require no special password or key for
unpacking, reading or copying.

\item Additional Terms.

``Additional permissions'' are terms that supplement the terms of this
License by making exceptions from one or more of its conditions.
Additional permissions that are applicable to the entire Program shall
be treated as though they were included in this License, to the extent
that they are valid under applicable law.  If additional permissions
apply only to part of the Program, that part may be used separately
under those permissions, but the entire Program remains governed by
this License without regard to the additional permissions.

When you convey a copy of a covered work, you may at your option
remove any additional permissions from that copy, or from any part of
it.  (Additional permissions may be written to require their own
removal in certain cases when you modify the work.)  You may place
additional permissions on material, added by you to a covered work,
for which you have or can give appropriate copyright permission.

Notwithstanding any other provision of this License, for material you
add to a covered work, you may (if authorized by the copyright holders of
that material) supplement the terms of this License with terms:
  \begin{enumerate}
  \item Disclaiming warranty or limiting liability differently from the
  terms of sections 15 and 16 of this License; or

  \item Requiring preservation of specified reasonable legal notices or
  author attributions in that material or in the Appropriate Legal
  Notices displayed by works containing it; or

  \item Prohibiting misrepresentation of the origin of that material, or
  requiring that modified versions of such material be marked in
  reasonable ways as different from the original version; or

  \item Limiting the use for publicity purposes of names of licensors or
  authors of the material; or

  \item Declining to grant rights under trademark law for use of some
  trade names, trademarks, or service marks; or

  \item Requiring indemnification of licensors and authors of that
  material by anyone who conveys the material (or modified versions of
  it) with contractual assumptions of liability to the recipient, for
  any liability that these contractual assumptions directly impose on
  those licensors and authors.
  \end{enumerate}

All other non-permissive additional terms are considered ``further
restrictions'' within the meaning of section 10.  If the Program as you
received it, or any part of it, contains a notice stating that it is
governed by this License along with a term that is a further
restriction, you may remove that term.  If a license document contains
a further restriction but permits relicensing or conveying under this
License, you may add to a covered work material governed by the terms
of that license document, provided that the further restriction does
not survive such relicensing or conveying.

If you add terms to a covered work in accord with this section, you
must place, in the relevant source files, a statement of the
additional terms that apply to those files, or a notice indicating
where to find the applicable terms.

Additional terms, permissive or non-permissive, may be stated in the
form of a separately written license, or stated as exceptions;
the above requirements apply either way.

\item Termination.

You may not propagate or modify a covered work except as expressly
provided under this License.  Any attempt otherwise to propagate or
modify it is void, and will automatically terminate your rights under
this License (including any patent licenses granted under the third
paragraph of section 11).

However, if you cease all violation of this License, then your
license from a particular copyright holder is reinstated (a)
provisionally, unless and until the copyright holder explicitly and
finally terminates your license, and (b) permanently, if the copyright
holder fails to notify you of the violation by some reasonable means
prior to 60 days after the cessation.

Moreover, your license from a particular copyright holder is
reinstated permanently if the copyright holder notifies you of the
violation by some reasonable means, this is the first time you have
received notice of violation of this License (for any work) from that
copyright holder, and you cure the violation prior to 30 days after
your receipt of the notice.

Termination of your rights under this section does not terminate the
licenses of parties who have received copies or rights from you under
this License.  If your rights have been terminated and not permanently
reinstated, you do not qualify to receive new licenses for the same
material under section 10.

\item Acceptance Not Required for Having Copies.

You are not required to accept this License in order to receive or
run a copy of the Program.  Ancillary propagation of a covered work
occurring solely as a consequence of using peer-to-peer transmission
to receive a copy likewise does not require acceptance.  However,
nothing other than this License grants you permission to propagate or
modify any covered work.  These actions infringe copyright if you do
not accept this License.  Therefore, by modifying or propagating a
covered work, you indicate your acceptance of this License to do so.

\item Automatic Licensing of Downstream Recipients.

Each time you convey a covered work, the recipient automatically
receives a license from the original licensors, to run, modify and
propagate that work, subject to this License.  You are not responsible
for enforcing compliance by third parties with this License.

An ``entity transaction'' is a transaction transferring control of an
organization, or substantially all assets of one, or subdividing an
organization, or merging organizations.  If propagation of a covered
work results from an entity transaction, each party to that
transaction who receives a copy of the work also receives whatever
licenses to the work the party's predecessor in interest had or could
give under the previous paragraph, plus a right to possession of the
Corresponding Source of the work from the predecessor in interest, if
the predecessor has it or can get it with reasonable efforts.

You may not impose any further restrictions on the exercise of the
rights granted or affirmed under this License.  For example, you may
not impose a license fee, royalty, or other charge for exercise of
rights granted under this License, and you may not initiate litigation
(including a cross-claim or counterclaim in a lawsuit) alleging that
any patent claim is infringed by making, using, selling, offering for
sale, or importing the Program or any portion of it.

\item Patents.

A ``contributor'' is a copyright holder who authorizes use under this
License of the Program or a work on which the Program is based.  The
work thus licensed is called the contributor's ``contributor version''.

A contributor's ``essential patent claims'' are all patent claims
owned or controlled by the contributor, whether already acquired or
hereafter acquired, that would be infringed by some manner, permitted
by this License, of making, using, or selling its contributor version,
but do not include claims that would be infringed only as a
consequence of further modification of the contributor version.  For
purposes of this definition, ``control'' includes the right to grant
patent sublicenses in a manner consistent with the requirements of
this License.

Each contributor grants you a non-exclusive, worldwide, royalty-free
patent license under the contributor's essential patent claims, to
make, use, sell, offer for sale, import and otherwise run, modify and
propagate the contents of its contributor version.

In the following three paragraphs, a ``patent license'' is any express
agreement or commitment, however denominated, not to enforce a patent
(such as an express permission to practice a patent or covenant not to
sue for patent infringement).  To ``grant'' such a patent license to a
party means to make such an agreement or commitment not to enforce a
patent against the party.

If you convey a covered work, knowingly relying on a patent license,
and the Corresponding Source of the work is not available for anyone
to copy, free of charge and under the terms of this License, through a
publicly available network server or other readily accessible means,
then you must either (1) cause the Corresponding Source to be so
available, or (2) arrange to deprive yourself of the benefit of the
patent license for this particular work, or (3) arrange, in a manner
consistent with the requirements of this License, to extend the patent
license to downstream recipients.  ``Knowingly relying'' means you have
actual knowledge that, but for the patent license, your conveying the
covered work in a country, or your recipient's use of the covered work
in a country, would infringe one or more identifiable patents in that
country that you have reason to believe are valid.

If, pursuant to or in connection with a single transaction or
arrangement, you convey, or propagate by procuring conveyance of, a
covered work, and grant a patent license to some of the parties
receiving the covered work authorizing them to use, propagate, modify
or convey a specific copy of the covered work, then the patent license
you grant is automatically extended to all recipients of the covered
work and works based on it.

A patent license is ``discriminatory'' if it does not include within
the scope of its coverage, prohibits the exercise of, or is
conditioned on the non-exercise of one or more of the rights that are
specifically granted under this License.  You may not convey a covered
work if you are a party to an arrangement with a third party that is
in the business of distributing software, under which you make payment
to the third party based on the extent of your activity of conveying
the work, and under which the third party grants, to any of the
parties who would receive the covered work from you, a discriminatory
patent license (a) in connection with copies of the covered work
conveyed by you (or copies made from those copies), or (b) primarily
for and in connection with specific products or compilations that
contain the covered work, unless you entered into that arrangement,
or that patent license was granted, prior to 28 March 2007.

Nothing in this License shall be construed as excluding or limiting
any implied license or other defenses to infringement that may
otherwise be available to you under applicable patent law.

\item No Surrender of Others' Freedom.

If conditions are imposed on you (whether by court order, agreement or
otherwise) that contradict the conditions of this License, they do not
excuse you from the conditions of this License.  If you cannot convey a
covered work so as to satisfy simultaneously your obligations under this
License and any other pertinent obligations, then as a consequence you may
not convey it at all.  For example, if you agree to terms that obligate you
to collect a royalty for further conveying from those to whom you convey
the Program, the only way you could satisfy both those terms and this
License would be to refrain entirely from conveying the Program.

\item Use with the GNU Affero General Public License.

Notwithstanding any other provision of this License, you have
permission to link or combine any covered work with a work licensed
under version 3 of the GNU Affero General Public License into a single
combined work, and to convey the resulting work.  The terms of this
License will continue to apply to the part which is the covered work,
but the special requirements of the GNU Affero General Public License,
section 13, concerning interaction through a network will apply to the
combination as such.

\item Revised Versions of this License.

The Free Software Foundation may publish revised and/or new versions of
the GNU General Public License from time to time.  Such new versions will
be similar in spirit to the present version, but may differ in detail to
address new problems or concerns.

Each version is given a distinguishing version number.  If the
Program specifies that a certain numbered version of the GNU General
Public License ``or any later version'' applies to it, you have the
option of following the terms and conditions either of that numbered
version or of any later version published by the Free Software
Foundation.  If the Program does not specify a version number of the
GNU General Public License, you may choose any version ever published
by the Free Software Foundation.

If the Program specifies that a proxy can decide which future
versions of the GNU General Public License can be used, that proxy's
public statement of acceptance of a version permanently authorizes you
to choose that version for the Program.

Later license versions may give you additional or different
permissions.  However, no additional obligations are imposed on any
author or copyright holder as a result of your choosing to follow a
later version.

\item Disclaimer of Warranty.

\begin{sloppypar}
 THERE IS NO WARRANTY FOR THE PROGRAM, TO THE EXTENT PERMITTED BY
 APPLICABLE LAW.  EXCEPT WHEN OTHERWISE STATED IN WRITING THE
 COPYRIGHT HOLDERS AND/OR OTHER PARTIES PROVIDE THE PROGRAM ``AS IS''
 WITHOUT WARRANTY OF ANY KIND, EITHER EXPRESSED OR IMPLIED,
 INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF
 MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.  THE ENTIRE
 RISK AS TO THE QUALITY AND PERFORMANCE OF THE PROGRAM IS WITH YOU.
 SHOULD THE PROGRAM PROVE DEFECTIVE, YOU ASSUME THE COST OF ALL
 NECESSARY SERVICING, REPAIR OR CORRECTION.
\end{sloppypar}

\item Limitation of Liability.

 IN NO EVENT UNLESS REQUIRED BY APPLICABLE LAW OR AGREED TO IN
 WRITING WILL ANY COPYRIGHT HOLDER, OR ANY OTHER PARTY WHO MODIFIES
 AND/OR CONVEYS THE PROGRAM AS PERMITTED ABOVE, BE LIABLE TO YOU FOR
 DAMAGES, INCLUDING ANY GENERAL, SPECIAL, INCIDENTAL OR CONSEQUENTIAL
 DAMAGES ARISING OUT OF THE USE OR INABILITY TO USE THE PROGRAM
 (INCLUDING BUT NOT LIMITED TO LOSS OF DATA OR DATA BEING RENDERED
 INACCURATE OR LOSSES SUSTAINED BY YOU OR THIRD PARTIES OR A FAILURE
 OF THE PROGRAM TO OPERATE WITH ANY OTHER PROGRAMS), EVEN IF SUCH
 HOLDER OR OTHER PARTY HAS BEEN ADVISED OF THE POSSIBILITY OF SUCH
 DAMAGES.

\item Interpretation of Sections 15 and 16.

If the disclaimer of warranty and limitation of liability provided
above cannot be given local legal effect according to their terms,
reviewing courts shall apply local law that most closely approximates
an absolute waiver of all civil liability in connection with the
Program, unless a warranty or assumption of liability accompanies a
copy of the Program in return for a fee.

\begin{center}
{\Large\sc End of Terms and Conditions}

\bigskip
How to Apply These Terms to Your New Programs
\end{center}

If you develop a new program, and you want it to be of the greatest
possible use to the public, the best way to achieve this is to make it
free software which everyone can redistribute and change under these terms.

To do so, attach the following notices to the program.  It is safest
to attach them to the start of each source file to most effectively
state the exclusion of warranty; and each file should have at least
the ``copyright'' line and a pointer to where the full notice is found.

{\footnotesize
\begin{verbatim}
<one line to give the program's name and a brief idea of what it does.>

Copyright (C) <textyear>  <name of author>

This program is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
(at your option) any later version.

This program is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with this program.  If not, see <https://www.gnu.org/licenses/>.
\end{verbatim}
}

Also add information on how to contact you by electronic and paper mail.

If the program does terminal interaction, make it output a short
notice like this when it starts in an interactive mode:

{\footnotesize
\begin{verbatim}
<program>  Copyright (C) <year>  <name of author>

This program comes with ABSOLUTELY NO WARRANTY; for details type `show w'.
This is free software, and you are welcome to redistribute it
under certain conditions; type `show c' for details.
\end{verbatim}
}

The hypothetical commands {\tt show w} and {\tt show c} should show
the appropriate
parts of the General Public License.  Of course, your program's commands
might be different; for a GUI interface, you would use an ``about box''.

You should also get your employer (if you work as a programmer) or
school, if any, to sign a ``copyright disclaimer'' for the program, if
necessary.  For more information on this, and how to apply and follow
the GNU GPL, see \texttt{https://www.gnu.org/licenses/}.

The GNU General Public License does not permit incorporating your
program into proprietary programs.  If your program is a subroutine
library, you may consider it more useful to permit linking proprietary
applications with the library.  If this is what you want to do, use
the GNU Lesser General Public License instead of this License.  But
first, please read \texttt{https://www.gnu.org/licenses/why-not-lgpl.html}.

\end{enumerate}

%\end{document}

### 2019-12-17 latex_user_manual/appendix/NetCDF_license.tex
Initial commit.

\fancyhead[LO]{NetCDF-LICENSE}
\fancyhead[RE]{NetCDF-LICENSE}
%\chapter*{\rlap{NetCDF-LICENSE}} \label{NetCDF_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{NetCDF-LICENSE}
\chapter{NetCDF-LICENSE} \label{NetCDF_LICENSE}

\parindent=0pt

\small

NetCDF (network Common Data Form) is a set of software libraries and
machine-independent data formats that support the creation, access,
and sharing of array-oriented scientific data.

NetCDF Java version 4.6.10 - netcdfAll-4.6.10.jar

Available on https://www.unidata.ucar.edu/downloads/netcdf/index.jsp

BSD 3-Clause License

Copyright (c) 1998-2018, University Corporation for Atmospheric Research/Unidata
All rights reserved.

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions are met:

* Redistributions of source code must retain the above copyright notice, this
  list of conditions and the following disclaimer.

* Redistributions in binary form must reproduce the above copyright notice,
  this list of conditions and the following disclaimer in the documentation
  and/or other materials provided with the distribution.

* Neither the name of the copyright holder nor the names of its
  contributors may be used to endorse or promote products derived from
  this software without specific prior written permission.

THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE LIABLE
FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR
SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER
CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY,
OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.

### 2019-12-17 latex_user_manual/appendix/SLF4J_license.tex
Initial commit.

\fancyhead[LO]{SLF4J-LICENSE}
\fancyhead[RE]{SLF4J-LICENSE}
%\chapter*{\rlap{SLF4J-LICENSE}} \label{SLF4J_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{SLF4J-LICENSE}
\chapter{SLF4J-LICENSE} \label{SLF4J_LICENSE}

\parindent=0pt

\small

SLF4J (Simple Logging Facade for Java) serves as a simple facade or
abstraction for various logging frameworks (e.g. java.util.logging,
logback, log4j) allowing the end user to plug in the desired logging
framework at deployment time.

SLF4J version 1.7.25 - slf4j-simple-1.7.25.jar
Available on https://www.slf4j.org/

Copyright (c) 2004-2017 QOS.ch
All rights reserved.

Permission is hereby granted, free  of charge, to any person obtaining
a  copy  of this  software  and  associated  documentation files  (the
"Software"), to  deal in  the Software without  restriction, including
without limitation  the rights to  use, copy, modify,  merge, publish,
distribute,  sublicense, and/or sell  copies of  the Software,  and to
permit persons to whom the Software  is furnished to do so, subject to
the following conditions:

The  above  copyright  notice  and  this permission  notice  shall  be
included in all copies or substantial portions of the Software.

THE  SOFTWARE IS  PROVIDED  "AS  IS", WITHOUT  WARRANTY  OF ANY  KIND,
EXPRESS OR  IMPLIED, INCLUDING  BUT NOT LIMITED  TO THE  WARRANTIES OF
MERCHANTABILITY,    FITNESS    FOR    A   PARTICULAR    PURPOSE    AND
NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE
LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION
OF CONTRACT, TORT OR OTHERWISE,  ARISING FROM, OUT OF OR IN CONNECTION
WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.

### 2019-12-17 latex_user_manual/appendix/WEKA_license.tex
Initial commit.

\fancyhead[LO]{WEKA-LICENSE}
\fancyhead[RE]{WEKA-LICENSE}
%\chapter*{\rlap{WEKA-LICENSE}} \label{WEKA_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{WEKA-LICENSE}
\chapter{WEKA-LICENSE} \label{WEKA_LICENSE}

\parindent=0pt

\small

Weka (Waikato Environment for Knowledge Analysis) is a Java tool that
incorporates several standard ML techniques into a software "workbench".

WEKA version 3.8.1 - weka.jar
Available on https://www.cs.waikato.ac.nz/ml/weka/

Copyright (C) 1998-2016  University of Waikato

The core WEKA system is distributed under the GNU public license.
Please read the the section entitled ``GNU GENERAL PUBLIC LICENSE`` \ref{GPL_LICENSE}.

Packages may be distributed under various licenses - check the
description of the package in question for license details.

### 2019-12-17 latex_user_manual/appendix/gettingMetData.tex
Initial commit.

\chapter{Getting meteorological data}\label{app:GettingMetData}

	\begin{onehalfspace}

		%	Formato con ejemplo y capturas:

		\section{Getting an annual text file from NDBC}

			The following steps shows the procedure to obtain an annual text file that contain the measurements collected by a buoy from NDBC, concretely the file of the year $2017$ corresponding to the buoy identified as \textit{Station 46001}, moored at Western Gulf of Alaska.

				\begin{itemize}
					%\setlength\itemsep{0.01cm}
					\item \textit{\textbf{Step 1}}: Open the web browser and visit the NDBC web page \cite{NOAA_1}.
					\item \textit{\textbf{Step 2}}: Navigate to the desired buoy and click on the year to download (see Fig. \ref{fig:selectingBuoy}).
						\begin{figure}[ht!]
							\centering
							\includegraphics[scale=0.48]{figures/selectingBuoy.png}
							\caption{Selecting the desired buoy and year.}
							\label{fig:selectingBuoy}
						\end{figure}
					\item \textit{\textbf{Step 3}}: In the option \textit{\textbf{Method Two}} click on the file (see Fig. \ref{fig:downloadingTextFile}) and its content will be shown.
						\begin{figure}[ht!]
							\centering
							\includegraphics[scale=0.48]{figures/downloadingTextFile.png}
							\caption{Downloading the annual text file.}
							\label{fig:downloadingTextFile}
						\end{figure}
					\item \textit{\textbf{Step 4}}: In the menu bar of the web browser select \textit{\textbf{File}} and \textit{\textbf{Save As}} to save the file in the computer.
				\end{itemize}

		\section{Getting a reanalysis data file from NNRP}

			%	Formato con ejemplo y capturas:

			The following steps shows the procedure to obtain a reanalysis data file from NNRP, concretely for the variable \textit{Air temperature} and corresponding to the year $2017$ and the reanalysis nodes $57.5$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W, $57.5$ N $\times$ $152.5$ W, $55.0$ N $\times$ $147.5$ W, $55.0$ N $\times$ $150.0$ W and $55.0$ N $\times$ $152.5$ W.

				\begin{itemize}
					%\setlength\itemsep{0.01cm}
					\item \textit{\textbf{Step 1}}: Open the web browser and visit the NNRP  web page \cite{NNRP}.
					\item \textit{\textbf{Step 2}}: Navigate to the \textit{\textbf{Surface}} section and click on it (see Fig. \ref{fig:selectingSurfaceSection}).
						\begin{figure}[ht!]
							\centering
							\includegraphics[scale=0.40]{figures/selectingSurfaceSection.png}
							\caption{Selecting \textit{\textbf{Surface}} section.}
							\label{fig:selectingSurfaceSection}
						\end{figure}
					\item \textit{\textbf{Step 3}}: Navigate to the \textit{\textbf{Update Schedule}} section, search for the statistic \textit{\textbf{4-times Daily}} of the desired variable and click on the image for creating a plot or subset (see Fig. \ref{fig:selectingVariable}).
						\begin{figure}[ht!]
							\centering
							\includegraphics[scale=0.55]{figures/selectingVariable.png}
							\caption{Selecting reanalysis variable.}
							\label{fig:selectingVariable}
						\end{figure}
					\item \textit{\textbf{Step 4}}: Navigate to the \textit{\textbf{Surface}} level and click on \textit{\textbf{Make plot or subset}} (see Fig. \ref{fig:selectingMakePlot}).
						\begin{figure}[ht!]
							\centering
							\includegraphics[scale=0.42]{figures/selectingMakePlot.png}
							\caption{Selecting \textit{\textbf{Make plot or subset}.}}
							\label{fig:selectingMakePlot}
						\end{figure}
					\item \textit{\textbf{Step 5}}: In the new web page that has been appeared do the following (see Fig. \ref{fig:selectingProperties.png}):
						\begin{itemize}
							\item In \textit{\textbf{Axis Dimensions}} type the desired sub-grid (spatial properties).
							\item In \textit{\textbf{Other dimension value(s)}} type the desired ranges of dates (temporal properties).
							\item In \textit{\textbf{Output options}} select the option \textit{Create a subset without making a plot}.
							\item Click on \textit{\textbf{Create Plot or Subset of Data}} button (at bottom of the web page) to generate the reanalysis data file with the desired properties.
						\end{itemize}
% 						\begin{figure}[ht!]
% 							\centering
% 							\includegraphics[scale=0.50]{figures/selectingProperties.png}
% 							\caption{Typing desired properties of the file to get.}
% 							\label{fig:selectingProperties.png}
% 						\end{figure}
					\item \textit{\textbf{Step 6}}: Finally click on \textit{\textbf{FTP a copy of the file}} button to download the reanalysis data file on the computer.
				\end{itemize}

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.50]{figures/selectingProperties.png}
					\caption{Typing desired properties of the file to get.}
					\label{fig:selectingProperties.png}
				\end{figure}

				\clearpage

	\end{onehalfspace}

%	Formato abreviado:

%	Follow these steps to obtain an annual text file that contain the measurements collected by a buoy from NDBC:
%		\begin{itemize}
%			\item \textit{\textbf{Step 1}}: Open the web browser and visit the NDBC web page \cite{NOAA_1}.
%			\item \textit{\textbf{Step 2}}: Navigate to the desired buoy and click on the year to download.
%			\item \textit{\textbf{Step 3}}: In the option \textit{\textbf{Method Two}} click on the file and its content will be shown.
%			\item \textit{\textbf{Step 4}}: In the menu bar of the web browser select \textit{\textbf{File}} and \textit{\textbf{Save As}} to save it in the computer.
%		\end{itemize}

%	Formato abreviado:

%	Follow these steps to obtain a reanalysis data file from NNRP:
%	
%		\begin{itemize}
%			\item \textit{\textbf{Step 1}}: Open the web browser and visit the NNRP  web page \cite{NNRP}.
%			\item \textit{\textbf{Step 2}}: Navigate to the \textit{\textbf{Surface}} section and click on it.
%			\item \textit{\textbf{Step 3}}: Navigate to the \textit{\textbf{Update Schedule}} section, search for the statistic \textit{\textbf{4-times Daily}} of the desired variable and click on the image for creating a plot or subset.
%			\item \textit{\textbf{Step 4}}: Navigate to the \textit{\textbf{Surface}} level and click on \textit{\textbf{Make plot or subset}}.
%			\item \textit{\textbf{Step 5}}: In the new web page that has been appeared do the following:
%				\begin{itemize}
%					\item In \textit{\textbf{Axis Dimensions}} type the desired grid (spatial properties).
%					\item In \textit{\textbf{Other dimension value(s)}} type the desired ranges of dates (temporal properties).
%					\item In \textit{\textbf{Output options}} select the option \textit{Create a subset without making a plot}.
%					\item Click on \textit{\textbf{Create Plot or Subset of Data}} button (bottom of the web page) to generate the reanalysis data file with the desired properties.
%				\end{itemize}
%			\item \textit{\textbf{Step 6}}: Finally click on \textit{\textbf{FTP a copy of the file}} button to download the reanalysis data file on the computer.
%		\end{itemize}

### 2019-12-17 latex_user_manual/chapters/acknowledgments.tex
Initial commit.

\chapter{Acknowledgments}

	\begin{onehalfspace}

		This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union. We also thank to NVIDIA Corporation for the transfer of computational resources for research works.

		The authors thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at https://www.esrl.noaa.gov/psd/, and to NOAA/NDBC by its data that were collected and made freely available.

		We also thank to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/caseStudy.tex
Initial commit.

\chapter{Case study}

	\begin{onehalfspace}

		This chapter will describe how the application works in a practical approach. To do so, an example showing how to create a fully processed (final) dataset starting from the raw data will be performed. The objective of this final dataset is to be used in ML algorithms to classify waves in the Gulf of Alaska depending on their height.

		\section{Case study}

		The meteorological data that will be used to perform this case study is described bellow:

			\begin{enumerate}
			\item The measurements obtained from 2013 to 2017 by the buoy (46001) placed in the Gulf of Alaska, which are provided by NDBC as annual text files. This data is publicly available at the NDBC website. 
			\item Complementary information collected from reanalysis data containing air temperature, pressure and two components of wind speed (South-North and West-East) measurements. This information will be collected from the four closest reanalysis nodes surrounding the geographical location of the buoy. This data is publicly available at the NNRP website and can be downloaded in NetCDF format.
			\end{enumerate}

			After gathering the information described above \footnote{Further instructions for downloading this data can be found in Appendix \ref{app:GettingMetData}.}, the researcher can open SPAMDA.

			In Figure \ref{fig:main_view} the main view can be found, in order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher will select the option \textcolor{blue}{\textit{Manage reanalysis data}} from the view.

			\begin{figure}[ht]
				\centering
				\includegraphics[scale=0.48]{figures/mainView.png}
				\caption{SPAMDA main view.}\label{fig:main_view}
			\end{figure}

			Then, the view represented in Figure \ref{fig:reanalysis} will appear. To enter the four reanalysis files (one per reanalysis variable) the researcher must click on the \textcolor{blue}{\textit{Add reanalysis file}} and a dialog box will appear for selecting them.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/manageReanalysisData_CS.png}
				\caption{Module Manage reanalysis data.}\label{fig:reanalysis}
			\end{figure}

			After the reanalysis files has been introduced in SPAMDA, this window can be closed and go back to the main view to continue entering the information related to the buoy under study. The researcher will now select \textcolor{blue}{\textit{Manage buoys data}} to open the view shown in Figure \ref{fig:manage_buoys}. In order to enter such data, click on the \textcolor{blue}{\textit{New}} button, then the window shown in Figure \ref{fig:entering_buoy} will pop-up.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/tabBuoys_CS.png}
				\caption{Tab \textit{Buoys}.}\label{fig:manage_buoys}
			\end{figure}

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/enteringBuoy_CS.png}
				\caption{Entering a new buoy.}\label{fig:entering_buoy}
			\end{figure}

			%Here the information about the buoy will be entered: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files. In this case, the files containing the data from year $2013$ to $2017$ are inserted. To do so, click on \textcolor{blue}{\textit{Add file}} button and a dialog box will appear for selecting them. Once the data has been introduced, it is necessary to click on \textcolor{blue}{\textit{Save}} button to insert the buoy in SPAMDA database, after that, the window can be closed. To proceed with the creation of the intermediate dataset, double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the next view (see Figure \ref{fig:show_datasets}).

			Here the information about the buoy will be entered: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files. In this case, the files containing the data from year $2013$ to $2017$ are inserted. To do so, click on the \textcolor{blue}{\textit{Add file}} button and a dialog box will appear for selecting them. Once the data has been introduced, it is necessary to click on the \textcolor{blue}{\textit{Save}} button to insert the buoy in SPAMDA database, after that, the window can be closed.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.47]{figures/tabDatasets_CS.png}
				\caption{Tab \textit{Datasets}.}\label{fig:show_datasets}
			\end{figure}

			%The next step would be to create an intermediate dataset, the researcher will double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the next view (see Figure \ref{fig:show_datasets}). In this view, the researcher can delete or consult a brief of each intermediate or pre-processed dataset by selecting it from the corresponding list and also create new ones. To proceed with the creation of the intermediate dataset, click on the \textcolor{blue}{\textit{New}} button and the view shown in Figure \ref{fig:intermediate} will appear.

			The next step would be to create an intermediate dataset, the researcher will double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the next view (see Figure \ref{fig:show_datasets}).

			To proceed with the creation of the intermediate dataset, click on the \textcolor{blue}{\textit{New}} button and the view shown in Figure \ref{fig:intermediate} will appear.

			%The next step would be to create an intermediate dataset including the desired annual text files of the buy. To do so, the researcher can click on \textcolor{blue}{\textit{New}} button and the view shown in Figure \ref{fig:intermediate} will appear.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.43]{figures/creatingIntermediateDataset_CS.png}
				\caption{New intermediate dataset view.}\label{fig:intermediate}
			\end{figure}

			Here the researcher can select the annual text files that he wants to include in the intermediate dataset by clicking on the \textcolor{blue}{\textit{$-$$>$}} button (click on the \textcolor{blue}{\textit{$<$$-$}} button to deselect a previously selected one). In this case, all the files introduced before which correspond to the buoy under study were selected. When the files selection is finished, \textcolor{blue}{\textit{Create}} button will be clicked in order to be able to introduce the description and the file name of the current intermediate dataset, and then clicking on the \textcolor{blue}{\textit{Save}} button, the creation process will start, and the application will show the status of such process.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/tabPreprocess_CS.png}
				\caption{Tab \textit{Pre-process}.}\label{fig:preprocess_data}
			\end{figure}

			%After that, in order to prepare the intermediate dataset to be able to be used by ML algorithms correctly, the dataset to be prepared is selected, and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (represented in Figure \ref{fig:preprocess_data}). In this tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.

			After that, in order to prepare the intermediate dataset to be able to be used by ML algorithms correctly, the dataset to be prepared is selected, and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (represented in Figure \ref{fig:preprocess_data}).

			%As mentioned at the beginning of this chapter, this case study will process the data to be ready to classify waves considering their height, so any missing data from wave height ($376$ values) was recovered and some filters were applied to the remaining variables. After finishing the pre-processing of the dataset, the researcher can click on the \textcolor{blue}{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.

			As mentioned at the beginning of this chapter, this case study will process the data to be ready to classify waves considering their height, so any missing data from wave height ($376$ values) and the remaining attributes are recovered, using the filter \textit{Replace missing values with symmetric $3$ hours mean}. Furthermore, the attributes MWD, DEWP, VIS and TIDE are removed from the dataset by applying the filter \textit{RemoveByName}, since the first two had more than $92$\% of missing data and the last two $100$\%. After finishing the pre-processing of the dataset, the researcher can click on the \textcolor{blue}{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.

			At this point, the researcher has registered the buoy in SPAMDA, then entered its raw data and selected the required data for the problem (intermediate dataset), finally, the data was pre-processed in order to be ready for its future use in ML algorithms. In order to achieve a more accurate description of the problem under study, a matching process can be carried out to merge the processed data from NDBC with the reanalysis data (also entered previously) from NNRP. The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/tabMatchingConfiguration_CS.png}
				\caption{Tab \textit{Matching configuration}.}\label{fig:matching_conf}
			\end{figure}

			In this view, the researcher can customise (or load) the parameters of the matching process according to his needs, and select the prediction task (described in Section \ref{sec:MatchingConf}) that the final dataset will be used in. For this example the pre-processed dataset created and the following parameters were selected:
			\begin{itemize}
				\setlength\itemsep{0.01cm}
				\item \textit{\textbf{Attribute to predict}}: WVHT.
				\item \textit{\textbf{Reanalysis data}}: Air, pressure, u-wind and v-wind.
				\item \textit{\textbf{Buoy attributes to be used as inputs}}: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP.
				\item \textit{\textbf{Reanalysis nodes to consider}}: $1$.
				\item \textit{\textbf{Number of final datasets}}: In this example that option is disabled because only 1 reanalysis node is considered.
				\item \textit{\textbf{Prediction task}}: Classification.
			\end{itemize}

			After configuring the matching process, the researcher can click on the \textcolor{blue}{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.40]{figures/tabFinalDatasets_CS.png}
				\caption{Tab \textit{Final datasets}.}\label{fig:final_dataset}
			\end{figure}

			As in the previous window \textit{Classification} was selected, now the researcher is able to add, modify or delete the thresholds (usually defined by an expert) used to discretise the output variable. After adding the necessary thresholds, the next step is to set the time horizon desired and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output one as explained in Section {\ref{sec:FinalDatasets}}. Then the researcher would click on the \textcolor{blue}{\textit{Update final dataset}} button to see the content of the final dataset which is shown in the bottom left corner and finally, after checking that everything is correct, the last step would be to select the name (and path) of the dataset file, its output format and click on the \textcolor{blue}{\textit{Create final datasets}} button. For this case study the following configuration was applied:
			\vspace{-0.25cm}
			\begin{table}[ht!]

				\caption{Defined thresholds.}
				\label{tab:thresholds}
				\footnotesize
				\centering

				\begin{tabular}{cm{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}

					\cline{1-5}

					\textbf{Class}&\textbf{Description}&\textbf{Inferior [}&\textbf{Superior )}&\\[0.20cm]

					\cline{1-5}

					Low & Low wave height & $0.36$ & $1.5$&\\[0.15cm]

					\cellcolor{gray090}Average & \cellcolor{gray090}Average wave height & \cellcolor{gray090}$1.5$ & \cellcolor{gray090}$2.5$&\\[0.15cm]

					Big & Big wave height & $2.5$ & $4.0$&\\[0.15cm]

					\cellcolor{gray090}Huge & \cellcolor{gray090}Huge wave height & \cellcolor{gray090}$4.0$ & \cellcolor{gray090}$9.9$&\\[0.02cm]

					\cline{1-5}

				\end{tabular}

			\end{table}

			\begin{itemize}
				\item Thresholds: see in Table \ref{tab:thresholds}.
				\item Prediction horizon: 6 hours
				\item Synchronisation: Disabled
			\end{itemize}

			At this point the final dataset would be created and stored in the computer of the researcher. Also there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Fig. \ref{fig:openigFinalDatasetWekaCE}.

			\begin{figure}[ht!]
				\centering
				\includegraphics[scale=0.39]{figures/openigFinalDatasetWekaCE.png}
				\caption{The final dataset opened with the environment Explorer of WEKA.}
				\label{fig:openigFinalDatasetWekaCE}
			\end{figure}

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/gettingStarted.tex
Initial commit.

\chapter{Getting started}

	\begin{onehalfspace}

		\section{System requirements}

			%SPAMDA has been developed in Java, therefore it is a multi-platform software tool. In this way, any computer able to install and run a Java Virtual Machine (JVM) will be enough for running the developed tool.

			SPAMDA has been developed in Java, therefore it is a multi-platform software tool. In this way, any computer having Java Virtual Machine (JVM) installed would be able to run the developed tool.

			As SPAMDA has been compiled for Java JRE 1.8v, Java version 8 needs to be installed in the system, which can be downloaded from \cite{Java} choosing the correct distribution depending on the system platform.

		\section{Installing SPAMDA}

			The process of installation of SPAMDA is quite easy and it does not require administrator permissions to carry it out. The following sections describe how to perform the installation process according to the system platform.

			\subsection{Installing on Linux}

				To install SPAMDA on Linux follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Download the SPAMDA software on the computer.
						\item \textit{\textbf{Step 2}}: Create a folder and copy the downloaded file inside it.
						\item \textit{\textbf{Step 3}}: Decompress the file.
					\end{itemize}

				After performing the above steps SPAMDA would be installed on the computer and ready to be run. Note that the process of installation creates all the folders and files necessary inside the folder created in \textit{\textbf{Step 2}}. In Fig. \ref{fig:installationOnLinux} is represented an example of installation in the folder named ``SPAMDA'', which will contain the software tool.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/installationOnLinux.png}
					\caption{Content of the folder after installing SPAMDA (Linux).}
					\label{fig:installationOnLinux}
				\end{figure}

				Following, the structures of each folder and file created as a result of the installation are described:

					\begin{itemize}
						\item \textbf{\textit{dist}}: Contains the binary distribution of SPAMDA, which consist of:
							\begin{itemize}
								\item \textbf{\textit{DB}}: Contains all the information managed by SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{buoysDatabase}}: Contains the database of the buoys.
										\item \textbf{\textit{finalDatasets}}: It is used as a default folder to save the final datasets.
										\item \textbf{\textit{id1}}: Contains the information of the buoy (annual text files, intermediate datasets, pre-processed datasets and matching configurations) entered as example.
										\item \textbf{\textit{reanalysisDatabase}}: Contains the database of the reanalysis data.
										\item \textbf{\textit{reanalysisFiles}}: Contains the reanalysis files entered through SPAMDA.
									\end{itemize}
								\item \textbf{\textit{help}}: Contains the user manual of SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{userManual.pdf}}: This is the user manual.
									\end{itemize}
								\item \textbf{\textit{SPAMDA.jar}}: This is the runnable file containing SPAMDA.
							\end{itemize}
						\item \textbf{\textit{COPYING}}: This file contains a copy of the license of the GNU GENERAL PUBLIC LICENSE.
						\item \textbf{\textit{LICENSE}}: This file contains a copy of the license of SPAMDA.
						\item \textbf{\textit{NetCDF-LICENSE}}: This file contains a copy of the license of the Library NetCDF Java version 4.6.10
						\item \textbf{\textit{README}}: This file contains the instructions for getting started with SPAMDA.
						\item \textbf{\textit{SLF4j-LICENSE}}: This file contains a copy of the license of the Library SLF4J version 1.7.25
						\item \textbf{\textit{WEKA-LICENSE}}: This file contains a copy of the license of the Library WEKA version 3.8.1
					\end{itemize}

				After installing SPAMDA, and in order to run it, open the ``dist'' folder (as shown in Fig. \ref{fig:runningSPAMDAonLinux}) and type the following command on the command-line of the terminal:

				\texttt{java -jar SPAMDA.jar}

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.50]{figures/runningSPAMDAonLinux.png}
					\caption{Running SPAMDA on Linux.}
					\label{fig:runningSPAMDAonLinux}
				\end{figure}

				After executing such command, the main view of SPAMDA represented in Figure \ref{fig:SPAMDAmainViewonLinux} will appear.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.50]{figures/mainView.png}
					\caption{SPAMDA main view (Linux).}
					\label{fig:SPAMDAmainViewonLinux}
				\end{figure}

			\subsection{Installing on Windows}

				To install SPAMDA on Windows follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Download the SPAMDA software on the computer.
						\item \textit{\textbf{Step 2}}: Create a folder and copy the downloaded file inside it.
						\item \textit{\textbf{Step 3}}: Decompress the file.
					\end{itemize}

				After performing the above steps SPAMDA would be installed on the computer and ready to be run. Note that the process of installation creates all the folders and files necessary inside the folder created in \textit{\textbf{Step 2}}. In Fig. \ref{fig:installationOnWindows} is represented an example of installation in the folder named ``SPAMDA'', which will contain the software tool.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.60]{figures/installationOnWindows.png}
					\caption{Content of the folder after installing SPAMDA (Windows).}
					\label{fig:installationOnWindows}
				\end{figure}

				Following, the structures of each folder and file created as a result of the installation are described:

					\begin{itemize}
						\item \textbf{\textit{dist}}: Contains the binary distribution of SPAMDA, which consist of:
							\begin{itemize}
								\item \textbf{\textit{DB}}: Contains all the information managed by SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{buoysDatabase}}: Contains the database of the buoys.
										\item \textbf{\textit{finalDatasets}}: It is used as a default folder to save the final datasets.
										\item \textbf{\textit{id1}}: Contains the information of the buoy (annual text files, intermediate datasets, pre-processed datasets and matching configurations) entered as example.
										\item \textbf{\textit{reanalysisDatabase}}: Contains the database of the reanalysis data.
										\item \textbf{\textit{reanalysisFiles}}: Contains the reanalysis files entered through SPAMDA.
									\end{itemize}
								\item \textbf{\textit{help}}: Contains the user manual of SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{userManual.pdf}}: This is the user manual.
									\end{itemize}
								\item \textbf{\textit{SPAMDA.jar}}: This is the runnable file containing SPAMDA.
							\end{itemize}
						\item \textbf{\textit{COPYING}}: This file contains a copy of the license of the GNU GENERAL PUBLIC LICENSE.
						\item \textbf{\textit{LICENSE}}: This file contains a copy of the license of SPAMDA.
						\item \textbf{\textit{NetCDF-LICENSE}}: This file contains a copy of the license of the Library NetCDF Java version 4.6.10
						\item \textbf{\textit{README}}: This file contains the instructions for getting started with SPAMDA.
						\item \textbf{\textit{SLF4j-LICENSE}}: This file contains a copy of the license of the Library SLF4J version 1.7.25
						\item \textbf{\textit{WEKA-LICENSE}}: This file contains a copy of the license of the Library WEKA version 3.8.1
					\end{itemize}

				After installing the software, and in order to run it, open the ``dist'' folder and double-click on the SPAMDA.jar file. Next, the main view of SPAMDA represented in figure \ref{fig:SPAMDAmainViewonWindows} will appear.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.80]{figures/mainViewonWindows.png}
					\caption{SPAMDA main view (Windows).}
					\label{fig:SPAMDAmainViewonWindows}
				\end{figure}

			\subsection{Installing on macOS}

				To install SPAMDA on macOS follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Download the SPAMDA software on the computer.
						\item \textit{\textbf{Step 2}}: Create a folder and copy the downloaded file inside it.
						\item \textit{\textbf{Step 3}}: Decompress the file.
					\end{itemize}

				After performing the above steps SPAMDA would be installed on the computer and ready to be run. Note that the process of installation creates all the folders and files necessary inside the folder created in \textit{\textbf{Step 2}}. In Fig. \ref{fig:installationOnMacOS} is represented an example of installation in the folder named ``SPAMDA'', which will contain the software tool.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.60]{figures/installationOnMacOS.png}
					\caption{Content of the folder after installing SPAMDA (macOS).}
					\label{fig:installationOnMacOS}
				\end{figure}

				Following, the structures of each folder and file created as a result of the installation are described:

					\begin{itemize}
						\item \textbf{\textit{dist}}: Contains the binary distribution of SPAMDA, which consist of:
							\begin{itemize}
								\item \textbf{\textit{DB}}: Contains all the information managed by SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{buoysDatabase}}: Contains the database of the buoys.
										\item \textbf{\textit{finalDatasets}}: It is used as a default folder to save the final datasets.
										\item \textbf{\textit{id1}}: Contains the information of the buoy (annual text files, intermediate datasets, pre-processed datasets and matching configurations) entered as example.
										\item \textbf{\textit{reanalysisDatabase}}: Contains the database of the reanalysis data.
										\item \textbf{\textit{reanalysisFiles}}: Contains the reanalysis files entered through SPAMDA.
									\end{itemize}
								\item \textbf{\textit{help}}: Contains the user manual of SPAMDA.
									\begin{itemize}
										\item \textbf{\textit{userManual.pdf}}: This is the user manual.
									\end{itemize}
								\item \textbf{\textit{SPAMDA.jar}}: This is the runnable file containing SPAMDA.
							\end{itemize}
						\item \textbf{\textit{COPYING}}: This file contains a copy of the license of the GNU GENERAL PUBLIC LICENSE.
						\item \textbf{\textit{LICENSE}}: This file contains a copy of the license of SPAMDA.
						\item \textbf{\textit{NetCDF-LICENSE}}: This file contains a copy of the license of the Library NetCDF Java version 4.6.10
						\item \textbf{\textit{README}}: This file contains the instructions for getting started with SPAMDA.
						\item \textbf{\textit{SLF4j-LICENSE}}: This file contains a copy of the license of the Library SLF4J version 1.7.25
						\item \textbf{\textit{WEKA-LICENSE}}: This file contains a copy of the license of the Library WEKA version 3.8.1
					\end{itemize}

				After installing the software, and in order to run it, open the ``dist'' folder and double-click on the SPAMDA.jar file. Next, the main view of SPAMDA represented in figure \ref{fig:SPAMDAmainViewonMacOS} will appear.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.50]{figures/mainViewonMacOS.png}
					\caption{SPAMDA main view (macOS).}
					\label{fig:SPAMDAmainViewonMacOS}
				\end{figure}

		\section{How to uninstall?}

			To uninstall SPAMDA just delete the folder in which the installation process was carried out.

			\begin{center}
				\begin{warningbox}{Warning}{13cm}
					This action will remove SPAMDA from your computer and any information entered in SPAMDA or generated by means of it will be lost.
				\end{warningbox}
			\end{center}

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/introduction.tex
Initial commit.

--- 2019-12-17 latex_manuscript/elsevier/2_Environmental_Modelling_and_Software/spamda.tex
@@
-% !TeX spellcheck = en_GB
-\documentclass[review]{elsarticle}
+\chapter{Introduction to SPAMDA}
 
-\usepackage{textcomp, gensymb}
-\usepackage{color}
-\usepackage[table]{xcolor}
-\usepackage{epstopdf}
-\usepackage{subfig}
-\usepackage[T1]{fontenc}
-\usepackage{lineno}
-\usepackage{amsmath}
-\usepackage{hyperref}
+	\begin{onehalfspace}
 
-\epstopdfsetup{outdir=./figures/}
+		\section{SPAMDA description}
 
-\definecolor{gray090}{gray}{0.90}
+			SPAMDA is a software tool for creating datasets with meteorological data from two well-known sources of information, \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
 
-\modulolinenumbers[5]
-
-\journal{Environmental Modelling \& Software}
-
-\bibliographystyle{elsarticle-num}
-
-\begin{document}
-
-	\begin{frontmatter}
-
-		\title{SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
-
-		\author[UCO]{Antonio M. G\'omez-Orellana}
-		\ead{am.gomez@uco.es}
-
-		\author[UCO]{Juan C. Fern\'andez\corref{JCFC}}
-		\cortext[JCFC]{Corresponding author.}
-		\ead{jfcaballero@uco.es}
-
-		\author[UCO]{Manuel Dorado-Moreno}
-		\ead{manuel.dorado@uco.es}
-
-		\author[UCO]{Pedro A. Guti\'errez}
-		\ead{pagutierrez@uco.es}
-
-		\author[UCO]{C\'esar Herv\'as-Mart\'inez}
-		\ead{chervas@uco.es}
-
-		\address[UCO]{Department of Computer Science and Numerical Analysis, University of Cordoba, C\'ordoba, Spain.}
-
-		\begin{abstract}
-			%Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
-Machine Learning (ML) techniques have become a valuable support in many research areas, improving the performance of traditional statistical procedures.
-However, such techniques require datasets containing information related to the topic under study, which are not always available in an appropriate format.
-Preparing these datasets implies a lot of time and effort by the researchers.
-This paper presents a software tool for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} (NDBC) and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project}.
-The datasets created by the software are ready to be used as inputs for ML techniques in prediction tasks (classification or regression).
-As the designed software is able to simplify the creation of the datasets and reduce the time needed for this task, it prevents researchers from performing repetitive technical work, allowing them to concentrate on the study of the meteorological aspects of the data.
-Therefore, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
-			Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
-Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation implies a lot of time and effort by the researchers.
-This paper presents a software for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-As the designed software is able to simplify the creation of the datasets (ready to be used as inputs for ML techniques in prediction tasks) and reduce the time needed for this task, researchers can benefit from it in order to achieve a more efficient exploitation and protection of the environment.
-		\end{abstract}
-
-		\begin{keyword}
-			Meteorological data\sep Reanalysis data\sep  Pre-processing\sep Creating datasets\sep  Prediction tasks\sep Marine energy
-		\end{keyword}
-
-	\end{frontmatter}
-
-	\linenumbers
-
-	\section{Introduction}
-
-		A better understanding of the environment is of vital importance for science, contributing not only to more efficient exploitation of natural resources but also to the development of new strategies aimed at its protection.
-In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g.
-climate change, agriculture, energy, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind energy assessment \cite{DVORAK20101244}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
-All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
-
-		On the other hand, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
-In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution.
-In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented.
-This tool can also be used to analyse the performance of the solar cells.
-A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{GUTIERREZ201369}, a tool based on \textit{Fatigue, Aerodynamics, Structures, and Turbulence} (FAST) code for performing design optimisation of offshore wind turbines is presented, which can analyse a massive group of cases by means of its parametric design capability.
-A design tool for architects based on \textit{Lighting and Thermal} (LT) method is used in \cite{BAKER2013156}, with the purpose of comparing and optimising design solutions in terms of energy and comfort performance.
-Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and an interactive web application, called the \textit{Stream Hydrology And Rainfall Knowledge System} (SHARKS), for analysis of hydrologic and meteorological time series data that can be used for providing valuable insights to urban hydrologic processes is presented in \cite{BRENDEL2019}.
-
-		Marine energy prediction is currently a hot topic where meteorological data is used in.
-Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{BHATTACHARYA2017157, BREDE201344}.
-Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed to transform this wave energy into electricity.
-WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them.
-Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction.
-For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-Additionally, wave predictions are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
-
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks \cite{srivastava2017large}, obtaining significant improvements in the performance of the results.
-ML methodologies can be used not only by experienced data and computer scientists but also by other researchers.
-For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms.
-In this way, ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction.
-The problem is that, in order to apply these methods, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge.
-Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.
-
-		The information to create these datasets can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
-Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
-Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a merging or matching process have to be carried out by the researchers to manually create the datasets with the needed information.
-Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
-In summary, many important details and different intermediate steps have to be considered when creating suitable datasets for ML techniques, resulting in an extremely tedious task.
-
-		The main purpose of this paper is to present a software tool able to prevent researchers from performing this repetitive work and greatly simplify all the steps involved in the creation of datasets.
-The meteorological data considered for the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
-The software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
-As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows the researchers focus on the study of the meteorological aspects of the observations.
-The datasets obtained are ready to be used as input for ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
-These datasets contain one or more meteorological variables as inputs and one variable as target (variable to be predicted).
-The format of the generated datasets will be \textit{Attribute-Relation File Format} (ARFF) \cite{WEKA_ARFF}, which is the one used by WEKA.
+			The datasets created with SPAMDA will be ready to use as input for Machine Learning (ML) techniques in classification or regression prediction tasks, although the researchers may use them in the way they deem suitable.
+These datasets will contain one or more meteorological variables as inputs and another one as target (variable to predict).
+The format of the datasets will be \textit{Attribute-Relation File Format} (ARFF) \cite{WEKA_ARFF} that it is used by the well-known tool \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA}, which provides a wide collection of ML algorithms.
 Besides, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling the researchers to use others tools.
 
-		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, combining meteorological data from NDBC and NNRP.
-The advantages that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
+			Some of the advantages that SPAMDA tool offers are briefly summarised below:
+
 				\begin{itemize}
@@
 
-		This paper is organised as follows: Section \ref{sec:DataSources} describes the sources of information used by SPAMDA for creating datasets.
-Section \ref{sec:SPAMDA} describes in detail the features of the software tool.
-Section \ref{sec:CaseStudy} shows a case study describing the use of SPAMDA in a practical approach.
-Section \ref{sec:Conclusions} provides the final conclusions and future work.
-
 	\section{Meteorological data sources}\label{sec:DataSources}
 
-		The data provided by the above-mentioned sources of information of SPAMDA is described below:
+		The data provided by the above-mentioned sources of information used by SPAMDA is briefly described below:
 
@@
 NDBC designs, develops, operates, and maintains a network of data collecting buoys (stations).
-The mission of the network is to collect real-time marine meteorological and oceanographic observations, such as $H_s$, dominant wave period, or wind speed and direction, among others.
+The mission of the network is to collect real-time marine meteorological and oceanographic observations, such as wave height, dominant wave period, or wind speed and direction, among others.
 
-			The buoys maintained by NDBC are deployed in the coastal and offshore waters around oceans and seas, and they are equipped with assorted sensors which allow them to perform different measurements.
-The information collected by the buoys is available on the NDBC website \cite{NOAA_1}, and it is divided into different groups.
-One of them corresponds to standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite {NOAA_2}.
-These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC (Universal Time Coordinated) and from $23$:$50$ 31th December of the previous desired year to $22$:$50$ 31th December of the desired year.
-In Table \ref{tab:measurementsDescription}, a comprehensive measurement description and the corresponding units are provided as a summary for the reader.
-A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Fig.
-\ref{fig:fragmentAnnualTexFile}.
-Each column corresponds to a meteorological variable or attribute, and each row or instance corresponds to the values of the measurements collected by the buoy for each attribute at a specific date and time.
+			The buoys maintained by NDBC are deployed in the coastal and offshore waters around oceans and seas, and are equipped with assorted sensors which allow them to perform different measurements.
+The information collected by the buoys is available in NDBC web page \cite{NOAA_1}, which is divided into different groups.
+One of them is the standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite {NOAA_2}.
+These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC and from $23$:$50$ 31th Dec of the previous desired year to $22$:$50$ 31th Dec of the desired year.
+In Table \ref{tab:measurementsDescription} a comprehensive measurements descriptions and units of such information is provided.
 
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[scale=0.50]{figures/FigureFragmentAnnualTextFile.png}
-				\caption{A fragment of an annual text file of the \textit{Station 46001}.}
-				\label{fig:fragmentAnnualTexFile}
-			\end{figure}
+			\clearpage
 
-			\begin{table}[!ht]
+			\begin{table}[ht!]
 
@@
 
-				\begin{tabular}{ccm{8.60cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
+				\begin{tabular}{ccm{11.00cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
 
 					\cline{1-4}
+
+					%\textbf{Variable}&\textbf{Units}&\textbf{Description}&\\[0.30cm]
 
@@
 
-					WDIR & degT & Wind direction (the direction the wind is coming from in degrees clockwise from true North) during the same period used for WSPD. \\
+					WDIR & degT & Wind direction (the direction the wind is coming from in degrees clockwise from true N) during the same period used for WSPD. \\
 
@@
 
-					ATMP & degC & Air temperature (Celsius degrees). &\\[0.10cm]
+					ATMP & degC & Air temperature (Celsius). &\\[0.10cm]
 
-					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature (Celsius degrees).
+					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature (Celsius).
 For buoys the depth is referenced to the hull's waterline.
@@
 
-					TIDE & ft & The water level in feet above or below MLLW. \\
+					%PTDY & hPa & Pressure Tendency is the direction (plus or minus) and the amount of pressure change (hPa) for a three hour period ending at the time of observation. \\
+
+					TIDE & ft & The water level in feet above or below Mean Lower Low Water (MLLW). \\
 
@@
 
-			Note that the data collected by the network of buoys may be incomplete due to diverse circumstances such as the weather conditions in which the buoys have to operate, failures or malfunctioning elements of the buoys, among others.
-Accordingly, it may be the situation that some of the measurements are completely missing (missing date or instance) or partially missing (some measurements not recorded), by a buoy or by a set of buoys, once in a while or over a period of time.
-It may be also possible that the measurements have been recorded at a time different from the expected one.
-This aspect have to be taken into account when creating the datasets.
-This casuistry is explained in detail in Appendix A.
+			\item NNRP provides three-dimensional global reanalysis of numerous meteorological variables (e.g.
+air temperature, U/V-wind, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$Z, $06$Z, $12$Z and $18$Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid.
+Weather observations are from different sources, such as ships, satellites and radar, among others.
 
-			\item NNRP provides three-dimensional global reanalysis of numerous meteorological variables (e.g.
-air temperature, components South-North and West-East of wind speed, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$ Z (Zulu time), $06$ Z, $12$ Z and $18$ Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid.
-Weather observations are from different sources, such as ships, satellites and radar, among others.
-Reanalysis data is created assimilating such observations using the same climate model throughout the entire reanalysis period in order to reduce the effects of modelling changes on climate statistics.
-Such information has become a substantial support of the needs of the research community, even more in locations where instrumental (real time) data is not available.
+			%The reanalysis data is available in NNRP web page \cite{NNRP}, which are accessible through the different sections.
+Such data can be fully (a global $2.5\degree$ x $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data which provides a description of what the data in each variable represents and the spatial and temporal properties of the data.
+Each reanalysis data file contains the estimated values by a mathematical model of one meteorological variable and for each reanalysis node.
+For a better understanding, in Fig.
+\ref{fig:subGrid} an approximate representation of a six sub-grid reanalysis nodes around the geographical localisation of a buoy (obtained from NDBC) is shown.
 
-			The reanalysis data is available in the NNRP website \cite{NNRP}, which it is accessible through different sections.
-Such data can be fully (a global $2.5\degree$ x $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data, which provides a description of the file contents and also includes the spatial and temporal properties of the data.
+			The reanalysis data is available in NNRP web page \cite{NNRP}, which are accessible through the different sections.
+Such data can be fully (a global $2.5\degree$ x $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data which provides a description of the file contents and also includes the spatial and temporal properties of the data.
 Each reanalysis file contains the values of a meteorological variable estimated by a mathematical model for each reanalysis node.
 For a better understanding, in Fig.
-\ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical location of a buoy (obtained from NDBC) is shown.
+\ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical localisation of a buoy (obtained from NDBC) is shown.
 
@@
 
-		Therefore, with both sources of information, which complement each other, and carrying out a matching process, SPAMDA will create datasets for prediction tasks.
-In this way, the dataset input variables will be one or more reanalysis variables from NNRP and one or more measurements from NDBC.
-The dataset output variable will always be one measurement from NDBC.
+		With both sources of information SPAMDA will create datasets for prediction tasks.
+In this way, the input variables of the dataset will be one or more reanalysis variables from NNRP and one or more measurements from NDBC.
+The output variable of the dataset will be one measurement from NDBC.
+Note that the output variable cannot be used as input also.
 
-	\section{SPAMDA}\label{sec:SPAMDA}
-
-		SPAMDA combines  meteorological information from NDBC and NNRP to obtain new datasets for oceanic and atmospheric studies.
-In order to do so, SPAMDA manages three different types of datasets, which will be described in detail in the following sections, but are briefly introduced bellow for giving the reader a better general understanding:
-			\begin{itemize}
-				\item \textit{Intermediate datasets}: They contain the meteorological observations from NDBC.
-				\item \textit{Pre-processed datasets}: They are obtained as a result of pre-processing tasks performed on the intermediate datasets.
-				\item \textit{Final datasets}: Created by merging an intermediate or pre-processed dataset (which contain the information from NDBC) with the reanalysis data from NNRP.
-This procedure is referenced in SPAMDA as matching process and will be carried out according to the study to be performed (classification or regression).
-			\end{itemize}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.32]{figures/FigureSPAMDA.eps}
-			\caption{Brief outline of the functionality provided by SPAMDA.}
-			\label{fig:SPAMDA}
-		\end{figure}
-		SPAMDA consists of three main functional modules, whose main features, represented in Fig. \ref{fig:SPAMDA}, are the following:
-		\begin{itemize}
-
-			\item \textit{Manage buoys data}: The aim of this module is to provide features for the management and analysis of the information related to the buoys from NDBC. This includes:
-			\begin{enumerate}
-				\item Entering and updating the information of each buoy.
-				\item Creation of intermediate datasets with the collected measurements.
-				\item Pre-processing tasks for obtaining the pre-processed datasets.
-				\item Matching process to merge the information from NDBC and NNRP
-				\item Creation of the final datasets accordingly to the ML technique to use (classification or regression).
-			\end{enumerate}
-
-			\item \textit{Manage reanalysis data}: This module is used for the management of the reanalysis data provided by the NNRP.
-In this way, the researchers can keep the reanalysis data files updated for their studies.
-Such files will be used, depending on the researchers needs, in the matching process when obtaining the final datasets.
-
-			\item \textit{Tools}: This module includes features for converting intermediate or pre-processed datasets to ARFF or CSV format and for opening ARFF files with WEKA software.
-
-		\end{itemize}
-
-		In the following subsections each integrated functional module is described in detail.
-
-			\subsection{Buoys}\label{sec:Buoys}
-
-				When a new buoy is included in SPAMDA the following information, which can be obtained from NDBC, is requested:
-				\begin{itemize}
-					\item \textit{Station ID}: An alphanumeric identifier that allows easy identification of the buoy.
-					\item \textit{Description}: A short description of the buoy.
-					\item \textit{Latitude}: North or South geographical localisation (degrees) of the buoy.
-					\item \textit{Longitude}: West or East geographical localisation (degrees) of the buoy.
-					\item \textit{Measurements files}: The above-mentioned annual text files of the standard meteorological information collected by the buoy and downloaded from the NDBC website.
-This will be used for the creation of the intermediate datasets.
-One file per year is expected.
-				\end{itemize}
-
-				For clarification, an example is presented in Fig. \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.42]{figures/FigureBuoys.eps}
-					\caption{Example of entering two buoys with its annual text files.}
-					\label{fig:buoys}
-				\end{figure}
-
-			\subsection{Datasets}\label{sec:Datasets}
-
-				Once a buoy has been included as described in Section \ref{sec:Buoys}, it is possible to create datasets with one or more annual text files, which are referenced in SPAMDA as intermediate datasets.
-In this module, the researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary ones.
-
-				When an intermediate dataset is created, it is associated with its corresponding buoy.
-Besides, a summary of its content is also created, providing relevant information such as the number of instances, the dates of the first and last measurements, the annual text files included, and the missing and duplicated dates.
-
-				An example where three intermediate datasets have been created is presented in Fig.
-\ref{fig:datasets}.
-The two intermediate datasets of the buoy ID1 contain meteorological data of different years, and the intermediate dataset of the buoy ID2 contains meteorological data of two years.
-For each buoy, as many intermediate datasets as needed can be created.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.40]{figures/FigureDatasets.eps}
-					\caption{Example of the creation of the intermediate datasets.}
-					\label{fig:datasets}
-				\end{figure}
-
-			\subsection{Pre-process} \label{sec:Preprocess}
-
-				Data pre-processing prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms.
-In this way, the quality of dat can be improved prior to the learning phase, by applying pre-processing tasks (filters).
-The result will be referenced as pre-processed datasets.
-
-				SPAMDA provides several filters grouped in three categories, \textit{Attribute}, \textit{Instance} and \textit{Recover missing data}, including the configuration of their parameters and a short description of them:
-
-				\begin{itemize}
-
-				 \item \textit{Attribute}: All theses filters can be applied to the attributes (variables of the buoy from NDBC) of the intermediate dataset.
-
-					\begin{itemize}
-						\item \textit{Normalize} \cite{WEKA_Filter_Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in the interval [0,1].
-						\item \textit{Remove} \cite{WEKA_Filter_Remove}: It removes an attribute or a range of them.
-						\item \textit{RemoveByName} \cite{WEKA_Filter_RemoveByName}: It removes attributes based on a regular expression matched against their names.
-						\item \textit{ReplaceMissingValues} \cite{WEKA_Filter_ReplaceMissingValues}: For each attribute, all the missing values will be replaced by the average value of the attribute.
-						\item \textit{ReplaceMissingWithUserConstant} \cite{WEKA_Filter_ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
-					\end{itemize}
-
-				 \item \textit{Instance}: All theses filters can be applied to the instances (hourly measurements of the buoy from NDBC) of the intermediate dataset.
-					\begin{itemize}
-						\item \textit{RemoveDuplicates} \cite{WEKA_Filter_RemoveDuplicates}: With this filter, all duplicated instances are removed.
-						\item \textit{RemoveWithValues} \cite{WEKA_Filter_RemoveWithValues}: This filter removes all the instances that match the attribute and the value supplied by the user.
-						\item \textit{SubsetByExpression} \cite{WEKA_Filter_SubsetByExpression}: It removes all the instances which do not match a user-specified expression.
-					\end{itemize}
-
-				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the intermediate dataset.
-					\begin{itemize}
-						\item \textit{Replace missing values with next nearest hour}: The missing values of each attribute are replaced with the next nearest non missing value.
-						\item \textit{Replace missing values with previous nearest hour}: This filter replaces the missing values of each attribute with the previous nearest non missing value.
-						\item \textit{Replace missing values with next $n$ hours mean}: The missing values of each attribute are replaced with the next $n$ nearest non missing values mean, where $n$ can be configured by the user.
-						\item \textit{Replace missing values with previous $n$ hours mean}: This filter replaces the missing values of each attribute in the intermediate dataset with the previous $n$ nearest non missing values mean.
-						\item \textit{Replace missing values with symmetric $n$ hours mean}: The missing values of each attribute in the intermediate dataset are replaced with the $n$ previous and $n$ next non missing values mean.
-					\end{itemize}
-
-				\end{itemize}
-
-				SPAMDA allows the researchers to undo the last filter applied or to restore the initial content of the intermediate dataset.
-Besides, the content and relevant statistical information (number of instances with missing values, minimum and maximum values, mean and standard deviation) of the intermediate and the pre-processed datasets can be visualised in this module.
-
-				Fig.
-\ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one.
-The intermediate dataset 1 of the buoy ID2 has been also pre-processed.
-\textit{Pre-processed dataset n} represents that the researchers can create as many pre-processed datasets as they consider opportune.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.36]{figures/FigurePreprocess.eps}
-					\caption{Example of the creation of pre-processed datasets.}
-					\label{fig:preprocess}
-				\end{figure}
-
-			\subsection{Matching configuration}
-			\label{sec:matching_conf}
-
-				In order to merge and format the data provided by the two sources of information described in Section \ref{sec:DataSources}, it is necessary to carry out a matching process.
-The matching procedure is performed using an intermediate or pre-processed dataset, which includes the measurements collected by a buoy from NDBC, and the needed reanalysis data files from NNRP.
-Note that SPAMDA is able to manage the NetCDF binary format for handling the information stored in the reanalysis files.
-
-				Such process merges the information of both sources that match on time, but, given that the measurements of the buoys are hourly collected from $00$:$50$ to $23$:$50$ UTC, and the reanalysis data is available every $6$ hours at $00$ Z, $06$ Z, $12$ Z and $18$ Z, the matching can only be carried every $6$ hours (discarding the rest of measurements from the buoy data).
-Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest buoy measurement (before or after) within a maximum of 60 minutes of difference.
-Finally, the matched instances of both sources will form the final datasets.
-
-				Fig.
-\ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year.
-In this way, only the instances from both sources that are linked with arrows (highlighted in green colour) will be used in the creation of the final datasets.
-Although the reanalysis dates have been presented in a human readable format, note that reanalysis dates are stored in hours from $01$-$01$-$1800$, and they have to be transformed for comparison taking into account the time zone.
-Such transformation is automatically done by SPAMDA when matching the instances.
-
-				The reader can check in Appendix A \ref{app:AppendixA} an example with a more complex case of the procedure.
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.38]{figures/FigureMatchingProcess.png}
-					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
-					\label{fig:matchingProcess}
-				\end{figure}
-
-				SPAMDA allows the researchers to perform a customisable matching process, for obtaining as many different versions of the same meteorological data as needed.
-				Prediction tasks are based on the estimation of the output attribute using the information provided by the input attributes.
-Depending on the task, the datasets must be prepared and configured differently:
-				\begin{itemize}
-					\item \textit{Classification}: The final datasets will be ready to use as input for ML classifiers, requiring a nominal output attribute, whose specific preparation is detailed in Section \ref{sec:FinalDatasets}.
-					\item \textit{Regression}: The final datasets will be ready to use as input for regression methods, requiring a real output attribute, whose preparation is also explained in Section \ref{sec:FinalDatasets}.
-					\item \textit{Direct matching}: In this case the inputs attributes have a direct correspondence with the output attribute, and it is not necessary to perform any additional preparation.
-Both input and target attributes are synchronised in time, in such a way that the final dataset is not intended for prediction purposes.
-For example, the final datasets may be used in lost data recovering tasks, in correlation studies, in descriptive analyses, etc.
-				\end{itemize}
-
-				The following parameters can be specified for the matching process:
-				\begin{itemize}
-
-					\item \textit{Flux of energy} \cite{FERNANDEZ201544}: When the $F_e$ is selected, it will be used as output.
-This attribute is not collected by the buoys, but it can be calculated from two wave parameters: $H_s$ and $T_e$, which are collected as WVHT and APD attributes, respectively, and were described in Table \ref{tab:measurementsDescription}.
-In this way, SPAMDA obtains the $F_e$ of each instance using the following equation:
-					\begin{equation}
-							F_e = 0.49 \cdot H^2_s \cdot T_e,
-							\label{eq:fluxOfEnergy}
-					\end{equation}
-
-					where $F_e$ is measured in kilowatts per meter, $H_s$ is measured in meters and $T_e$ is measured in seconds.
-Note also that $F_e$ is defined in Eq.
-\ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.
-
-					\item \textit{Attribute to predict}: Instead of using $F_e$, researchers can select any of the attributes collected by the buoys as output (e.g.
-significant wave height, WVHT, wind direction, WDIR, sea level pressure, PRES, etc.).
-Therefore, they can conduct different studies by selecting one attribute or other.
-
-					\item \textit{Reanalysis data files}: In order to have a possible better description of the problem under study, more than one reanalysis variable can be considered as input.
-Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Fig.
-\ref{fig:subGrid}) for each variable of reanalysis.
-
-					In that sense, the reanalysis data files must have the same spatial and temporal properties but related to different variables.
-SPAMDA simplifies this task by showing the reanalysis data files that are compatibles each other, and checking that the selection made by the researches meets that condition.
-
-					\item \textit{Buoys attributes}: In addition to the reanalysis variables, the final datasets will also include the selected attributes as inputs (of the intermediate or pre-processed dataset used), providing a possible better characterisation of the problem under study, although it will depend on how correlated the attributes are.
-
-					\item \textit{Include missing dates}: As above-mentioned, the information collected by a buoy may be incomplete due to measurements not recorded by it.
-As a consequence, the matching of instances between both sources of information may not be possible (missing dates).
-In that situation, the researchers can consider two options: 1) discard the instances affected or 2) include them.
-In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft\textit{?}\guillemotright.
-
-					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Fig.
-\ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
-In this way, the researcher can:
-
-						\begin{itemize}
-
-							\item Consider all the reanalysis nodes contained in each file: in this case, the information provided by each reanalysis node contained in each selected reanalysis data file will be used.
-
-							\item Consider only some of the reanalysis nodes contained in each file: in this case, only the information of the $N$ closest reanalysis nodes to the buoy will be used  ($N$ given by the user).
-To do that, SPAMDA uses the \textit{Haversine} equation \cite{Haversine_2009} to calculate the distance from each reanalysis node to the location of the buoy and obtain the closest ones.
-Haversine equation is also known as the great circle distance and performs calculation from main point to destination point with a trigonometric function using latitude and longitude:
-								\begin{eqnarray}
-									\label{eq:Haversine}
-									d(p_0,p_j) & = & \arccos(\sin(lat_0)\cdot \sin(lat_j) \nonumber \\
-									& & \cdot \cos(lon_0-lon_j) + \cos(lat_0) \\
-									& & \cdot \cos(lat_j)), \nonumber 
-								\end{eqnarray}								
-							where $p_0$ is the buoy geographical location, $p_j$ stands for the location of each reanalysis node, and $lat$ and $lon$ are the latitude and longitude of the points, respectively.
-						\end{itemize}
-
-					\item \textit{Number of final datasets}: Depending on the number of nearest reanalysis nodes to consider, the number of final datasets to create and the content of them can be configured according to the following options:
-						\begin{itemize}
-
-							\item \textit{One (using weighted mean of the $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of the reanalysis data used (one per selected reanalysis data file).
-This weighted mean is obtained by SPAMDA and uses Eq.
-\ref{eq:Haversine} to obtain the distance from each reanalysis node to the location of the buoy.
-Once the distances have been calculated they are inverted and normalised as follows:
-								\begin{linenomath*}
-									\begin{equation}
-										w_i=\frac{\sum_{j=1}^{N} d(p_0,p_j)}{d(p_0,p_i)}, ~~i=1, \ldots, N.
-										\label{eq:weightedMean}
-									\end{equation}
-								\end{linenomath*}
-
-							With these weights, a weighted mean of each variable of reanalysis is obtained for each of the $N$ nodes.
-Therefore, the closest reanalysis nodes to the localisation of the buoy will provide more information.
-
-							Considering as example the two nearest reanalysis nodes represented in Fig.
-\ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.
-
-							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by the researcher will be created.
-Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.
-
-							In this case, and considering as example the four closest reanalysis nodes (see Fig.
-\ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.
-
-						\end{itemize}
-
-				\end{itemize}
-
-				Once the matching parameters have been described, for a better understanding of them,  Fig.
-\ref{fig:directMatching} presents an example of the matched information considering the data shown in Fig.
-\ref{fig:matchingProcess} and using the following matching configuration\footnote{Note that the date is shown just for better understanding, but it will not be included in the final dataset.}:
-					\begin{itemize}
-						\item Variable WVHT as attribute to predict.
-						\item Variable Pres as reanalysis input attribute.
-						\item Variable WSDP as buoy input attribute.
-						\item Not including missing dates.
-						\item Considering the closest reanalysis node.
-						\item Task to be used: \textit{Direct matching}.
-					\end{itemize}
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.49]{figures/FigureDirectMatching.png}
-					\caption{An example of the matched information for \textit{Direct matching}.}
-					\label{fig:directMatching}
-				\end{figure}
-
-			\subsection{Final datasets} \label{sec:FinalDatasets}
-
-				Once the matching process has been performed with the desired configuration, it is necessary to prepare the matched information for the desired prediction task (\textit{Regression} or \textit{Classification}), obtaining as a result the final datasets.
-Remember that \textit{Direct matching}, as it was described in Section \ref{sec:matching_conf}, performs a direct correspondence between the attributes used as inputs and the output one, and it is not necessary to carry out any preparation.
-
-				SPAMDA allows the researchers to make such preparation by means of the following options:
-
-					\begin{itemize}
-
-						\item \textit{Prediction horizon} (Classification and Regression): This option indicates the time gap for moving backward the attribute to predict (output attribute).
-In this way, the input attributes (variables of the buoy and reanalysis data) will be used to predict the output attribute in a specific future time (e.g.
-+6h, +12h, +18h, +1 day, etc.).
-
-						The minimum interval for increasing and decreasing the prediction horizon is $6$h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out.
-Therefore, for each increment of the prediction horizon, an instance is lost of the dataset is lost (as this future information is not available).
-As the minimum prediction horizon is $6$h, at least one instance will be lost.
-The relation between the inputs and the attribute to predict will be defined as follows:
-						\begin{linenomath*}
-							\begin{equation}
-								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t}),
-								\label{eq:noSynchronisingRD}
-							\end{equation}
-						\end{linenomath*}
-						where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to be predicted, $\mathbf{b}_t$ is the vector containing the selected NDBC variables and $\mathbf{r}_t$ is the vector containing the selected reanalysis variables.
-In this way and considering the matched information shown in Fig.
-\ref{fig:directMatching}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
-
-						Optionally, the reanalysis variables can be synchronised with the attribute to predict.
-Given that these variables are estimated by a mathematical model, we can obtain very good future estimations, which can improve the performance of the results.
-In this case, the relation between the inputs and the attribute to predict would be:
-						\begin{linenomath*}
-							\begin{equation}
-								o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t+\Delta t}).
-								\label{eq:synchronisingRD}
-							\end{equation}
-						\end{linenomath*}
-
-						Note that the selected NDBC variables as input cannot be synchronised with the attribute to predict.
-
- 						For better understanding, considering the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of one \textit{Regression} dataset is shown in Fig.
-\ref{fig:regressionNoSync}.
-As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one).
-The options considered for the preparation of each final dataset are the following:
- 							\begin{itemize}
- 								\item Do not synchronise the reanalysis data (see Eq. \ref{eq:noSynchronisingRD} for the relation between the inputs and the output).
- 								\item A prediction horizon of $6$h.
- 							\end{itemize}
-
-							\begin{figure}[ht!]
-								\centering
-								\includegraphics[scale=0.49]{figures/FigureRegressionNoSync.png}
-								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and without synchronisation.}
-								\label{fig:regressionNoSync}
-							\end{figure}
-
- 						Note that, due to prediction horizon is $6$h, the values of WVHT attribute are moved backward one instance (up).
-As a consequence, the last instance ($2017$/$12$/$31$ $18$:$00$) is lost and is not included in the final dataset.
-Besides, and because the reanalysis data has not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
-\ref{eq:noSynchronisingRD}).
-
- 						Moreover, considering again the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of the same dataset but applying synchronisation (see Eq.
-\ref{eq:synchronisingRD}) is shown in Fig.
-\ref{fig:regressionSync}.
-
-							\begin{figure}[ht!]
-								\centering
-								\includegraphics[scale=0.49]{figures/FigureRegressionSync.png}
-								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and with synchronisation.}
-								\label{fig:regressionSync}
-							\end{figure}
-
-						Again, and due to the prediction horizon selected ($6$h), the values of the WVHT attribute are moved backward one instance (up) and the last instance ($2017$/$12$/$31$ $18$:$00$) is not included in the final dataset.
-But now, the values of the Pres variable are also moved backward one instance (due to the synchronisation).
-Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Eq.
-\ref{eq:synchronisingRD}).
-
-						\item \textit{Thresholds of the output attribute} (Classification): Since the values of the variables collected by the buoys are real numbers, it is necessary to discretise them (convert them from real to nominal values) for the attribute selected as output (attribute to be predicted).
-SPAMDA allows the researchers to perform this process by defining the necessary classes with their thresholds, which will be used to carry out such discretisation.
-
-						Considering again the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
-\ref{fig:prediction}.
-The options considered for the preparation of the final dataset are the following:
-							\begin{itemize}
-								\item Do not synchronise the reanalysis data.
-								\item A prediction horizon of $6$h.
-								\item The thresholds shown in Table \ref{tab:thresholds}.
-							\end{itemize}
-
-						\begin{table}[!h]
-
-							\caption{Thresholds for the classification example represented in Fig. \ref{fig:prediction}}
-							\label{tab:thresholds}
-							\footnotesize
-							\centering
-
-							\begin{tabular}{cm{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
-
-								\cline{1-5}
-
-								\textbf{Class}&\textbf{Description}&\textbf{Inferior [}&\textbf{Superior )}&\\[0.20cm]
-
-								\cline{1-5}
-
-								Low & Low wave height & $0.36$ & $1.5$&\\[0.15cm]
-
-								\cellcolor{gray090}Average & \cellcolor{gray090}Average wave height & \cellcolor{gray090}$1.5$ & \cellcolor{gray090}$2.5$&\\[0.15cm]
-
-								Big & Big wave height & $2.5$ & $4.0$&\\[0.15cm]
-
-								\cellcolor{gray090}Huge & \cellcolor{gray090}Huge wave height & \cellcolor{gray090}$4.0$ & \cellcolor{gray090}$9.9$&\\[0.15cm]
-
-								\cline{1-5}
-
-							\end{tabular}
-
-						\end{table}
-
-						\begin{figure}[ht!]
-							\centering
-							\includegraphics[scale=0.49]{figures/FigureClassification.png}
-							\caption{An example of the creation a \textit{Classification} dataset, with a prediction horizon of $6$h and without synchronisation.}
-							\label{fig:prediction}
-						\end{figure}
-
-						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable, because its values have been discretised according to the thresholds.
-Besides, and due to the $6$h prediction horizon, the last instance is lost ($2017$/$12$/$31$ $18$:$00$) and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up).
-As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
-\ref{eq:noSynchronisingRD}).
-
-					\end{itemize}
-
-				The content of the final datasets, obtained as the result of the preparation of the matched data, can be visualised to check everything before saving them on disk.
-Such preparation can be performed as many times as required and considering the different options in each moment.
-Although the date will not be included in the final datasets, it can be shown to properly check the matching.
-
-				Finally, it is necessary to define the output configuration to create the final datasets:			
-				\begin{itemize}
-
-					\item \textit{Output path file}: Name of the final datasets and folder to save them on disk.
-
-					\item \textit{Final datasets format}:
-
-						\begin{itemize}
-
-							\item \textit{ARFF}: \textit{Attribute-Relation File Format} \cite{WEKA_ARFF}, which is used by WEKA.
-SPAMDA allows the researchers to directly open the final datasets in the Explorer environment of WEKA (in the same context of work), enabling them to choose the most appropriate ML method to tackle the problem under study.
-
-							\item \textit{CSV}: \textit{Comma-Separated Values}. This format is included in order to consider other different tasks of software tools.
-
-						\end{itemize}
-
-				\end{itemize}
-
-				A text file that summarises the configuration used in matching process and in the preparation of the matched data is also generated.
-It can be saved and loaded, enabling the researchers to resume their studies at any other time.
-
-			\subsection{Manage reanalysis data}
-
-				As mentioned in Section \ref{sec:DataSources}, the reanalysis data files provided by NNRP contain the estimated values by a mathematical model of one meteorological variable.
-
-				In this module (see Fig.
-\ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones.
-Besides, useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final time, latitude and longitude.
-All these fields summarise the temporal and spatial properties of the data.
-Thus, the researcher can quickly and easily identify each reanalysis file entered in SPAMDA.
-
-				An example where two reanalysis data files have been entered in SPAMDA is shown in Fig. \ref{fig:manageReanalisys}.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.45]{figures/FigureManageReanalisys.eps}
-					\caption{Example of entering two reanalysis data files.}
-					\label{fig:manageReanalisys}
-				\end{figure}
-
-			\subsection{Tools}
-
-				SPAMDA also contains another module that provides two utilities: one of them is \textit{Dataset converter} used for converting the desired intermediate or pre-processed datasets to ARFF or CSV formats; the other utility can be used for opening ARFF files with WEKA Explorer environment, which is useful for easily checking the results of different configurations of the pre-processing.
-
-	\section{Case study}\label{sec:CaseStudy}
-
-		Although the software includes a user manual, this section will describe how the application works in a practical approach.
-To do so, an example showing how to create a fully processed dataset (final dataset), starting from the raw data, will be described.
-The objective of this final dataset is to be used with ML algorithms to classify waves in the Gulf of Alaska depending on their height.
-The data collected to perform this example is:
-		\begin{enumerate}
-		\item The measurements obtained from 2013 to 2017 by the buoy with ID 46001, placed in the Gulf of Alaska, which are provided by NDBC as annual text files.
-This data is publicly available at the NDBC website.
-		\item Complementary information collected from reanalysis data containing air temperature, pressure and two components of wind speed measurements (South-North and West-East).
-This information will be collected from the four closest reanalysis nodes surrounding the geographical location of the buoy.
-This data is publicly available at the NNRP website and can be downloaded in NetCDF format.
-		\end{enumerate}
-
-		After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.
-
-		In Figure \ref{fig:main_view}, the main window is shown.
-In order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher has to select the option \textcolor{blue}{\textit{Manage reanalysis data}}.
-
-		\begin{figure}[ht]
-			\centering
-			\includegraphics[width=0.60\textwidth]{figures/FigureMain_view.png}
-			\caption{SPAMDA main window.}\label{fig:main_view}
-		\end{figure}
-
-		Then, the window of Figure \ref{fig:reanalysis} is shown.
-Here, using the buttons located at the bottom, it is possible to add, delete or consult any data from the different reanalysis files.
-After the information has been introduced in the application, this window can be closed and the user can go back to the main window to continue entering the information related to the buoy under study.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.89\textwidth]{figures/FigureManage_reanalysis_data.png}
-			\caption{\textit{Manage reanalysis data} window.}\label{fig:reanalysis}
-		\end{figure}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.89\textwidth]{figures/FigureManage_buoys.png}
-			\caption{\textit{Buoys} tab.}\label{fig:manage_buoys}
-		\end{figure}
-
-		The researcher has to select \textcolor{blue}{\textit{Manage buoys data}} to open the window shown in Figure \ref{fig:manage_buoys}.
-In this tab, the researcher can consult, modify, add or delete different data related to the buoy.
-In order to enter such data, click on the \textcolor{blue}{\textit{New}} button, and then the window shown in Figure \ref{fig:new_buoy} pops up.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.70\textwidth]{figures/FigureNew_buoy.png}
-			\caption{\em{Entering a new buoy} window.}\label{fig:new_buoy}
-		\end{figure}
-
-		Here the information about the buoy has to be included: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files.
-In this case, the files containing the data from year $2013$ to $2017$ are inserted by clicking on the \textcolor{blue}{\textit{Add file}} button.
-Once the data has been introduced, it is necessary to click on the \textcolor{blue}{\textit{Save}} button to insert the buoy in SPAMDA database.
-After that, the window can be closed.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.90\textwidth]{figures/FigureDatasets.png}
-			\caption{\textit{Datasets} tab.}\label{fig:show_datasets}
-		\end{figure}
-
-		To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab to switch to the corresponding view (see Figure \ref{fig:show_datasets}).
-In this view, the researcher can delete or consult a summary of each intermediate or pre-processed dataset by selecting it from the corresponding list.
-It can also create new ones.
-To proceed with the creation of the intermediate dataset, the user clicks on the \textcolor{blue}{\textit{New}} button, and the view shown in Figure \ref{fig:intermediate} appears.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.70\textwidth]{figures/FigureNew_intermediate_dataset.png}
-			\caption{\textit{New intermediate dataset} view.}\label{fig:intermediate}
-		\end{figure}
-
-		Here the researcher can select the annual text files to be included in the intermediate dataset.
-In this case, all the files introduced before, which correspond to the buoy under study, are selected.
-When the file selection is finished, \textcolor{blue}{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textcolor{blue}{\textit{Save}} button, the creation process starts, showing the status of the process during it.
-
-		After that, in order to prepare the intermediate dataset, the dataset is selected, and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}).
-In this tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.90\textwidth]{figures/FigurePreprocess.png}
-			\caption{\textit{Pre-process} tab.}\label{fig:preprocess_data}
-		\end{figure}
-
-		Here the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter.
-As mentioned earlier, this case study is focused on classifying waves considering their height, so any missing data from wave height ($376$ values) and the remaining attributes are recovered, using the filter \textit{Replace missing values with symmetric $3$ hours mean}.
-Furthermore, the attributes MWD, DEWP, VIS and TIDE are removed from the dataset by applying the filter \textit{RemoveByName}, since the first two had more than $92$\% of missing data and the last two $100$\%.
-After finishing the pre-processing of the dataset, the researcher can click on the \textcolor{blue}{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.
-
-		At this point, the researcher has registered the buoy in SPAMDA, then entered its raw data and selected the required data for the problem (intermediate dataset).
-Finally, the data has been pre-processed in order to be ready for its future use in ML algorithms.
-In order to achieve a more accurate description of the problem under study, a matching process can be carried out to merge the processed data from NDBC with the reanalysis data (also entered previously) from NNRP.
-The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.90\textwidth]{figures/FigureMatching_configuration.png}
-			\caption{\textit{Matching configuration} tab.}\label{fig:matching_conf}
-		\end{figure}
-
-		In this view, the researcher can customise (or load) the parameters of the matching process according to their needs and select the prediction task  (described in Section \ref{sec:matching_conf}) that the final dataset will be used for.
-In this example, the following parameters were selected:
-		\begin{itemize}
-			\item Attribute to predict: WVHT.
-			\item Reanalysis data: Air, pressure, u-wind and v-wind.
-			\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP.
-			\item Reanalysis nodes to consider: $1$.
-			\item Number of final datasets: In this example that option is disabled, because only one reanalysis node is considered.
-			\item Prediction task: Classification.
-		\end{itemize}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.90\textwidth]{figures/FigureFinal_datasets.png}
-			\caption{\textit{Final datasets} tab.}\label{fig:final_dataset}
-		\end{figure}
-
-		After configuring the matching process, the researcher can click on the \textcolor{blue}{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
-Given that, in the previous window, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable.
-After this, the next step is to set the time horizon desired and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output, as explained in Section {\ref{sec:FinalDatasets}}.
-Then the researcher can click on the \textcolor{blue}{\textit{Update final dataset}} button to see the content shown in the bottom left corner.
-Finally, after checking that everything is correct, the last step would be to select the name (and path) of the dataset file and its output format and click on the \textcolor{blue}{\textit{Create final datasets}} button.
-For this case study, the following configuration was applied:
-		\begin{itemize}
-			\item Thresholds: see Table \ref{tab:thresholds}.
-			\item Prediction horizon: 6 hours
-			\item Synchronisation: Disabled
-		\end{itemize}
-
-		At this point the final dataset would be created and stored in the computer of the researcher.
-Also, there is an option to open the dataset with WEKA (after creating it) in order to perform a first classification approach or a preliminary study of the data structure, as shown in Fig.
-\ref{fig:openigFinalDatasetWeka}.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka.png}
-			\caption{Final dataset opened with the environment Explorer of WEKA.}
-			\label{fig:openigFinalDatasetWeka}
-		\end{figure}
-
-	\section{Conclusions}\label{sec:Conclusions}
-
-		A software tool for creating datasets using meteorological data from NDBC and NNRP has been presented.
-These datasets will be ready to use as input for ML techniques in prediction tasks (classification or regression).
-As a result, the researchers will benefit from a great support when carrying out their oceanic and atmospheric studies, related to energy and environmental modelling.
-Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.
-In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
-
-		In order to improve SPAMDA, some future work could be focused on new functional modules for managing meteorological data of different formats \cite{NOAA_3}, so that the developed tool can be extended to any other research, new pre-processing functionalities such as filters to analyse the correlation between attributes or new functional modules for recovering missing values using nearby buoys data \cite{DuranRosal2016}.
-Furthermore, the developed software could manage other sources of reanalysis data (with different spatial and temporal resolution), and new output formats for the datasets which could be used as input by other tools for ML such as KEEL (\textit{Knowledge Extraction based on Evolutionary Learning}) \cite{AlcalFdez2009KEELAS}.
-However, such new functionalities can be developed with a reasonable effort to be able to manage each particular casuistry.
-For example, when dealing with incomplete data, interpreting different data and files structures or carrying out the matching process of two environmental data sources.
-
-	\section*{Acknowledgments}
-
-		This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union.
-We also thank to NVIDIA Corporation for the transfer of computational resources for research works.
-
-		The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.
-
-	\section*{References}
-
-		\bibliography{bibfile}
-
-	\section*{Appendix A. Managing incomplete data}\label{app:AppendixA}
-
-		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.
-
-		The measurements collected by the buoys may be incomplete or recorded at a different time than the expected one, due to the weather conditions in which the buoys have to operate.
-To illustrate this casuistry, the following examples are shown in Fig.
-\ref{fig:measurements}:
-		\begin{itemize}
-			\item In the instance marked with a), the measurement of $17$:$50$ was collected at $17$:$45$, $5$ minutes earlier.
-			\item In the instance marked with b), the measurement of $23$:$50$ was collected at $23$:$30$, $20$ minutes earlier.
-			\item In the instance marked with c), the measurement of $05$:$50$ is duplicated.
-			\item In the instance marked with d), the measurement of $11$:$50$ is missing (missing date or instance).
-			\item In the instance marked with e), the measurement of $17$:$50$ and $18$:$50$ are missing (missing dates or instances).
-			\item Missing values highlighted in red colour.
-		\end{itemize}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.45]{figures/FigureMeasurements.png}
-			\caption{A fragment of an annual text file with different missing value examples.}
-			\label{fig:measurements}
-		\end{figure}
-
-		SPAMDA has been designed to tackle these situations, and it informs the researchers of any incidence found while reading the annual text files for creating the intermediate datasets.
-For the case of measurements that were recorded at a different time than expected, it has been established a time gap of 6 minutes ($10\%$ of an hour).
-Therefore, if the time difference exceeds such value the date will be considered as an unexpected.
-
-		Fig.
-\ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Fig \ref{fig:measurements}.
-Note that the instance marked with a) has not been informed by SPAMDA as an unexpected date because its time difference is less than $6$ minutes.
-Depending on the affected attribute, NDBC uses a specific value \cite{NOAA_3} to indicate the presence of lost data (e.g.
-$99$ for VIS and TIDE attributes, $999$ for DEWP, MWD and WDIR, etc.).
-SPAMDA interprets these specific values and, after creating the intermediate dataset, the researchers can check if it contains missing values by visualising its statistical information or content.
-Remember that SPAMDA provides several filters for recovering missing data, which were described in subsection \ref{sec:Datasets}.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.40]{figures/FigureCreatingDataset.png}
-			\caption{Status of the creation of the intermediate dataset for the example of Fig \ref{fig:measurements}.}
-			\label{fig:creatingDataset}
-		\end{figure}
-
-		SPAMDA takes into account this casuistry when carrying out the matching process.
-An example is given in Fig.
-\ref{fig:matchingMeasurements}.
-As above-mentioned, the matching process is performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference.
-However, in the instance marked with $e)$, given that the measurements dates $01$/$05$/$2017$ $17$:$50$ and $01$/$05$/$2017$ $18$:$50$ are missing, the reanalysis date $01$/$05$/$2017$ $18$:$00$ cannot be matched with buoy data (this date is highlighted in mauve colour in the Figure).
-Depending on the selection made by the researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.36]{figures/FigureMatchingMeasurements.png}
-			\caption{Matching the measurements (left) and the reanalysis data (right).}
-			\label{fig:matchingMeasurements}
-		\end{figure}
-
-\end{document}
+	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/ns_Antecedentes.tex
Initial commit.

\chapter{Antecedentes}

	\begin{onehalfspace}

		Se puede considerar como antecedente de referencia la herramienta \textit{Scientific Processing of Meteorological Data} (SPMD). Así mismo, se han buscado aplicaciones que generen conjuntos de datos creados a partir de la información obtenida del NDBC y NNRP, pero no se tiene conocimiento de su existencia. Sin embargo, sí se han encontrado algunas aplicaciones que trabajan con ficheros en formato NetCDF.

		\section{ncdump}

			La herramienta \textit{ncdump} \cite{Ncdump2017} permite generar una representación del contenido de un fichero NetCDF empleando \textit{Common Data form Language} (CDL), la cual puede ser visualizada por pantalla, editada o servir de entrada para generar un nuevo fichero.

			Dispone de diferentes opciones con las cuales es posible configurar la salida de dicha representación, así como seleccionar qué contenidos del fichero se desean procesar.

			Es muy versátil para analizar el contenido de este tipo de ficheros y almacenarlo en un formato de texto plano, consiguiendo de este modo reducir la complejidad a una dimensión.

			La ejecución de la misma se realiza desde la línea de comandos y no dispone de interfaz gráfica. En la figura \ref{figura:ncdump} se muestra el uso de \textit{ncdump} para obtener el contenido de la cabecera de un fichero NetCDF. En ella se puede observar que la dimensionalidad del fichero es de 1465 (número de mediciones) x 2 (latitud Norte: 57.5\degree \space y 55.0\degree) x 2 (longitud Este: 210\degree \space y 212.5\degree).

			\begin{figure}
				\centering
			   \includegraphics[scale=0.41]{img/antecedentes/ncdump.png}
				\caption{Uso de \textit{ncdump} para mostrar la cabecera de un fichero NetCDF}
				\label{figura:ncdump}
			\end{figure}

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/ns_installation.tex
Initial commit.

\chapter{Installing SPADNN}

	\begin{onehalfspace}

		The process of installation of SPADNN is quite easy and it does not require administrator permissions to carry it out. Once the compressed file with SPADNN has been saved on the computer, the installation process can be performed just doing the following steps:

			\begin{enumerate}

				\item Create a folder where holding the software tool.
				\item Copy the compressed file inside the folder.
				\item Decompress the file by double click on it.

			\end{enumerate}

		After performing the above steps SPADNN has been installed on the computer. Therefore, and in order to run it, just open the ``dist'' folder and type the following sentence on the command line:

		\texttt{java -jar "SPADNN"}

		\section{How to uninstall?}

			Note that this action will remove SPADNN from your computer and any information entered or generated by the software tool will be lost.

			To uninstall SPADNN just delete the folder in which the installation process was carried out.

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/ns_requirements.tex
Initial commit.

\chapter{System requirements}

	\begin{onehalfspace}

		SPADNN has been developed in Java and it is a multi-platform software tool. In this way, any computer able to install and run a Java Virtual Machine (JVM) will be enough for running the developed tool. As SPADNN has been compiled for Java JRE 1.8v the Java version 8 needs to be installed in the system, which can be downloaded from [\textcolor{red}{poner enlace}] choosing the correct distribution depending on the system platform.

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/chapters/usingSPAMDA.tex
Initial commit.

\chapter{Using SPAMDA}

	\begin{onehalfspace}

		SPAMDA has been designed to greatly simplify all the steps involved in the creation of datasets with information from the sources mentioned in Section \ref{sec:DataSources}, thus the researcher can create as different datasets of the same meteorological data as needed, in a quick and efficient manner. For this purpose, SPAMDA manages three different types of datasets that are briefly introduced bellow:
			\begin{itemize}
				\item \textit{Intermediate datasets}: Which will contain the meteorological observations from NDBC.
				\item \textit{Pre-processed datasets}: Obtained as a result of pre-processing tasks performed on the intermediate datasets.
				\item \textit{Final datasets}: Created by merging an intermediate or pre-processed dataset with the reanalysis data (referenced as matching process) and according to the needs of the study to perform (classification or regression).
			\end{itemize}

		SPAMDA consists of the following three main functional modules:
			\begin{itemize}
				\item \textbf{\textit{Manage buoys data}}
				\item \textbf{\textit{Manage reanalysis data}}
				\item \textbf{\textit{Tools}}
			\end{itemize}

		\begin{figure}[ht!]
			\centering
			\includegraphics[scale=0.5]{figures/mainView.png}
			\caption{SPAMDA main view.}
			\label{fig:mainView}
		\end{figure}

		Such functional modules, which will be described in detail in the following sections, are accessible through the main view of SPAMDA represented in Fig. \ref{fig:mainView}.

		\section{Manage buoys data}

			The aim of this module is to provide the features for the management and analysis of the information related to the buoys from NDBC, since such information is entered in SPAMDA until it is used by the researchers for conducting their studies. Such management and analysis involves:
				\begin{itemize}
					\item Entering and updating the information of each buoy.
					\item The creation of the intermediate datasets with the collected measurements.
					\item Pre-processing tasks for obtaining the pre-processed datasets.
					\item The matching process to merge the information from NDBC and NNRP.
					\item The creation of the final datasets accordingly to the ML technique to use.
				\end{itemize}

			The following sections describe the organisation of this module.

			\subsection{Buoys}

				The \textit{Buoys} tab, which is represented in Fig. \ref{fig:tabBuoys}, allows the researchers to enter and update the information of each buoy. When entering a new buoy the following information, which can be obtained from NDBC, is requested:

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/tabBuoys.png}
					\caption{Tab \textit{Buoys}.}
					\label{fig:tabBuoys}
				\end{figure}

				\begin{itemize}
					\item \textit{\textbf{Station ID}}: An alphanumeric identifier that allows the researchers to easily identify the buoy.
					\item \textit{\textbf{Description}}: A short description of the buoy.
					\item \textit{\textbf{Latitude}}: North or South geographical localisation (degrees) of the buoy. 
					\item \textit{\textbf{Longitude}}: West or East geographical localisation (degrees) of the buoy.
					\item \textit{\textbf{Measurements files}}: The above-mentioned annual text files of the standard meteorological information collected by the buoy and downloaded from NDBC web page, which will be used for the creation of the intermediate datasets. The researchers will add to the buoy one file per year and as many as needed. Remember that such files are available in NDBC web page \cite{NOAA_1}.
				\end{itemize}

				%To enter a new buoy click on \textcolor{blue}{\textit{New}} button, after typing the required information click on \textcolor{blue}{\textit{Save}} button for inserting the buoy. An example of entering a new boy is represented in Fig. \ref{fig:enteringBuoy}.

				To enter a new buoy follow the next steps:

					\begin{itemize}\itemsep0.02cm
						\item \textit{\textbf{Step 1}}: Click on the \textcolor{blue}{\textit{New}} button and the view represented in Fig. \ref{fig:enteringBuoy} will be displayed. (The remaining steps are related to such view).
						\item \textit{\textbf{Step 2}}: Type the required information about the buoy. To enter a new annual text file of the buoy click on the \textcolor{blue}{\textit{Add file}} button or click on the \textcolor{blue}{\textit{Delete file}} button to delete the selected one. By clicking on the \textcolor{blue}{\textit{Clear}} button it is possible to clean the form that request the information.
						\item \textit{\textbf{Step 3}}: Click on the \textcolor{blue}{\textit{Save}} button to insert the buoy.
					\end{itemize}

				%An example of entering a new boy is represented in Fig. \ref{fig:enteringBuoy}.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.41]{figures/enteringBuoy.png}
					\caption{Entering a new buoy.}
					\label{fig:enteringBuoy}
				\end{figure}

				In the same way, by clicking on the \textcolor{blue}{\textit{View/Modify}} button it is possible to view or modify the data relating to the selected buoy. To delete a buoy click on the \textcolor{blue}{\textit{Delete}} button.

				\begin{center}
					\begin{warningbox}{Warning}{12cm}
						Deleting a buoy will remove all the data related to the buoy.
					\end{warningbox}
				\end{center}

			\subsection{Datasets}\label{sec:Datasets}

				%The \textit{Datasets} tab, which is represented in Fig. \ref{fig:tabDatasets}, allows the researchers to manage the intermediate datasets needed for their studies and relating to each buoy.

				The \textit{Datasets} tab, which is represented in Fig. \ref{fig:tabDatasets}, allows the researchers to manage the intermediate datasets of each buoy, which are the baseline for their studies.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/tabDatasets.png}
					\caption{Tab \textit{Datasets}.}
					\label{fig:tabDatasets}
				\end{figure}

				Once a buoy has been entered it is possible to create intermediate datasets using one ore more annual text files (added previously), which contain the measurements collected by the buoy. Note that such measurements may be incomplete or recorded at a different time than the expected one due to the weather conditions in which the buoys have to operate. SPAMDA has been designed to tackle such situation and it informs the researchers of any incidence found while reading the annual text files for creating the intermediate datasets.

				When an intermediate dataset is created, it is associated with its corresponding buoy, enabling the researchers to identify which intermediate datasets belong to each buoy. Besides, a summary of the content of the intermediate dataset is also created, providing relevant information about its content such as number of instances, date of first and last measurement, annual text files included, missing and duplicated dates.

				To proceed with the creation of an intermediate dataset follow the next steps:

					\begin{itemize}\itemsep0.02cm
						\item \textit{\textbf{Step 1}}: Select the desired buoy from the list shown on the left side.
						\item \textit{\textbf{Step 2}}: Click on the \textcolor{blue}{\textit{New}} button, then the view represented in Fig. \ref{fig:creatingIntermediateDataset} will be displayed. (The remaining steps are related to such view).
							\begin{figure}[ht!]
								\centering
								\includegraphics[scale=0.38]{figures/creatingIntermediateDataset.png}
								\caption{New intermediate dataset view.}
								\label{fig:creatingIntermediateDataset}
							\end{figure}
						\item \textit{\textbf{Step 3}}: Select the annual text files to use by clicking on the \textcolor{blue}{\textit{$-$$>$}} button. To deselect a previously selected file click on the \textcolor{blue}{\textit{$<$$-$}} button.
						\item \textit{\textbf{Step 4}}: When finished the selection click on the \textcolor{blue}{\textit{Create}} button.
						\item \textit{\textbf{Step 5}}: Type the description and the name of the intermediate dataset and click on the \textcolor{blue}{\textit{Save}} button to start the process of creation.
					\end{itemize}

				After that, SPAMDA will show the status of such process and the incidences that were found in the data, when it finished click on the \textcolor{blue}{\textit{Ok}} button. Note that the process can be cancelled by clicking on the \textcolor{blue}{\textit{Cancel creation}} button.

				The view that shows the status of the process of creation of an intermediate dataset for the buoy identified as \textit{Station 46001} using the annual text files of the years $2013$, $2014$, $2015$, $2016$ and $2017$ is represented in Fig. \ref{fig:statusCreationIntermediateDataset}.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.39]{figures/statusCreationIntermediateDataset.png}
						\caption{Status of the creation of the intermediate dataset.}
						\label{fig:statusCreationIntermediateDataset}
					\end{figure}

				%As it is shown in Fig. \ref{fig:tabDatasets} when a buoy is selected the intermediate datasets belonging to it are displayed, similarly occurs with the pre-processed datasets (which will be described in Section \ref{sec:Preprocess}) when selecting an intermediate dataset. Besides, on the right side is showed the summary of the intermediate or pre-processed dataset selected.

				As it is shown in Fig. \ref{fig:tabDatasets} when a buoy is selected, the intermediate datasets belonging to it are displayed, similarly, if one of these intermediate datasets  is selected, the pre-processed datasets (described in Section \ref{sec:Preprocess}) obtained from it are displayed. Besides, on the right side the summary of the intermediate or pre-processed dataset selected is shown.

				By clicking on the corresponding \textcolor{blue}{\textit{Delete}} button, the intermediate or pre-processed dataset selected will be removed.

				\begin{center}
					\begin{warningbox}{Warning}{13cm}
						Deleting an intermediate dataset will also remove all the pre-processed datasets belonging to it.
					\end{warningbox}
				\end{center}

				On the other hand, when clicking on the corresponding \textcolor{blue}{\textit{Open}} button the application will be redirected to the \textit{Pre-process} tab (see Section \ref{sec:Preprocess}) and the selected dataset (intermediate or pre-processed) will be opened for being pre-processed.

			\subsection{Pre-process}\label{sec:Preprocess}

				%The \textit{Preprocess} tab, which is represented in Fig. \ref{fig:tabPreprocess}, provides data pre-processing, which is useful for transforming raw data (intermediate datasets) into \textit{clean} and \textit{tidy} data. In this way, the quality of data can be improved prior to computational learning by ML techniques.

				The \textit{Pre-process} tab, which is represented in Fig. \ref{fig:tabPreprocess}, allows to perform data pre-processing, which prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms. In this way, the quality of data can be improved prior to computational learning.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/tabPreprocess.png}
					\caption{Tab \textit{Pre-process}.}
					\label{fig:tabPreprocess}
				\end{figure}

				Once an intermediate dataset has been created it is possible to apply the necessary pre-processing tasks (filters) to enhance the data quality. In the \textcolor{blue}{\textit{Statistical information}} tab relevant data about each attribute of the opened dataset (the one is currently being pre-processed) such as number of instances with missing values, minimum and maximum values, mean and standard deviation is shown. Providing the researchers the capacity to evaluate the pre-processing being performed.

				SPAMDA provides several configurable filters grouped in three categories, \textit{Attribute}, \textit{Instance} and \textit{Recover missing data}:

				\begin{itemize}
					\item \textit{Attribute}: All these filters can be applied to the attributes of the opened dataset.
						\begin{itemize}
							\item \textit{Normalize} \cite{WEKA_Filter_Normalize}: This filter normalises all numeric values of each attribute. The resulting values are by default in [0,1] for the data used to compute the normalisation intervals.
							\item \textit{Remove} \cite{WEKA_Filter_Remove}: It removes an attribute or a range of them.
							\item \textit{RemoveByName} \cite{WEKA_Filter_RemoveByName}: It allows to remove attributes based on a regular expression matched against their names.
							\item \textit{ReplaceMissingValues} \cite{WEKA_Filter_ReplaceMissingValues}: For each attribute all the missing values will be replaced with its mean.
							\item \textit{ReplaceMissingWithUserConstant} \cite{WEKA_Filter_ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
						\end{itemize}
				 \item \textit{Instance}: All these filters can be applied to the instances (hourly measurements) of the opened dataset.
					\begin{itemize}
						\item \textit{RemoveDuplicates} \cite{WEKA_Filter_RemoveDuplicates}: With this filter all duplicated instances are removed.
						\item \textit{RemoveWithValues} \cite{WEKA_Filter_RemoveWithValues}: This filter removes all the instances that match on the attribute and value user-supplied.
						\item \textit{SubsetByExpression} \cite{WEKA_Filter_SubsetByExpression}: It removes all the instances which don't match on a user-specified expression.
					\end{itemize}
				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the opened dataset.
					\begin{itemize}
						\item \textit{Replace missing values with next nearest hour}: The missing values of each attribute are replaced with the next nearest non missing value.
						\item \textit{Replace missing values with previous nearest hour}: This filter replaces the missing values of each attribute with the previous nearest non missing value.
						\item \textit{Replace missing values with next $n$ hours mean}: The missing values of each attribute are replaced with the next $n$ nearest (configurable) non missing values mean. Note that these values may not coincide with the next $n$ hours.
						\item \textit{Replace missing values with previous $n$ hours mean}: This filter replaces the missing values of each attribute with the previous $n$ nearest non missing values mean. Note that these values may not coincide with the previous $n$ hours.
						\item \textit{Replace missing values with symmetric $n$ hours mean}: The missing values of each attribute are replaced with the $n$ previous and $n$ next non missing values mean. Note that these values may not coincide with the symmetric $n$ hours.
					\end{itemize}
				\end{itemize}

				%As a result of the data pre-processing it is possible to create new datasets, which are referenced as pre-processed datasets. Besides, a pre-processed dataset can be also pre-processed again enabling the researchers to resume such task at any other time.

				Once the pre-processing has been performed it is possible to save the resulting dataset, which will be referenced as a pre-processed dataset in SPAMDA. Besides, a pre-processed dataset can be also pre-processed again enabling the researchers to resume such task at any other time.

				To apply a filter follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Open the desired intermediate or pre-processed dataset by clicking on either the \textcolor{blue}{\textit{Open dataset}} button or the corresponding \textcolor{blue}{\textit{Open}} button of tab \textit{Datasets}.
						\item \textit{\textbf{Step 2}}: Select one of the available filters.
						\item \textit{\textbf{Step 3}}: Configure the filter if necessary by clicking on the \textcolor{blue}{\textit{Configure}} button.
						\item \textit{\textbf{Step 4}}: Apply the filter by clicking on the \textcolor{blue}{\textit{Apply}} button.
					\end{itemize}

				As shown in Fig. \ref{fig:tabPreprocess} the pre-processing tasks performed on the opened dataset are displayed on the right side. SPAMDA allows the researchers to undo the last filter applied (\textcolor{blue}{\textit{Undo}} button) or to restore the initial content of the dataset (\textcolor{blue}{\textit{Restore data}} button).

				To create a pre-processed dataset follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Click on the \textcolor{blue}{\textit{Save}} button, then the view represented in Fig. \ref{fig:creatingPreProcessedDataset} will be displayed. (The remaining step is related to such view).
							\begin{figure}[ht!]
								\centering
								\includegraphics[scale=0.40]{figures/creatingPreProcessedDataset.png}
								\caption{New pre-processed dataset view.}
								\label{fig:creatingPreProcessedDataset}
							\end{figure}
						\item \textit{\textbf{Step 2}}: Type the description and the name of the pre-processed dataset and click on the \textcolor{blue}{\textit{Save}} button.
					\end{itemize}

				When saving the pre-processed dataset it will be associated with its corresponding intermediate dataset.

				Moreover, it is also possible to visualise the content of the opened dataset, enabling the researchers to easily identify if any of the instances is missing, duplicated, incomplete (missing values) or was recorded at a different time than the expected one, as mentioned in Section \ref{sec:Datasets}. To proceed with such action click on the \textcolor{blue}{\textit{Content}} tab and use the buttons \textcolor{blue}{\textit{$>$}} and \textcolor{blue}{\textit{$<$}} to check the possible affected instances as it is shown in Fig. \ref{fig:contentOpenedDataset}.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/contentOpenedDataset.png}
					\caption{Visualising the content of the opened dataset.}
					\label{fig:contentOpenedDataset}
				\end{figure}

			\subsection{Matching configuration}\label{sec:MatchingConf}

				The \textit{Matching configuration} tab, which is represented in Fig. \ref{fig:tabMatchingConfiguration}, allows the researchers to customise the parameters of the matching process, which is necessary to carry out in order to merge and format the data provided by the two sources of information described in Section \ref{sec:DataSources}.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/tabMatchingConfiguration.png}
					\caption{Tab \textit{Matching configuration}.}
					\label{fig:tabMatchingConfiguration}
				\end{figure}

				The matching procedure is performed using an intermediate or pre-processed dataset, which includes the measurements collected by a buoy from NDBC, and also the corresponding reanalysis data files from NNRP. Note that SPAMDA is able to manage the NetCDF binary format for handling the information stored in such reanalysis files.

				%Such process merges the information of both sources that matches on time, but due to that the measurements of the buoys are hourly collected from $00$:$50$ to $23$:$50$ UTC, and the reanalysis data is available every $6$ hours at $00$Z, $06$Z, $12$Z and $18$Z, the matching can only be carried out each $6$ hours (discarding the unused measurements from the buoy data). Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference. Finally, the matched instances of both sources will result in the final datasets.

				Such process merges the information of both sources that match on time, but due to the measurements of the buoys are hourly collected from $00$:$50$ to $23$:$50$ UTC, and the reanalysis data is available every $6$ hours at $00$Z, $06$Z, $12$Z and $18$Z, the matching can only be carried out each $6$ hours (discarding the unused measurements from the buoy data). Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference. Finally, the matched instances of both sources will result in the final datasets.

				SPAMDA allows the researchers to perform a customisable matching process, through which the researchers can easily obtain as different final datasets of the same meteorological data as needed, allowing them to consider different factors of the problem under study. These final datasets can be used for classification or regression prediction tasks, or direct matching. Prediction tasks are used to estimate the value of the output attribute in a concrete future using the information provided by the input attributes. Depending on the task to use, the final datasets must be prepared and configured in a specific way:

				\begin{itemize}
					\item \textit{Classification}: The final datasets will be ready to use as input in classification methods, which require a nominal output attribute and whose specific preparation is explained in Section \ref{sec:FinalDatasets}.
					\item \textit{Regression}: The final datasets will be ready to use as input in regression methods, which require a real output attribute and whose preparation is also explained in Section \ref{sec:FinalDatasets}.
					\item \textit{Direct matching}: In this case the inputs attributes have a direct correspondence with the output attribute, and it is not necessary to perform any additional preparation. For example, the final datasets may be used in lost data recovering tasks, or in the way the researchers consider suitable according to the problem under study.
				\end{itemize}

				Following, the parameters that can be configured in the matching process are described:

				\begin{itemize}

					\item \textit{Flux of energy} \cite{FERNANDEZ201544}: When the $F_e$ is selected, it will be used as output. This attribute is not collected by the buoys, but it can be calculated from two wave parameters: $H_s$ and $T_e$, which are collected as WVHT and APD attributes respectively. In this way, SPAMDA will obtain the $F_e$ of each instance using the following equation:

						\begin{equation}
							F_e = 0.49 \cdot H^2_s \cdot T_e
							\label{eq:fluxOfEnergy}
						\end{equation}

					where $F_e$ is measured in kilowatts per meter, $H_s$ is measured in meters and $T_e$ is measured in seconds. Note also that $F_e$ is defined in Eq. \ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.

					\item \textit{Attribute to predict}: Instead of using the $F_e$, the researchers can select any of the attributes collected by the buoys as output (e.g. significant wave height (WVHT), wind direction (WDIR), sea level pressure (PRES), etc.). Therefore, they can focus on different studies by selecting an attribute or other.

					\item \textit{Reanalysis data files}: In order to have a more accurate description of the problem under study, more than one reanalysis variable can be considered as input. Remember that these files have to be previously downloaded by the researcher from the website of the NNRP \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Fig. \ref{fig:subGrid}) for each variable of reanalysis. In that sense, the reanalysis data files must have the same spatial and temporal properties but relating to different variables each other.

					To select the needed reanalysis data files click on the corresponding \textcolor{blue}{\textit{Add/Modify}} button and the view represented in Fig. \ref{fig:selectingReanalysisDataFiles} will be displayed. SPAMDA facilitates this task by showing in cyan colour the reanalysis data files that are compatibles each other when selecting or clicking on a file. Click on the \textcolor{blue}{\textit{Confirm selection}} button when finished and SPAMDA will check that the selected files meet such condition.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.40]{figures/selectingReanalysisDataFiles.png}
						\caption{Selecting reanalysis data files.}
						\label{fig:selectingReanalysisDataFiles}
					\end{figure}

					\item \textit{Buoys attributes}: In addition to the reanalysis variables, the final datasets will also include the selected attributes as inputs (of the intermediate or pre-processed dataset used), providing a possible better characterisation of the problem under study, although it will depend on how correlated the attributes are.

					In the same way, in order to select the attributes of the buoy click on the corresponding \textcolor{blue}{\textit{Add/Modify}} button and the view represented in Fig. \ref{fig:selectingBuoysAttributes} will be displayed showing the available attributes of the buoy.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.40]{figures/selectingBuoysAttributes.png}
						\caption{Selecting buoys attributes.}
						\label{fig:selectingBuoysAttributes}
					\end{figure}

					Once done the selection click on the \textcolor{blue}{\textit{Confirm selection}} button.

					\item \textit{Include missing dates}: As mentioned in Section \ref{sec:Datasets}, the information collected by a buoy may be incomplete due to measurements not recorded by it. As a consequence, the matching of instances between both sources of information may mismatch (missing dates). In that situation, the researchers can consider two options: 1) discard the instances affected or 2) include them. In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft\textit{?}\guillemotright.

					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Fig. \ref{fig:subGrid}, the reanalysis data files may contain information of several reanalysis nodes. In this way, this parameter allows the researchers to choose from:

						\begin{itemize}

							\item Consider all the reanalysis nodes: in this case, all the information of each selected reanalysis data file will be used.

							\item Consider only some of the reanalysis nodes: in this case, only the information of the $N$ closets reanalysis nodes (configurable) to the buoy will be used. To do that, SPAMDA uses the \textit{Haversine} formula \cite{Haversine_2009} to calculate the distance from each reanalysis node to the localisation of the buoy and obtain the closest ones. Haversine formula is also known as great circle distance, this formula perform calculation from main point to destination point with trigonometric function by using latitude and longitude. Haversine formula is calculated as follows:
							  \begin{linenomath*}
									\begin{equation}
										d(p_0,p_j)=\arccos(\sin(lat_0)\cdot \sin(lat_j)\cdot \cos(lon_0-lon_j) + \cos(lat_0) \cdot \cos(lat_j)),
										\label{eq:Haversine}
									\end{equation}
								\end{linenomath*}
							where $p_0$ is the buoy geographical localisation, $p_j$ stands for the location of each reanalysis node, and $lat$ and $lon$ are the latitude and longitude of the points, respectively.

						\end{itemize}

					\item \textit{Number of final datasets}: Depending on the number of the nearest reanalysis nodes to consider, the number of the final datasets to create and therefore the content of them can be configured according to the following options:

						\begin{itemize}
							%\item \textit{One (using weighted mean of $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of reanalysis used (one per selected reanalysis data file). This weighted mean takes into account the distance from each reanalysis node to the localisation of the buoy, and it is calculated by SPAMDA using the formula defined in Eq. \ref{eq:weightedMean}. Therefore, the closest reanalysis nodes to the localisation of the buoy will provide more information.

							\item \textit{One (using weighted mean of $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of reanalysis used (one per selected reanalysis data file). This weighted mean is obtained by SPAMDA and uses the distance (using the formula defined in Eq. \ref{eq:Haversine}) from each reanalysis node to the localisation of the buoy. Once the distances have been calculated they are inverted and normalised as follows:

								\begin{linenomath*}
									\begin{equation}
										w_i=\frac{\sum_{j=1}^{N} d(p_0,p_j)}{d(p_0,p_i)}, ~~i=1, \ldots, N.
										\label{eq:weightedMean}
									\end{equation}
								\end{linenomath*}

							After calculating these weights, they are applied to obtain a weighted mean of each variable of reanalysis. Therefore, the closest reanalysis nodes to the localisation of the buoy will provide more information.

							Considering as example the two nearest reanalysis nodes represented in Fig. \ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.

							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as number of nearest $N$ reanalysis nodes configured by the researcher will be created. Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.

							In this case, and considering as example the four closest reanalysis nodes (see Fig. \ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, then only four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.

						\end{itemize}

				\end{itemize}

				To proceed with the matching process follow the next steps:

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Select the desired buoy from the shown on the left side.
						\item \textit{\textbf{Step 2}}: Select the intermediate or pre-processed dataset to use.
						\item \textit{\textbf{Step 3}}: Configure the above-mentioned matching parameters and click on the \textcolor{blue}{\textit{Run}} button to start the matching process.
					\end{itemize}

				Then, SPAMDA will show the status of such process and the missing dates that were found, when it finished click on the \textcolor{blue}{\textit{OK}} button and the application will be redirected to the \textit{Final datasets} tab (see Section \ref{sec:FinalDatasets}) to perform the preparation of the matched data.

				In Fig. \ref{fig:statusMatchingProcess} is represented the status of a matching process using the configuration showed in Fig. \ref{fig:tabMatchingConfiguration}.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.40]{figures/statusMatchingProcess.png}
						\caption{Status of the matching process.}
						\label{fig:statusMatchingProcess}
					\end{figure}

				Instead of typing all the required parameters, it is possible to load a previously matching configuration saved. To do that click on the \textcolor{blue}{\textit{Load config.}} button, then the view represented in Fig. \ref{fig:selectingMatchingConfiguration} will be displayed for selecting the configuration to be loaded (Section \ref{sec:FinalDatasets} describes how to save the configuration used in the matching process and for the preparation of the matched data).

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/selectingMatchingConfiguration.png}
					\caption{Load matching configuration.}
					\label{fig:selectingMatchingConfiguration}
				\end{figure}

			\subsection{Final datasets}\label{sec:FinalDatasets}

				The \textit{Final datasets} tab, which is represented in Fig. \ref{fig:tabFinalDatasets}, permits the researchers to prepare the matched data for the desired prediction task (\textit{Regression} or \textit{Classification}), obtaining as a result the final datasets. Remember that \textit{Direct matching}, as it was described in Section \ref{sec:MatchingConf}, performs a direct correspondence between the attributes used as inputs and the output one, and it is not necessary to carry out any preparation.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/tabFinalDatasets.png}
					\caption{Tab \textit{Final datasets}.}
					\label{fig:tabFinalDatasets}
				\end{figure}

				SPAMDA allows the researchers to make such preparation by means of the following options:

					\begin{itemize}
						\item \textit{Prediction horizon} (Classification and Regression): This option indicates the time gap for moving backward the output attribute. In this way, the input attributes (variables of the buoy and reanalysis data) will be used to predict the output attribute in a concrete future (e.g. 6h, 12h, 18h, 1 day, etc.).

						The minimum interval for increasing and decreasing the prediction horizon is $6$h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out. Therefore, for each increment of the prediction horizon an instance is lost from the end of the final datasets. As the minimum prediction horizon is $6$h at least one instance will be lost. The relation between the inputs and the attribute to predict will be defined as follows:

						\begin{linenomath*}
								\begin{equation}
									o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t})
								\end{equation}
							\end{linenomath*}

						Where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to predict, $\mathbf{b}$ is the vector containing the selected NDBC variables and $\mathbf{r}$ is the vector containing the selected reanalysis variables. Optionally, the reanalysis variables can be synchronised with the attribute to predict. Given that such variables are estimated by a mathematical model, it is allowed to use these future values, which could improve the performance of the results. In this case, the relation between the inputs and the attribute to predict would be:

							\begin{linenomath*}
								\begin{equation}
									o_{t+\Delta t}=\phi(\mathbf{b}_t,\mathbf{r}_{t+\Delta t})
								\end{equation}
							\end{linenomath*}

						\item \textit{Thresholds of the output attribute} (Classification): Since the values of the variables collected by the buoys are real numbers, it is necessary to discretise (convert from real to nominal values) the selected attribute as output. SPAMDA allows the researchers to perform this process by defining the necessary classes with their thresholds, which will be used against the values of the output attribute to carry out such discretisation.
					\end{itemize}

				Follow these steps to proceed with the preparation of the final datasets:

%					\begin{itemize}
%						\item \textit{\textbf{Step 1}}: Define the classes and its thresholds (\textit{Classification}).
%						\item \textit{\textbf{Step 2}}: Define the prediction horizon (\textit{Classification and Regression}).
%						\item \textit{\textbf{Step 3}}: Click on \textcolor{blue}{\textit{Update final dataset}} button.
%					\end{itemize}

					\begin{itemize}
						\item \textit{\textbf{Step 1}}: Define the classes and their thresholds (\textit{Classification}).
							\begin{itemize}
								\item To do that, use the buttons \textcolor{blue}{\textit{Add}}, \textcolor{blue}{\textit{Modify threshold}} or \textcolor{blue}{\textit{Delete}} to add a new threshold, modify or delete the selected one respectively.
							\end{itemize}
						\item \textit{\textbf{Step 2}}: Define the prediction horizon (\textit{Classification and Regression}).
							\begin{itemize}
								\item To do that, use the buttons \textcolor{blue}{\textit{$-$}} or \textcolor{blue}{\textit{$+$}} to decrease or increase the prediction horizon, and select or deselect the option \textcolor{blue}{\textit{Synchronise the reanalysis data}} depending on needs.
							\end{itemize}
						\item \textit{\textbf{Step 3}}: Click on the \textcolor{blue}{\textit{Update final dataset}} button to take the new configuration of the preparation.
					\end{itemize}

				Such preparation can be performed as many times as required and considering the typed options in each moment.

				As shown in Fig. \ref{fig:tabFinalDatasets}, the content of the final datasets obtained as a result of the custom preparation of the matched data, can be visualised enabling the researchers to check the final datasets before saving them on disk. Although the date will not be included in the final datasets, by selecting the \textcolor{blue}{\textit{Show date}} option it can be shown in order to know the dates of each instance of the intermediate or pre-processed dataset used, that matched with the reanalysis data. Moreover, by clicking on the \textcolor{blue}{\textit{$>$}} and \textcolor{blue}{\textit{$<$}} buttons it is possible to check the dates that were mismatched.

				Finally, and before creating the final datasets, it is necessary to define the output configuration:

				\begin{itemize}

					\item \textit{Output file}: Name of the final datasets and folder to save them on disk.

					\item \textit{Final datasets format}:

						\begin{itemize}

							\item \textit{ARFF}: \textit{Attribute-Relation File Format} \cite{WEKA_ARFF} which is used by the tool WEKA. SPAMDA allows the researchers to open the final datasets created with this format by running the environment Explorer of WEKA (in the same context of work), enabling they to choose the most appropriate ML method to tackle the problem under study. To do that select the option \textcolor{blue}{\textit{Open final datasets with WEKA}}.

							\item \textit{CSV}: \textit{Comma-Separated Values}. With this format the researchers may use the final datasets in the way they deem suitable.

						\end{itemize}

				\end{itemize}

				Once finished the preparation, by clicking on the \textcolor{blue}{\textit{Create final datasets}} button the final datasets will be saved on disk in the selected folder, an them will be opened by the tool WEKA, as represented in Fig. \ref{fig:openigFinalDatasetWeka}, if the researchers selected such option.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.375]{figures/openigFinalDatasetWeka.png}
					\caption{Opening with WEKA the final dataset created.}
					\label{fig:openigFinalDatasetWeka}
				\end{figure}

				%Besides, a text file that summarises the configuration used in the matching process and for the preparation of the matched data is also generated in addition to the final datasets.

				SPAMDA allows the researchers to save the configuration used in the matching process and for the preparation of the matched data, enabling the researchers to resume their studies at any other time. To do that click on the \textcolor{blue}{\textit{Save config.}} button, then the view represented in Fig. \ref{fig:creatingMatchingConfiguration} will be displayed. After typing the description and the file name click on the \textcolor{blue}{\textit{Save}} button to save the configuration, which will be associated with the buoy used.

				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/creatingMatchingConfiguration.png}
					\caption{New matching configuration.}
					\label{fig:creatingMatchingConfiguration}
				\end{figure}

		\section{Manage reanalysis data}

			This module, which is represented in Fig. \ref{fig:manageReanalysisData}, allows the management of the reanalysis data provided by NNRP. In this way, the researchers can keep up to date the reanalysis files needed for their studies. Remember that such data is available in NNRP web page \cite{NNRP}.
				\begin{figure}[ht!]
					\centering
					\includegraphics[scale=0.40]{figures/manageReanalysisData.png}
					\caption{Module Manage reanalysis data.}
					\label{fig:manageReanalysisData}
				\end{figure}

			To enter a new reanalysis file just click on the \textcolor{blue}{\textit{Add reanalysis file}} button and select the desired file, in order to delete a reanalysis file just click on the \textcolor{blue}{\textit{Delete reanalysis file}} button.

			As it is shown in Fig. \ref{fig:manageReanalysisData} useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final: time, latitude and longitude; which summarises the temporal and spatial properties of the data. Thus the researcher can quickly and easily identify each reanalysis file entered in SPAMDA.

		\section{Tools}

			This module includes two utilities, one for converting intermediate or pre-processed datasets to ARFF or CSV format and the other one for opening ARFF files with WEKA. Both features are accessible through the \textcolor{blue}{\textit{Tools}} option of the menu bar showed in Fig. \ref{fig:mainView}.

			\subsection{Datasets converter}

				This utility permits the researchers to convert the desired intermediate or pre-processed datasets to ARFF and CSV format. In this way, the researchers can use these converted datasets as they consider opportune.

				To convert an intermediate or pre-processed dataset click on the \textcolor{blue}{\textit{Dataset converter}} option and the view represented in Fig. \ref{fig:convertDatasets} will be displayed. On the left side are shown the buoys entered in SPAMDA, by clicking on one of them its intermediate datasets will be shown on the top side. Similarly, by clicking on one intermediate dataset its pre-processed datasets will be shown on the bottom side. Once selected the desired dataset (intermediate or pre-processed) just click on the \textcolor{blue}{\textit{Convert to ARFF}} or \textcolor{blue}{\textit{Convert to CSV}} button, and a dialog box will appear asking for the name of the target file that will be created as a result of the conversion.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.40]{figures/convertDatasets.png}
						\caption{Utility Dataset converter.}
						\label{fig:convertDatasets}
					\end{figure}

			\subsection{Open ARFF file with WEKA}

				This other utility allows the researchers to open ARFF files by running the environment Explorer of WEKA in the same context of work, enabling they to resume experiments with previously created final datasets.

				To open an ARFF file click on the \textcolor{blue}{\textit{Open ARFF file with WEKA}} option and the view represented in Fig. \ref{fig:openARFFfile} will be displayed.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.50]{figures/openARFFfile.png}
						\caption{Utility Open ARFF file with WEKA.}
						\label{fig:openARFFfile}
					\end{figure}

				To search for and select an ARFF file click on the \textcolor{blue}{\textit{Browse...}} button, when finished click on the \textcolor{blue}{\textit{Open file}} button for opening it and the view represented in Fig. \ref{fig:openingARFFFileWEKA} will be displayed.

					\begin{figure}[ht!]
						\centering
						\includegraphics[scale=0.50]{figures/openingARFFFileWEKA.png}
						\caption{Opening an ARFF file with WEKA.}
						\label{fig:openingARFFFileWEKA}
					\end{figure}

				Remember that it is also possible to open the final datasets when creating them by selecting the option \textcolor{blue}{\textit{Open final datasets with WEKA}} (see Fig. \ref{fig:tabFinalDatasets}).

	\end{onehalfspace}

### 2019-12-17 latex_user_manual/license/licenseDoc.tex
Initial commit.

\thispagestyle{empty}

\small

\vspace*{\fill} % El texto aparecerá en la parte inferior de la página.

\begin{quote}

	SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets.

	This is the version 1.0 of the SPAMDA manual.

	Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/

	\hspace{1.005cm}Authors:

	\hspace{1.30cm}Antonio Manuel Gómez Orellana, Juan Carlos Fernández Caballero,

	\hspace{1.30cm}Manuel Dorado Moreno, Pedro Antonio Gutiérrez Peña and

	\hspace{1.30cm}César Hervás Martínez.

	\begin{quote} % El texto aparecerá con un margen izquierdo y derecho de 1cm.
		Permission is granted to copy, distribute and/or modify this document 
		under the terms of the GNU Free Documentation License, Version 1.3 or any later
		version published by the Free Software Foundation; with no Invariant Sections, 
		with no Front-Cover Texts, with the Front-Cover logo, and with no Back-Cover Texts.
		\newline
		\newline
		A copy of the license is included in the appendix entitled "GNU Free Documentation License" \ref{GFDL_LICENSE}. If not, see <http://www.gnu.org/licenses/>.

		\vspace{0.50cm}
		Contact information:

		\begin{quote}
			Juan Carlos Fernández Caballero, PhD.

			email: jfcaballero@uco.es

			Address: University of Córdoba, Department of Computer Science and Numerical Analysis, Rabanales Campus, AYRNA Research Group, Einstein Building, 3rd floor. Road Madrid-Cádiz, Km 396-A. 14071 - Córdoba (Spain)
		\end{quote}

	\end{quote}

\end{quote}

### 2019-12-17 latex_user_manual/preamble.tex
Initial commit.

% Contenido del preámbulo del documento LaTeX

\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{tabulary}
\usepackage[table]{xcolor}
\usepackage[T1]{fontenc}		% Para poner los símbolos << y >>
\usepackage{setspace}
\usepackage{pstricks}
\usepackage{fancyhdr}
\usepackage{tocbibind}			% Incluye bibliografía en índice
\usepackage{natbib}
\bibliographystyle{unsrt}		% Estilo de la bibliografía.
\usepackage{fancyvrb}			% Para usar Verbatim
\usepackage{bookmark}			% Para insertar la bibliografía al mismo nivel que el resto de capítulos.
\usepackage{textcomp} 			% Para usar \textquotesingle
\usepackage{pdfpages}
\usepackage{epstopdf}
\usepackage{tikz} 				% Para posicionar imágenes.
\usepackage{gensymb} 			% Para poner el símbolo de los grados.

\usepackage{changepage}			% Para usar {adjustwidth}
\usepackage{tocbibind}			% Incluye bibliografía en índice

\usepackage{indentfirst}		% Indenta el primer párrafo de cada sección.

\usepackage{lineno}
\usepackage{caption}
\usepackage{appendix}

\usepackage{tcolorbox}

% Macro para los mensajes de tipo Warning.
\newtcolorbox{warningbox}[2]
{
	title = \textbf{#1},
	text width=#2,
	colback=red!4!white,
	colframe=orange!65!white
}

% Color gris para las filas de la tabla.
\definecolor{gray090}{gray}{0.90}

% Carpeta en la que se crean los .pdf a partir de las imágenes .eps
\epstopdfsetup{outdir=./figures_converted_to_pdf/}

\DeclareGraphicsExtensions{.eps}

% Espaciado de párrafo.
\parskip=5pt

% Indentación a la izquierda en cada comienzo de párrafo.
\parindent=30pt

% Márgenes del documento.
\usepackage[left=3.5cm, right=2.5cm, top=2.5cm, bottom=3.5cm]{geometry}

% Configuración del formato para la cabecera de los títulos.
\makeatletter
\def\LigneVerticale{\vrule height 4cm depth 2cm\hspace{0.1cm}\relax}

\def\LignesVerticales{%
  \let\LV\LigneVerticale\LV\LV\LV\LV\LV\LV\LV\LV\LV\LV}

\def\GrosCarreAvecUnChiffre#1{%
  \rlap{\vrule height 0.8cm width 1cm depth 0.2cm}%
  \rlap{\hbox to 1cm{\hss\mbox{\white #1}\hss}}%
  \vrule height 0pt width 1cm depth 0pt}

\def\@makechapterhead#1{\hbox{%
    \huge
    \LignesVerticales
    \hspace{-0.5cm}%
    \GrosCarreAvecUnChiffre{\thechapter}
    \hspace{0.4cm}\hbox{#1}%
}\par\vskip 2cm}

%\def\@makeschapterhead#1{\hbox{%
%    \huge
%    \LignesVerticales
%    \hspace{0.5cm}
%    \hbox{#1}%
%}\par\vskip 2cm}

% Configuración del formato de cabecera y pie de página.
\pagestyle{fancy}
\fancyhf{}							% Elimina la configuración actual de la cabecera.
\fancyfoot{}						% Elimina la configuración actual del pie de página.
\fancyhead[LO]{\leftmark}	 	% Páginas impares ........: en la parte izquierda del encabezado aparecerá el nombre del capítulo.
\fancyhead[RE]{\rightmark}	 	% Páginas pares ..........: en la parte derecha del encabezado aparecerá el nombre de la sección.
\fancyfoot[RO,LE]{\thepage} 	% Númeración de páginas ..: impares en la derecha, pares en la izquierda.

% Para no incluir el encabezado en páginas en blanco.
\makeatletter
  \def\cleardoublepage{\clearpage\if@twoside \ifodd\c@page\else
  \vspace*{\fill}
    \thispagestyle{empty}
    \newpage
    \if@twocolumn\hbox{}\newpage\fi\fi\fi}
\makeatother

% Comando para insertar una página vacía completa.
\newcommand{\paginavaciacompleta}{\newpage{\thispagestyle{empty}\cleardoublepage}}

\hyphenation{matching}

### 2019-12-17 src/controller/ControllerSPAMDA.java

@@ -1,0 +1,177 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
[…]

### 2019-12-17 src/controller/ControllerViewAbout.java

@@ -1,0 +1,180 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/controller/ControllerViewBuoyData.java

@@ -1,0 +1,1228 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/controller/ControllerViewDatasetFileConverter.java

@@ -1,0 +1,789 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/controller/ControllerViewIndex.java

@@ -1,0 +1,386 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/controller/ControllerViewLoadMatchingFile.java

@@ -1,0 +1,762 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/controller/ControllerViewManageReanalysisData.java

@@ -1,0 +1,737 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/controller/ControllerViewNewDatasetFile.java

@@ -1,0 +1,750 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/controller/ControllerViewNewMatchingFile.java

@@ -1,0 +1,320 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/controller/ControllerViewNewPreprocessedDatasetFile.java

@@ -1,0 +1,528 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/controller/ControllerViewOpenARFFfile.java

@@ -1,0 +1,274 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/controller/ControllerViewOpenDatasetFile.java

@@ -1,0 +1,884 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/controller/ControllerViewSelectBuoyAttributes.java

@@ -1,0 +1,580 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/controller/ControllerViewSelectReanalysisFiles.java

@@ -1,0 +1,825 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/controller/package-info.java

@@ -1,0 +1,66 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/icons/package-info.java

@@ -1,0 +1,65 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/model/BuoyDatabase.java

@@ -1,0 +1,543 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
[…]

### 2019-12-17 src/model/DatasetFile.java

@@ -1,0 +1,334 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
[…]

### 2019-12-17 src/model/HistoricalFile.java

@@ -1,0 +1,394 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/model/Instance.java

@@ -1,0 +1,196 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/model/InstancesWeka.java

@@ -1,0 +1,325 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
[…]

### 2019-12-17 src/model/MatchingFinalDataset.java

@@ -1,0 +1,218 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/model/PreprocessDatasetFile.java

@@ -1,0 +1,694 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/model/ReanalysisDatabase.java

@@ -1,0 +1,564 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/model/ReanalysisFile.java

@@ -1,0 +1,432 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/model/RecoveringFilters.java

@@ -1,0 +1,419 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
[…]

### 2019-12-17 src/model/package-info.java

@@ -1,0 +1,65 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/utils/Utils.java

@@ -1,0 +1,871 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
+	Notice of license in the file SLF4J-LICENSE
+-Library WEKA, version 3.8.1
[…]

### 2019-12-17 src/utils/WekaExplorer.java

@@ -1,0 +1,194 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/utils/package-info.java

@@ -1,0 +1,65 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
[…]

### 2019-12-17 src/view/About.java

@@ -1,0 +1,333 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
[…]

### 2019-12-17 src/view/BuoyData.java

@@ -1,0 +1,596 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/DatasetFileConverter.java

@@ -1,0 +1,450 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/Index.java

@@ -1,0 +1,273 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
[…]

### 2019-12-17 src/view/LoadMatchingFile.java

@@ -1,0 +1,425 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/ManageBuoys.java

@@ -1,0 +1,3675 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/ManageReanalysisData.java

@@ -1,0 +1,328 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/NewDatasetFile.java

@@ -1,0 +1,540 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/NewMatchingFile.java

@@ -1,0 +1,293 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/NewPreprocessedDatasetFile.java

@@ -1,0 +1,303 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/OpenARFFfile.java

@@ -1,0 +1,290 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
[…]

### 2019-12-17 src/view/OpenDatasetFile.java

@@ -1,0 +1,412 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/RunCreationDataset.java

@@ -1,0 +1,1012 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/view/RunMatching.java

@@ -1,0 +1,1969 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/view/SelectBuoyAttributes.java

@@ -1,0 +1,363 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/SelectReanalysisFiles.java

@@ -1,0 +1,384 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewAbout.java

@@ -1,0 +1,139 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewBuoyData.java

@@ -1,0 +1,283 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewDatasetFileConverter.java

@@ -1,0 +1,194 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewIndex.java

@@ -1,0 +1,138 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewLoadMatchingFile.java

@@ -1,0 +1,177 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewManageBuoys.java

@@ -1,0 +1,1497 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewManageReanalysisData.java

@@ -1,0 +1,149 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewNewDatasetFile.java

@@ -1,0 +1,235 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewNewMatchingFile.java

@@ -1,0 +1,134 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewNewPreprocessedDatasetFile.java

@@ -1,0 +1,143 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewOpenARFFfile.java

@@ -1,0 +1,145 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewOpenDatasetFile.java

@@ -1,0 +1,175 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewSelectBuoyAttributes.java

@@ -1,0 +1,167 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/view/interfaces/InterfaceViewSelectReanalysisFiles.java

@@ -1,0 +1,180 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
[…]

### 2019-12-17 src/view/interfaces/package-info.java

@@ -1,0 +1,66 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/view/package-info.java

@@ -1,0 +1,65 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/xml/DatasetInformation.java

@@ -1,0 +1,560 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2019-12-17 src/xml/MatchingInformation.java

@@ -1,0 +1,808 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
[…]

### 2019-12-17 src/xml/XMLFile.java

@@ -1,0 +1,227 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
[…]

### 2019-12-17 src/xml/package-info.java

@@ -1,0 +1,66 @@
+/*
+SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets
+
+Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+	Authors: 
+		Antonio Manuel Gomez Orellana, Juan Carlos Fernandez Caballero,
+		Manuel Dorado Moreno, Pedro Antonio Gutierrez Peña and 
+		Cesar Hervas Martinez.
+
+This program is free software: you can redistribute it and/or modify it under the
+terms of the GNU General Public License as published by the Free Software Foundation,
+either version 3 of the License, or (at your option) any later version.
+
+This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
+without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+See the GNU General Public License for more details.
+
+You should have received a copy of the GNU General Public License along with this program,
+in the file COPYING. If not, see <http://www.gnu.org/licenses/>.
+
+Additional permissions under GNU GPL version 3 section 7:
+1. Redistributions of source code, with or without modification, must retain
+the above full copyright notice as author attributions.
+
+2. Redistributions in binary form and/or the use of the documentation,
+with or without modification, must reproduce the above full copyright notice
+as author attributions in the documentation and/or materials provided with
+the distribution.
+
+3. Modified versions of source code and/or documentation, as well as binary
+distributions, must be marked in reasonable ways as different from the original version.
+
+4. Neither name of copyright holders nor the names of its contributors may be used
+to endorse or promote products derived from this software for publicity purposes
+without specific prior written permission.
+
+5. Redistribution and/or use of source code, binary format and documentation,
+with or without modification, could require indemnification of licensors
+and authors by anyone who conveys the material (or modified versions of it)
+with contractual assumptions of liability to the recipient, for any liability
+that these contractual assumptions directly impose on those licensors and authors.
+
+SPAMDA uses some external libraries. You can see their respective notices about license,
+copyright and disclaimer in the following files. For a more complete information about
+such licenses, see the distributions provided by their authors:
+-Library NetCDF Java, version 4.6.10
+	Notice of license in the file NetCDF-LICENSE
+-Library SLF4J, version 1.7.25
[…]

### 2020-02-06 latex_manuscript/elsevier/4_energy_conversion_and_management/spamda.tex
Sending to energy conversion and management journal. Latex and letters files.

--- 2019-12-17 latex_manuscript/elsevier/3_Journal_of_Environmental_Management/spamda.tex
@@
 
-\journal{Journal of Environmental Management}
+\journal{Energy Conversion and Management}
 
@@
 In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g.
-climate change, agriculture, energy, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: patterns in transmission of air pollution \cite{LU2019109377}, directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind energy assessment \cite{DVORAK20101244}, analysis of ecological typologies of large marine areas \cite{DELAHOZ201859}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+energy, climate change, agriculture, etc.).
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: photovoltaic power forecasting \cite{WANG2019443}, directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind and wave exploitation \cite{WEISS201843}, impact of atmospheric components on clear-sky solar irradiance estimation \cite{ANTONANZASTORRES2016122}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{PERERA2005291}, a modelling tool for water allocation and management is presented.
-An interactive system for multidisciplinary wind turbine site selection is developed in \cite{RAFIEE2018107}.
+In \cite{LIU2018516}, a software package to evaluate the performance of floating renewable energy systems is presented.
+A new wind turbine emulator for analysing the tower shadow and yaw error effects is developed in \cite{MOHAMMADI2018378}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and an intuitive tool for stormwater runoff evaluation and management is developed in \cite{CHEN201645}.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and a predictive tool to assess the energy performance in agrifood industries is developed in \cite{NUNES2014758}.
 
@@
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
+In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing offshore oil and gas platforms \cite{OLIVEIRAPINTO2019556} or seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
 WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them.
@@
 
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{MARTINSSILVA2019109368, TORREGROSSA20181061}.
+		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{TU20102938, WANG2019111799}.
 ML methodologies can be used not only by experienced computer scientists but also by other researchers.
@@
 			\item It facilitates data management and well-organised storage of the datasets.
-			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for environmental research.
+			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy and environmental research.
 			\item It includes an user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
@@
 These datasets will be ready to use as input for ML techniques in prediction tasks (classification or regression).
-As a result, the researchers will benefit from significant support when carrying out environmental modelling related to atmospheric, oceanic or energy studies, among others.
+As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
 Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.

### 2020-02-14 latex_user_manual/appendix/GPL3_SPAMDA_license.tex
Building the version of the paper for the journal "Advances in Engineering Software"

@@
 
-Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
+Copyright (c) 2017-2020 by AYRNA Research Group. https://www.uco.es/ayrna/
 	\begin{quote}
@@
 
+\vspace{0.50cm}
 Contact information:
 
-Juan Carlos Fernández Caballero, PhD.
+% \begin{quote}
+Antonio Manuel Gomez Orellana.
+
+email: am.gomez@uco.es
+
+Address: University of Córdoba, Department of Computer Science and Numerical Analysis, Rabanales Campus, AYRNA Research Group, Einstein Building, 3rd floor.
+Road Madrid-Cádiz, Km 396-A.
+14071 - Córdoba (Spain)
+% \end{quote}
+
+\vspace{0.50cm}
+
+% \begin{quote}
+Juan Carlos Fernández Caballero.
 
@@
 
-Address: University of Cordoba, Department of Computer Science
-and Numerical Analysis, AYRNA Research Group, Rabanales Campus,
-Einstein Building, 3rd floor. Road Madrid-Cadiz, Km 396-A.
-14071 - Cordoba (Spain).
+Address: University of Córdoba, Department of Computer Science and Numerical Analysis, Rabanales Campus, AYRNA Research Group, Einstein Building, 3rd floor.
+Road Madrid-Cádiz, Km 396-A.
+14071 - Córdoba (Spain)
+% \end{quote}

### 2020-02-27 latex_manuscript/elsevier/5_Advances_in_Engineering_Software/spamda.tex
Sending to Advances in Engineering Software journal. Latex and letters files.

--- 2020-02-06 latex_manuscript/elsevier/4_energy_conversion_and_management/spamda.tex
@@
 
-\journal{Energy Conversion and Management}
+\journal{Advances in Engineering Software}
 
@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: photovoltaic power forecasting \cite{WANG2019443}, directional analysis of sea storms \cite{LAFACE201545}, wind power ramp events prediction \cite{DORADOMORENO2017428}, offshore wind and wave exploitation \cite{WEISS201843}, impact of atmospheric components on clear-sky solar irradiance estimation \cite{ANTONANZASTORRES2016122}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: optimum layout design of onshore wind farms \cite{KALLIORAS20158}, directional analysis of sea storms \cite{LAFACE201545}, risk assessment of port areas exposed to intense wind actions \cite{REPETTO201829}, wind power ramp events prediction \cite{DORADOMORENO2017428}, structural health monitoring of dams \cite{KANG201960}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{LIU2018516}, a software package to evaluate the performance of floating renewable energy systems is presented.
-A new wind turbine emulator for analysing the tower shadow and yaw error effects is developed in \cite{MOHAMMADI2018378}.
+In \cite{KRATZIG201333}, a software tool to analysis solar updraft power plant design is presented.
+A review of software tools used for heating, ventilation, and air conditioning (HVAC) systems research, which are employed to study and reduce the energy consumption in buildings, is presented in \cite{OROSA2011846}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and a predictive tool to assess the energy performance in agrifood industries is developed in \cite{NUNES2014758}.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Su et al.
+\cite{SU20167} developed a large scale marine hydrological environmental data-oriented visualization software to simulate and show properties and change process of marine water environmental factors efficiently and intuitively.
 
@@
 
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{TU20102938, WANG2019111799}.
+		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{ZHOU20191,NOURANI2012127,GUO201816}.
 ML methodologies can be used not only by experienced computer scientists but also by other researchers.

### 2020-03-05 latex_manuscript/elsevier/5_Advances_in_Engineering_Software/spamda.tex
Sending to Advances in Engineering Software journal. Latex and letters files.

@@
 
-		This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union.
+		This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
 We also thank to NVIDIA Corporation for the transfer of computational resources for research works.

### 2020-03-11 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/spamda.tex
Sending to Sustainable Energy Technologies and Assessments. Latex and letters files.

--- 2020-03-05 latex_manuscript/elsevier/5_Advances_in_Engineering_Software/spamda.tex
@@
 
-\journal{Advances in Engineering Software}
+\journal{Sustainable Energy Technologies and Assessments}
 
@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: optimum layout design of onshore wind farms \cite{KALLIORAS20158}, directional analysis of sea storms \cite{LAFACE201545}, risk assessment of port areas exposed to intense wind actions \cite{REPETTO201829}, wind power ramp events prediction \cite{DORADOMORENO2017428}, structural health monitoring of dams \cite{KANG201960}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{KRATZIG201333}, a software tool to analysis solar updraft power plant design is presented.
-A review of software tools used for heating, ventilation, and air conditioning (HVAC) systems research, which are employed to study and reduce the energy consumption in buildings, is presented in \cite{OROSA2011846}.
+In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed.
+A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Su et al.
-\cite{SU20167} developed a large scale marine hydrological environmental data-oriented visualization software to simulate and show properties and change process of marine water environmental factors efficiently and intuitively.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
+\cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.
 
@@
 
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{ZHOU20191,NOURANI2012127,GUO201816}.
+		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816}.
 ML methodologies can be used not only by experienced computer scientists but also by other researchers.

### 2020-03-16 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/spamda.tex
Sending to Sustainable Energy Technologies and Assessments

@@
 
-		\title{SPAMDA: A software tool to manage meteorological data for environmental modelling}
+		%\title{SPAMDA: A software tool to manage meteorological data for environmental modelling}
+
+		\title{A new software tool to facilitate environmental modelling procedures based on meteorological data management}
 
@@
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modeling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
 \cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.

### 2020-04-17 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/2 - envio/spamda.tex
Reorganización de los envíos, de la cover letter y la carta de respuesta al revisor.
Envío a la revista: Sustainable_Energy_Technologies_and_Assessments

--- 2020-03-16 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/spamda.tex
@@
 		\begin{abstract}
-			Meteorological data play an important role for the comprehension of the environment, and they are extensively used to perform environmental learning.
-Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation implies a lot of time and effort by the researchers.
-This paper presents a software for creating datasets using meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-As the designed software is able to simplify the creation of the datasets (ready to be used as inputs for ML techniques in prediction tasks) and reduce the time needed for this task, researchers can benefit from it in order to achieve a more efficient exploitation and protection of environmental resources.
+			Meteorological data are extensively used to perform environmental learning.
+Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
+This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
+Such datasets can be created using buoys and reanalysis data by customisables procedures, in terms of predictive and objective variables and temporal resolution, to be used with ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
 		\end{abstract}
 
+%		\begin{keyword}
+%			Meteorological data\sep Reanalysis data\sep Pre-processing\sep Creating datasets\sep Environmental prediction\sep Marine energy
+%		\end{keyword}
+
 		\begin{keyword}
-			Meteorological data\sep Reanalysis data\sep Pre-processing\sep Creating datasets\sep Environmental prediction\sep Marine energy
+			Meteorological data\sep Reanalysis data\sep Creating datasets\sep Environmental prediction\sep Marine energy
 		\end{keyword}
@@
 
-		The main purpose of this paper is to present a software tool able to prevent researchers from performing this repetitive work and greatly simplify all the steps involved in the creation of datasets.
-The meteorological data considered for the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
+		The main purpose of this paper is to present a new open source tool for the creation of datasets integrated by meteorological variables from two sources of information.
+Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed, and it also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps.
+The meteorological data used by the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
 The software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
@@
 
-		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, combining meteorological data from NDBC and NNRP.
-The advantages that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
-		\begin{itemize}
-			\item The generation of datasets becomes a very easy and customizable task, by means of the selection of different input parameters.
-			\item The created datasets can be easily used by ML tools.
-			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks.
+		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP.
+The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
+		\begin{itemize}
+
+			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
+			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious and pre-processing procedures that implies a lot of time and effort.
+			\item It manages the extensive casuistry of the data integration which can lead to incomplete datasets.
+			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
 			\item It provides information about the quality and quantity of the data.
-			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets, such as time and geographical coordinates conversion or missing values handling, among others.
 			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
 			\item It facilitates data management and well-organised storage of the datasets.
+			\item The created datasets can be easily used by ML tools.
 			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy and environmental research.
@@
 		\end{itemize}
+
+		Therefore, the functionalities and characteristics that SPAMDA offers make it a novel support tool to researchers when carrying out environmental modelling related to energy, design of production systems as WECs (that need a proper prediction of waves in order to maximise the wave energy extraction) and atmospheric or oceanic studies, among others.
 
@@
 
-		A software tool for creating datasets using meteorological data from NDBC and NNRP has been presented.
-These datasets will be ready to use as input for ML techniques in prediction tasks (classification or regression).
+		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
+These datasets, that will be ready to use as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
 As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
-Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.
+Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.
 In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.

### 2020-04-21 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/2 - envio/spamda.tex
Envío a la revista: Sustainable_Energy_Technologies_and_Assessments

@@
 This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-Such datasets can be created using buoys and reanalysis data by customisables procedures, in terms of predictive and objective variables and temporal resolution, to be used with ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
+Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution.
+These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
 		\end{abstract}
@@
 		\begin{itemize}
-
 			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
-			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious and pre-processing procedures that implies a lot of time and effort.
-			\item It manages the extensive casuistry of the data integration which can lead to incomplete datasets.
-			\item It avoids possible researcher errors in the intermediate steps of the process of creation of the datasets, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
+			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
+			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
+			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
 			\item It provides information about the quality and quantity of the data.
@@
 			\item The created datasets can be easily used by ML tools.
-			\item Its modular design allows the implementation of new functional modules for managing meteorological data from others sources for renewable energy and environmental research.
-			\item It includes an user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
+			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
+			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
 			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
@@
 		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
-These datasets, that will be ready to use as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
+These datasets, that will be ready to be used as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
 As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.

### 2020-09-09 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Adaptación a la plantilla de Energies para un Special Issue

--- 2020-04-21 latex_manuscript/elsevier/6_Sustainable_Energy_Technologies_and_Assessments/2 - envio/spamda.tex
@@
-% !TeX spellcheck = en_GB
-\documentclass[review]{elsarticle}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+
+%=================================================================
+\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+
+% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% biologics, jmp, eng, jor, nursrep, biophysica, gastroent, jox, adolescents, hygiene, taxonomy, business, nanomanufacturing, geography, compoundsacoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, applmech, applnano, applsci, arts, asc, asi, atmosphere, atoms, automation, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomedinformatics, biomimetics, biomolecules, biosensors, bloods, brainsci, breath, buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, cyber, dairy, data, dentistry, dermatopathology, designs, diabetology, diagnostics, digital, diseases, diversity, drones, earth, econometrics, ecologies, economies, education, ejbc, ejihpe, electricity, electrochem, electronicmat, electronics, endocrines, energies, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, fuels, futureinternet, futurephys, galaxies, games, gardens, gases, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrogen, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, journalmedia, jpm, jrfm, jsan, land, languages, laws, life, literature, livers, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, obesities, oceans, ohbm, osteology, optics, organics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, surfaces, surgeries, suschem, sustainability, world, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj, women
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{xx}
+\issuenum{1}
+\articlenumber{5}
+\pubyear{2020}
+\copyrightyear{2020}
+%\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+%\updates{yes} % If there is an update available, un-comment this line
+
+%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
+%\continuouspages{yes}
+
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
 
@@
 \usepackage[T1]{fontenc}
-\usepackage{lineno}
-\usepackage{amsmath}
-\usepackage{hyperref}
 
 \epstopdfsetup{outdir=./figures/}
-
 \definecolor{gray090}{gray}{0.90}
 
-\modulolinenumbers[5]
-
-\journal{Sustainable Energy Technologies and Assessments}
-
-\bibliographystyle{elsarticle-num}
-
-\begin{document}
-
-	\begin{frontmatter}
-
-		%\title{SPAMDA: A software tool to manage meteorological data for environmental modelling}
-
-		\title{A new software tool to facilitate environmental modelling procedures based on meteorological data management}
-
-		\author[UCO]{Antonio M. G\'omez-Orellana\corref{AMGO}}
-		\cortext[AMGO]{Corresponding author.}
-		\ead{am.gomez@uco.es}
-
-		\author[UCO]{Juan C. Fern\'andez\corref{JCFC}}
-		\cortext[JCFC]{Corresponding author.}
-		\ead{jfcaballero@uco.es}
-
-		\author[UCO]{Manuel Dorado-Moreno}
-		\ead{manuel.dorado@uco.es}
-
-		\author[UCO]{Pedro A. Guti\'errez}
-		\ead{pagutierrez@uco.es}
-
-		\author[UCO]{C\'esar Herv\'as-Mart\'inez}
-		\ead{chervas@uco.es}
-
-		\address[UCO]{Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}
-
-		\begin{abstract}
-			Meteorological data are extensively used to perform environmental learning.
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Nuevo titulo relacionado con el Special Issue. El actual es SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0001-8849-6036} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-9773-6783} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0002-2657-776X} % Add \orcidC{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0003-4564-1816} % Add \orcidD{} behind the author's name
+
+% Authors, for the paper (add full first names)
+%\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+\Author{Antonio Manuel G\'omez-Orellana $^{1}$*, Juan Carlos Fern\'andez $^{1}$*\orcidA{}, Manuel Dorado-Moreno $^{1}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1}$*\orcidD{}}
+
+% Authors, for metadata in PDF
+%\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Antonio Manuel G\'omez-Orellana, Juan Carlos Fern\'andez, Manuel Dorado-Moreno, Pedro Antonio Guti\'errez and C\'esar Herv\'as-Mart\'inez}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+%\address{%
+%$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
+%$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+\address{%
+$^{1}$ \quad Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}
+
+% Contact information of the corresponding author
+%\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+\corres{Correspondence: am.gomez@uco.es (A.M.G.-O); jfcaballero@uco.es (J.C.F.); manuel.dorado@uco.es (M.D.-M.); pagutierrez@uco.es (P.A.G.); chervas@uco.es (C.H.-M.) }
+
+% Current address and/or shared authorship
+%\firstnote{Current address: Affiliation 3} 
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{Meteorological data are extensively used to perform environmental learning.
 Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
@@
 Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution.
-These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.
-		\end{abstract}
-
-%		\begin{keyword}
-%			Meteorological data\sep Reanalysis data\sep Pre-processing\sep Creating datasets\sep Environmental prediction\sep Marine energy
-%		\end{keyword}
-
-		\begin{keyword}
-			Meteorological data\sep Reanalysis data\sep Creating datasets\sep Environmental prediction\sep Marine energy
-		\end{keyword}
-
-	\end{frontmatter}
-
-	\linenumbers
+These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
+
+% Keywords
+\keyword{Meteorological data; Reanalysis data; Creating datasets; Environmental prediction; Marine energy; AQUI TAMBIEN HABRA QUE VER QUE PONEMOS ???????????}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%\setcounter{secnumdepth}{4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\setcounter{section}{-1} %% Remove this when starting to work on the template.
 
@@
 
-	\section*{Acknowledgments}
-
-		This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
-We also thank to NVIDIA Corporation for the transfer of computational resources for research works.
-
-		The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.
-
     \section*{Additional material}
@@
 
-	\section*{References}
-
-		\bibliography{bibfile}
-
-	\section*{Appendix A. Managing incomplete data}\label{app:AppendixA}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work reported.}
+
+\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
+We also thank to NVIDIA Corporation for the transfer of computational resources for research works.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\acknowledgments{The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}
+
+\conflictsofinterest{The authors declare no conflict of interest.
+The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\abbreviations{The following abbreviations are used in this manuscript:\\
+
+% \noindent 
+% \begin{tabular}{@{}ll}
+% MDPI & Multidisciplinary Digital Publishing Institute\\
+% DOAJ & Directory of open access journals\\
+% TLA & Three letter acronym\\
+% LD & linear dichroism
+% \end{tabular}}
+
+\noindent 
+\begin{tabular}{@{}ll}
+$F_e$ & Flux of energy\\
+$H_s$ & Significant wave height\\
+$T_e$ & Wave energy period\\
+$p_0$ & Geographical location of the buoy\\
+$p_j$ & Geographical location of each reanalysis node\\
+$lat$ & Latitude of the point\\
+$lon$ & Longitude of the point\\
+$o_{t}$ & The attribute to be predicted at the time instant to study\\
+$\Delta t$ & The prediction horizon\\
+$\mathbf{b}_t$ & The vector containing the selected NDBC variables\\
+$\mathbf{r}_{t}$ & The vector containing the selected reanalysis variables\\
+\end{tabular}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+\appendixtitles{yes}
+
+\appendix
+\section{Managing incomplete data}\label{app:AppendixA}
 		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+%\externalbibliography{yes}
+%\bibliography{your_external_BibTeX_file}
+\externalbibliography{yes}
+\bibliography{bibfile}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+%\begin{thebibliography}{999}
+% Reference 1
+%\bibitem[Author1(year)]{ref-journal}
+%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+%\bibitem[Author2(year)]{ref-book}
+%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+%\end{thebibliography}
+
+% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2020-09-15 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para el Special Issue de Energies

@@
 % Full title of the paper (Capitalized)
-\Title{Nuevo titulo relacionado con el Special Issue. El actual es SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets}
+\Title{\textcolor{red}{A Software Tool for building Suitable Datasets for Softcomputing and Machine Learning Techniques from Meteorological Data Integration}}
 
@@
 \abstract{Meteorological data are extensively used to perform environmental learning.
-Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
+\textcolor{red}{Soft Computing (SC) and} Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
 This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
 Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution.
-These datasets can be used by ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
+These datasets can be used by \textcolor{red}{SC and} ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
 
 % Keywords
-\keyword{Meteorological data; Reanalysis data; Creating datasets; Environmental prediction; Marine energy; AQUI TAMBIEN HABRA QUE VER QUE PONEMOS ???????????}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
+\keyword{Meteorological data; Reanalysis data; Environmental prediction; Marine energy; \textcolor{red}{Renewable Energy resource evaluation};  \textcolor{red}{Suitable Datasets for user models};}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
 
@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415} \textcolor{red}{Poner o sustituir aquí por alguna referencia de Energies}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
 
-		On the other hand, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
+		\textcolor{red}{Once good quality and well-formed data are obtained, these can be used to extract information and building prediction models that explain the behavior of a certain problem.
+Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modeling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data, you will not obtain quality models.}
+
+		\textcolor{red}{Continuing with this line}, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
 In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution.
@@
 
-		Currently, and as a support to traditional study procedures, Machine Learning (ML) techniques \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816}.
-ML methodologies can be used not only by experienced computer scientists but also by other researchers.
+		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816} \textcolor{red}{(Sustituir algunas de las 5 referencias anteriores por algunas de Energies donde se apliquen tecnicas de SC y ML)}.
+\textcolor{red}{SC and} ML methodologies can be used not only by experienced computer scientists but also by other researchers.
 For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML algorithms.
@@
 		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP.
-The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
+The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
 		\begin{itemize}
@@
 			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
+\textcolor{red}{Decir que permite un estudio previo o sondeo antes de hacer estudios más exhaustivos}
 			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
@@
 			\item It provides information about the quality and quantity of the data.
-			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.
+			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.\textcolor{red}{Hay que quitarle importancia al tema del preprocesamiento, para que no nos comparen con eso.}
 			\item It facilitates data management and well-organised storage of the datasets.
@@
 These aspects have to be taken into account when creating the datasets.
-This casuistry is explained in detail in Appendix A \ref{app:AppendixA}.
+This casuistry is explained in detail in Appendix \ref{app:AppendixA}.
 
@@
 
+				\textcolor{red}{Aquí en esta subsección hay que intentar darle poca importancia respecto a que el software no se centra en los filtros de preprocesamiento}
+
 				Data pre-processing prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms.
@@
 
-				The reader can check in Appendix A \ref{app:AppendixA} an example with a more complex case of the procedure.
+				The reader can check in Appendix \ref{app:AppendixA} an example with a more complex case of the procedure.
 				\begin{figure}[ht!]
@@
 \section{Managing incomplete data}\label{app:AppendixA}
+
+        \textcolor{red}{¿Como podriamos ponerle otro titulo para que no se asemeje a manual de usuario?
+Tenemos que mirar a ver si podemos tambien cambiar el titulo de alguna otra seccion en cuanto a ese fin.}
+
 		In this appendix, we describe how SPAMDA deals with incomplete data when creating intermediate datasets and performing the matching process.

### 2020-09-22 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para el Special Issue de Energies

@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, assessment and control of wind turbine \cite{DAHHANI2018167}, wind power ramp events prediction \cite{DORADOMORENO2017428}, evaluation of salinity-gradient energy harvesting \cite{REYESMENDOZA2020100645}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415} \textcolor{red}{Poner o sustituir aquí por alguna referencia de Energies}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, \textcolor{red}{estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}}, wind power ramp events prediction \cite{DORADOMORENO2017428}, \textcolor{red}{sea surface temperature prediction \cite{He2020}}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 
+		\textcolor{red}{de aqui quitar dos sobre la 9, la 12 y la 13}
 		\textcolor{red}{Continuing with this line}, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.

### 2020-09-28 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para Special Issue Energies

@@
 % Full title of the paper (Capitalized)
-\Title{\textcolor{red}{A Software Tool for building Suitable Datasets for Softcomputing and Machine Learning Techniques from Meteorological Data Integration}}
+\Title{\textcolor{red}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case study for predicting significant wave height in the Gulf of Alaska}}
 
@@
 		\textcolor{red}{Once good quality and well-formed data are obtained, these can be used to extract information and building prediction models that explain the behavior of a certain problem.
-Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modeling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data, you will not obtain quality models.}
-
-		\textcolor{red}{de aqui quitar dos sobre la 9, la 12 y la 13}
+Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modeling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.}
+
 		\textcolor{red}{Continuing with this line}, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
-In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution.
+\textcolor{red}{In \cite{Neeraj2020} a software package in R called ``ForecastTB'' is developed to compare the accuracy of different forecasting methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modeling.}
+		%In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. 
 		In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented.
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed.
+		\textcolor{red}{In \cite{Roberta2020} a new software named ``Storage LCA Tool`` is presented as tool for comparing PCM - phase change materials- storage systems  with traditional systems that do not entail energy storage, being beneficial in order to support decision-making on energy concepts for buildings.}
+		%In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed. 
 		A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}.
@@
 		Marine energy prediction is currently a hot topic where meteorological data is used in.
-Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{BHATTACHARYA2017157, BREDE201344}.
+Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \textcolor{red}{\cite{en12050787}}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
+Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \textcolor{red}{\cite{en12091657, BREDE201344}}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
@@
 Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction.
+As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \textcolor{red}{\cite{en80910370}}.
 For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-Additionally, wave predictions are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
-
-		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816} \textcolor{red}{(Sustituir algunas de las 5 referencias anteriores por algunas de Energies donde se apliquen tecnicas de SC y ML)}.
+Additionally, wave predictions \textcolor{red}{\cite{en11010011}} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
+
+		\textcolor{red}{VOY POR AQUI!!!!!!}
+
+		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816} \textcolor{red}{(Sustituir algunas de las 5 referencias anteriores por algunas de Energies donde se apliquen tecnicas de SC y ML, principalmente por la 31 y la 32)}.
 \textcolor{red}{SC and} ML methodologies can be used not only by experienced computer scientists but also by other researchers.
@@
 
-								\textbf{Class}&\textbf{Description}&\textbf{Inferior [}&\textbf{Superior )}&\\[0.20cm]
+								\textbf{Class}&\textbf{Description}&\textbf{Lower end [}&\textbf{Upper end )}&\\[0.20cm]
 
@@
 
-	\section{Case study}\label{sec:CaseStudy}
-
+	%\section{Case study}\label{sec:CaseStudy}
+	\section{\textcolor{red}{A case study applied to Gulf of Alaska}}\label{sec:CaseStudy}		
 		This section will describe how the application works in a practical approach.
@@
 			Although it has been performed a first classification approach, to illustrate the use of datasets created with SPAMDA for environmental modelling employing ML algorithms, both models have obtained a good performance despite the problem tackled is to predict wave height six hours in advance.
+
+			\textcolor{red}{Quizas deberiamos poner algun parrafo mas sobre las conclusiones de la experimentacion, mirar el paper mio de ``Significant wave height and energy flux range forecast with machine learning classifiers'' por si sacar algo.}
 

### 2020-09-28 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para Special Issue Energies

@@
 % Full title of the paper (Capitalized)
-\Title{\textcolor{red}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case study for predicting significant wave height in the Gulf of Alaska}}
+\Title{\textcolor{red}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case  study for predicting significant wave height and flux of energy}}
 
@@
 % Keywords
-\keyword{Meteorological data; Reanalysis data; Environmental prediction; Marine energy; \textcolor{red}{Renewable Energy resource evaluation};  \textcolor{red}{Suitable Datasets for user models};}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
+\keyword{Environmental Prediction; \textcolor{red}{Renewable Energy Resource Evaluation}; Meteorological Data; Reanalysis Data; Marine Energy;   \textcolor{red}{Suitable Datasets for Soft Computing Models}}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
 
@@
 
-		\textcolor{red}{VOY POR AQUI!!!!!!}
-
-		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \cite{Alpaydin:2004:IML:1036287,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{DELFANI2019100559,NASRUDDIN201948,GUO201816} \textcolor{red}{(Sustituir algunas de las 5 referencias anteriores por algunas de Energies donde se apliquen tecnicas de SC y ML, principalmente por la 31 y la 32)}.
+		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \textcolor{red}{\cite{Sang-Yong2014,Bishop:2006:PRM:1162264}} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \textcolor{red}{\cite{Fi-John2019,Mosavi2019,GUO201816}}.
 \textcolor{red}{SC and} ML methodologies can be used not only by experienced computer scientists but also by other researchers.
@@
 Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.
+
+		\textcolor{red}{VOY POR AQUI!!!!!!}
 
@@
 			Now, it will be described how to obtain wave classification models using the final dataset previously created with SPAMDA.
-The modelling will be performed using WEKA as ML tool, which can be opened through SPAMDA, as shown in Fig.
+The modelling will be performed using WEKA as \textcolor{red}{SC and} ML tool, which can be opened through SPAMDA, as shown in Fig.
 \ref{fig:openigFinalDatasetWeka}.
@@
 
-			\textcolor{red}{Quizas deberiamos poner algun parrafo mas sobre las conclusiones de la experimentacion, mirar el paper mio de ``Significant wave height and energy flux range forecast with machine learning classifiers'' por si sacar algo.}
+			\textcolor{red}{Quizas deberiamos poner algun parrafo mas sobre las conclusiones de la experimentacion, mirar el paper mio de ``Significant wave height and energy flux range forecast with machine learning classifiers'' por si sacar algo.
+Y quizas deberiamos hacer la experimentacion tambien con el flujo de energia en forma de regresion, ya que el special issue se titula ``Soft Computing Techniques in Energy System''.
+¿Esta el dataset en el repositorio?}
 
@@
 
-\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}
+\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.
+\textcolor{red}{ESTO DE LAS CONTRIBUCIONES HAY QUE VER SI LO CAMBIAMOS PARA QUE NO SEA TAN LARGO Y CON TANTA MORRALLA Y COMO PONERLO PARA QUE TE PUEDA VENIR BIEN POR SI LO NECESITASES PARA EL MASTER}}
 
@@
 
+%\funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
+We also thank to NVIDIA Corporation for the transfer of computational resources for research works.}
+
 \funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
-We also thank to NVIDIA Corporation for the transfer of computational resources for research works.}
+\textcolor{red}{VER SI HAY QUE ACTUALIZARLO Y PONER ALGUN OTRO}}
 

### 2020-09-29 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Cambios para Special Issue Energies

@@
 
-		\textcolor{red}{VOY POR AQUI!!!!!!}
-
-		The information to create these datasets can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
+		The information to create these datasets \textcolor{red}{related with MRE} can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
 Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
-Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by the researchers to manually create the datasets with the needed information.
-Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
-In summary, many important details and different intermediate steps have to be considered when creating suitable datasets for ML techniques, resulting in an extremely tedious task.
+Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document and \textcolor{red}{in which the main feature of the tool presented in this work is focused}, has to be carried out by the researchers to manually create the datasets with the needed information.
+Moreover, depending on the subject and the \textcolor{red}{SC and }ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
+In summary, many important details and different intermediate steps have to be considered when creating suitable datasets, \textcolor{red}{especially when there is data integration}, resulting in an extremely tedious task.
 
 		The main purpose of this paper is to present a new open source tool for the creation of datasets integrated by meteorological variables from two sources of information.
-Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed, and it also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps.
+Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed, and it also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps\textcolor{red}{, at least as a preliminary study of certain areas}.
 The meteorological data used by the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
@@
 As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows the researchers focus on the study of the meteorological aspects of the observations.
-The datasets obtained are ready to be used as input for ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
+The datasets obtained are ready to be used as input for \textcolor{red}{SC and }ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
 These datasets contain one or more meteorological variables as inputs and one variable as target (variable to be predicted).
@@
 
-		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP.
-The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
-		\begin{itemize}
-			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
-			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
-\textcolor{red}{Decir que permite un estudio previo o sondeo antes de hacer estudios más exhaustivos}
+		Up to now, and to the best of our knowledge, this is the first software tool \textcolor{red}{with GUI (Graphical User Interface) }addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP \textcolor{red}{and the casuistry that it entails}.
+\textcolor{blue}{(desde ''Up to now...
+hasta ...entails lo quitaría)} The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
+		\begin{itemize}
+			%\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
+			\item The generation of datasets \textcolor{red}{with SPAMDA} becomes a very easy and customisable task by means of the selection of different input parameters, such as predictive and objective variables, classification and regression, output discretisation \textcolor{red}{(even for ordinal regression)} or prediction horizon, among others.
+			\item The created datasets can be easily used by \textcolor{red}{SC and }ML tools.
+			\item It provides information about the quality and quantity of the data.
+			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures \textcolor{red}{as scripts}, that imply a lot of time and effort, \textcolor{red}{at least in early stages of the research}.
+\textcolor{red}{Por tanto, esta herramienta podría ser tambíen de utilidad en estudios previos realizados sobre diferentes zonas oceánicas y costeras where some kind of environmental prediction is needed.}
+			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), \textcolor{red}{different data collection times, }among others.
 			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
-			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
-			\item It provides information about the quality and quantity of the data.
-			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.\textcolor{red}{Hay que quitarle importancia al tema del preprocesamiento, para que no nos comparen con eso.}
+			\item \textcolor{red}{Possibility of using one or more reanalysis points close to the study location.}
+			\item It includes different pre-processing tasks, such as normalisation and missing data recovery, \textcolor{red}{aunque esa no es la característica en la que se centra la herramienta, sino en el preprocesamiento y cauistica generada por la integración de los datos provenientes del NDBC y NNRP.
+Un preprocesamiento mucho más amplio se puede realizar posteriormente con otras herramientas, incluso con el propio módulo de Weka que se integra en SPAMDA.}
 			\item It facilitates data management and well-organised storage of the datasets.
-			\item The created datasets can be easily used by ML tools.
-			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
+			\item \textcolor{red}{Providing the source code to the scientific community and }its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
 			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
@@
 
-				\textcolor{red}{Aquí en esta subsección hay que intentar darle poca importancia respecto a que el software no se centra en los filtros de preprocesamiento}
+				\textcolor{red}{Aquí en esta subsección hay que intentar darle poca importancia respecto a que el software no se centra en los filtros de preprocesamiento, como se indico al final de la introduccion}
 
@@
 Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.
+\textcolor{red}{No se si deberiamos resaltar un poco esta funcionalidad, porque no se explota mucho en el paper creo, o al menos ni lo leen o ni la ven.}
 
@@
 Y quizas deberiamos hacer la experimentacion tambien con el flujo de energia en forma de regresion, ya que el special issue se titula ``Soft Computing Techniques in Energy System''.
-¿Esta el dataset en el repositorio?}
+¿Esta el dataset en el repositorio?.
+Esto es lo que nos llevaria mas tiempo, el resto del paper no debemos tocarlo mucho mas de lo que tenemos anotado.}
 

### 2020-11-17 latex_manuscript/MDPI/2_energies_special_issue/template.tex
Letters for Speccial Issue Energies

--- 2020-09-28 latex_manuscript/MDPI/2_energies_special_issue/template.tex
@@
 % Full title of the paper (Capitalized)
-\Title{\textcolor{red}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case  study for predicting significant wave height and flux of energy}}
+\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case  study for predicting significant wave height}
 
@@
 \abstract{Meteorological data are extensively used to perform environmental learning.
-\textcolor{red}{Soft Computing (SC) and} Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
+Soft Computing (SC) and Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
 This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
 Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution.
-These datasets can be used by \textcolor{red}{SC and} ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
+These datasets can be used by SC and ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
 
 % Keywords
-\keyword{Environmental Prediction; \textcolor{red}{Renewable Energy Resource Evaluation}; Meteorological Data; Reanalysis Data; Marine Energy;   \textcolor{red}{Suitable Datasets for Soft Computing Models}}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
+\keyword{Environmental Prediction; Renewable Energy Resource Evaluation; Meteorological Data; Reanalysis Data; Marine Energy; Soft Computing}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
 
@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, \textcolor{red}{estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}}, wind power ramp events prediction \cite{DORADOMORENO2017428}, \textcolor{red}{sea surface temperature prediction \cite{He2020}}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}, wind power ramp events prediction \cite{DORADOMORENO2017428}, sea surface temperature prediction \cite{He2020}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
 
-		\textcolor{red}{Once good quality and well-formed data are obtained, these can be used to extract information and building prediction models that explain the behavior of a certain problem.
-Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modeling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.}
-
-		\textcolor{red}{Continuing with this line}, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
-\textcolor{red}{In \cite{Neeraj2020} a software package in R called ``ForecastTB'' is developed to compare the accuracy of different forecasting methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modeling.}
+		Once quality and well-formed data are obtained, these can be used to extract information and build prediction models that explain the behavior of a certain problem.
+Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modelling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.
+
+		Continuing with this line, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
+In \cite{Neeraj2020}, a software package in R called ``ForecastTB'' is developed to compare the accuracy of different forecasting methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modelling.
 		%In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. 
@@
 A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-		\textcolor{red}{In \cite{Roberta2020} a new software named ``Storage LCA Tool`` is presented as tool for comparing PCM - phase change materials- storage systems  with traditional systems that do not entail energy storage, being beneficial in order to support decision-making on energy concepts for buildings.}
+		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool`` is presented for comparing PCM - phase change materials- storage systems  with traditional systems that do not entail energy storage, being beneficial in order to support decision-making on energy concepts for buildings.
 		%In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed. 
@@
 		Marine energy prediction is currently a hot topic where meteorological data is used in.
-Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \textcolor{red}{\cite{en12050787}}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \textcolor{red}{\cite{en12091657, BREDE201344}}.
+Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
+Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{en12091657, BREDE201344}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
@@
 Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \textcolor{red}{\cite{en80910370}}.
+As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \cite{en80910370}.
 For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-Additionally, wave predictions \textcolor{red}{\cite{en11010011}} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
-
-		Currently, and as a support to traditional study procedures, \textcolor{red}{SC and ML techniques} \textcolor{red}{\cite{Sang-Yong2014,Bishop:2006:PRM:1162264}} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \textcolor{red}{\cite{Fi-John2019,Mosavi2019,GUO201816}}.
-\textcolor{red}{SC and} ML methodologies can be used not only by experienced computer scientists but also by other researchers.
+Additionally, wave predictions \cite{en11010011} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
+
+		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{Fi-John2019,Mosavi2019,GUO201816}.
+SC and ML methodologies can be used not only by experienced computer scientists but also by other researchers.
 For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML  algorithms.
@@
 
-		\textcolor{red}{VOY POR AQUI!!!!!!}
-
-		The information to create these datasets can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
+		The information to create these datasets related with MRE can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
 Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
 Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by the researchers to manually create the datasets with the needed information.
-Moreover, depending on the subject and the ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
-In summary, many important details and different intermediate steps have to be considered when creating suitable datasets for ML techniques, resulting in an extremely tedious task.
+Given that such process is of great relevance and has an extensive casuistry, the present work has been specially focused on it.
+Moreover, depending on the subject and the SC and ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
+In summary, many important details and different intermediate steps have to be considered when creating suitable datasets, specially when data integration is required, resulting in an extremely tedious task.
 
 		The main purpose of this paper is to present a new open source tool for the creation of datasets integrated by meteorological variables from two sources of information.
-Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed, and it also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps.
+Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed.
+It also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps, at least as a preliminary study in certain areas where some kind of environmental prediction is needed.
 The meteorological data used by the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
-The software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
+The open source software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
 As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows the researchers focus on the study of the meteorological aspects of the observations.
-The datasets obtained are ready to be used as input for ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
+The datasets obtained are ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
 These datasets contain one or more meteorological variables as inputs and one variable as target (variable to be predicted).
@@
 
-		Up to now, and to the best of our knowledge, this is the first software tool addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP.
-The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
-		\begin{itemize}
-			\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
-			\item It makes the researcher focus on environmental modelling, without having to worry about mechanical tasks, avoiding laborious pre-processing procedures that imply a lot of time and effort.
-\textcolor{red}{Decir que permite un estudio previo o sondeo antes de hacer estudios más exhaustivos}
-			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets.
-			\item It avoids possible researcher errors in the intermediate steps of the process, such as time and geographical coordinates conversion or missing values handling (dates or measurements not recorded), among others.
+		In order to address the problem previously discussed, meteorological data integration from NDBC and NNRP and the casuistry that it entails, SPAMDA offers to the researchers novelties and functionalities that will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
+
+		%Up to now, and to the best of our knowledge, this is the first software tool \textcolor{red}{with GUI (Graphical User Interface) }addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP \textcolor{red}{and the casuistry that it entails}.
+\textcolor{blue}{(desde ''Up to now...
+hasta ...entails lo quitaría)} The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
+		\begin{itemize}
+			%\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
+			\item The generation of datasets becomes a very easy and customisable task by means of the selection of different input parameters, such as predictive and objective variables, classification and regression, output discretisation (useful for ordinal regression) or prediction horizon, among others.
+			\item The created datasets can be easily used by SC and ML tools.
 			\item It provides information about the quality and quantity of the data.
-			\item It includes different pre-processing tasks, such as normalisation and missing data recovery.\textcolor{red}{Hay que quitarle importancia al tema del preprocesamiento, para que no nos comparen con eso.}
+			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a lot of time and effort in early stages of the research.
+			\item It avoids possible researcher errors in the intermediate steps of the process, such as geographical coordinates conversion, missing values handling (dates or measurements not recorded) or different temporal resolution of the data collected, among others.
+			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets, described in Appendix \ref{app:AppendixA}.
+			\item Possibility of selecting one or more reanalysis nodes near the localisation under study.
+			\item It also includes some basic pre-processing tasks, such as normalisation and missing data recovery.
 			\item It facilitates data management and well-organised storage of the datasets.
-			\item The created datasets can be easily used by ML tools.
 			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
@@
 
-				\textcolor{red}{Aquí en esta subsección hay que intentar darle poca importancia respecto a que el software no se centra en los filtros de preprocesamiento}
-
 				Data pre-processing prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms.
-In this way, the quality of dat can be improved prior to the learning phase, by applying pre-processing tasks (filters).
+In this way, the quality of data can be improved prior to the learning phase, by applying pre-processing tasks (filters).
 The result will be referenced as pre-processed datasets.
@@
 
+				Nevertheless, further pre-processing tasks can be performed after obtaining the final datasets by means of the Explorer environment of WEKA or other tools.
+
 			\subsection{Matching configuration}
@@
 
-				In order to use the data provided by the two sources of information described in Section \ref{sec:DataSources}, it is necessary to carry out a data integration process, denominated as the matching process in this document, to merge and format such data.
+				The automatic integration of the data provided by the two sources of information described in Section \ref{sec:DataSources}, to merge and format such data, is denominated as the matching process in this document.
+Such process is one of the most powerful and remarkable features of this software tool due to its great relevance and extensive casuistry.
+In this sense, SPAMDA has been developed to provide great flexibility to the researchers.
+
 				The matching procedure is performed using an intermediate or pre-processed dataset, which includes the measurements collected by a buoy from NDBC, and the needed reanalysis data files from NNRP.
@@
 									\begin{equation}
-										w_i=\frac{\sum_{j=1}^{N} d(p_0,p_j)}{d(p_0,p_i)}, ~~i=1, \ldots, N.
+										w_i=\frac{d(p_0,p_i)}{\sum_{j=1}^{N} d(p_0,p_j)}, ~~i=1, \ldots, N.
 										\label{eq:weightedMean}
@@
 Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.
+In this way, researchers can perform comparison studies depending on the reanalysis node considered, to achieve better performance for the problem under study.
 
@@
 	%\section{Case study}\label{sec:CaseStudy}
-	\section{\textcolor{red}{A case study applied to Gulf of Alaska}}\label{sec:CaseStudy}		
+	\section{A case study applied to Gulf of Alaska}\label{sec:CaseStudy}		
 		This section will describe how the application works in a practical approach.
@@
 
-			After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.
+			%After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.
+In Figure \ref{fig:main_view}, the main window is shown.
+In order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher has to select the option \textcolor{blue}{\textit{Manage reanalysis data}}.
+
+			After gathering the information described above, the researcher can open SPAMDA.
 In Figure \ref{fig:main_view}, the main window is shown.
@@
 			Now, it will be described how to obtain wave classification models using the final dataset previously created with SPAMDA.
-The modelling will be performed using WEKA as \textcolor{red}{SC and} ML tool, which can be opened through SPAMDA, as shown in Fig.
+The modelling will be performed using WEKA as SC and ML tool, which can be opened through SPAMDA, as shown in Fig.
 \ref{fig:openigFinalDatasetWeka}.
@@
 
-			\textcolor{red}{Quizas deberiamos poner algun parrafo mas sobre las conclusiones de la experimentacion, mirar el paper mio de ``Significant wave height and energy flux range forecast with machine learning classifiers'' por si sacar algo.
-Y quizas deberiamos hacer la experimentacion tambien con el flujo de energia en forma de regresion, ya que el special issue se titula ``Soft Computing Techniques in Energy System''.
-¿Esta el dataset en el repositorio?}
-
 		\subsection{Important remarks}
@@
 		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
-These datasets, that will be ready to be used as input for ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
+These datasets, that will be ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
 As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
-Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the ML technique to use), it avoids errors and reduces the time needed.
+Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the SC or ML technique to use), it avoids errors and reduces the time needed.
 In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
+
+		The case study illustrates how SPAMDA can be used by the researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
 
@@
 
-\authorcontributions{Conceptualization, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Formal analysis, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Pedro Antonio Gutiérrez and César Hervás-Martínez; Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Project administration, Pedro Antonio Gutiérrez and César Hervás-Martínez; Resources, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.
-\textcolor{red}{ESTO DE LAS CONTRIBUCIONES HAY QUE VER SI LO CAMBIAMOS PARA QUE NO SEA TAN LARGO Y CON TANTA MORRALLA Y COMO PONERLO PARA QUE TE PUEDA VENIR BIEN POR SI LO NECESITASES PARA EL MASTER}}
+\authorcontributions{Conceptualization, Formal analysis and Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Project administration, Resources and Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}
 
@@
 
-\funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.
-\textcolor{red}{VER SI HAY QUE ACTUALIZARLO Y PONER ALGUN OTRO}}
+\funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.}
 
@@
 \appendix
-\section{Managing incomplete data}\label{app:AppendixA}
-
-        \textcolor{red}{¿Como podriamos ponerle otro titulo para que no se asemeje a manual de usuario?
-Tenemos que mirar a ver si podemos tambien cambiar el titulo de alguna otra seccion en cuanto a ese fin.}
+\section{Managing the casuistry of incomplete data}\label{app:AppendixA}
 

### 2020-11-30 latex_manuscript/MDPI/2_energies_special_issue/envio-2/responseToReviewers/letter-response-reviewers-energies-1023087.tex
Adicion de la carta de respuesta a los revisores en latex

% !TeX spellcheck = en_GB
%\documentclass[a4paper,10pt]{article}
\documentclass[a4paper,twoside,11pt]{article}

\usepackage[utf8x]{inputenc}
%\usepackage[latin9]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage{palatino}
\usepackage{verbatim}
\usepackage{color}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{eurosym}
\usepackage{url}
% \usepackage[tight]{subfigure}
\usepackage{fancyhdr}
\usepackage{fancybox}
\usepackage{fancyvrb}
\usepackage{a4wide}
\usepackage[authoryear]{natbib}
\usepackage[misc,geometry]{ifsym} 
%\usepackage{wrapfig}
%\usepackage{graphicx}

\usepackage{graphicx}
\usepackage[caption=false]{subfig}

% Text color
\usepackage{color}
% For colors in Tables
\usepackage{colortbl} 
% Multirow
\usepackage{multirow}
% Long Tables
\usepackage{longtable} 
% Dash hline in tables
\usepackage{arydshln}
\usepackage{tcolorbox}
\usepackage[parfill]{parskip}
\usepackage{lineno}
\usepackage{bm}
\usepackage[mathscr]{euscript}

%\usepackage{graphicx}
%\usepackage[caption=false]{subfig}

\usepackage[linktoc=all,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}

%\setcounter{secnumdepth}{4}
%\setlength{\topmargin}{0cm}
%\setlength{\textheight}{23cm}
%\setlength{\textwidth}{16cm}
%\setlength{\oddsidemargin}{0.5cm}
%\setlength{\evensidemargin}{-0.4cm}
%\setlength{\marginparsep}{0cm}
%\setlength{\marginparwidth}{0cm}
%\setlength{\parskip}{4mm}
%\setlength{\baselineskip}{3cm}
%\setlength{\headsep}{1.3cm}
%\setlength{\unitlength}{1cm}
%%\setlength{\footskip}{1 cm}
%%\setlength{\headheight}{0cm}
%%\setlength{\headsep}{0cm}
%\setlength{\tolerance}{1000}
%\setlength{\pretolerance}{2000}
%\pagestyle{headings}

%\newcommand\myeq{\stackrel{\text{\tiny def}}{=}}

\newcommand{\myAuthors}{Antonio Manuel Gómez-Orellana$^1$, Juan Carlos Fernández$^1$, Manuel Dorado-Moreno$^1$, Pedro Antonio Gutiérrez$^1$ and César Hervás-Martínez$^1$.}

\newcommand{\myAuthorsShort}{A. M. Gómez-Orellana$^1$, J. C. Fernández$^1$, M. Dorado-Moreno$^1$, P. A. Gutiérrez$^1$ and C. Hervás-Martínez$^1$.}

\newcommand{\myEmail}{\Letter \quad  am.gomez@uco.es, jfcaballero@uco.es, manuel.dorado@uco.es, pagutierrez@uco.es, chervas@uco.es}

\newcommand{\myTitle}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height}

\newcommand{\myShortTitle}{Response to reviewers}

\newcommand{\myJournal}{Energies, Special Issue on ``Soft Computing Techniques in Energy System''}

\newcommand{\myRefPaper}{\textbf{Paper reference: energies-1023087}}

\newcommand{\myDepartament}{$^1$Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, Cordoba, Spain.}

\definecolor{gray090}{gray}{0.90}

%%%%%%%%%%%% headings setup %%%%%%%%%%%%
\pagestyle{fancy}
% clear styles
\fancyhf{}
% odd pages -> chapter name
\fancyhead[LO]{\leftmark}
% even pages
%\fancyhead[RE]{\myJournal -- \myShortTitle}
\fancyhead[RE]{\myJournal}
% odd/even pages
\fancyhead[RO,LE]{\thepage}

\renewcommand{\sectionmark}[1]{\markboth{\textbf{\thesection. #1}}{}}
%\renewcommand{\subsectionmark}[1]{\markright{\textbf{\thesection. #1}}}

% line under headings
\renewcommand{\headrulewidth}{0.6pt}
% line over foot
%\renewcommand{\footrulewidth}{0.6pt}

% increases heading weight
\setlength{\headheight}{1.5\headheight}
\fancyfoot{}

\newcommand{\newtext}[1]{\textcolor{red}{#1}}

% New revisor comment command
\newcounter{ecomments}[section]

\newcommand{\ecomment}[1]
{
	\stepcounter{comments}
%	\addcontentsline{toc}{subsection}{Editor \arabic{section} Comment \arabic{comments}}
	\begin{tcolorbox}[colback=black!5,colframe=white!45!black,title=Comment \arabic{comments}]
		#1
	\end{tcolorbox}
}

% New revisor comment command
\newcounter{comments}[section]

\newcommand{\rcomment}[1]
{
	\stepcounter{comments}
%	\addcontentsline{toc}{subsection}{Reviewer \arabic{section} Comment \arabic{comments}}
	\vspace{0.6cm}
	\begin{tcolorbox}[colback=black!5,colframe=white!45!black,title=Comment \arabic{comments}]
		#1
	\end{tcolorbox}
}

\begin{document}

\thispagestyle{plain}

\begin{center}
	{\LARGE\myTitle} \vspace{0.5cm} \\
	{\LARGE\myJournal} \vspace{0.5cm} \\
	{\Large\myRefPaper} \vspace{0.5cm} \\
% 	\today \vspace{0.5cm} \\
	\myAuthors \vspace{0.5cm} \\
	{\myDepartament} \vspace{0.5cm} \\
	{\myEmail}
\end{center}

%\tableofcontents

\section{Responses to the Editor}
We would like to thank the journal's Editor for giving us the opportunity to improve our paper with this review. Following the Editor's indications, we have prepared a revised version of the paper, where we have addressed the points required by the reviewers to obtain an improved version of the manuscript.

Please find below the responses to the reviewers' comments.

\section{Responses to reviewers' comments}
We sincerely thank the anonymous reviewers for their careful reviews, constructive comments and hints to improve the impact of this work. We honestly believe that the quality of the paper has increased after addressing the comments and issues raised by the reviewers.

Most important changes in the paper have been highlighted in \textcolor{blue}{blue color} in the revised version of the manuscript. Please, find below detailed responses to the comments.

Additionally, in order to improve the english language and style we have carried out a proof-reading.

% The most important changes made in the first revision of the paper concerning the initial version are  \textcolor{blue}{marked in blue}, while the changes made in this second revision  \textcolor{red}{are marked in red}. Also, this letter response attempts to clarify the questions asked by the reviewers.

% Most important changes in the paper have been highlighted in \textcolor{red}{red} color in the revised version of the manuscript. Please, find below detailed responses to the comments.

% Additionally, in order to improve the paper's readability, we have carried out a proof-reading, where different typos and weird expressions have been corrected.

\newpage
\section{Reviewer \#1}
\addtocounter{section}{-2}

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
This work develops a framework to collect, integrate, and preprocess meteorological observation data from NDBC and NNRP.
Moreover, this framework uses machine learning techniques to do predictions based on these pre-processed datasets. This work
saves researchers from tedious and repetitive data collection and pre-processing work. Also, the use of machine learning in this framework is very useful given that NDBC and NNRP contain missing data in their observation datasets.

\vspace{0.5cm}
My only comment is that the wave prediction from the machine learning technique is not well validated.
}

\textbf{Response:}
{
Thank you very much for your review, providing very interesting comments and suggestions.

Bla, bla, bla. \textbf{These clarifications are mainly described in the paper on page 1, lines 1-33, on page 2, lines 95-112, on page 3, lines 136-151 and 165-181, as well as in the ``Conclusions'' Section}.
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% Another Reviewer %%%%%%%%%%%%%%%%%%%%
\newpage
\clearpage
\addtocounter{section}{+2}
\section{Reviewer \#2}
\addtocounter{section}{-2}

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment{
Dear authors,
Thank you for submitting your paper to the Energies. I think your paper is an informative paper that can be published after a major revision.

\vspace{0.5cm}
It is not clear what the research gap that the paper is addressing. What is the objective of this paper? Please clarify somewhere clearly all your contributions.
}

\textbf{Response:}
{
First of all, we would like to thank Reviewer \#2 for his/her suggestions to improve the paper.

In our understanding, we think that this work contributes several ideas to the field of Machine Learning (ML):

\begin{itemize}
	\item Bla, bla, bla,
	\item Bla, bla, bla,
	\item Bla, bla, bla,
\end{itemize}

Bla, bla, bla, \textbf{This clarification has been rewritten in page 2, lines 53-61.}
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
The literature review is not goal-oriented. The process should be as follows:

\vspace{0.5cm}
i) Critical evaluation of the literature; ii) identifying the gap based on this critical evaluation of the literature; iii) proposing your hypothesis to address the identified gap; iv) posing the appropriate and relevant research question based on your proposed hypothesis, and finally explaining your proposed method to answer this research question. Therefore, you will have a systematic way of conducting your research. Right now, the literature review section has no clear objective.
}

\textbf{Response:}
{
Following the instructions of the reviewer, bla, bla, bla. \textbf{These clarifications have been included in the paper on page 11, lines 646-660, in the new Figure 6, and in the ``Conclusions'' Section, lines 672-682.}
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
A thorough editorial check and English improvement are needed. Please kindly proofread the entire manuscript.

}

\textbf{Response:}
{
A complete revision of the English has been carried out in this review step.
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
The conclusion part is also needed to be revised; which questions are answered, what is the value/originality/contribution of the paper, how the proposed method answers the research questions that previous methods are not able to answer?
}

\textbf{Response:}
{
Bla, bla, bla,

Ok, good point. There are two main reasons for choosing
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
It feels you need a king of aggregating results somewhere clearer.

}

\textbf{Response:}
{
Bla, bla, bla. The reviewer is right about this comment on input variables.....
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
Please provide more explanation for Figure 21. Also, there are too many figures in the paper. Please reconsider combining some of them together.

}
\textbf{Response:}
{
Bl, bla, bla. Following this suggestion by the reviewer, we have changed...
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
An adequate literature review and a clear gap identification have been tried to be conducted. However, authors have ignored some research that has been done in the area. I strongly recommend the authors to provide a more comprehensive literature review in the introduction section. The following papers are recommended:

\begin{itemize}
    \item Significant wave height forecasting via an extreme learning machine model integrated with improved complete ensemble empirical mode decomposition. Renewable and Sustainable Energy Reviews, 104, 281-295. 
    \item A wavelet-Particle swarm optimization-Extreme learning machine hybrid modeling for significant wave height prediction. Ocean Engineering, 213, 107777.
    \item Managing computational complexity using surrogate models: a critical review. Research in Engineering Design, 31(3), 275-298.
    \item Near real-time significant wave height forecasting with hybridized multiple linear regression algorithms. Renewable and Sustainable Energy Reviews, 132, 110003.
    \item Significant wave height and energy flux prediction for marine energy applications: A grouping genetic algorithm–Extreme Learning Machine approach. Renewable Energy, 97, 380-389.
    \item Outlook on biofuels in future studies: A systematic literature review. Renewable and Sustainable Energy Reviews, 134, 110326.
    \item Statistical models for improving significant wave height predictions in offshore operations. Ocean Engineering, 206, 107249.
    \item Regional ocean wave height prediction using sequential learning neural networks. Ocean Engineering, 129, 605-612.
    \item Ensemble of surrogates and cross-validation for rapid and accurate predictions using small data sets. AI EDAM, 33(4), 484-501.
    \item Prediction of significant wave height; comparison between nested grid numerical model, and machine learning models of artificial neural networks, extreme learning and support vector machines. Engineering Applications of Computational Fluid Mechanics, 14(1), 805-817.
\end{itemize}
}

\textbf{Response:}
{
Bla, bla, bla,
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
If you can, please make a small comparison between what did you do and what others did before, as a conclusion.

}

\textbf{Response:}
{
Bla, bla, bla,
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
The abstract is not deep enough and Is not well prepared. Please try to re-write it better. The problem should be clearly stated and the gap in which you are going to address the need to be clarified. Simply explain your contributions and key findings.
}

\textbf{Response:}
{
Bla, bla, bla. The reviewer is completely right. We have rewritten the abstract to better describe the application tackled.
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment
{
There are some errors in your reference list. Please check and fix the errors.

}

\textbf{Response:}
{
Bla, bla, bla. Thanks for this comment. The distribution....

\vspace{0.5cm}
We would like to thank you again for the excellent review carried out to our work.
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% Another Reviewer %%%%%%%%%%%%%%%%%%%%
\newpage
\clearpage

\addtocounter{section}{+2}
\section{Reviewer \#3}
\addtocounter{section}{-2}

%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
\rcomment{
The Authors present a new open source tool for the creation of datasets integrated by meteorological variables from two sources
of information. Basically, a user-friendly software has been developed and described on the basis of pivotal parameters, carefully selected for the specific study of the wave height in the Gulf of Alaska. The statistical model can be in principle extended to other meteorological measurements and monitoring.

\vspace{0.5cm}
This study is very technical and of interest for a specific audience. By the way, I don’t see any link between the topic of the manuscript and the journal Energies.

\vspace{0.5cm}
The software developed it is clearly presented as useful tool for meteorological application and no application to energy saving, production, conversion or similar it is presented.

\vspace{0.5cm}
Thus, I recommend to submit the menuscript to a more specialized journal.
}

\textbf{Response:}
{
First of all, we would like to thank Reviewer \#3 for his/her suggestions to improve the submitted manuscript.

\vspace{0.5cm}
VAMOS A VER PEDAZO DE HIJO DE PUTA, TE HAS QUERIDO QUITAR EL LEER LA REVISION DE UN PLUMAZO
}
%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2020-12-23 latex_manuscript/MDPI/2_energies_special_issue/envio-2/responseToReviewers/letter-response-reviewers-energies-1023087.tex
Version para el segundo envio a Energies Special Issue

@@
 \usepackage[english]{babel}
-
+\usepackage{gensymb}
 \usepackage{palatino}
@@
 %\usepackage{wrapfig}
-%\usepackage{graphicx}
 
@@
 \usepackage[linktoc=all,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}
+%\usepackage[linktoc=page,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}
 
@@
 
-\newcommand{\myTitle}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height}
+\newcommand{\myTitle}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height and Energy Flux}
 
@@
 
-\newcommand{\myJournal}{Energies, Special Issue on ``Soft Computing Techniques in Energy System''}
+\newcommand{\myJournal}{\textbf{Energies}}
 
@@
 
-%\tableofcontents
+\tableofcontents
 
 \section{Responses to the Editor}
-We would like to thank the journal's Editor for giving us the opportunity to improve our paper with this review.
-Following the Editor's indications, we have prepared a revised version of the paper, where we have addressed the points required by the reviewers to obtain an improved version of the manuscript.
+We would like to thank the Editor for giving us the opportunity to improve our paper with this review and resubmission.
+Following all indications, we have prepared a revised version of the paper, where we have addressed the points required by the reviewers to obtain an improved version of the manuscript.
+
+%\textcolor{red}{Summary of changes}
 
@@
 
-Additionally, in order to improve the english language and style we have carried out a proof-reading.
+Additionally, in order to improve the English quality and style we have carried out a proof-reading.
+% Additionally, in order to improve the paper's readability, we have carried out a proof-reading, where different typos and weird expressions have been corrected.
+%Additionally, in order to improve the english language and style we have carried out a review.
 
@@
 
-% Most important changes in the paper have been highlighted in \textcolor{red}{red} color in the revised version of the manuscript. Please, find below detailed responses to the comments.
-
-% Additionally, in order to improve the paper's readability, we have carried out a proof-reading, where different typos and weird expressions have been corrected.
-
-\newpage
+%\newpage
 \section{Reviewer \#1}
@@
 {
-Thank you very much for your review, providing very interesting comments and suggestions.
+Thank you very much for your review and for your comment, which we try to clarify below.
 
-Bla, bla, bla.
-\textbf{These clarifications are mainly described in the paper on page 1, lines 1-33, on page 2, lines 95-112, on page 3, lines 136-151 and 165-181, as well as in the ``Conclusions'' Section}.
+This work does not focus on carrying out an extensive experimentation and validation of results obtained by different ML and SC techniques, which should be combined with statistical tests to check if significant differences exist.
+On the contrary, this work presents a very versatile software tool that facilitates the automated building of meteorological datasets from which researchers can perform prediction studies and apply them to different tasks related to coastal and ocean engineering and to marine energy prediction.
+
+Nevertheless, although it is not the primary objective, this software tool also allows the use of ML and SC methods available in WEKA from SPAMDA itself or from an independent WEKA instance.
+In this way, researchers can carry out a preliminary study or even a complete study on the prediction problem which they are working on.
+Moreover, SPAMDA also gives the possibility to save the integrated datasets in \texttt{.csv} format, and the researchers could use these datasets with other ML and SC tools that they consider appropriate.
+
+%\textcolor{red}{TODO: Cambiar este párrafo por el definitivo sobre la ampliación del caso de estudio.
+We have also extended the case study of this paper with an additional experiment regarding the short-term prediction of the energy flux generated by the waves in a location in the Gulf of Alaska.
+This prediction is configurable respect to the prediction horizon, which could provide a prediction of how much energy could be obtained through WECs, for example, within the next 6 hours.
+The Reviewer \# 1 can view this extension in the \textbf{red lines and new figures in Section ``4.
+A case study applied to Gulf of Alaska''.
+Some lines (red lines) in Sections ``3.4.
+Matching configuration'' and ``3.5.
+Final datasets'' also had to be rewritten for this purpose, as well as unifying several figures.}}
+
+In any case, to further study the validation as requested by the reviewer, \textbf{we have extended Section ''4.
+A case study applied to Gulf of Alaska`` with a short-term forecasting problem of the energy flux generated by the waves in a specific location of that zone}.
+Besides, \textbf{Sections ''3.4.
+Matching configuration``} and \textbf{''3.5.
+Final datasets``} have been also rewritten for this purpose.
+In this way, it would be possible to perform prediction studies 6h in advance (note that the prediction horizon is customisable) of the amount of marine energy that could be extracted by the WECs.
+Given that this work does not focus on proposing new models or methodologies, a more extended validation or a comparison study of the results obtained in such example does not make sense.
+However, typical performance metrics as CCR (Accuracy), Kappa, RMSE (Root Mean Squared Error) or Correlation coefficient are shown.
+
+We would like to thank you again for the suggestion and review carried out to our work.
 }
@@
 %%%%%%%%%%%%%%%%%%%% Another Reviewer %%%%%%%%%%%%%%%%%%%%
-\newpage
+%\newpage
 \clearpage
@@
 
-In our understanding, we think that this work contributes several ideas to the field of Machine Learning (ML):
+The reviewer is right about this comment, in the original version, we did not adequately indicated the purpose of this work.
+We think that this work can be useful as a support for applications in the fields of marine energy, engineering and environment studies, by creating well-constructed datasets ready to be used with ML and SC prediction techniques.
+Datasets can be used both in classification and regression problems based on the integration of NDBC observations and NNRP reanalysis data.
 
+From our experience, manuscripts in marine energy and engineering that use ML and SC algorithms for prediction tasks do not provide software tools for incorporating and integrating the two data sources, considering the casuistry involved.
+These studies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, artificial neural networks, etc.), normally applying custom-made implementations or scripts in different programming languages.
+We have not found tools for building datasets ready to be used with any of those methodologies or algorithms by an automated and very versatile process.
+In this sense, it is convenient to indicate that the objective of these works is different than that of our manuscript.
+Although our software allows the use of ML and SC algorithms available in WEKA, datasets can also be exported to \texttt{.csv} format to researchers can use other tools appart from WEKA, and we neither focus on those methodologies nor on improving them.
+
+Some of these clarifications have been included in \textbf{Section ``1.
+Introduction" (lines 112-129 - ''The main purpose ...
+tools``)}, and others have been added to \textbf{''Abstract`` (lines 10-19)} and \textbf{Section ``5.
+Conclusions`` (lines 759-766 - ''The aim ...
+regression``, lines 781-782 - ''Given that ...
+carried out``)}, so that the reader can better understand the scope of our work.
+\textbf{Comments 4 and 9} suggested by Reviewer \#2 are also related to these clarifications and have been included in \textbf{''Abstract``} and \textbf{Section ''5.
+Conclusions''}.
+
+At the end of \textbf{Section ``1.
+Introduction'' (lines 133-168 - ``The generation ...
+system'')}, we have tried to summarize in a general way the contributions of SPAMDA, which are specified and detailed throughout the paper.
+Below, we comment on some contributions that have also been taken into account in this revised version of the manuscript (\textbf{highlighted in blue color}):
 \begin{itemize}
-	\item Bla, bla, bla,
-	\item Bla, bla, bla,
-	\item Bla, bla, bla,
+
+	\item SPAMDA allows preliminary studies of missing values (dates or measurements not recorded) in buoys managed by NDBC, so that researchers have an idea of the quality of data observed by buoys and whether these would be suitable for the work to be carried out.
+Sometimes, buoys do not collect observations temporarily due to mechanical problems or weather conditions, so, depending on the problem to be solved, such raw data could not be appropriate for accurate modelling.
+Even so, SPAMDA allows data integration taking into account such missing values when needed by researchers.
+More details about how this task is performed can be found in \textbf{Section ``2.
+Meteorological data sources''  (lines 208-215 - ``Note ...Appendix A.''), Section ``3.2.
+Datasets'' (lines 283-286 - ``When an ...
+dates.''), Section ``3.3.
+Pre-process'' (lines 315-328 - ``Recover ...values mean.''), Section ``3.4.
+Matching configuration'' (lines 398-403  - ``Include missing ...as \guillemotleft\textit{?}\guillemotright''), and Section ``Appendix A.
+Managing the casuistry of incomplete data''}.
+
+	\item Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoys measurements, such as normalisation and missing data recovery (see \textbf{Section ``3.3.
+Pre-process'' (lines 323-328 - ``Replace missing...values mean'')} for more details).
+
+	\item Building as many different datasets from the same meteorological data as needed, through a customisable matching process with respect to the input variables from NNRP or the output variable from NDBC, and ready to be used in regression and classification prediction tasks (more details in \textbf{Section ``3.4.
+Matching configuration'' and Section ''3.5.
+Final datasets``}).
+
+	\item The possibility of using several reanalysis nodes near to the NDBC buoy under study, which could provide a better description of the problem to achieve more accurate models (more details can be found in \textbf{Section ''3.4.
+Matching configuration`` (lines 404-417 - ''Nearest reanalysis ...respectively.``)}).
+
+	\item Estimation of the amount of energy flux that can be produced at different prediction horizons, for example, 6h, 12h, 18h, 24h, ...
+in advance (more details in \textbf{Section ''3.5.
+Final datasets`` (lines 463-502 - ''Prediction horizon ...
+(${t+\Delta t}$ in Eq.
+5)``)}.
+Although this work does not focus on model performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.
+
+	\item This software would be provided to the scientific community as a free and open source tool, so that any researcher could access its source code and extend its functionality (the github repository is shown in \textbf{Section "Additional material"}).
 \end{itemize}
 
-Bla, bla, bla, \textbf{This clarification has been rewritten in page 2, lines 53-61.}
+From an applicability point of view, SPAMDA could be used in coastal and ocean engineering applications and marine energy prediction, such as:
+\begin{itemize}
+    \item Estimation of the energy supply sources in hybrid energy systems [3], based on the amount of energy that can be obtained by a marine energy system in the next 6h or 12h, for example.
+
+    \item Regulation of WECs to avoid malfunction or breakage [27,28], depending on the significant wave height and/or energy flux expected, as well as the possibility of reconfiguring them in order to maximise the wave energy extraction.
+
+    \item Decision-making in exploitation and environmental protection for the construction of marine energy storage plants, future strategies on renewable energy and coastal planning [52].
+
+    \item Support tool for the prediction of the energy that could be obtained from a certain maritime location, to know whether or not it is convenient to install WECs as power supply in marine structures, such as offshore oil and gas platforms or seawater desalination plants [24, 25].
+
+    \item Design of offshore structures and ports [53].
+
+    \item Decision-making and risk assessment about operational works in the sea [54], security systems for structures or naval security [55].
+\end{itemize}
+
+Some of these clarifications about the applicability of SPAMDA have been included in \textbf{Section ''1. Introduction`` (lines 169-182 - ''Therefore ... security``).} 
 }
@@
 {
-Following the instructions of the reviewer, bla, bla, bla.
-\textbf{These clarifications have been included in the paper on page 11, lines 646-660, in the new Figure 6, and in the ``Conclusions'' Section, lines 672-682.}
+Thank you very much for questioning this fact and allowing us to clarify it.
+
+Although we have cited several software tools related to energy prediction and environmental modelling, as well as works that use ML and SC methodologies in marine energy studies (now much more complete with the references provided by Reviewer \#2 in \textbf{Comment 7}), we have not made a critical evaluation of the state-of-the-art because, to our knowledge, there are no open source software tools available to the scientific community, that perform the data integration provided by SPAMDA with the functionalities described in this work.
+
+The reading order of the contents shown in \textbf{Section "1.
+Introduction"} has been carried out with the intention that is commented below, and the background has been reinforced with the interesting references suggested by Reviewer \#2 \textbf{(lines 37-39, 66, 74, 81, 86-94)}:
+\begin{enumerate}
+    \item Firstly, readers are introduced to the importance of having quality and well-formed meteorological data and observations to perform studies of environmental comprehension, conservation, modelling and prediction, providing examples of this.
+It is also indicated that ML and SC methodologies, despite managing uncertainty, need adequate data to obtain accurate models \textbf{(lines 23-43 - ``A better ...
+quality models.'')}.
+
+    \item Secondly, it is explained that researchers need special purpose software to be able to perform data adaptation tasks and to carry out studies related to energy and environmental modelling  \textbf{(lines 44-59 - ``Continuing with ...
+solar tracker'')}.
+
+    \item Thirdly, marine energy is introduced as one of the most important renewable and sustainable energy sources today, where meteorological data is used for prediction purposes.
+WECs are also introduced as the main mechanical devices for capturing and converting energy from waves.
+It is concluded that for both the correct design and use of WECs and for other applications related to marine energy, it is very useful to have accurate prediction models, for which it is necessary to have both data and techniques that use them \textbf{(lines 60-77 - ``Marine ...\textit{wave energy period} ($T_e$).'')}.
+
+    \item Fourthly, it is explained that ML and SC methodologies are supporting traditional statistical modelling procedures and that there are tools, such as Weka, that can be used even by researchers who are not specialized in the field of artificial intelligence or in the use of programming languages.
+Some studies, related to marine energy, using ML algorithms for wave height and energy flux prediction, are also mentioned.
+In this part, it is emphasized that datasets are  essential both for these tools that offer an user-friendly interface for researchers not specialized in ML and for more specific works, and that, usually, they are not publicly available in an appropriate format.
+This causes them to present inconsistencies that prevent their treatment, and a great deal of effort and time is spent on solving them, which the software tool developed avoids  \textbf{(lines 78-111 - ``Currently ...
+tedious task.'')}.
+
+    \item Next, it is indicated that the main purpose of this work is to solve in an automatic, versatile and customisable way, the integration of observations from NDBC and reanalysis data from NNRP to build datasets (address the gap), without focusing on the ML and SC methodologies that use them.
+To do that, a software tool has been developed that manages the casuistry involved in such data integration for marine energy prediction tasks \textbf{(lines 112-129 - ``The main ...
+tools'')}.
+
+    \item Finally, the functionalities and the general contributions of the developed tool are described, as well as some examples of applicability \textbf{(lines 130-168 - ``In order to ...
+system'', 169-182  - ``Therefore ...
+security [55].'')}.
+The remaining sections of the manuscript explain in detail the contributions of SPAMDA and how they are carried out, ending with a case study in the Gulf of Alaska tackling significant wave height and energy flux prediction.
+\end{enumerate}
 }
@@
 {
-A complete revision of the English has been carried out in this review step.
+Thanks for your suggestion, a revision of the English language has been carried out in this version of the manuscript.
 }
 %%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
 
@@
 {
-Bla, bla, bla,
-
-Ok, good point. There are two main reasons for choosing
+Following the instructions of the reviewer, we have modified \textbf{Section ``1.
+Introduction''} and   \textbf{Section ``5.
+Conclusions''}, so that readers will better understand the contribution of the developed tool.
 }
@@
 {
-Bla, bla, bla. The reviewer is right about this comment on input variables.....
+We would like to thank the reviewer for her/his suggestion.
+As this work is not a study that presents a prediction methodology related to marine energy, and whose results must be compared with other methodologies in the state-of-the-art, there is no a Section ``Results''.
+
+Nevertheless, a case study is carried out to predict significant wave height and energy flux in the Gulf of Alaska through data integration from NDBC and NNRP using SPAMDA.
+Besides, \textbf{we have extended the case study} of the initial version of this paper with experimentation about energy flux prediction that reinforces the usefulness of this software, without comparing the results achieved by the different algorithms applied, because this is not the purpose of this work.
+Only typical performance metrics as CCR (Accuracy), Kappa, RMSE (Root Mean Squared Error) or Correlation coefficient are shown.
+
+Such experimental extension has been included in \textbf{Section ``4.
+A case study applied to Gulf of Alaska''}.
+Besides, \textbf{Sections ``3.4.
+Matching configuration''} and \textbf{``3.5.
+Final datasets''} \textbf{have been rewritten} for this purpose trying to add as few figures as possible and unifying existing ones, as the Reviewer \#2 also suggests in \textbf{Comment 6}.
 }
+
 %%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {
-Bl, bla, bla. Following this suggestion by the reviewer, we have changed...
+Many thanks for the suggestion.
+We have clarified the explanation of \textbf{Figure 21, now Figure 20  (lines 642-657 - ``After configuring ...
+ARFF'')}.
+By extending the case study the number of figures has increased, but \textbf{we have combined Figures 8 and 9}.
 }
@@
 {
-Bla, bla, bla,
+We have taken into account and included all of them in \textbf{Section ``1.
+Introduction'' (lines 37-39, 66, 74, 81, 86-94)} of this new revised version of the manuscript.
+The following one was already in the original version:
+
+[40] Kumar, N.K.; Savitha, R.; Al Mamun, A.
+''Regional ocean wave height prediction using sequential learning neural networks``.
+Ocean Engineering, vol.
+129, pp.
+605–612, 2017.
+doi:10.1016/j.oceaneng.2016.10.033.
+
 }
@@
 
+\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {
-Bla, bla, bla,
+Many thanks for the appreciation.
+To our knowledge, we can not make a comparison with other applications as there are no other software tools with the functionality and characteristics presented in this work.
+There may be console programs or scripts that use NDBC or NNRP data to perform some type of application on data, but we have not found in the literature any open source software available to the scientific community that performs the data integration process described in this paper.
+Moreover, we have not found any software that exhibits the functionalities and configurations available in SPAMDA, not only for the data integration but also to manage buoys and reanalysis data in a simple and versatile way, creating datasets in  a proper format to be easily used with ML and SC algorithms.
 }
@@
 {
-Bla, bla, bla. The reviewer is completely right. We have rewritten the abstract to better describe the application tackled.
+Thanks for this comment, we have rewritten the \textbf{''Abstract`` (lines 10-19)}, trying not to exceed  the allowed size.
+We hope that, with this version of the abstract, readers can better understand what this work offers.
 }
@@
 There are some errors in your reference list. Please check and fix the errors.
-
 }
@@
 {
-Bla, bla, bla. Thanks for this comment. The distribution....
+The reviewer is completely right, we have revised the bibliography and corrected the errors detected.
+Please, excuse us if there is an omission in any type of character or compound name that could be due to the latex bibliographic format established by the journal.
+
+These are the references that have been modified:
+
+[3] Shivam, K.; Tzou, J.C.; Wu, S.C.
+Multi-Objective Sizing Optimization of a Grid-Connected Solar–Wind Hybrid System Using Climate Classification: A Case Study of Four Locations in Southern Taiwan.
+Energies 2020, 13, 2505.
+doi:10.3390/en13102505
+
+[4] Dorado-Moreno, M.; Cornejo-Bueno, L.; Gutiérrez, P.A.; Prieto, L.; Hervás-Martínez, C.; Salcedo-Sanz, S.
+Robust estimation of wind power ramp events with reservoir computing.
+Renewable Energy 2017, 111, 428-437.
+doi:10.1016/j.renene.2017.04.016
+
+[5] He, Q.; Zha, C.; Song, W.; Hao, Z.; Du, Y.; Liotta, A.; Perra, C.
+Improved Particle Swarm Optimization for Sea Surface Temperature Prediction.
+Energies 2020, 13, 1369.
+\newline doi:10.3390/en13061369
+
+[7] da Silva, V.d.P.R.; Araújo e Silva, R.; Cavalcanti, E.P.; Braga Campos, C.; Vieira de Azevedo, P.; Singh, V.P.; Rodrigues Pereira, E.R.
+Trends in solar radiation in NCEP/NCAR database and measurements in northeastern Brazil.
+Solar Energy 2010, 84, 1852.1862.
+\newline doi:10.1016/j.solener.2010.07.011
+
+[11] Dhanraj Bokde, N.; Mundher Yaseen, Z.; Bruun Andersen, G.
+ForecastTB-An R Package as a Test-Bench for Time Series Forecasting-Application of Wind Speed and Solar Radiation Modeling.
+Energies 2020, 13, 2578.
+doi:10.3390/en13102578
+
+[14] Di Bari, R.; Horn, R.; Nienborg, B.; Klinker, F.; Kieseritzky, E.; Pawelz, F.
+The Environmental Potential of Phase Change Materials in Building Applications.
+A Multiple Case Investigation Based on Life Cycle Assessment and Building Simulation.
+Energies 2020, 13, 3045.
+doi:10.3390/en13123045
+
+[15] Astiaso García, D.; Bruschi, D.
+A risk assessment tool for improving safety standards and emergency management in Italian onshore wind farms.
+Sustainable Energy Technologies and Assessments 2016, 18, 48.58.
+doi:10.1016/j.seta.2016.09.009
+
+[16] Raabe, A.L.A.; Klein, A.H.d.F.; González, M.; Medina, R.
+MEPBAY and SMC: Software tools to support different operational levels of headland-bay beach in coastal engineering projects.
+Coastal Engineering 2010, 57, 213.226.
+doi:10.1016/j.coastaleng.2009.10.008
+
+[17] Motahhir, S.; EL Hammoumi, A.; EL Ghzizal, A.; Derouich, A.
+Open hardware/software test bench for solar tracker with virtual instrumentation.
+Sustainable Energy Technologies and Assessments 2019, 31, 9-16.
+doi:10.1016/j.seta.2018.11.003
+
+[18] Cascajo, R.; García, E.; Quiles, E.; Correcher, A.; Morant, F.
+Integration of Marine Wave Energy Converters into Seaports: A Case Study in the Port of Valencia.
+Energies 2019, 12, 787.
+doi:10.3390/en12050787
+
+[20] De Jong, M.; Hoppe, T.; Noori, N.
+City Branding, Sustainable Urban Development and the Rentier State.
+How do Qatar, Abu Dhabi and Dubai present Themselves in the Age of Post Oil and Global Warming?
+Energies 2019, 12, 1657.
+doi:10.3390/en12091657
+
+[21] Brede, M.; de Vries, B.J.M.
+The energy transition in a climate-constrained world: Regional vs.
+global optimization.
+Environmental Modelling \& Software 2013, 44, 44-61.
+\newline doi:10.1016/j.envsoft.2012.07.011
+
+[25] Fernández Prieto, L.; Rodríguez Rodríguez, G.; Schallenberg Rodríguez, J.
+Wave energy to power a desalination plant in the north of Gran Canaria Island: Wave resource, socioeconomic and environmental assessment.
+Journal of Environmental Management 2019, 231, 546-551.
+doi:10.1016/j.jenvman.2018.10.071
+
+[38] Frank, E.; Hall, M.A.; Witten, I.H. The WEKA Workbench. Online Appendix for Data Mining: Practical Machine Learning Tools and Techniques, 2016
+
+[49] Kalnay, E.; Kanamitsu, M.; Kistler, R.; Collins, W.; Deaven, D.; Gandin, L.; Iredell, M.; Saha, S.; White, G.; Woollen, J.; Zhu, Y.; Leetmaa, A.; Reynolds, R.; Chelliah, M.; Ebisuzaki, W.; Higgins, W.; Janowiak, J.; Mo, K.C.; Ropelewski, C.; Wang, J.; Jenne, R.; Joseph, D.
+The NCEP/NCAR 40-Year Reanalysis Project.
+Bulletin of the American Meteorological Society 1996, 77, 437-471.
+doi:10.1175/1520-0477(1996)077<0437:TNYRP>2.0.CO;2
+
+[62] Quinlan, J.R. C4. 5: Programs for machine learning; Morgan Kaufmann, 1992.
+
+[63] Breiman, L. Random forests. Machine learning 2001, 45, 5-32. doi:10.1023/A:1010933404324
+
+[64] Cortes, C.; Vapnik, V. Support-vector networks. Machine learning 1995, 20, 273-297. doi:10.1007/BF00994018
+
+[68] Alcalá-Fdez, J.; Sánchez, L.; García, S.; del Jesús, M.J.; Ventura, S.; Garrell, J.M.; Otero, J.; Romero, C.; Bacardit, J.; Rivas, V.M.; Fernández, J.C.; Herrera, F.
+KEEL: a software tool to assess evolutionary algorithms for data mining problems.
+Soft Comput.
+2009, 13, 307-318.
+doi:10.1007/s00500-008-0323-y
 
@@
 \vspace{0.5cm}
-Thus, I recommend to submit the menuscript to a more specialized journal.
+Thus, I recommend to submit the manuscript to a more specialized journal.
 }
@@
 {
-First of all, we would like to thank Reviewer \#3 for his/her suggestions to improve the submitted manuscript.
+First of all, we would like to thank Reviewer \#3 for his/her suggestions.
 
-\vspace{0.5cm}
-VAMOS A VER PEDAZO DE HIJO DE PUTA, TE HAS QUERIDO QUITAR EL LEER LA REVISION DE UN PLUMAZO
-}
+In the initial version we did not adequately explain what the purpose of this work is and its relation to energy, we hope that now it is better clarified and readers can understand the scope of our work.
+In this revised version of the manuscript we have tried to clarify the contributions of this work, emphasizing them mainly in \textbf{''Abstract`` (lines 10-19)} and \textbf{Sections ''1.
+Introduction`` (lines 130-168 - ''In order to ...
+system``, 169-182 - ''Therefore ...
+security [55])} and \textbf{``5.
+Conclusions`` (lines 759-766 - ''The aim ...
+regression``, lines 779-782 - ''The case ...
+carried out``)}.
+Besides, in \textbf{Section ''1.
+Introduction`` (lines 60-77)} the relation of using this tool (to create datasets to model waves or flux of energy) with WECs, is discussed.
+
+Prediction studies using ML and SC algorithms do not provide tools to incorporate and integrate, considering the casuistry that it entails, the two data sources used by SPAMDA.
+Such studies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, artificial neural networks, etc) on data, using custom-made implementations or scripts developed in some programming language, but they do not allow to build datasets, ready to be used with any of those methodologies or algorithms, through an automated and versatile process.
+
+SPAMDA allows the creation of datasets to estimate the amount of energy flux that can be generated by waves at different prediction horizons.
+These datasets are well-formed and ready to be used with ML and SC prediction techniques.
+Therefore, we think that this work can be useful as a support for applications related to marine energy saving, production and conversion processes.
+
+In that sense, and to better clarify the applicability and the relation to energy of this work, \textbf{we have extended Section ''4.
+A case study applied to Gulf of Alaska`` with a short-term forecasting problem of the energy flux generated by the waves in a specific location of that zone}.
+Besides, \textbf{Sections ''3.4.
+Matching configuration``} and \textbf{''3.5.
+Final datasets``} have been also rewritten for the purpose of flux of energy.
+In this way, it would be possible to perform prediction studies 6h in advance (the prediction horizon is customisable) of the amount of marine energy that could be extracted by WECs.
+Given that this work does not focus on models performance, a validation or comparison study of the results obtained in such example has not been carried out.
+
+Having said that, we think that this software tool could be used in coastal and ocean engineering applications and marine energy prediction, such as (included in \textbf{Section ''1.
+Introduction``, lines 169-182 - ''Therefore ...
+security [55]''}):
+\begin{itemize}
+    \item EStimation of the energy supply sources in hybrid energy systems [3], based on the amount of energy that can be obtained by a marine energy system in the next 6h or 12h, for example.
+
+    \item Regulation of WECs to avoid malfunction or breakage [27,28], depending on the significant wave height and/or energy flux expected, as well as the possibility of reconfiguring them in order to maximise the wave energy extraction.
+
+    \item Decision-making in exploitation and environmental protection for the construction of marine energy storage plants, future strategies on renewable energy and coastal planning [52].
+
+    \item Support tool for the prediction of the energy that could be obtained from a certain maritime location, to know whether or not it is convenient to install WECs as power supply in marine structures, such as offshore oil and gas platforms or seawater desalination plants [24, 25].
+
+    \item Design of offshore structures and ports [53].
+
+    \item Decision-making and risk assessment about operational works in the sea [54], security systems for structures or naval security [55].
+\end{itemize}
+
+We would like to thank you again for the suggestions and review carried out to our work.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-01-07 latex_manuscript/MDPI/2_energies_special_issue/envio-3/latex/energies-1069865.tex
Minor Changes Energies Special Issue

--- 2020-11-17 latex_manuscript/MDPI/2_energies_special_issue/template.tex
@@
+% !TeX spellcheck = en_GB
 %  LaTeX support: latex@mdpi.com 
@@
 % Full title of the paper (Capitalized)
-\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A case  study for predicting significant wave height}
+\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height \textcolor{blue}{and Energy Flux}}
 
@@
 \abstract{Meteorological data are extensively used to perform environmental learning.
-Soft Computing (SC) and Machine Learning (ML) techniques, a valuable support in many research areas, require datasets containing information related to the topic under study, which are not always available in an appropriate format, and its preparation and pre-processing implies a lot of time and effort by the researchers.
+Soft Computing (SC) and Machine Learning (ML) techniques represent a valuable support in many research areas, but require datasets containing information related to the topic under study.
+Such datasets are not always available in an appropriate format and its preparation and pre-processing implies a lot of time and effort by researchers.
 This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-Such datasets can be created using buoys and reanalysis data by customisable procedures, in terms of predictive and objective variables and temporal resolution.
-These datasets can be used by SC and ML methodologies for prediction tasks (classification or regression) that support improvement of sustainability energy production, design of production systems as WECs or environmental modelling, among others.}
+Such datasets can be created using buoys and reanalysis data through customisable procedures, in terms of temporal resolution, predictive and objective variables, \textcolor{blue}{ and can be used by SC and ML methodologies for prediction tasks (classification or regression).
+The objective is providing the research community with an automated and versatile system for the casuistry that entails well-formed and quality data integration, potentially leading to better prediction models.
+The software tool can be used as a supporting tool for coastal and ocean engineering applications, sustainable energy production or environmental modelling; as well as for decision making in the design and construction of coastal protection structures, marine transportation, offshore industry, ocean energy converters and efficient operation of offshore and coastal engineering activities.
+Finally, to illustrate the applicability of the proposed tool, a case study to classify waves depending on their significant height and to predict energy flux in the Gulf of Alaska is presented.
+}
+%Therefore, it does not make a validation or comparison study of the performance of state-of-the-art prediction algorithms.
+}
 
@@
 		Once quality and well-formed data are obtained, these can be used to extract information and build prediction models that explain the behavior of a certain problem.
+\textcolor{blue}{The choice of the appropriate model, either in engineering problems or in any other problem, is also an important factor in addition to data \cite{Alizadeh2019, Alizadeh2020}}.
 Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modelling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.
@@
 		Continuing with this line, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
-In \cite{Neeraj2020}, a software package in R called ``ForecastTB'' is developed to compare the accuracy of different forecasting methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modelling.
-		%In \cite{CAMARGONOGUEIRA2016361}, a software tool for designing solar water heating systems is developed, which simulates different situations and finds the best technical and economical solution. 
+In \cite{Neeraj2020}, a software package in R called ``ForecastTB'' is developed to compare the accuracy of different prediction methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modelling.
 		In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented.
@@
 		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool`` is presented for comparing PCM - phase change materials- storage systems  with traditional systems that do not entail energy storage, being beneficial in order to support decision-making on energy concepts for buildings.
-		%In \cite{SHEHADEH2020100644}, a load management application to maximize the used energy of thermal systems and reduce the costs and emissions is developed. 
 		A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}.
@@
 Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{en12091657, BREDE201344}.
+Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \textcolor{blue}{\cite{en12091657, BREDE201344, Alizadeh2020a}}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
@@
 Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves, in order to maximise the wave energy extraction \cite{en80910370}.
+As a consequence of this complexity, many aspects of WEC design, deployment and operation \textcolor{blue}{\cite{CROWLEY2018159, Abdelkhalik2016, 6898109}} need a proper prediction of waves \textcolor{blue}{\cite{en11010011, Kaloop2020}}, in order to maximise the wave energy extraction \cite{en80910370}.
 For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-Additionally, wave predictions \cite{en11010011} are also helpful for designing offshore structures \cite{CHATZIIOANNOU2017126}, operational works in the sea \cite{DALGIC2015211}, providing technical information to the coastal and coral reef planners in developing countries \cite{CALLAGHAN2018123}, etc.
-
-		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results \cite{Fi-John2019,Mosavi2019,GUO201816}.
+
+		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results, \textcolor{blue}{ either in engineering \cite{Alizadeh2020}, energy or environmental problems} \cite{Fi-John2019,Mosavi2019,GUO201816}.
 SC and ML methodologies can be used not only by experienced computer scientists but also by other researchers.
 For example, the well-known \textit{Waikato Environment for Knowledge Analysis} (WEKA) \cite{WEKA} software tool provides researchers with a wide collection of ML  algorithms.
-In this way, ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction.
-The problem is that, in order to apply these methods, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge.
+ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction.
+\textcolor{blue}{In \cite{Ali2020}, a reliable ML model based on multiple linear regression and covariant-weighted least square estimation for $H_s$ modelling is presented in order to forecast half-hourly nearly real-time significant wave height values.
+In \cite{Cornejo-Bueno2016}, an  approach for feature selection problems is developed and applied to a specific problem of $H_s$ and $F_e$ prediction in oceanic buoys,  obtaining excellent results in terms of prediction quality.
+In \cite{Emmanouil2020}, a Bayesian Network system provides an useful tool for the decision making process of installation and maintenance operations in offshore wind farms using predictions of $H_s$, among others.
+In \cite{Shamshirband2020}, several ML methods are implemented and compared for the prediction of $H_s$ in the Persian Gulf, the extreme learning machine (ELM) providing the best results.} The problem is that, in order to apply \textcolor{blue}{ML and SC techniques}, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge.
 Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.
@@
 Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
-Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by the researchers to manually create the datasets with the needed information.
+Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by researchers to manually create the datasets with the needed information.
 Given that such process is of great relevance and has an extensive casuistry, the present work has been specially focused on it.
@@
 The open source software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
-As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows the researchers focus on the study of the meteorological aspects of the observations.
-The datasets obtained are ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), although the researchers can use them for other purposes.
+As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows researchers focus on the study of the meteorological aspects of the observations.
+The datasets obtained are ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), although researchers can use them for other purposes.
 These datasets contain one or more meteorological variables as inputs and one variable as target (variable to be predicted).
 The format of the generated datasets will be \textit{Attribute-Relation File Format} (ARFF) \cite{WEKA_ARFF}, which is the one used by WEKA.
-Besides, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling the researchers to use others tools.
-
-		In order to address the problem previously discussed, meteorological data integration from NDBC and NNRP and the casuistry that it entails, SPAMDA offers to the researchers novelties and functionalities that will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
-
-		%Up to now, and to the best of our knowledge, this is the first software tool \textcolor{red}{with GUI (Graphical User Interface) }addressing the problem previously discussed, integrating meteorological data from NDBC and NNRP \textcolor{red}{and the casuistry that it entails}.
-\textcolor{blue}{(desde ''Up to now...
-hasta ...entails lo quitaría)} The novelties and functionalities that SPAMDA offers to the researchers will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below \textcolor{red}{Esto hay que revisarlo}:
-		\begin{itemize}
-			%\item The generation of datasets becomes a very easy and customisable task, by means of the selection of different input parameters, such as predictive and objective variables, classification and regression ML techniques, output discretisation or prediction horizon, among others.
+Besides, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling researchers to use others tools.
+
+		In order to address the problem previously discussed, meteorological data integration from NDBC and NNRP and the casuistry that it entails, SPAMDA offers to researchers novelties and functionalities that will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
+
+		\begin{itemize}
 			\item The generation of datasets becomes a very easy and customisable task by means of the selection of different input parameters, such as predictive and objective variables, classification and regression, output discretisation (useful for ordinal regression) or prediction horizon, among others.
+
 			\item The created datasets can be easily used by SC and ML tools.
+
+			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a lot of time and effort in early stages of the research.
+
+			\item It avoids possible researcher errors in the intermediate steps of the process, such as geographical coordinates conversion, missing values handling (dates or measurements not recorded) or different temporal resolution of the data collected, among others.
+
 			\item It provides information about the quality and quantity of the data.
-			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a lot of time and effort in early stages of the research.
-			\item It avoids possible researcher errors in the intermediate steps of the process, such as geographical coordinates conversion, missing values handling (dates or measurements not recorded) or different temporal resolution of the data collected, among others.
+\textcolor{blue}{SPAMDA allows preliminary studies of missing values (dates or measurements not recorded) in buoys managed by NDBC, so that the researcher can have an idea of the quality of the data recorded by the buoys and about their suitability for the intended purpose.
+In any case, SPAMDA allows data integration taking into account such missing values when needed by the user.}
+
+			\item \textcolor{blue}{Estimation of the amount of energy flux that can be produced at different prediction horizons: short-term, mid-term or long-term.
+Although this work does not focus on models performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.}
+
 			\item It manages the extensive casuistry of data integration which can lead to incomplete datasets, described in Appendix \ref{app:AppendixA}.
-			\item Possibility of selecting one or more reanalysis nodes near the localisation under study.
-			\item It also includes some basic pre-processing tasks, such as normalisation and missing data recovery.
+
+			\item Possibility of selecting one or more reanalysis nodes near the localisation under study, \textcolor{blue}{which could provide a better description of the problem to achieve more accurate models.}
+
+			\item \textcolor{blue}{Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoys measurements, such as normalisation and missing data recovery.}
+
 			\item It facilitates data management and well-organised storage of the datasets.
-			\item Its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
+\textcolor{blue}{Environmental studies in different geographical locations can be carried out by merely introducing and using other collected data.}
+
+			\item SPAMDA is distributed as open source tool, its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
+
 			\item It includes a user-friendly GUI, facilitating and greatly simplifying data management, and it is integrated with the Explorer environment of WEKA.
+
 			\item It is multi-platform, and it can be used on any computer with Java regardless of the operating system.
@@
 
-		Therefore, the functionalities and characteristics that SPAMDA offers make it a novel support tool to researchers when carrying out environmental modelling related to energy, design of production systems as WECs (that need a proper prediction of waves in order to maximise the wave energy extraction) and atmospheric or oceanic studies, among others.
+		\textcolor{blue}{Therefore, the functionalities and characteristics that SPAMDA offers make it a supporting tool for researchers, which could be used in applications related to coastal and ocean engineering, and also in marine energy prediction.
+In \cite{Kumar2020}, the estimation of energy supply sources in hybrid energy systems is based on the amount of energy that can be obtained by a marine energy system within a prediction horizon.
+Regulation of WECs to avoid malfunction or breakage, depending on the significant wave height and/or energy flux expected, as well as the possibility of reconfiguring them in order to maximise the wave energy extraction, is studied in \cite{CROWLEY2018159, Abdelkhalik2016}.
+The prediction of the energy that could be obtained from a certain maritime location is considered in \cite{OLIVEIRAPINTO2019556, FERNANDEZPRIETO2019546} in order to know whether it is convenient to install WECs as power supply in marine structures, such as offshore oil and gas platforms or seawater desalination plants.
+In \cite{Ali2019}, significant wave height forecasting is applied for decision-making in exploitation and environmental protection for the construction of marine energy storage plants, future strategies on renewable energy and coastal planning.
+Other examples of application are: design of offshore structures and ports \cite{CHATZIIOANNOU2017126}, decision-making and risk assessment about operational works in the sea \cite{DALGIC2015211}, security systems for structures or naval security \cite{Spaulding2020}.}
 
@@
 				\centering
-				\includegraphics[scale=0.50]{figures/FigureFragmentAnnualTextFile.png}
+				\includegraphics[scale=0.56]{figures/FigureFragmentAnnualTextFile.png}
 				\caption{A fragment of an annual text file of the \textit{Station 46001}.}
@@
 				\centering
-				\includegraphics[scale=0.52]{figures/FigureSubGrid.jpg}
+				\includegraphics[scale=0.55]{figures/FigureSubGrid.jpg}
 				\caption{Example of a six sub-grid reanalysis nodes around the \textit{Station 46001}.}
@@
 			\centering
-			\includegraphics[scale=0.32]{figures/FigureSPAMDA.eps}
+			\includegraphics[scale=0.33]{figures/FigureSPAMDA.eps}
 			\caption{Brief outline of the functionality provided by SPAMDA.}
@@
 				\item Pre-processing tasks for obtaining the pre-processed datasets.
-				\item Matching process to merge the information from NDBC and NNRP
+				\item Matching process to merge the information from NDBC and NNRP.
 				\item Creation of the final datasets accordingly to the ML technique to use (classification or regression).
@@
 			\item \textit{Manage reanalysis data}: This module is used for the management of the reanalysis data provided by the NNRP.
-In this way, the researchers can keep the reanalysis data files updated for their studies.
-Such files will be used, depending on the researchers needs, in the matching process when obtaining the final datasets.
+In this way, researchers can keep the reanalysis data files updated for their studies.
+Such files will be used, depending on researchers needs, in the matching process when obtaining the final datasets.
 
@@
 					\centering
-					\includegraphics[scale=0.42]{figures/FigureBuoys.eps}
+					\includegraphics[scale=0.4]{figures/FigureBuoys.eps}
 					\caption{Example of entering two buoys with its annual text files.}
@@
 				Once a buoy has been included as described in Section \ref{sec:Buoys}, it is possible to create datasets with one or more annual text files, which are referenced in SPAMDA as intermediate datasets.
-In this module, the researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary ones.
+In this module, researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary ones.
 
@@
 
-				SPAMDA allows the researchers to undo the last filter applied or to restore the initial content of the intermediate dataset.
+				SPAMDA allows researchers to undo the last filter applied or to restore the initial content of the intermediate dataset.
 Besides, the content and relevant statistical information (number of instances with missing values, minimum and maximum values, mean and standard deviation) of the intermediate and the pre-processed datasets can be visualised in this module.
@@
 The intermediate dataset 1 of the buoy ID2 has been also pre-processed.
-\textit{Pre-processed dataset n} represents that the researchers can create as many pre-processed datasets as they consider opportune.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.36]{figures/FigurePreprocess.eps}
+\textit{Pre-processed dataset n} represents that researchers can create as many pre-processed datasets as they consider opportune.
+
+				\begin{figure}[ht!]
+					\centering
+					\includegraphics[scale=0.40]{figures/FigurePreprocess.eps}
 					\caption{Example of the creation of pre-processed datasets.}
@@
 Such process is one of the most powerful and remarkable features of this software tool due to its great relevance and extensive casuistry.
-In this sense, SPAMDA has been developed to provide great flexibility to the researchers.
+In this sense, SPAMDA has been developed to provide great flexibility to researchers.
 
@@
 					\centering
-					\includegraphics[scale=0.38]{figures/FigureMatchingProcess.png}
+					\includegraphics[scale=0.43]{figures/FigureMatchingProcess.png}
 					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
@@
 
-				SPAMDA allows the researchers to perform a customisable matching process, for obtaining as many different versions of the same meteorological data as needed.
+				SPAMDA allows researchers to perform a customisable matching process, for obtaining as many different versions of the same meteorological data as needed.
 				Prediction tasks are based on the estimation of the output attribute using the information provided by the input attributes.
@@
 As a consequence, the matching of instances between both sources of information may not be possible (missing dates).
-In that situation, the researchers can consider two options: 1) discard the instances affected or 2) include them.
+In that situation, researchers can consider two options: 1) discard the instances affected or 2) include them.
 In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft\textit{?}\guillemotright.
@@
 \ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
-In this way, the researcher can:
+In this way, researchers can:
 
@@
 
-							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by the researcher will be created.
+							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by researcher will be created.
 Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.
@@
 
-				Once the matching parameters have been described, for a better understanding of them,  Fig.
-\ref{fig:directMatching} presents an example of the matched information considering the data shown in Fig.
-\ref{fig:matchingProcess} and using the following matching configuration\footnote{Note that the date is shown just for better understanding, but it will not be included in the final dataset.}:
-					\begin{itemize}
-						\item Variable WVHT as attribute to predict.
+				\textcolor{blue}{Once the matching parameters have been described, for a better understanding of them,  Fig.
+\ref{fig:directMatching} presents an example of the data integration considering the data shown in Fig.
+\ref{fig:matchingProcess} and using the following configuration\footnote{Note that the date is shown just for a better understanding, but it will not be included in the final dataset.}:}
+					\begin{itemize}
+						\item \textcolor{blue}{Attribute to predict: variable WVHT (Fig. \ref{fig:directMatchingWave}) / flux of energy (Fig. \ref{fig:directMatchingFlux})}.
 						\item Variable Pres as reanalysis input attribute.
@@
 
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.49]{figures/FigureDirectMatching.png}
-					\caption{An example of the matched information for \textit{Direct matching}.}
-					\label{fig:directMatching}
+% 				\begin{figure}[ht!]
+% 					\centering
+% 					\includegraphics[scale=0.49]{figures/FigureDirectMatching.png}
+% 					\caption{An example of the matched information for \textit{Direct matching}.}
+% 					\label{fig:directMatching}
+% 				\end{figure}
+
+				\begin{figure}[ht!]
+					\centering
+					\subfloat[attribute to predict: WVHT\label{fig:directMatchingWave}]{\includegraphics[width=0.4\textwidth]{figures/FigureDirectMatching.png}}\hfill
+					\subfloat[attribute to predict: flux of energy\label{fig:directMatchingFlux}]{\includegraphics[width=0.47\textwidth]{figures/FigureDirectMatching_EF.png}}\hfill
+					\caption{\textcolor{blue}{Example of data integration for \textit{Direct matching}.}} \label{fig:directMatching}
 				\end{figure}
@@
 
-				SPAMDA allows the researchers to make such preparation by means of the following options:
+				SPAMDA allows researchers to make such preparation by means of the following options:
 
@@
 						The minimum interval for increasing and decreasing the prediction horizon is $6$h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out.
-Therefore, for each increment of the prediction horizon, an instance is lost of the dataset is lost (as this future information is not available).
+Therefore, for each increment of the prediction horizon, an instance of the dataset is lost (as this future information is not available).
 As the minimum prediction horizon is $6$h, at least one instance will be lost.
@@
 						where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to be predicted, $\mathbf{b}_t$ is the vector containing the selected NDBC variables and $\mathbf{r}_t$ is the vector containing the selected reanalysis variables.
-In this way and considering the matched information shown in Fig.
-\ref{fig:directMatching}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
+\textcolor{blue}{In this way and considering the matched information shown in Fig.
+\ref{fig:directMatchingWave}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.}
 
@@
 
- 						For better understanding, considering the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of one \textit{Regression} dataset is shown in Fig.
+ 						\textcolor{blue}{For the sake of clarity, considering the matched information shown in Fig.
+\ref{fig:directMatchingWave}, an example of building a dataset for a \textit{Regression} task is shown in Fig.
 \ref{fig:regressionNoSync}.
 As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one).
-The options considered for the preparation of each final dataset are the following:
+The options considered for the preparation of each final dataset are the following:}
  							\begin{itemize}
@@
 
-							\begin{figure}[ht!]
-								\centering
-								\includegraphics[scale=0.49]{figures/FigureRegressionNoSync.png}
-								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and without synchronisation.}
-								\label{fig:regressionNoSync}
-							\end{figure}
+% 							\begin{figure}[ht!]
+% 								\centering
+% 								\includegraphics[scale=0.49]{figures/FigureRegressionNoSync.png}
+% 								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and without synchronisation.}
+% 								\label{fig:regressionNoSync}
+% 							\end{figure}
 
@@
 
- 						Moreover, considering again the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of the same dataset but applying synchronisation (see Eq.
+ 						\textcolor{blue}{Moreover, considering again the matched information shown in Fig.
+\ref{fig:directMatchingWave}, an example of the creation of the same dataset but applying synchronisation (see Eq.
 \ref{eq:synchronisingRD}) is shown in Fig.
-\ref{fig:regressionSync}.
-
-							\begin{figure}[ht!]
-								\centering
-								\includegraphics[scale=0.49]{figures/FigureRegressionSync.png}
-								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and with synchronisation.}
-								\label{fig:regressionSync}
+\ref{fig:regressionSync}.}
+
+% 							\begin{figure}[ht!]
+% 								\centering
+% 								\includegraphics[scale=0.49]{figures/FigureRegressionSync.png}
+% 								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and with synchronisation.}
+% 								\label{fig:regressionSync}
+% 							\end{figure}
+
+							%TODO: Aañdir rectángulo a la columna de la variable Pres.
+
+							\begin{figure}[ht!]
+								\centering
+								\subfloat[without synchronisation\label{fig:regressionNoSync}]{\includegraphics[width=0.39\textwidth]{figures/FigureRegressionNoSync.png}}\hfill
+								\subfloat[with synchronisation\label{fig:regressionSync}]{\includegraphics[width=0.39\textwidth]{figures/FigureRegressionSync.png}}\hfill
+								\caption{\textcolor{blue}{Example of the creation of a \textit{Regression} dataset with a prediction horizon of $6$h.}} \label{fig:regression}
 							\end{figure}
@@
 						\item \textit{Thresholds of the output attribute} (Classification): Since the values of the variables collected by the buoys are real numbers, it is necessary to discretise them (convert them from real to nominal values) for the attribute selected as output (attribute to be predicted).
-SPAMDA allows the researchers to perform this process by defining the necessary classes with their thresholds, which will be used to carry out such discretisation.
-
-						Considering again the matched information shown in Fig.
-\ref{fig:directMatching}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
-\ref{fig:prediction}.
-The options considered for the preparation of the final dataset are the following:
+SPAMDA allows researchers to perform this process by defining the necessary classes with their thresholds, which will be used to carry out such discretisation.
+
+						\textcolor{blue}{Considering again the matched information shown in Fig.
+\ref{fig:directMatchingWave}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
+\ref{fig:prediction}.} The options considered for the preparation of the final dataset are the following:
 							\begin{itemize}
@@
 							\centering
-							\includegraphics[scale=0.49]{figures/FigureClassification.png}
+							\includegraphics[scale=0.6]{figures/FigureClassification.png}
 							\caption{An example of the creation a \textit{Classification} dataset, with a prediction horizon of $6$h and without synchronisation.}
@@
 
-						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable, because its values have been discretised according to the thresholds.
+						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable, because its values have been discretised according to the thresholds (usually defined by an expert).
 Besides, and due to the $6$h prediction horizon, the last instance is lost ($2017$/$12$/$31$ $18$:$00$) and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up).
@@
 							\item \textit{ARFF}: \textit{Attribute-Relation File Format} \cite{WEKA_ARFF}, which is used by WEKA.
-SPAMDA allows the researchers to directly open the final datasets in the Explorer environment of WEKA (in the same context of work), enabling them to choose the most appropriate ML method to tackle the problem under study.
+SPAMDA allows researchers to directly open the final datasets in the Explorer environment of WEKA (in the same context of work), enabling them to choose the most appropriate ML method to tackle the problem under study.
 
@@
 				A text file that summarises the configuration used in matching process and in the preparation of the matched data is also generated.
-It can be saved and loaded, enabling the researchers to resume their studies at any other time.
+It can be saved and loaded, enabling researchers to resume their studies at any other time.
 
@@
 All these fields summarise the temporal and spatial properties of the data.
-Thus, the researcher can quickly and easily identify each reanalysis file entered in SPAMDA.
+Thus, researcher can quickly and easily identify each reanalysis file entered in SPAMDA.
 
@@
 	\section{A case study applied to Gulf of Alaska}\label{sec:CaseStudy}
-		This section will describe how the application works in a practical approach.
+		%This section will describe how the application works in a practical approach.
 To do so, an example showing how to create a fully processed dataset (final dataset), starting from the raw data, will be described.
@@
 
-		\subsection{Obtaining the final dataset}\label{sec:ObtainingFinalDataset}
-
-			The data collected to perform this example is:
+		\textcolor{blue}{This section describes how SPAMDA works in a practical approach showing two examples to create fully processed datasets (final datasets) starting from the raw data.
+The objective of these final datasets is to be used with SC and ML algorithms for environmental modelling, in this case, to classify waves depending on their height and to predict energy flux in the Gulf of Alaska.}
+
+		\textcolor{blue}{On the one hand, wave classification is tackled as a multi-class classification problem, given that any continuous variable can be discretised in different classes.
+Such waves modelling can be applied with different purposes, such as missing buoy data reconstruction, extreme significant wave heights detection or decision-making and risk assessment about operational works in the sea.}
+
+		\textcolor{blue}{On the other hand, energy flux prediction is addressed as a regression problem.
+Energy flux prediction is related to marine energy and it is useful to characterise the wave energy production from WECs facilities, which could be injected into the electric network or supplied to existing marine platforms.}
+
+		%
+		\textcolor{blue}{\subsection{Gathering the information and introducing it in SPAMDA}}\label{sec:ObtainingFinalDataset}
+
+			The data collected to perform this \textcolor{blue}{case study} is:
 			\begin{enumerate}
@@
 
-			After gathering the information described above, the researcher can open SPAMDA.
-In Figure \ref{fig:main_view}, the main window is shown.
-In order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher has to select the option \textcolor{blue}{\textit{Manage reanalysis data}}.
+			After gathering the information described above, researchers can open SPAMDA.
+In Fig.
+\ref{fig:main_view}, the main view is shown.
+In order to input the reanalysis data which will be used in further steps for creating the final dataset, researchers has to select the option \textbf{\textit{Manage reanalysis data}}.
 
@@
 				\centering
-				\includegraphics[width=0.60\textwidth]{figures/FigureMain_view.png}
-				\caption{SPAMDA main window.}\label{fig:main_view}
-			\end{figure}
-
-			Then, the window of Figure \ref{fig:reanalysis} is shown.
+				\includegraphics[width=0.5\textwidth]{figures/FigureMain_view.png}
+				\caption{SPAMDA main view.}\label{fig:main_view}
+			\end{figure}
+
+			Then, the view of Fig.
+\ref{fig:reanalysis} is shown.
 Here, using the buttons located at the bottom, it is possible to add, delete or consult any data from the different reanalysis files.
-Once the information has been introduced in the application, this window can be closed and the user can go back to the main window to continue entering the information related to the buoy under study.
+Once the information has been introduced in the application, this view can be closed and the user can go back to the main view to continue entering the information related to the buoy under study.
 
@@
 				\includegraphics[width=0.89\textwidth]{figures/FigureManage_reanalysis_data.png}
-				\caption{\textit{Manage reanalysis data} window: downloaded files containing the four closest reanalysis nodes.}\label{fig:reanalysis}
+				\caption{\textit{Manage reanalysis data} view: downloaded files containing the four closest reanalysis nodes.}\label{fig:reanalysis}
 			\end{figure}
@@
 
-			After that, the researcher has to select \textcolor{blue}{\textit{Manage buoys data}} to open the window shown in Figure \ref{fig:manage_buoys}, where several tabs are available.
-In \textcolor{blue}{\textit{Buoys}} tab, the researcher can consult, modify, add or delete different data related to the buoy.
-
-			In order to enter such data, click on the \textcolor{blue}{\textit{New}} button, and then the window shown in Figure \ref{fig:new_buoy} pops up.
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.70\textwidth]{figures/FigureNew_buoy.png}
-				\caption{\textit{New buoy} window: information of the buoy ID 46001.}\label{fig:new_buoy}
+			After that, the researcher has to select \textbf{\textit{Manage buoys data}} to open the view shown in Fig.
+\ref{fig:manage_buoys}, where several tabs are available.
+In \textbf{\textit{Buoys}} tab, the researcher can consult, modify, add or delete different data related to the buoy.
+
+			In order to enter such data, click on the \textbf{\textit{New}} button, and then the view shown in Fig. \ref{fig:new_buoy} pops up.
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.65\textwidth]{figures/FigureNew_buoy.png}
+				\caption{\textit{New buoy} view: information of the buoy ID 46001.}\label{fig:new_buoy}
 			\end{figure}
@@
 			Here the information about the buoy has to be included: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files.
-In this case, the files containing the data from year $2013$ to $2017$ are inserted by clicking on the \textcolor{blue}{\textit{Add file}} button.
-Once the data has been introduced, it is necessary to click on the \textcolor{blue}{\textit{Save}} button to insert the buoy in SPAMDA database.
-After that, the window can be closed.
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.90\textwidth]{figures/FigureDatasets.png}
+In this case, the files containing the data from year $2013$ to $2017$ are inserted by clicking on the \textbf{\textit{Add file}} button.
+Once the data has been introduced, it is necessary to click on the \textbf{\textit{Save}} button to insert the buoy in SPAMDA database.
+After that, the view can be closed.
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.83\textwidth]{figures/FigureDatasets.png}
 				\caption{\textit{Datasets} tab: intermediate datasets of the buoy ID 46001.}\label{fig:show_datasets}
@@
 
-			To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textcolor{blue}{\textit{Datasets}} tab (see Figure \ref{fig:manage_buoys}) to switch to the corresponding view (see Figure \ref{fig:show_datasets}).
+			To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textbf{\textit{Datasets}} tab (see Fig.
+\ref{fig:manage_buoys}) to switch to the corresponding view (see Fig.
+\ref{fig:show_datasets}).
 In this view, the researcher can delete or consult a summary of each intermediate or pre-processed dataset by selecting it from the corresponding list.
 It can also create new ones.
-To proceed with the creation of the intermediate dataset, the user clicks on the \textcolor{blue}{\textit{New}} button, and the view shown in Figure \ref{fig:intermediate} appears.
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.70\textwidth]{figures/FigureNew_intermediate_dataset.png}
+To proceed with the creation of the intermediate dataset, the user clicks on the \textbf{\textit{New}} button, and the view shown in Fig.
+\ref{fig:intermediate} appears.
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.67\textwidth]{figures/FigureNew_intermediate_dataset.png}
 				\caption{\textit{New intermediate dataset} view: creating the intermediate dataset with $5$ annual text files.}\label{fig:intermediate}
@@
 
-			Here the researcher can select the annual text files to be included in the intermediate dataset, by clicking on the \textcolor{blue}{\textit{->}} and \textcolor{blue}{\textit{<-}} buttons.
+			Here the researcher can select the annual text files to be included in the intermediate dataset, by clicking on the \textbf{\textit{->}} and \textbf{\textit{<-}} buttons.
 In this case, all the files introduced before, which correspond to the buoy under study, are selected.
-When the file selection is finished, \textcolor{blue}{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textcolor{blue}{\textit{Save}} button, the creation process starts, showing the status of the process during it.
-After that, in order to prepare the intermediate dataset, the dataset is selected (see Figure \ref{fig:show_datasets}), and then the button \textcolor{blue}{\textit{Open}} is clicked to jump to the tab \textcolor{blue}{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}).
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.90\textwidth]{figures/FigurePreprocess.png}
+When the file selection is finished, \textbf{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textbf{\textit{Save}} button, the creation process starts, showing the status of the process during it.
+After that, in order to prepare the intermediate dataset, the dataset is selected (see Fig.
+\ref{fig:show_datasets}), and then the button \textbf{\textit{Open}} is clicked to jump to the tab \textbf{\textit{Pre-process}} (shown in Fig.
+\ref{fig:preprocess_data}).
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.84\textwidth]{figures/FigurePreprocess.png}
 				\caption{\textit{Pre-process} tab: preprocessing the created intermediate dataset.}\label{fig:preprocess_data}
@@
 
-			In \textcolor{blue}{\textit{Pre-process}} tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
+			In \textbf{\textit{Pre-process}} tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
 Here the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter.
@@
 Furthermore, the attributes MWD, DEWP, VIS and TIDE are removed from the dataset by applying the filter \textit{RemoveByName}, since the first two had more than $92$\% of missing data and the last two $100$\%.
-After finishing the pre-processing of the dataset, the researcher can click on the \textcolor{blue}{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.
+After finishing the pre-processing of the dataset, the researcher can click on the \textbf{\textit{Save}} button, to introduce the description and file name for the current pre-processed dataset.
 
@@
 				\centering
-				\includegraphics[width=0.90\textwidth]{figures/FigureMatching_configuration.png}
-				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files.}\label{fig:matching_conf}
-			\end{figure}
-
-			The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf}.
+				\includegraphics[width=0.84\textwidth]{figures/FigureMatching_configuration.png}
+				\caption{\textcolor{blue}{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (waves classification)}.}\label{fig:matching_conf_wc}
+			\end{figure}
+
+			%The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf_wc}.
 In this view, the researcher can customise (or load) the parameters of the matching process according to their needs and select the prediction task  (described in Section \ref{sec:matching_conf}) that the final dataset will be used for.
 In this example, the following parameters were selected:
+
+			\textcolor{blue}{The next step is to customise (or load) the parameters of the matching process according to the problem being studied and to select the prediction task (described in Section \ref{sec:matching_conf}) that the final dataset will be used for, in this case, waves classification or energy flux prediction.}
+			%
+			\textcolor{blue}{\subsection{Waves classification}}
+
+			\textcolor{blue}{As mentioned above, the objective of the final dataset is to be used with SC and ML algorithms to classify waves depending on their significant height.
+Following sections describe the procedure of performing the data integration provided by SPAMDA, modelling wave height by using classification algorithms available in WEKA.}
+			%
+			\textcolor{blue}{\subsubsection{Obtaining the final dataset}}
+
+			\textcolor{blue}{By clicking on the \textbf{\textit{Matching configuration}} tab, the view shown in Fig.
+\ref{fig:matching_conf_wc} will be opened.
+In this view, the researcher can configure the parameters of the data integration process.
+For this problem, the following parameters were selected:}
+
 			\begin{itemize}
@@
 
-			After configuring the matching process, the researcher can click on the \textcolor{blue}{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
-Given that, in the previous window, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable.
-After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output, as explained in Section {\ref{sec:FinalDatasets}}.
-Then the researcher can click on the \textcolor{blue}{\textit{Update final dataset}} button to see the content shown in the bottom left corner.
-Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textcolor{blue}{\textit{Create final datasets}} button.
-For this case study, the following configuration was applied:
+			After configuring the matching process, the researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Fig.
+\ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
+Given that, in the previous view \textcolor{blue}{(Fig.
+\ref{fig:matching_conf_wc})}, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable \textcolor{blue}{(top left of Fig.
+\ref{fig:final_dataset})}.
+After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output \textcolor{blue}{(top right of Fig.
+\ref{fig:final_dataset})}, as explained in Section {\ref{sec:FinalDatasets}}.
+Then the researcher can click on the \textbf{\textit{Update final dataset}} button to see the content shown in the bottom left corner \textcolor{blue}{(NDBC observations, NNRP variables, missing values, dates)}.
+Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textbf{\textit{Create final datasets}} button \textcolor{blue}{(bottom right of Fig.
+\ref{fig:final_dataset})}.
+For this \textcolor{blue}{example}, the following configuration was applied:
 			\begin{itemize}
 				\item Thresholds: see Table \ref{tab:thresholds}.
-				\item Prediction horizon: 6 hours
-				\item Synchronisation: Disabled
+				\item Prediction horizon: 6 hours.
+				\item Synchronisation: Disabled.
+				\item \textcolor{blue}{Final dataset format: ARFF.}
 			\end{itemize}
@@
 Concretely, the final dataset consists of $7302$ instances and whose distribution is represented in Table \ref{tab:datasetDistribution}.
-
-				\begin{table}[!h]
-
+			%
+				\begin{table}[!h]%
 					\caption{Distribution of instances of the final dataset}
@@
 					\centering
-
 					\begin{tabular}{cc}
-
 						\cline{1-2}
-
 						\textbf{Year}&\textbf{Number of instances}\\
-
 						\cline{1-2}
-
 						$2013$&$1460$\\
@@
 						$2017$&$1458$\\
-
 						\cline{1-2}
-
 						&$7302$\\
-
 						\cline{1-2}
-
 					\end{tabular}
-
 				\end{table}
-
-		\subsection{Obtaining models with ML algorithms}
-
-			Now, it will be described how to obtain wave classification models using the final dataset previously created with SPAMDA.
+		%
+		\textcolor{blue}{\subsubsection{Obtaining classification models with ML algorithms}}%
+
+			Now, the process to obtain wave classification models is described using the final dataset previously created with SPAMDA.
 The modelling will be performed using WEKA as SC and ML tool, which can be opened through SPAMDA, as shown in Fig.
 \ref{fig:openigFinalDatasetWeka}.
-Nevertheless, as mentioned above, the researcher can create the final dataset in CSV format in order to use other ML tools, such as KEEL, Python or R, among others.
+Nevertheless, as mentioned above, the researcher can create the final dataset in CSV format in order to use any other ML tools, such as KEEL, Python or R, among others.
 
@@
 				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka.png}
-				\caption{Final dataset opened with the environment Explorer of WEKA.}
+				\caption{\textcolor{blue}{Final dataset opened with the environment Explorer of WEKA (waves classification)}.}
 				\label{fig:openigFinalDatasetWeka}
@@
 In this way, years from $2013$ to $2015$ will be used for the training phase ($4380$ instances) whereas $2016$ and $2017$ years will be used for the test phase ($2922$ instances).
-Previously to the learning phase, the attributes are normalised to avoid that some attributes could be more important than others because their values could be on a larger scale.
-
-			The classification algorithms that will be considered for the wave modelling are Logistic Regression \cite{hosmer2013applied}, C4.5 \cite{quinlan2014c4}, Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}, which will be applied with the default values of the parameters configured in WEKA.
-Given that Logistic Regression and C4.5 algorithms are deterministic, it is only necessary to launch one execution for each one.
-However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so in this case 30 executions for each one will be carried out.
+Previous to the learning phase, the attributes are normalised to avoid that some attributes dominate others because of a larger scale.
+
+			The classification algorithms that will be considered for wave modelling are Logistic Regression \cite{hosmer2013applied}, C4.5 \cite{quinlan2014c4}, Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}, which will be applied with the default values of the parameters provided by WEKA.
+Given that Logistic Regression and C4.5 algorithms are deterministic, only one run will be considered for each one.
+However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so, in this case, 30 executions for each one will be carried out.
 The results obtained are shown in Table \ref{tab:results}.
@@
 			As can be seen, Random Forest and Multilayer Perceptron algorithms have achieved similar accuracy, but the performance of the latter is slightly better.
-
-			Although it has been performed a first classification approach, to illustrate the use of datasets created with SPAMDA for environmental modelling employing ML algorithms, both models have obtained a good performance despite the problem tackled is to predict wave height six hours in advance.
+Although this is an illustrative classification example using datasets built with SPAMDA, both models have obtained good performance, despite the problem tackled is difficult (prediction is approached six hours in advance).
+
+			\textcolor{blue}{\subsection{Energy flux prediction}}
+
+			\textcolor{blue}{As mentioned above, the final dataset of this example is also used with SC and ML algorithms to predict flux of energy.
+Following sections explain the process of performing the data integration provided by SPAMDA to build the final dataset, modelling the flux of energy by using regression algorithms available in WEKA.}
+
+			\textcolor{blue}{\subsubsection{Obtaining the final dataset}}
+
+			\textcolor{blue}{The researcher can configure the parameters of the data integration by clicking on the \textbf{\textit{Matching configuration}} tab.
+For this problem, as shown in Fig.
+\ref{fig:matching_conf_ef}, the following parameters were selected:}
+
+			\begin{itemize}
+				\item \textcolor{blue}{Attribute to predict: Flux of energy.}
+				\item \textcolor{blue}{Reanalysis data: Air, pressure, u-wind and v-wind.}
+				\item \textcolor{blue}{Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription}).}
+				\item \textcolor{blue}{Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).}
+				\item \textcolor{blue}{Number of final datasets: In this example, this option is disabled, because only one reanalysis node is considered.}
+				\item \textcolor{blue}{Prediction task: Regression.}
+			\end{itemize}
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.85\textwidth]{figures/FigureMatching_configuration_EF.png}
+				\caption{\textcolor{blue}{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (energy flux prediction)}.}\label{fig:matching_conf_ef}
+			\end{figure}
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.85\textwidth]{figures/FigureFinal_datasets_EF.png}
+				\caption{\textcolor{blue}{\textit{Final datasets} tab: content of the final dataset created after data integration.}}\label{fig:final_dataset_EF}
+			\end{figure}
+
+			\textcolor{blue}{After configuring the parameters of the matching process, the next step is to define the final dataset structure according to the selected prediction task.
+Researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Fig.
+\ref{fig:final_dataset_EF}.
+Note that, the thresholds for discretising the output variable (top left of Fig.
+\ref{fig:final_dataset_EF}) are disabled due to, in this case, energy flux prediction is a regression problem.}
+
+			\textcolor{blue}{By default, the time horizon is set to $6$ hours, that is, the energy flux prediction will be performed $6$ hours in advance (top right of Fig.
+\ref{fig:final_dataset_EF}), but researchers can increase such time horizon depending on their needs.
+The synchronisation (in time) of reanalysis variables with the output (explained in Section {\ref{sec:FinalDatasets}}) can be set in this view.
+By clicking on the \textbf{\textit{Update final dataset}} button, researchers can preview the content of the final dataset (bottom left corner of Fig.
+\ref{fig:final_dataset_EF}).
+Finally, the last step would be to set the name, path and output format (CSV or ARFF) of the dataset file, and then the user should click on the \textbf{\textit{Create final datasets}} button (bottom right of Fig.
+\ref{fig:final_dataset_EF}).
+For this example, the following configuration was applied:}
+
+			\begin{itemize}
+				\item \textcolor{blue}{Prediction horizon: 6 hours.}
+				\item \textcolor{blue}{Synchronisation: Disabled.}
+				\item \textcolor{blue}{Final dataset format: ARFF.}
+			\end{itemize}
+
+			\textcolor{blue}{After that, the final dataset would be created and stored in the computer of the researcher according to the introduced configuration, ready to be used as input for SC and ML techniques to tackle the problem of energy flux prediction.
+The number of instances ($7302$) and the distribution of the final dataset (Table \ref{tab:datasetDistribution}) are the same as in the previous example (waves classification) since the data used to create the final dataset and the time horizon selected (6h) are the same.}
+
+		\textcolor{blue}{\subsubsection{Obtaining prediction models with ML algorithms}}
+
+			\textcolor{blue}{In this example, WEKA is used as SC and ML tool to obtain energy flux prediction models, as shown in Fig.
+\ref{fig:openigFinalDatasetWeka_EF}.
+Nonetheless, the final dataset can be created in CSV format so that the researcher can use any other SC and ML tool.}
+
+			\begin{figure}[ht!]
+				\centering
+				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka_EF.png}
+				\caption{\textcolor{blue}{Final dataset opened with the environment Explorer of WEKA (energy flux prediction).}}
+				\label{fig:openigFinalDatasetWeka_EF}
+			\end{figure}
+
+			\textcolor{blue}{For this problem, the same partitioning scheme used in the wave classification problem is considered (60\% train / 40\% test), that is, years from $2013$ to $2015$ for the training phase ($4380$ instances) and years $2016$ and $2017$ for the test phase ($2922$ instances).
+Again, the attributes are normalised prior to the learning phase.}
+
+			\textcolor{blue}{To perform the energy flux modelling, one execution will be run for the deterministic algorithm Linear Regression \cite{Bishop:2006:PRM:1162264}, whereas $30$ executions will be considered for the stochastic ones: Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}.
+Table \ref{tab:results_EF} shows the results obtained by each algorithm using the default values of the algorithm parameters provided by WEKA.}
+
+				\begin{table}[!h]
+
+					\caption{\textcolor{blue}{Results (mean$\pm$SD) obtained by the algorithms}}
+					\label{tab:results_EF}
+					\footnotesize
+					\centering
+
+					\begin{tabular}{m{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
+
+						\cline{1-4}
+
+						\textbf{Algorithm}&\textbf{Root mean squared error}&\textbf{Correlation coefficient}&\\[0.20cm]
+
+						\cline{1-4}
+
+						Linear Regression & $29.6368$ & $0.7296$&\\[0.15cm]
+
+						\cellcolor{gray090}Random Forest &\cellcolor{gray090} $\mathbf{23.4353} \pm \mathbf{0.1313}$ &\cellcolor{gray090} $\textit{0.8408} \pm \textit{0.0021}$&\\[0.15cm]
+
+						Support Vector Machine & $31.3008 \pm 0.1197$ & $0.7275 \pm 0.0015$&\\[0.15cm]
+
+						\cellcolor{gray090}Multilayer Perceptron &\cellcolor{gray090} $\textit{27.1151} \pm \textit{7.5536}$ &\cellcolor{gray090} $\mathbf{0.8444} \pm \mathbf{0.0193}$&\\[0.15cm]
+
+						\cline{1-4}
+
+					\end{tabular}
+
+				\end{table}
+
+			\textcolor{blue}{As can be checked, Random Forest has achieved the best performance in terms of Root mean squared error.
+The standard deviation of the results obtained by Multilayer Perceptron indicates that this algorithm may has been slightly affected by its stochastic component.
+However, both Multilayer Perceptron and Random Forest have obtained an excellent Correlation coefficient.}
+
+			\textcolor{blue}{The objective of this case study has been to illustrate the use of datasets created with SPAMDA to address an energy flux prediction problem.
+An exhaustive comparison of state-of-the-art regression algorithms is not the purpose of this paper.
+However, note that Multilayer Perceptron and Random Forest algorithms have achieved very good results despite the fact that the energy flux prediction has been performed with a time horizon of 6h.}
 
@@
 
-			In this section, it has been described how to use SPAMDA to create a final dataset with the aim of classifying waves.
+			In this section, it has been described how to use SPAMDA to create \textcolor{blue}{final datasets with the aim of classifying waves and predicting flux of energy}.
 However, using the same data described in Section \ref{sec:ObtainingFinalDataset} the researcher can quickly address other objectives or different studies by merely tailoring the matching configuration of the data integration process.
-For example, longer-term wave prediction can be addressed by changing the time horizon, wave modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
+For example, longer-term wave \textcolor{blue}{or energy flux} prediction can be addressed by changing the time horizon, waves modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
 
@@
 
-			As SPAMDA performs all data processing and management to create the datasets, it not only prevents the researchers from performing repetitive tasks but also prevents them from making possible errors.
-In this way, the researchers can focus on the studies they are carrying out.
+			As SPAMDA performs all data processing and management to create the datasets, it not only prevents researchers from performing repetitive tasks but also prevents them from making possible errors.
+In this way, researchers can focus on the studies they are carrying out.
 
@@
 		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
-These datasets, that will be ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), are easily obtained by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
-As a result, the researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
+\textcolor{blue}{The aim of the tool presented in this work is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.}
+
+		\textcolor{blue}{Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).} These datasets can be easily obtained with SPAMDA by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
+As a result, researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
 Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as entering the meteorological information, managing with the incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the SC or ML technique to use), it avoids errors and reduces the time needed.
-In this way, the researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
-
-		The case study illustrates how SPAMDA can be used by the researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
+In this way, researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
+
+		The case study \textcolor{blue}{described in Section \ref{sec:CaseStudy}} illustrates how SPAMDA can be used by researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
+\textcolor{blue}{The case study also covers an example of energy flux prediction, to predict the wave energy that could be exploited by WECs facilities six hours in advance, although such time horizon is customisable.
+Given that this work does not focus on models performance, a more extensive validation or comparison study of the results obtained in both examples has not been carried out.}
 
@@
 
-\acknowledgments{The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
+\acknowledgments{The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the WEKA (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
 
@@
 			\centering
-			\includegraphics[scale=0.45]{figures/FigureMeasurements.png}
+			\includegraphics[scale=0.47]{figures/FigureMeasurements.png}
 			\caption{A fragment of an annual text file with different missing value examples.}
@@
 
-		SPAMDA has been designed to tackle these situations, and it informs the researchers of any incidence found while reading the annual text files for creating the intermediate datasets.
+		SPAMDA has been designed to tackle these situations, and it informs researchers of any incidence found while reading the annual text files for creating the intermediate datasets.
 For the case of measurements that were recorded at a different time than expected, it has been established a time gap of 6 minutes ($10\%$ of an hour).
@@
 $99$ for VIS and TIDE attributes, $999$ for DEWP, MWD and WDIR, etc.).
-SPAMDA interprets these specific values and, after creating the intermediate dataset, the researchers can check if it contains missing values by visualising its statistical information or content.
+SPAMDA interprets these specific values and, after creating the intermediate dataset, researchers can check if it contains missing values by visualising its statistical information or content.
 Remember that SPAMDA provides several filters for recovering missing data, which were described in subsection \ref{sec:Datasets}.
@@
 			\centering
-			\includegraphics[scale=0.40]{figures/FigureCreatingDataset.png}
+			\includegraphics[scale=0.47]{figures/FigureCreatingDataset.png}
 			\caption{Status of the creation of the intermediate dataset for the example of Fig \ref{fig:measurements}.}
@@
 However, in the instance marked with $e)$, given that the measurements dates $01$/$05$/$2017$ $17$:$50$ and $01$/$05$/$2017$ $18$:$50$ are missing, the reanalysis date $01$/$05$/$2017$ $18$:$00$ cannot be matched with buoy data (this date is highlighted in mauve colour in the Figure).
-Depending on the selection made by the researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.36]{figures/FigureMatchingMeasurements.png}
+Depending on the selection made by researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.
+
+		\begin{figure}[ht!]
+			\centering
+			\includegraphics[scale=0.4]{figures/FigureMatchingMeasurements.png}
 			\caption{Matching the measurements (left) and the reanalysis data (right).}

### 2021-01-11 latex_manuscript/MDPI/2_energies_special_issue/envio-3/coverLetter-responseToReviewers/cover-letter-response-reviewers-energies-1069865.tex
Minor Changes Energies Special Issue
Version final para el tercer envio Special Issue Energies

--- 2020-11-30 latex_manuscript/MDPI/2_energies_special_issue/envio-2/responseToReviewers/letter-response-reviewers-energies-1023087.tex
@@
 
-\newcommand{\myJournal}{Energies, Special Issue on ``Soft Computing Techniques in Energy System''}
+\newcommand{\myJournal}{\textbf{Energies}}
 
-\newcommand{\myRefPaper}{\textbf{Paper reference: energies-1023087}}
+\newcommand{\myRefPaper}{\textbf{Paper reference: energies-1069865}}
 
@@
 
-%\tableofcontents
+\tableofcontents
 
 \section{Responses to the Editor}
-We would like to thank the journal's Editor for giving us the opportunity to improve our paper with this review.
-Following the Editor's indications, we have prepared a revised version of the paper, where we have addressed the points required by the reviewers to obtain an improved version of the manuscript.
+We would like to thank the Editor and the anonymous reviewers for this new round of review. We have addressed the final points required by the reviewers.
+
+%\textcolor{red}{Summary of changes}
 
@@
 We sincerely thank the anonymous reviewers for their careful reviews, constructive comments and hints to improve the impact of this work.
-We honestly believe that the quality of the paper has increased after addressing the comments and issues raised by the reviewers.
 
-Most important changes in the paper have been highlighted in \textcolor{blue}{blue color} in the revised version of the manuscript. Please, find below detailed responses to the comments.
+Most important changes highlighted in \textcolor{blue}{blue color} in the previous version of the manuscript have already been set to the default color (\textbf{black color}).
+Lines highlighted in \textcolor{blue}{blue color} correspond to the final points of this new revision.
+Please, find below detailed responses to the comments.
 
-Additionally, in order to improve the english language and style we have carried out a proof-reading.
+% Additionally, in order to improve the English quality and style we have carried out a proof-reading.
+% Additionally, in order to improve the paper's readability, we have carried out a proof-reading, where different typos and weird expressions have been corrected.
+%Additionally, in order to improve the english language and style we have carried out a review.
 
@@
 {
-This work develops a framework to collect, integrate, and preprocess meteorological observation data from NDBC and NNRP.
-Moreover, this framework uses machine learning techniques to do predictions based on these pre-processed datasets. This work
-saves researchers from tedious and repetitive data collection and pre-processing work.
-Also, the use of machine learning in this framework is very useful given that NDBC and NNRP contain missing data in their observation datasets.
+Dear authors,
+Thank you for your efforts in revising the manuscript.
+I believe you did a great job in the revision and the changes alleviated my concerns regarding the manuscript.
+Therefore, I recommend its publication.
 
 \vspace{0.5cm}
-My only comment is that the wave prediction from the machine learning technique is not well validated.
+Looking forward to seeing the published version of your paper.
+
+\vspace{0.5cm}
+Good luck
 }
@@
 {
-Thank you very much for your review, providing very interesting comments and suggestions.
-
-Bla, bla, bla.
-\textbf{These clarifications are mainly described in the paper on page 1, lines 1-33, on page 2, lines 95-112, on page 3, lines 136-151 and 165-181, as well as in the ``Conclusions'' Section}.
+Thank you very much for the excellent review carried out to our work.
 }
@@
 \rcomment{
-Dear authors,
-Thank you for submitting your paper to the Energies. I think your paper is an informative paper that can be published after a major revision.
-
-\vspace{0.5cm}
-It is not clear what the research gap that the paper is addressing. What is the objective of this paper? Please clarify somewhere clearly all your contributions.
+The manuscript is in a good shape. The manuscript is suitable for publication in its current form.
 }
@@
 {
-First of all, we would like to thank Reviewer \#2 for his/her suggestions to improve the paper.
-
-In our understanding, we think that this work contributes several ideas to the field of Machine Learning (ML):
-
-\begin{itemize}
-	\item Bla, bla, bla,
-	\item Bla, bla, bla,
-	\item Bla, bla, bla,
-\end{itemize}
-
-Bla, bla, bla, \textbf{This clarification has been rewritten in page 2, lines 53-61.}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-The literature review is not goal-oriented. The process should be as follows:
-
-\vspace{0.5cm}
-i) Critical evaluation of the literature; ii) identifying the gap based on this critical evaluation of the literature; iii) proposing your hypothesis to address the identified gap; iv) posing the appropriate and relevant research question based on your proposed hypothesis, and finally explaining your proposed method to answer this research question.
-Therefore, you will have a systematic way of conducting your research.
-Right now, the literature review section has no clear objective.
-}
-
-\textbf{Response:}
-{
-Following the instructions of the reviewer, bla, bla, bla.
-\textbf{These clarifications have been included in the paper on page 11, lines 646-660, in the new Figure 6, and in the ``Conclusions'' Section, lines 672-682.}
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-A thorough editorial check and English improvement are needed. Please kindly proofread the entire manuscript.
-
-}
-
-\textbf{Response:}
-{
-A complete revision of the English has been carried out in this review step.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-The conclusion part is also needed to be revised; which questions are answered, what is the value/originality/contribution of the paper, how the proposed method answers the research questions that previous methods are not able to answer?
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla,
-
-Ok, good point. There are two main reasons for choosing
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-It feels you need a king of aggregating results somewhere clearer.
-
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla. The reviewer is right about this comment on input variables.....
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-Please provide more explanation for Figure 21. Also, there are too many figures in the paper. Please reconsider combining some of them together.
-
-}
-\textbf{Response:}
-{
-Bl, bla, bla. Following this suggestion by the reviewer, we have changed...
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-An adequate literature review and a clear gap identification have been tried to be conducted.
-However, authors have ignored some research that has been done in the area.
-I strongly recommend the authors to provide a more comprehensive literature review in the introduction section.
-The following papers are recommended:
-
-\begin{itemize}
-    \item Significant wave height forecasting via an extreme learning machine model integrated with improved complete ensemble empirical mode decomposition.
-Renewable and Sustainable Energy Reviews, 104, 281-295.
-    \item A wavelet-Particle swarm optimization-Extreme learning machine hybrid modeling for significant wave height prediction. Ocean Engineering, 213, 107777.
-    \item Managing computational complexity using surrogate models: a critical review. Research in Engineering Design, 31(3), 275-298.
-    \item Near real-time significant wave height forecasting with hybridized multiple linear regression algorithms. Renewable and Sustainable Energy Reviews, 132, 110003.
-    \item Significant wave height and energy flux prediction for marine energy applications: A grouping genetic algorithm–Extreme Learning Machine approach. Renewable Energy, 97, 380-389.
-    \item Outlook on biofuels in future studies: A systematic literature review. Renewable and Sustainable Energy Reviews, 134, 110326.
-    \item Statistical models for improving significant wave height predictions in offshore operations. Ocean Engineering, 206, 107249.
-    \item Regional ocean wave height prediction using sequential learning neural networks. Ocean Engineering, 129, 605-612.
-    \item Ensemble of surrogates and cross-validation for rapid and accurate predictions using small data sets. AI EDAM, 33(4), 484-501.
-    \item Prediction of significant wave height; comparison between nested grid numerical model, and machine learning models of artificial neural networks, extreme learning and support vector machines.
-Engineering Applications of Computational Fluid Mechanics, 14(1), 805-817.
-\end{itemize}
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla,
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-If you can, please make a small comparison between what did you do and what others did before, as a conclusion.
-
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla,
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-The abstract is not deep enough and Is not well prepared.
-Please try to re-write it better.
-The problem should be clearly stated and the gap in which you are going to address the need to be clarified.
-Simply explain your contributions and key findings.
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla. The reviewer is completely right. We have rewritten the abstract to better describe the application tackled.
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%
-\rcomment
-{
-There are some errors in your reference list. Please check and fix the errors.
-
-}
-
-\textbf{Response:}
-{
-Bla, bla, bla. Thanks for this comment. The distribution....
-
-\vspace{0.5cm}
 We would like to thank you again for the excellent review carried out to our work.
@@
 \rcomment{
-The Authors present a new open source tool for the creation of datasets integrated by meteorological variables from two sources
-of information.
-Basically, a user-friendly software has been developed and described on the basis of pivotal parameters, carefully selected for the specific study of the wave height in the Gulf of Alaska.
-The statistical model can be in principle extended to other meteorological measurements and monitoring.
+The manuscript is related to an interesting topic and based on open database.
 
 \vspace{0.5cm}
-This study is very technical and of interest for a specific audience. By the way, I don’t see any link between the topic of the manuscript and the journal Energies.
+The structure is organized in a good way. Few adjustments are requested to introduction and conclusions.
 
 \vspace{0.5cm}
-The software developed it is clearly presented as useful tool for meteorological application and no application to energy saving, production, conversion or similar it is presented.
+First of all, in the abstract openess of data is mentioned but it is not discussed deeply in stating the background.
+See recent articles published in top-ranked Journals such as Energy, Energies and others.
+As example, consider
 
 \vspace{0.5cm}
-Thus, I recommend to submit the menuscript to a more specialized journal.
+https://doi.org/10.1016/j.energy.2020.118803
+
+\vspace{0.5cm}
+https://doi.org/10.3390/en13226095
+
+\vspace{0.5cm}
+The research question formulated by the authors to fill the identified research gap should open the conclusions section as
+well as reporting limitation and replicability potential of the answer.
+
+\vspace{0.5cm}
+Please check carefully English.
 }
@@
 {
-First of all, we would like to thank Reviewer \#3 for his/her suggestions to improve the submitted manuscript.
+First of all, we would like to thank Reviewer \#3 for his/her suggestions.
 
-\vspace{0.5cm}
-VAMOS A VER PEDAZO DE HIJO DE PUTA, TE HAS QUERIDO QUITAR EL LEER LA REVISION DE UN PLUMAZO
-}
+We have mentioned the two sources of data from which the data integration using SPAMDA is carried out in \textbf{Section ''1.
+Introduction`` (lines 117-122 - ''The meteorological data ...
+build datasets``)}, but we have not made an extensive explanation of these data because they are well known.
+Data provided by the NDBC and the NNRP are available for the scientific community and there are manuals and detailed descriptions on the websites of both information sources.
+Even so, we have dedicated \textbf{Section ``2.
+Meteorological data sources''} to briefly discuss both sources of data and how they are used in SPAMDA, supporting this section with bibliographical references and website addresses.
+%Figure 1 and Table 1 are also included as additional explanations of both data sources.
+
+We sincerely believe that the paper [11] indicated by the reviewer underlines the importance of having well-formatted and high-quality data to be able to use them with modelling techniques, as well as the importance of both data and modelling tools being available to the scientific community and decision-makers.
+This reinforces our work, so we have introduced this reference in \textbf{Section ''1.
+Introduction`` (lines 37-40 - ''In this sense ...
+decision-makers
+[11]``)}.
+The reference [25] also provides background to our work, since it reflects the importance of the correct arrangement of WECs for a better obtaining of power output and energy.
+This reference has been introduced in \textbf{Section ''1.
+Introduction`` (line 69)}.
+
+[11] Manfren, M.; Nastasi, B.; Groppi, D.; Astiaso Garcia, D.
+Open data and energy analytics - An analysis of essential information for energy system planning, design and operation.
+Energies 2020, 13(9), 2334.
+
+[25] Amini, E.; Golbaz, D.; Amini, F.; Majidi Nezhad, M.; Neshat, M.; Astiaso Garcia, D. A Parametric Study of Wave Energy Converter Layouts in Real Wave Models. Energies 2020, 13(22), 6095.
+
+Finally, we have modified Section ''5.
+Conclusions`` to begin with the research gap.
+The replicability of data used in Section ''4.
+A case study applied to Gulf of Alaska`` or any other dataset created with SPAMDA, can be obtained by providing the parameters entered in the tool or simply by providing the final dataset after the last data integration step.
+These clarifications have been included in \textbf{Section ''5.
+Conclusions`` (lines 755-763 - ''Studies on marine energy ...
+among others``, lines 780-783 - ''The final datasets ...
+SC or ML tool used``)}.
+
+%As a conclusion, we believe that open data, open science, open innovation and, more specifically, open energy modelling principles are crucial to create an effective science-policy-market interaction in energy and sustainability transitions.
+Research efforts in energy modelling should be oriented to the identification of solutions that represent good compromises between connectivity and diversity and to the definition of appropriate hierarchies of data and models, for example by using linked open data schema and meta-models.
+
+We would like to thank you again for the suggestions and review carried out to our work.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%% Comment %%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-01-11 latex_manuscript/MDPI/2_energies_special_issue/envio-3/latex/energies-1069865.tex
Version final para el tercer envio Special Issue Energies

@@
 % Full title of the paper (Capitalized)
-\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height \textcolor{blue}{and Energy Flux}}
+\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height and Energy Flux}
 
@@
 Such datasets are not always available in an appropriate format and its preparation and pre-processing implies a lot of time and effort by researchers.
-This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two well-known sources of information: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
-Such datasets can be created using buoys and reanalysis data through customisable procedures, in terms of temporal resolution, predictive and objective variables, \textcolor{blue}{ and can be used by SC and ML methodologies for prediction tasks (classification or regression).
+This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two data sources: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
+Such datasets can be created using buoys and reanalysis data through customisable procedures, in terms of temporal resolution, predictive and objective variables, and can be used by SC and ML methodologies for prediction tasks (classification or regression).
 The objective is providing the research community with an automated and versatile system for the casuistry that entails well-formed and quality data integration, potentially leading to better prediction models.
-The software tool can be used as a supporting tool for coastal and ocean engineering applications, sustainable energy production or environmental modelling; as well as for decision making in the design and construction of coastal protection structures, marine transportation, offshore industry, ocean energy converters and efficient operation of offshore and coastal engineering activities.
-Finally, to illustrate the applicability of the proposed tool, a case study to classify waves depending on their significant height and to predict energy flux in the Gulf of Alaska is presented.
-}
-%Therefore, it does not make a validation or comparison study of the performance of state-of-the-art prediction algorithms.
-}
+The software tool can be used as a supporting tool for coastal and ocean engineering applications, sustainable energy production or environmental modelling; as well as for decision making in the design and building of coastal protection structures, marine transport, ocean energy converters and well-planned running of offshore and coastal engineering activities.
+Finally, to illustrate the applicability of the proposed tool, a case study to classify waves depending on their significant height and to predict energy flux in the Gulf of Alaska is presented.}
 
@@
 energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: estimation of global solar radiation based on sunshine duration \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}, wind power ramp events prediction \cite{DORADOMORENO2017428}, sea surface temperature prediction \cite{He2020}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: global solar radiation estimation \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}, wind power ramp events prediction \cite{DORADOMORENO2017428}, sea surface temperature prediction \cite{He2020}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 		Once quality and well-formed data are obtained, these can be used to extract information and build prediction models that explain the behavior of a certain problem.
-\textcolor{blue}{The choice of the appropriate model, either in engineering problems or in any other problem, is also an important factor in addition to data \cite{Alizadeh2019, Alizadeh2020}}.
+The choice of the appropriate model, either in engineering problems or in any other problem, is also an important factor in addition to data \cite{Alizadeh2019, Alizadeh2020}.
+\textcolor{blue}{In this sense, public and research organisms are increasingly promoting the use of open and robust datasets to boost policies coherent with the environmental exploitation, protection and conservation, as well as modelling tools available to the scientific community and decision-makers \cite{Manfren2020}}.
 Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modelling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.
@@
 		Continuing with this line, special purpose software is usually developed to help researchers to advance in their studies related to energy and environmental modelling, becoming a great support for decision-making in the exploitation and protection of the environment.
-In \cite{Neeraj2020}, a software package in R called ``ForecastTB'' is developed to compare the accuracy of different prediction methods as related to the characteristics of a time series dataset, presenting the software as a stepping stone in ML automation modelling.
-		In \cite{LO2015293}, an integrated simulation tool for the optimum design of bifacial solar panel with reflectors is presented.
-This tool can also be used to analyse the performance of the solar cells.
-A framework for data integration of offshore wind farms is implemented in \cite{NGUYEN2013150} in order to facilitate data exchange and improve operation and maintenance practices.
-		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool`` is presented for comparing PCM - phase change materials- storage systems  with traditional systems that do not entail energy storage, being beneficial in order to support decision-making on energy concepts for buildings.
+In \cite{Neeraj2020}, a software package in R called ``ForecastTB'' is developed for comparing the performance of distinct prediction methods, presenting the software as a stepping stone in ML automation modelling.
+		In \cite{LO2015293}, an integrated simulation tool to optimise the design of bifacial solar panel with reflectors is presented.
+This tool can also be applied to study the efficiency of the solar cells.
+A framework for integrating information from offshore wind farms is implemented in \cite{NGUYEN2013150} in order to ease data interchange and enhance operation and maintenance practices.
+		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool`` is presented for comparing PCM - phase change materials- storage systems with conventional systems that do not involve energy storage, being beneficial for supporting decision-making on energy concepts for buildings.
 		A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), to support different operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
+\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), for supporting distinct operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
 \cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.
@@
 Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \textcolor{blue}{\cite{en12091657, BREDE201344, Alizadeh2020a}}.
+Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{en12091657, BREDE201344, Alizadeh2020a}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing offshore oil and gas platforms \cite{OLIVEIRAPINTO2019556} or seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
-WECs are mechanical devices that convert kinetic energy into electrical energy by means of either the vertical oscillation of waves or the linear motion of them.
+In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed, \textcolor{blue}{even arranged in array form \cite{Amini2020}}, to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing offshore oil and gas platforms \cite{OLIVEIRAPINTO2019556} or seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
+WECs are mechanical devices that convert kinetic energy into electrical energy through either the vertical oscillation or the linear motion of waves.
 Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
-As a consequence of this complexity, many aspects of WEC design, deployment and operation \textcolor{blue}{\cite{CROWLEY2018159, Abdelkhalik2016, 6898109}} need a proper prediction of waves \textcolor{blue}{\cite{en11010011, Kaloop2020}}, in order to maximise the wave energy extraction \cite{en80910370}.
-For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most important wave parameters in this regard: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-
-		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results, \textcolor{blue}{ either in engineering \cite{Alizadeh2020}, energy or environmental problems} \cite{Fi-John2019,Mosavi2019,GUO201816}.
+As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves \cite{en11010011, Kaloop2020}, in order to maximise the wave energy extraction \cite{en80910370}.
+For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most relevant wave parameters related to this aspect: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
+
+		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results, either in engineering \cite{Alizadeh2020}, energy or environmental problems \cite{Fi-John2019,Mosavi2019,GUO201816}.
 SC and ML methodologies can be used not only by experienced computer scientists but also by other researchers.
@@
 ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction.
-\textcolor{blue}{In \cite{Ali2020}, a reliable ML model based on multiple linear regression and covariant-weighted least square estimation for $H_s$ modelling is presented in order to forecast half-hourly nearly real-time significant wave height values.
-In \cite{Cornejo-Bueno2016}, an  approach for feature selection problems is developed and applied to a specific problem of $H_s$ and $F_e$ prediction in oceanic buoys,  obtaining excellent results in terms of prediction quality.
-In \cite{Emmanouil2020}, a Bayesian Network system provides an useful tool for the decision making process of installation and maintenance operations in offshore wind farms using predictions of $H_s$, among others.
-In \cite{Shamshirband2020}, several ML methods are implemented and compared for the prediction of $H_s$ in the Persian Gulf, the extreme learning machine (ELM) providing the best results.} The problem is that, in order to apply \textcolor{blue}{ML and SC techniques}, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge.
+In \cite{Ali2020}, a reliable ML model based on multiple linear regression and covariant-weighted least square estimation for $H_s$ modelling is presented in order to predict significant wave height 30 minutes in advance.
+In \cite{Cornejo-Bueno2016}, an  approach for feature selection problems is developed and applied for $H_s$ and $F_e$ prediction in oceanic buoys, obtaining very good results.
+In \cite{Emmanouil2020}, a Bayesian Network system provides a helpful tool to support decision-making process of installation and maintenance operations in offshore wind farms using predictions of $H_s$, among others.
+In \cite{Shamshirband2020}, several ML methods are implemented and compared for the prediction of $H_s$ in the Persian Gulf, the extreme learning machine (ELM) providing the best results.
+The problem is that, in order to apply ML and SC techniques, it is essential to obtain datasets with relevant information about the issue under study, used to infer knowledge.
 Usually, these datasets are not publicly available in a friendly format, and their generation is the first step needed.
@@
 Given that such process is of great relevance and has an extensive casuistry, the present work has been specially focused on it.
-Moreover, depending on the subject and the SC and ML technique to be applied, or even if the researcher considers other factors in order to improve the results or have more in-depth conclusions, the datasets would have to be updated afterwards.
+Moreover, depending on the subject and the SC and ML technique to be applied, or even if the researcher considers other factors in order to enhance the performance obtained or have more in-depth conclusions, the datasets would have to be updated afterwards.
 In summary, many important details and different intermediate steps have to be considered when creating suitable datasets, specially when data integration is required, resulting in an extremely tedious task.
@@
 
-			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a lot of time and effort in early stages of the research.
+			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a great deal of time and endeavour in early stages of the research.
 
@@
 			\item It provides information about the quality and quantity of the data.
-\textcolor{blue}{SPAMDA allows preliminary studies of missing values (dates or measurements not recorded) in buoys managed by NDBC, so that the researcher can have an idea of the quality of the data recorded by the buoys and about their suitability for the intended purpose.
-In any case, SPAMDA allows data integration taking into account such missing values when needed by the user.}
-
-			\item \textcolor{blue}{Estimation of the amount of energy flux that can be produced at different prediction horizons: short-term, mid-term or long-term.
-Although this work does not focus on models performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.}
+SPAMDA allows preliminary studies of missing values (dates or measurements not recorded) in buoys managed by NDBC, so that the researcher can have an idea of the quality of the data recorded by the buoys and about their suitability for the intended purpose.
+In any case, SPAMDA allows data integration taking into account such missing values when needed by the user.
+
+			\item Estimation of the amount of energy flux that can be produced at different prediction horizons: short-term, mid-term or long-term.
+Although this work does not focus on models performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.
 
@@
 
-			\item Possibility of selecting one or more reanalysis nodes near the localisation under study, \textcolor{blue}{which could provide a better description of the problem to achieve more accurate models.}
-
-			\item \textcolor{blue}{Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoys measurements, such as normalisation and missing data recovery.}
+			\item Possibility of selecting one or more reanalysis nodes near the localisation under study, which could provide a better description of the problem to achieve more accurate models.
+
+			\item Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoys measurements, such as normalisation and missing data recovery.
 
 			\item It facilitates data management and well-organised storage of the datasets.
-\textcolor{blue}{Environmental studies in different geographical locations can be carried out by merely introducing and using other collected data.}
+Environmental studies in different geographical locations can be carried out by merely introducing and using other collected data.
 
@@
 
-		\textcolor{blue}{Therefore, the functionalities and characteristics that SPAMDA offers make it a supporting tool for researchers, which could be used in applications related to coastal and ocean engineering, and also in marine energy prediction.
+		Therefore, the functionalities and characteristics that SPAMDA offers make it a supporting tool for researchers, which could be used in applications related to coastal and ocean engineering, and also in marine energy prediction.
 In \cite{Kumar2020}, the estimation of energy supply sources in hybrid energy systems is based on the amount of energy that can be obtained by a marine energy system within a prediction horizon.
@@
 In \cite{Ali2019}, significant wave height forecasting is applied for decision-making in exploitation and environmental protection for the construction of marine energy storage plants, future strategies on renewable energy and coastal planning.
-Other examples of application are: design of offshore structures and ports \cite{CHATZIIOANNOU2017126}, decision-making and risk assessment about operational works in the sea \cite{DALGIC2015211}, security systems for structures or naval security \cite{Spaulding2020}.}
+Other examples of application are: design of offshore structures and ports \cite{CHATZIIOANNOU2017126}, decision-making and risk assessment about operational works in the sea \cite{DALGIC2015211}, security systems for structures or naval security \cite{Spaulding2020}.
 
@@
 
-			\item NDBC is a part of the \textit{National Weather Service} (NWS).
-NDBC designs, develops, operates, and maintains a network of data collecting buoys (stations).
-The mission of the network is to collect real-time marine meteorological and oceanographic observations, such as $H_s$, dominant wave period, or wind speed and direction, among others.
-
-			The buoys maintained by NDBC are deployed in the coastal and offshore waters around oceans and seas, and they are equipped with assorted sensors which allow them to perform different measurements.
+			\item NDBC belongs to the \textit{National Weather Service} (NWS) and operates and supports a network of marine and ocean buoys that record data.
+The mission of the network is to record marine and ocean meteorological data, such as $H_s$, dominant wave period, or wind speed and direction, among others.
+
+			The buoys maintained by NDBC are located in coastal and offshore waters, and they are provided with specific sensors and devices which allow them to perform measurements.
 The information collected by the buoys is available on the NDBC website \cite{NOAA_1}, and it is divided into different groups.
@@
 
-				\caption{Measurements descriptions and units of each meteorological variable or attribute collected by the buoys.}
+				\caption{Measurements descriptions and units of each meteorological variable or attribute collected by the buoys (a detailed description can be found in NDBC website \cite{NOAA_4})}.
 				\label{tab:measurementsDescription}
@@
 
-				\begin{tabular}{ccm{8.60cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
+				\begin{tabular}{ccm{7.0cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
 
@@
 
-					WDIR & degT & Wind direction (the direction the wind is coming from in degrees clockwise from true North) during the same period used for WSPD. \\
-
-					\cellcolor{gray090}WSPD & \cellcolor{gray090} m/s & \cellcolor{gray090} Wind speed (m/s) averaged over an eight-minute period for buoys and a two-minute period for land stations.
-Reported Hourly.
-\\
-
-					GST &  m/s & Peak 5 or 8 second gust speed (m/s) measured during the eight-minute or two-minute period. \\
-
-					\cellcolor{gray090} WVHT & \cellcolor{gray090} m & \cellcolor{gray090} Significant wave height (meters) is calculated as the average of the highest one-third of all of the wave heights during the 20-minute sampling period.
-\\
-
-					DPD & sec & Dominant wave period (seconds) is the period with the maximum wave energy. \\
-
-					\cellcolor{gray090}APD & \cellcolor{gray090} sec & \cellcolor{gray090} Average wave period (seconds) of all waves during the 20-minute period. \\
-
-					MWD & degT & The direction from which the waves at the dominant period (DPD) are coming.
-The units are degrees from true North, increasing clockwise, with North as 0 (zero) degrees and East as 90 degrees.
-\\
-
-					\cellcolor{gray090}PRES & \cellcolor{gray090} hPa & \cellcolor{gray090} Sea level pressure (hPa).
-For C-MAN sites and Great Lakes buoys, the recorded pressure is reduced to sea level using the method described in NWS Technical Procedures Bulletin 291 (11/14/80).
-\\
-
-					ATMP & degC & Air temperature (Celsius degrees). &\\[0.10cm]
-
-					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature (Celsius degrees).
-For buoys the depth is referenced to the hull's waterline.
-For fixed platforms it varies with tide, but is referenced to, or near Mean Lower Low Water (MLLW).\\
-
-					DEWP & degC & Dewpoint temperature taken at the same height as the air temperature measurement. \\
-
-					\cellcolor{gray090}VIS & \cellcolor{gray090} nmi & \cellcolor{gray090} Station visibility (nautical miles). Note that buoy stations are limited to reports from 0 to 1.6 nmi. \\
-
-					TIDE & ft & The water level in feet above or below MLLW. \\
+					WDIR & degT & The direction the wind is coming from true North. \\
+
+					\cellcolor{gray090}WSPD & \cellcolor{gray090} m/s & \cellcolor{gray090} The speed of the wind. \\
+
+					GST &  m/s & Peak of gust speed. \\
+
+					\cellcolor{gray090} WVHT & \cellcolor{gray090} m & \cellcolor{gray090} Significant wave height. \\
+
+					DPD & sec & Dominant wave period (maximum wave energy). \\
+
+					\cellcolor{gray090}APD & \cellcolor{gray090} sec & \cellcolor{gray090} Average wave period of all waves. \\
+
+					MWD & degT & The direction from which the waves at the dominant period are coming. \\
+
+					\cellcolor{gray090}PRES & \cellcolor{gray090} hPa & \cellcolor{gray090} Sea level pressure. \\
+
+					ATMP & degC & Air temperature. &\\[0.10cm]
+
+					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature.\\
+
+					DEWP & degC & Dewpoint temperature. \\
+
+					\cellcolor{gray090}VIS & \cellcolor{gray090} nmi & \cellcolor{gray090} Visibility of the station. \\
+
+					TIDE & ft & The water level. \\
 
@@
 
-			\item NNRP provides three-dimensional global reanalysis of numerous meteorological variables (e.g.
-air temperature, components South-North and West-East of wind speed, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$ Z (Zulu time), $06$ Z, $12$ Z and $18$ Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid.
+			\item NNRP provides three-dimensional global reanalysis of numerous meteorological observations (e.g.
+components  \textit{Zonal} and \textit{Meridional} of the velocity of the wind, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$ Z (Zulu time), $06$ Z, $12$ Z and $18$ Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid.
 Weather observations are from different sources, such as ships, satellites and radar, among others.
-Reanalysis data is created assimilating such observations using the same climate model throughout the entire reanalysis period in order to reduce the effects of modelling changes on climate statistics.
+Reanalysis data is created assimilating such observations employing the same climate model along the whole period of reanalysis in order to decrease the impact of modelling changes on climate statistics.
 Such information has become a substantial support of the needs of the research community, even more in locations where instrumental (real time) data is not available.
@@
 Each reanalysis file contains the values of a meteorological variable estimated by a mathematical model for each reanalysis node.
-For a better understanding, in Fig.
-\ref{fig:subGrid} an approximate representation of a sub-grid containing six reanalysis nodes around the geographical location of a buoy (obtained from NDBC) is shown.
+For the sake of clarity, in Fig.
+\ref{fig:subGrid}, an example to approximately illustrate a sub-grid containing six nodes of reanalysis surrounding the geographic localisation of a buoy (obtained from NDBC) is shown.
 
@@
 				\includegraphics[scale=0.55]{figures/FigureSubGrid.jpg}
-				\caption{Example of a six sub-grid reanalysis nodes around the \textit{Station 46001}.}
+				\caption{Sub-grid representation of six nodes of reanalysis surrounding the \textit{Station 46001}.}
 				\label{fig:subGrid}
@@
 		SPAMDA combines  meteorological information from NDBC and NNRP to obtain new datasets for oceanic and atmospheric studies.
-In order to do so, SPAMDA manages three different types of datasets, which will be described in detail in the following sections, but are briefly introduced bellow for giving the reader a better general understanding:
+In order to do so, SPAMDA manages three different types of datasets which are described in the following sections, but are briefly introduced bellow for giving the reader a better general understanding:
 			\begin{itemize}
@@
 					\item \textit{Longitude}: West or East geographical localisation (degrees) of the buoy.
-					\item \textit{Measurements files}: The above-mentioned annual text files of the standard meteorological information collected by the buoy and downloaded from the NDBC website.
+					\item \textit{Measurements files}: The above-mentioned annual text files of the standard meteorological information recorded by the buoy and downloaded from the NDBC website.
 This will be used for the creation of the intermediate datasets.
@@
 				Data pre-processing prepares the raw data (intermediate datasets) to be able to be treated correctly by ML algorithms.
-In this way, the quality of data can be improved prior to the learning phase, by applying pre-processing tasks (filters).
+This action can enhance the quality of data before the learning phase, by applying pre-processing tasks (filters).
 The result will be referenced as pre-processed datasets.
@@
 				SPAMDA allows researchers to undo the last filter applied or to restore the initial content of the intermediate dataset.
-Besides, the content and relevant statistical information (number of instances with missing values, minimum and maximum values, mean and standard deviation) of the intermediate and the pre-processed datasets can be visualised in this module.
+Besides, the content and relevant statistical information of the intermediate and the pre-processed datasets can be visualised in this module, for example: minimum and maximum values, mean, standard deviation or even the number of instances with missing values.
 
@@
 
-				Such process merges the information of both sources that match on time, but, given that the measurements of the buoys are hourly collected from $00$:$50$ to $23$:$50$ UTC, and the reanalysis data is available every $6$ hours at $00$ Z, $06$ Z, $12$ Z and $18$ Z, the matching can only be carried every $6$ hours (discarding the rest of measurements from the buoy data).
+				Such process merges the information of both sources that match on time, but, given that the reanalysis data is available with a minimum time horizon of $6$ hours at $00$ Z, $06$ Z, $12$ Z and $18$ Z, and the measurements of the buoys are recorded at hourly intervals, from $00$:$50$ to $23$:$50$ UTC, the matching can only be carried every $6$ hours (discarding the rest of measurements from the buoy data).
 Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest buoy measurement (before or after) within a maximum of 60 minutes of difference.
@@
 					\item \textit{Flux of energy} \cite{FERNANDEZ201544}: When the $F_e$ is selected, it will be used as output.
-This attribute is not collected by the buoys, but it can be calculated from two wave parameters: $H_s$ and $T_e$, which are collected as WVHT and APD attributes, respectively, and were described in Table \ref{tab:measurementsDescription}.
-In this way, SPAMDA obtains the $F_e$ of each instance using the following equation:
+This attribute is not collected by the buoys, but there are two parameters from which it can be computed: $H_s$ and $T_e$, which are collected as WVHT and APD attributes, respectively, and were described in Table \ref{tab:measurementsDescription}.
+In this way, SPAMDA obtains the $F_e$ (measured in kilowatts per meter) of each instance using the following equation:
 					\begin{equation}
@@
 
-					where $F_e$ is measured in kilowatts per meter, $H_s$ is measured in meters and $T_e$ is measured in seconds.
-Note also that $F_e$ is defined in Eq.
-\ref{eq:fluxOfEnergy} as an average energy flux ($H_s$ is a kind of average wave height), though for simplicity it will be referred just as flux of energy.
+					where $H_s$ is measured in meters and $T_e$ in seconds.
+$F_e$ is referred to as flux of energy but it is defined as an average energy flux because $H_s$ is an average wave height (see descriptions of the measurements on the NDBC website).
 
@@
 
-							\item Consider only some of the reanalysis nodes contained in each file: in this case, only the information of the $N$ closest reanalysis nodes to the buoy will be used  ($N$ given by the user).
-To do that, SPAMDA uses the \textit{Haversine} equation \cite{Haversine_2009} to calculate the distance from each reanalysis node to the location of the buoy and obtain the closest ones.
-Haversine equation is also known as the great circle distance and performs calculation from main point to destination point with a trigonometric function using latitude and longitude:
+							\item Consider only some of the reanalysis nodes contained in each file: in this case, the information used is only that corresponding to the closest nodes to the buoy (the number of nodes, $N$, is indicated by the user).
+To do that, SPAMDA uses the \textit{Haversine} equation \cite{Haversine_2009} (or the great-circle distance) to calculate the distance from the location of the buoy to each node of reanalysis and obtain the closest ones.
+Haversine equation performs calculation from main point to destination point with a trigonometric function:
 								\begin{eqnarray}
@@
 								\end{eqnarray}
-							where $p_0$ is the buoy geographical location, $p_j$ stands for the location of each reanalysis node, and $lat$ and $lon$ are the latitude and longitude of the points, respectively.
+							where $p_0$ is the geographical location of the buoy and $p_j$ is the position of each node. Finally, $lat$ and $lon$ represent the latitude and longitude of the positions of the points.
 						\end{itemize}
@@
 This weighted mean is obtained by SPAMDA and uses Eq.
-\ref{eq:Haversine} to obtain the distance from each reanalysis node to the location of the buoy.
-Once the distances have been calculated they are inverted and normalised as follows:
+\ref{eq:Haversine} to calculate the distance from the geographical position of the buoy to each node of reanalysis.
+Once the distances have been computed, they are normalised and inverted as shown in the following equation:
 								\begin{linenomath*}
@@
 
-							With these weights, a weighted mean of each variable of reanalysis is obtained for each of the $N$ nodes.
-Therefore, the closest reanalysis nodes to the localisation of the buoy will provide more information.
+							Then, with these calculated weights, a weighted mean of each variable of reanalysis is obtained for each of the $N$ nodes.
+In this way, the closest reanalysis nodes to the geographical position of the buoy will provide more information.
 
@@
 
-				\textcolor{blue}{Once the matching parameters have been described, for a better understanding of them,  Fig.
+				Once the matching parameters have been described, for a better understanding of them,  Fig.
 \ref{fig:directMatching} presents an example of the data integration considering the data shown in Fig.
-\ref{fig:matchingProcess} and using the following configuration\footnote{Note that the date is shown just for a better understanding, but it will not be included in the final dataset.}:}
-					\begin{itemize}
-						\item \textcolor{blue}{Attribute to predict: variable WVHT (Fig. \ref{fig:directMatchingWave}) / flux of energy (Fig. \ref{fig:directMatchingFlux})}.
+\ref{fig:matchingProcess} and using the following configuration\footnote{Note that the date is shown just for a better understanding, but it will not be included in the final dataset.}:
+					\begin{itemize}
+						\item Attribute to predict: variable WVHT (Fig. \ref{fig:directMatchingWave}) / flux of energy (Fig. \ref{fig:directMatchingFlux}).
 						\item Variable Pres as reanalysis input attribute.
@@
 
-% 				\begin{figure}[ht!]
-% 					\centering
-% 					\includegraphics[scale=0.49]{figures/FigureDirectMatching.png}
-% 					\caption{An example of the matched information for \textit{Direct matching}.}
-% 					\label{fig:directMatching}
-% 				\end{figure}
-
 				\begin{figure}[ht!]
@@
 					\subfloat[attribute to predict: flux of energy\label{fig:directMatchingFlux}]{\includegraphics[width=0.47\textwidth]{figures/FigureDirectMatching_EF.png}}\hfill
-					\caption{\textcolor{blue}{Example of data integration for \textit{Direct matching}.}} \label{fig:directMatching}
+					\caption{Example of data integration for \textit{Direct matching}.} \label{fig:directMatching}
 				\end{figure}
@@
 As the minimum prediction horizon is $6$h, at least one instance will be lost.
-The relation between the inputs and the attribute to predict will be defined as follows:
+The relation between the inputs and the output (attribute to predict) is defined as follows:
 						\begin{linenomath*}
@@
 						\end{linenomath*}
-						where $t$ represents the time instant to study and $\Delta t$ the prediction horizon; $o$ is the attribute to be predicted, $\mathbf{b}_t$ is the vector containing the selected NDBC variables and $\mathbf{r}_t$ is the vector containing the selected reanalysis variables.
-\textcolor{blue}{In this way and considering the matched information shown in Fig.
-\ref{fig:directMatchingWave}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.}
+						where $t$ is the time instant to study, $\Delta t$ is the prediction horizon, $o$ is the attribute to be predicted, $\mathbf{b}_t$ represents the vector that contains the selected NDBC variables, and, finally, $\mathbf{r}_t$ represents the vector that contains the selected reanalysis variables.
+In this way and considering the matched information shown in Fig.
+\ref{fig:directMatchingWave}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
 
@@
 
- 						\textcolor{blue}{For the sake of clarity, considering the matched information shown in Fig.
+ 						For the sake of clarity, considering the matched information shown in Fig.
 \ref{fig:directMatchingWave}, an example of building a dataset for a \textit{Regression} task is shown in Fig.
@@
 As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one).
-The options considered for the preparation of each final dataset are the following:}
+The options considered for the preparation of each final dataset are the following:
  							\begin{itemize}
@@
  							\end{itemize}
-
-% 							\begin{figure}[ht!]
-% 								\centering
-% 								\includegraphics[scale=0.49]{figures/FigureRegressionNoSync.png}
-% 								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and without synchronisation.}
-% 								\label{fig:regressionNoSync}
-% 							\end{figure}
 
@@
 
- 						\textcolor{blue}{Moreover, considering again the matched information shown in Fig.
+ 						Moreover, considering again the matched information shown in Fig.
 \ref{fig:directMatchingWave}, an example of the creation of the same dataset but applying synchronisation (see Eq.
 \ref{eq:synchronisingRD}) is shown in Fig.
-\ref{fig:regressionSync}.}
-
-% 							\begin{figure}[ht!]
-% 								\centering
-% 								\includegraphics[scale=0.49]{figures/FigureRegressionSync.png}
-% 								\caption{An example of the creation of a \textit{Regression} dataset, with a prediction horizon of $6$h and with synchronisation.}
-% 								\label{fig:regressionSync}
-% 							\end{figure}
-
-							%TODO: Aañdir rectángulo a la columna de la variable Pres.
+\ref{fig:regressionSync}.
 
@@
 								\subfloat[with synchronisation\label{fig:regressionSync}]{\includegraphics[width=0.39\textwidth]{figures/FigureRegressionSync.png}}\hfill
-								\caption{\textcolor{blue}{Example of the creation of a \textit{Regression} dataset with a prediction horizon of $6$h.}} \label{fig:regression}
+								\caption{Example of the creation of a \textit{Regression} dataset with a prediction horizon of $6$h.} \label{fig:regression}
 							\end{figure}
@@
 
-						\textcolor{blue}{Considering again the matched information shown in Fig.
+						Considering again the matched information shown in Fig.
 \ref{fig:directMatchingWave}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
-\ref{fig:prediction}.} The options considered for the preparation of the final dataset are the following:
+\ref{fig:prediction}.
+The options considered for the preparation of the final dataset are the following:
 							\begin{itemize}
@@
 
-	%\section{Case study}\label{sec:CaseStudy}
 	\section{A case study applied to Gulf of Alaska}\label{sec:CaseStudy}
-		%This section will describe how the application works in a practical approach.
-To do so, an example showing how to create a fully processed dataset (final dataset), starting from the raw data, will be described.
-The objective of this final dataset is to be used with ML algorithms for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
-In this case, the problem will be tackled as a multi-class classification problem, given that any continuous variable can be discretised in different classes.
-Such waves modelling can be applied with different purposes, energy flux range forecasting, missing buoy data reconstruction, or extreme significant wave heights detection, among others.
-
-		\textcolor{blue}{This section describes how SPAMDA works in a practical approach showing two examples to create fully processed datasets (final datasets) starting from the raw data.
-The objective of these final datasets is to be used with SC and ML algorithms for environmental modelling, in this case, to classify waves depending on their height and to predict energy flux in the Gulf of Alaska.}
-
-		\textcolor{blue}{On the one hand, wave classification is tackled as a multi-class classification problem, given that any continuous variable can be discretised in different classes.
-Such waves modelling can be applied with different purposes, such as missing buoy data reconstruction, extreme significant wave heights detection or decision-making and risk assessment about operational works in the sea.}
-
-		\textcolor{blue}{On the other hand, energy flux prediction is addressed as a regression problem.
-Energy flux prediction is related to marine energy and it is useful to characterise the wave energy production from WECs facilities, which could be injected into the electric network or supplied to existing marine platforms.}
-
-		%
-		\textcolor{blue}{\subsection{Gathering the information and introducing it in SPAMDA}}\label{sec:ObtainingFinalDataset}
-
-			The data collected to perform this \textcolor{blue}{case study} is:
+
+		This section describes how SPAMDA works in a practical approach showing two examples to create fully processed datasets (final datasets) starting from the raw data.
+The objective of these final datasets is to be used with SC and ML algorithms for environmental modelling, in this case, to classify waves depending on their height and to predict energy flux in the Gulf of Alaska.
+
+		On the one hand, wave classification is addressed as a multi-class approach, given that a continuous attribute can be discretised, using different thresholds, in distinct classes.
+Such waves modelling can be applied with different purposes, such as missing buoy data reconstruction, extreme significant wave heights detection or decision-making and risk assessment about operational works in the sea.
+
+		On the other hand, the prediction of the energy flux is addressed as a regression problem.
+Energy flux prediction is related to marine energy and it is useful to characterise the wave energy production from WECs facilities, which could be injected into the electric network or supplied to existing marine platforms.
+
+		\subsection{Gathering the information and introducing it in SPAMDA}\label{sec:ObtainingFinalDataset}
+
+			The data collected to perform this case study is:
 			\begin{enumerate}
@@
 This data is publicly available at the NDBC website.
-			\item Complementary information collected from reanalysis data containing air temperature (air), pressure (pres) and two components of wind speed measurements: South-North (vwind) and West-East (uwind).
-This information will be collected from the four closest reanalysis nodes surrounding the geographical location of the buoy.
-This data is publicly available at the NNRP website and can be downloaded in NetCDF format.
+			\item Complementary information collected from reanalysis data containing air temperature (air), pressure (pres) and two components of wind speed measurements, South-North (vwind) and West-East (uwind).
+This information can be downloaded from the NNRP website in NetCDF format for the four closest nodes of reanalysis surrounding the position of the buoy.
 Concretely, the closest reanalysis nodes downloaded are $57.5$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150$ W, $55$ N $\times$ $147.5$ W and $55$ N $\times$ $150$ W.
@@
 			\end{enumerate}
-
-			%After gathering the information described above\footnote{Further instructions for downloading this data can be found in the user manual of the application.}, the researcher can open SPAMDA.
-In Figure \ref{fig:main_view}, the main window is shown.
-In order to input the reanalysis data which will be used in further steps for creating the final dataset, the researcher has to select the option \textcolor{blue}{\textit{Manage reanalysis data}}.
 
@@
 				\includegraphics[width=0.84\textwidth]{figures/FigurePreprocess.png}
-				\caption{\textit{Pre-process} tab: preprocessing the created intermediate dataset.}\label{fig:preprocess_data}
+				\caption{\textit{Pre-process} tab: pre-processing the created intermediate dataset.}\label{fig:preprocess_data}
 			\end{figure}
@@
 Finally, the data has been pre-processed in order to be ready for its future use in ML algorithms.
-In order to achieve a more accurate description of the problem under study, a matching process can be carried out to merge the processed data from NDBC with the reanalysis data (also entered previously) from NNRP.
+Then, a data integration process can be carried out to merge the processed data from NDBC with the reanalysis data (also included previously) from NNRP.
 
@@
 				\includegraphics[width=0.84\textwidth]{figures/FigureMatching_configuration.png}
-				\caption{\textcolor{blue}{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (waves classification)}.}\label{fig:matching_conf_wc}
-			\end{figure}
-
-			%The next step is to click on the \textcolor{blue}{\textit{Matching configuration}} tab, to open the view shown in  Figure \ref{fig:matching_conf_wc}.
-In this view, the researcher can customise (or load) the parameters of the matching process according to their needs and select the prediction task  (described in Section \ref{sec:matching_conf}) that the final dataset will be used for.
-In this example, the following parameters were selected:
-
-			\textcolor{blue}{The next step is to customise (or load) the parameters of the matching process according to the problem being studied and to select the prediction task (described in Section \ref{sec:matching_conf}) that the final dataset will be used for, in this case, waves classification or energy flux prediction.}
-			%
-			\textcolor{blue}{\subsection{Waves classification}}
-
-			\textcolor{blue}{As mentioned above, the objective of the final dataset is to be used with SC and ML algorithms to classify waves depending on their significant height.
-Following sections describe the procedure of performing the data integration provided by SPAMDA, modelling wave height by using classification algorithms available in WEKA.}
-			%
-			\textcolor{blue}{\subsubsection{Obtaining the final dataset}}
-
-			\textcolor{blue}{By clicking on the \textbf{\textit{Matching configuration}} tab, the view shown in Fig.
+				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (waves classification).}\label{fig:matching_conf_wc}
+			\end{figure}
+
+			The next step is to customise (or load) the parameters of the matching process according to the problem being studied and to select the prediction task (described in Section \ref{sec:matching_conf}) that the final dataset will be used for, in this case, waves classification or energy flux prediction.
+
+			\subsection{Waves classification}
+
+			As mentioned above, the objective of the final dataset is to be used with SC and ML algorithms to classify waves depending on their significant height.
+Following sections describe the procedure of performing the data integration provided by SPAMDA, modelling wave height by using classification algorithms available in WEKA.
+
+			\subsubsection{Obtaining the final dataset}
+
+			By clicking on the \textbf{\textit{Matching configuration}} tab, the view shown in Fig.
 \ref{fig:matching_conf_wc} will be opened.
 In this view, the researcher can configure the parameters of the data integration process.
-For this problem, the following parameters were selected:}
+For this problem, the following parameters were selected:
 
@@
 				\item Reanalysis data: Air, pressure, u-wind and v-wind.
-				\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription}).
+				\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription} or descriptions of the measurements in the NDBC website).
 				\item Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).
@@
 \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
-Given that, in the previous view \textcolor{blue}{(Fig.
-\ref{fig:matching_conf_wc})}, \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable \textcolor{blue}{(top left of Fig.
-\ref{fig:final_dataset})}.
-After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output \textcolor{blue}{(top right of Fig.
-\ref{fig:final_dataset})}, as explained in Section {\ref{sec:FinalDatasets}}.
-Then the researcher can click on the \textbf{\textit{Update final dataset}} button to see the content shown in the bottom left corner \textcolor{blue}{(NDBC observations, NNRP variables, missing values, dates)}.
-Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textbf{\textit{Create final datasets}} button \textcolor{blue}{(bottom right of Fig.
-\ref{fig:final_dataset})}.
-For this \textcolor{blue}{example}, the following configuration was applied:
+Given that, in the previous view (Fig.
+\ref{fig:matching_conf_wc}), \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable (top left of Fig.
+\ref{fig:final_dataset}).
+After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output (top right of Fig.
+\ref{fig:final_dataset}), as explained in Section {\ref{sec:FinalDatasets}}.
+Then the researcher can click on the \textbf{\textit{Update final dataset}} button to see the content shown in the bottom left corner (NDBC observations, NNRP variables, missing values, dates).
+Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textbf{\textit{Create final datasets}} button (bottom right of Fig.
+\ref{fig:final_dataset}).
+For this example, the following configuration was applied:
 			\begin{itemize}
@@
 				\item Synchronisation: Disabled.
-				\item \textcolor{blue}{Final dataset format: ARFF.}
+				\item Final dataset format: ARFF.
 			\end{itemize}
@@
 		%
-		\textcolor{blue}{\subsubsection{Obtaining classification models with ML algorithms}}%
+		\subsubsection{Obtaining classification models with ML algorithms}
 
@@
 				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka.png}
-				\caption{\textcolor{blue}{Final dataset opened with the environment Explorer of WEKA (waves classification)}.}
+				\caption{Final dataset opened with the environment Explorer of WEKA (waves classification).}
 				\label{fig:openigFinalDatasetWeka}
@@
 However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so, in this case, 30 executions for each one will be carried out.
-The results obtained are shown in Table \ref{tab:results}.
+Table \ref{tab:results} shows the results of this experimentation.
 
@@
 
-			\textcolor{blue}{\subsection{Energy flux prediction}}
-
-			\textcolor{blue}{As mentioned above, the final dataset of this example is also used with SC and ML algorithms to predict flux of energy.
-Following sections explain the process of performing the data integration provided by SPAMDA to build the final dataset, modelling the flux of energy by using regression algorithms available in WEKA.}
-
-			\textcolor{blue}{\subsubsection{Obtaining the final dataset}}
-
-			\textcolor{blue}{The researcher can configure the parameters of the data integration by clicking on the \textbf{\textit{Matching configuration}} tab.
+			\subsection{Energy flux prediction}
+
+			As mentioned above, the final dataset of this example is also used with SC and ML algorithms to predict flux of energy.
+Following sections explain the process of performing the data integration provided by SPAMDA to build the final dataset, modelling the flux of energy by using regression algorithms available in WEKA.
+
+			\subsubsection{Obtaining the final dataset}
+
+			The researcher can configure the parameters of the data integration by clicking on the \textbf{\textit{Matching configuration}} tab.
 For this problem, as shown in Fig.
-\ref{fig:matching_conf_ef}, the following parameters were selected:}
-
-			\begin{itemize}
-				\item \textcolor{blue}{Attribute to predict: Flux of energy.}
-				\item \textcolor{blue}{Reanalysis data: Air, pressure, u-wind and v-wind.}
-				\item \textcolor{blue}{Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription}).}
-				\item \textcolor{blue}{Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).}
-				\item \textcolor{blue}{Number of final datasets: In this example, this option is disabled, because only one reanalysis node is considered.}
-				\item \textcolor{blue}{Prediction task: Regression.}
+\ref{fig:matching_conf_ef}, the following parameters were selected:
+
+			\begin{itemize}
+				\item Attribute to predict: Flux of energy.
+				\item Reanalysis data: Air, pressure, u-wind and v-wind.
+				\item Buoy attributes to be used as inputs: WDIR, WSPD, GST, DPD, APD, PRES, ATMP and WTMP (see Table \ref{tab:measurementsDescription} or descriptions of the measurements in the NDBC website).
+				\item Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).
+				\item Number of final datasets: In this example, this option is disabled, because only one reanalysis node is considered.
+				\item Prediction task: Regression.
 			\end{itemize}
@@
 				\includegraphics[width=0.85\textwidth]{figures/FigureMatching_configuration_EF.png}
-				\caption{\textcolor{blue}{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (energy flux prediction)}.}\label{fig:matching_conf_ef}
+				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (energy flux prediction).}\label{fig:matching_conf_ef}
 			\end{figure}
@@
 				\includegraphics[width=0.85\textwidth]{figures/FigureFinal_datasets_EF.png}
-				\caption{\textcolor{blue}{\textit{Final datasets} tab: content of the final dataset created after data integration.}}\label{fig:final_dataset_EF}
-			\end{figure}
-
-			\textcolor{blue}{After configuring the parameters of the matching process, the next step is to define the final dataset structure according to the selected prediction task.
+				\caption{\textit{Final datasets} tab: content of the final dataset created after data integration.}\label{fig:final_dataset_EF}
+			\end{figure}
+
+			After configuring the parameters of the matching process, the next step is to define the final dataset structure according to the selected prediction task.
 Researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Fig.
@@
 Note that, the thresholds for discretising the output variable (top left of Fig.
-\ref{fig:final_dataset_EF}) are disabled due to, in this case, energy flux prediction is a regression problem.}
-
-			\textcolor{blue}{By default, the time horizon is set to $6$ hours, that is, the energy flux prediction will be performed $6$ hours in advance (top right of Fig.
+\ref{fig:final_dataset_EF}) are disabled due to, in this case, energy flux prediction is a regression problem.
+
+			By default, the time horizon is set to $6$ hours, that is, the energy flux prediction will be performed $6$ hours in advance (top right of Fig.
 \ref{fig:final_dataset_EF}), but researchers can increase such time horizon depending on their needs.
@@
 \ref{fig:final_dataset_EF}).
-For this example, the following configuration was applied:}
-
-			\begin{itemize}
-				\item \textcolor{blue}{Prediction horizon: 6 hours.}
-				\item \textcolor{blue}{Synchronisation: Disabled.}
-				\item \textcolor{blue}{Final dataset format: ARFF.}
-			\end{itemize}
-
-			\textcolor{blue}{After that, the final dataset would be created and stored in the computer of the researcher according to the introduced configuration, ready to be used as input for SC and ML techniques to tackle the problem of energy flux prediction.
-The number of instances ($7302$) and the distribution of the final dataset (Table \ref{tab:datasetDistribution}) are the same as in the previous example (waves classification) since the data used to create the final dataset and the time horizon selected (6h) are the same.}
-
-		\textcolor{blue}{\subsubsection{Obtaining prediction models with ML algorithms}}
-
-			\textcolor{blue}{In this example, WEKA is used as SC and ML tool to obtain energy flux prediction models, as shown in Fig.
+For this example, the following configuration was applied:
+
+			\begin{itemize}
+				\item Prediction horizon: 6 hours.
+				\item Synchronisation: Disabled.
+				\item Final dataset format: ARFF.
+			\end{itemize}
+
+			After that, the final dataset would be created and stored in the computer of the researcher according to the introduced configuration, ready to be used as input for SC and ML techniques to tackle the problem of energy flux prediction.
+The number of instances ($7302$) and the distribution of the final dataset (Table \ref{tab:datasetDistribution}) are the same as in the previous example (waves classification) since the data used to create the final dataset and the time horizon selected (6h) are the same.
+
+		\subsubsection{Obtaining prediction models with ML algorithms}
+
+			In this example, WEKA is used as SC and ML tool to obtain energy flux prediction models, as shown in Fig.
 \ref{fig:openigFinalDatasetWeka_EF}.
-Nonetheless, the final dataset can be created in CSV format so that the researcher can use any other SC and ML tool.}
+Nonetheless, the final dataset can be created in CSV format so that the researcher can use any other SC and ML tool.
 
@@
 				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka_EF.png}
-				\caption{\textcolor{blue}{Final dataset opened with the environment Explorer of WEKA (energy flux prediction).}}
+				\caption{Final dataset opened with the environment Explorer of WEKA (energy flux prediction).}
 				\label{fig:openigFinalDatasetWeka_EF}
@@
 
-			\textcolor{blue}{For this problem, the same partitioning scheme used in the wave classification problem is considered (60\% train / 40\% test), that is, years from $2013$ to $2015$ for the training phase ($4380$ instances) and years $2016$ and $2017$ for the test phase ($2922$ instances).
-Again, the attributes are normalised prior to the learning phase.}
-
-			\textcolor{blue}{To perform the energy flux modelling, one execution will be run for the deterministic algorithm Linear Regression \cite{Bishop:2006:PRM:1162264}, whereas $30$ executions will be considered for the stochastic ones: Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}.
-Table \ref{tab:results_EF} shows the results obtained by each algorithm using the default values of the algorithm parameters provided by WEKA.}
+			For this problem, the same partitioning scheme used in the wave classification problem is considered (60\% train / 40\% test), that is, years from $2013$ to $2015$ for the training phase ($4380$ instances) and years $2016$ and $2017$ for the test phase ($2922$ instances).
+Again, the attributes are normalised prior to the learning phase.
+
+			To perform the energy flux modelling, one execution will be run for the deterministic algorithm Linear Regression \cite{Bishop:2006:PRM:1162264}, whereas $30$ executions will be considered for the stochastic ones: Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}.
+Table \ref{tab:results_EF} shows the experimental results obtained using the default values for the parameters of the algorithms provided by WEKA.
 
@@
 
-					\caption{\textcolor{blue}{Results (mean$\pm$SD) obtained by the algorithms}}
+					\caption{Results (mean$\pm$SD) obtained by the algorithms}
 					\label{tab:results_EF}
@@
 
-			\textcolor{blue}{As can be checked, Random Forest has achieved the best performance in terms of Root mean squared error.
+			As can be checked, Random Forest has achieved the best performance for the Root mean squared error.
 The standard deviation of the results obtained by Multilayer Perceptron indicates that this algorithm may has been slightly affected by its stochastic component.
-However, both Multilayer Perceptron and Random Forest have obtained an excellent Correlation coefficient.}
-
-			\textcolor{blue}{The objective of this case study has been to illustrate the use of datasets created with SPAMDA to address an energy flux prediction problem.
-An exhaustive comparison of state-of-the-art regression algorithms is not the purpose of this paper.
-However, note that Multilayer Perceptron and Random Forest algorithms have achieved very good results despite the fact that the energy flux prediction has been performed with a time horizon of 6h.}
+However, both Multilayer Perceptron and Random Forest have obtained an excellent Correlation coefficient.
+
+			In this case study, the use of datasets created with SPAMDA has been shown to address an energy flux prediction problem.
+An exhaustive comparison of regression algorithms is not the purpose of this work.
+However, note that Multilayer Perceptron and Random Forest algorithms have achieved very good results despite the fact that the energy flux prediction has been performed with a time horizon of 6h.
 
@@
 
-			In this section, it has been described how to use SPAMDA to create \textcolor{blue}{final datasets with the aim of classifying waves and predicting flux of energy}.
+			In this section, it has been described how to use SPAMDA to create final datasets with the aim of classifying waves and predicting flux of energy.
 However, using the same data described in Section \ref{sec:ObtainingFinalDataset} the researcher can quickly address other objectives or different studies by merely tailoring the matching configuration of the data integration process.
-For example, longer-term wave \textcolor{blue}{or energy flux} prediction can be addressed by changing the time horizon, waves modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
+For example, longer-term wave or energy flux prediction can be addressed by changing the time horizon, waves modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
 
@@
 
-		A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
-\textcolor{blue}{The aim of the tool presented in this work is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.}
-
-		\textcolor{blue}{Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).} These datasets can be easily obtained with SPAMDA by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
+		%A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
+The aim of the tool presented in this work is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.
+
+		%Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
+These datasets can be easily obtained with SPAMDA by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
 As a result, researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
@@
 
-		The case study \textcolor{blue}{described in Section \ref{sec:CaseStudy}} illustrates how SPAMDA can be used by researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
-\textcolor{blue}{The case study also covers an example of energy flux prediction, to predict the wave energy that could be exploited by WECs facilities six hours in advance, although such time horizon is customisable.
-Given that this work does not focus on models performance, a more extensive validation or comparison study of the results obtained in both examples has not been carried out.}
+		\textcolor{blue}{Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc.) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
+In this sense, a new open source tool named SPAMDA has been presented in this work, with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP.
+The aim of the tool is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.}
+
+		Such datasets can be easily obtained with SPAMDA by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
+As a result, researchers will benefit from significant support when carrying out environmental modelling related to energy, atmospheric or oceanic studies, among others.
+Moreover, given that SPAMDA simplifies all the intermediate steps involved in the creation of datasets and manages the extensive casuistry of the data integration (such as specifying the meteorological information, managing incomplete data, pre-processing tasks, the customisable matching process to merge the data and the preparation of the datasets according to the SC or ML technique to use), it avoids errors and reduces the time needed.
+In this way, researchers will be able to have more in-depth analysis, which could result in more complete conclusions about the issue under study.
+
+		The case study described in Section \ref{sec:CaseStudy} illustrates how SPAMDA can be used by researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
+The case study also covers an example of energy flux prediction, to predict the wave energy that could be exploited by WECs facilities six hours in advance, although such time horizon is customisable.
+Given that this work does not focus on models performance, a more extensive validation or comparison study of the results obtained in both examples has not been carried out.
+\textcolor{blue}{The final datasets obtained with SPAMDA can be replicated by researchers using the same meteorological data from NDBC and NNRP (publicly available) and applying the same parameters for the pre-processing tasks and the data integration process.
+After that, the models and results obtained, using such final datasets, will depend on the SC or ML tool used.}
 
@@
 
-\conflictsofinterest{The authors declare no conflict of interest.
+%\conflictsofinterest{The authors declare no conflict of interest.
 The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results.}
 
+\conflictsofinterest{The authors declare no conflict of interest.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
+%\abbreviations{The following abbreviations are used in this manuscript:\\
+\abbreviations{Abbreviations used in this manuscript:\\
 

### 2021-01-11 latex_manuscript/MDPI/2_energies_special_issue/envio-3/originality-report/originality-report-energies-1069865.tex
Version final para el tercer envio Special Issue Energies

% !TeX spellcheck = en_GB
%\documentclass[a4paper,10pt]{article}
\documentclass[a4paper,twoside,11pt]{article}

\usepackage[utf8x]{inputenc}
%\usepackage[latin9]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{gensymb}
\usepackage{palatino}
\usepackage{verbatim}
\usepackage{color}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{eurosym}
\usepackage{url}
% \usepackage[tight]{subfigure}
\usepackage{fancyhdr}
\usepackage{fancybox}
\usepackage{fancyvrb}
\usepackage{a4wide}
\usepackage[authoryear]{natbib}
\usepackage[misc,geometry]{ifsym} 
%\usepackage{wrapfig}

\usepackage{graphicx}
\usepackage[caption=false]{subfig}

% Text color
\usepackage{color}
% For colors in Tables
\usepackage{colortbl} 
% Multirow
\usepackage{multirow}
% Long Tables
\usepackage{longtable} 
% Dash hline in tables
\usepackage{arydshln}
\usepackage{tcolorbox}
\usepackage[parfill]{parskip}
\usepackage{lineno}
\usepackage{bm}
\usepackage[mathscr]{euscript}

%\usepackage{graphicx}
%\usepackage[caption=false]{subfig}

\usepackage[linktoc=all,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}
%\usepackage[linktoc=page,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}

%\setcounter{secnumdepth}{4}
%\setlength{\topmargin}{0cm}
%\setlength{\textheight}{23cm}
%\setlength{\textwidth}{16cm}
%\setlength{\oddsidemargin}{0.5cm}
%\setlength{\evensidemargin}{-0.4cm}
%\setlength{\marginparsep}{0cm}
%\setlength{\marginparwidth}{0cm}
%\setlength{\parskip}{4mm}
%\setlength{\baselineskip}{3cm}
%\setlength{\headsep}{1.3cm}
%\setlength{\unitlength}{1cm}
%%\setlength{\footskip}{1 cm}
%%\setlength{\headheight}{0cm}
%%\setlength{\headsep}{0cm}
%\setlength{\tolerance}{1000}
%\setlength{\pretolerance}{2000}
%\pagestyle{headings}

%\newcommand\myeq{\stackrel{\text{\tiny def}}{=}}

\newcommand{\myAuthors}{Antonio Manuel Gómez-Orellana$^1$, Juan Carlos Fernández$^1$, Manuel Dorado-Moreno$^1$, Pedro Antonio Gutiérrez$^1$ and César Hervás-Martínez$^1$.}

\newcommand{\myAuthorsShort}{A. M. Gómez-Orellana$^1$, J. C. Fernández$^1$, M. Dorado-Moreno$^1$, P. A. Gutiérrez$^1$ and C. Hervás-Martínez$^1$.}

\newcommand{\myEmail}{\Letter \quad  am.gomez@uco.es, jfcaballero@uco.es, manuel.dorado@uco.es, pagutierrez@uco.es, chervas@uco.es}

\newcommand{\myTitle}{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height and Energy Flux}

\newcommand{\myShortTitle}{Response to reviewers}

\newcommand{\myJournal}{\textbf{Reference: energies-1069865}}

\newcommand{\myRefPaper}{\textbf{Originality Report}}

\newcommand{\myDepartament}{$^1$Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, Cordoba, Spain.}

\definecolor{gray090}{gray}{0.90}

%%%%%%%%%%%% headings setup %%%%%%%%%%%%
\pagestyle{fancy}
% clear styles
\fancyhf{}
% odd pages -> chapter name
\fancyhead[LO]{\leftmark}
% even pages
%\fancyhead[RE]{\myJournal -- \myShortTitle}
\fancyhead[RE]{\myJournal}
% odd/even pages
\fancyhead[RO,LE]{\thepage}

\renewcommand{\sectionmark}[1]{\markboth{\textbf{\thesection. #1}}{}}
%\renewcommand{\subsectionmark}[1]{\markright{\textbf{\thesection. #1}}}

% line under headings
\renewcommand{\headrulewidth}{0.6pt}
% line over foot
%\renewcommand{\footrulewidth}{0.6pt}

% increases heading weight
\setlength{\headheight}{1.5\headheight}
\fancyfoot{}

\newcommand{\newtext}[1]{\textcolor{red}{#1}}

% New revisor comment command
\newcounter{ecomments}[section]

\newcommand{\ecomment}[1]
{
	\stepcounter{comments}
%	\addcontentsline{toc}{subsection}{Editor \arabic{section} Comment \arabic{comments}}
	\begin{tcolorbox}[colback=black!5,colframe=white!45!black,title=Comment \arabic{comments}]
		#1
	\end{tcolorbox}
}

% New revisor comment command
\newcounter{comments}[section]

\newcommand{\rcomment}[1]
{
	\stepcounter{comments}
%	\addcontentsline{toc}{subsection}{Reviewer \arabic{section} Comment \arabic{comments}}
	\vspace{0.6cm}
	\begin{tcolorbox}[colback=black!5,colframe=white!45!black,title=Comment \arabic{comments}]
		#1
	\end{tcolorbox}
}

\begin{document}

\thispagestyle{plain}

\begin{center}
	{\LARGE\myTitle} \vspace{0.5cm} \\
	{\LARGE\myJournal} \vspace{0.5cm} \\
	{\Large\myRefPaper} \vspace{0.5cm} \\
% 	\today \vspace{0.5cm} \\
	\myAuthors \vspace{0.5cm} \\
	{\myDepartament} \vspace{0.5cm} \\
	{\myEmail}
\end{center}

%\tableofcontents

\section*{Responses to the Editor}
We would like to thank the Editor for the suggestions related to the duplication rate of the manuscript. We have addressed the most the proposed changes (lines highlighted in \textcolor{red}{red color}) regarding the primary sources listed in the originality report that has been provided to us.

There are similarities and duplicities that we cannot change, for example, in the names of institutions, websites, author names, descriptions of meteorological variables, bibliographic references and even Energies template headings, among others, so we understand that this rate can be reduced but not completely removed.

\textbf{The version of the manuscript sent to the Reviewers is the same as the one sent to the Editor, but similarity changes are not highlighted in \textcolor{red}{red color}, only changes suggested by Reviewers are highlighted in \textcolor{blue}{blue color}.}

The new version of the manuscript with the changes highlighted in red and blue color is \underline{\textbf{attached to this document}}.

We hope that in this new version of the paper the duplication rate has been reduced.

\end{document}

### 2021-01-16 latex_manuscript/MDPI/2_energies_special_issue/envio final/latex/energies-1069865-for proofreading.tex
Borrador final para publicacion

--- 2021-01-11 latex_manuscript/MDPI/2_energies_special_issue/envio-3/latex/energies-1069865.tex
@@
 %=================================================================
-\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[energies,article,accept,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
+In addition,  line numbering will be removed.
 Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
@@
 % The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
 
@@
 \makeatother
-\pubvolume{xx}
+\pubvolume{1}
 \issuenum{1}
-\articlenumber{5}
-\pubyear{2020}
-\copyrightyear{2020}
-%\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
+\articlenumber{1}
+\pubyear{2021}
+\copyrightyear{2021}
+%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
+\datereceived{} 
+\dateaccepted{} 
+\datepublished{} 
+\hreflink{https://doi.org/} % Please use \linebreak if need to do line break
 %\updates{yes} % If there is an update available, un-comment this line
@@
 \usepackage{color}
-\usepackage[table]{xcolor}
+%\usepackage[table]{xcolor}
 \usepackage{epstopdf}
-\usepackage{subfig}
+\usepackage[labelformat=simple]{subfig}
+\renewcommand\thesubfigure{\normalfont(\textbf{\alph{subfigure}})}
+
 \usepackage[T1]{fontenc}
@@
 % Full title of the paper (Capitalized)
-\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height and Energy Flux}
-
+\Title{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A~Case Study for Predicting Significant Wave Height and Energy~Flux}
+%Attention AE/ME.
+The following layout issues have not been checked by the English Editing Department and must be carefully verified by the AE/Layout Department: All callout issues, bold usage of callouts, and references to callouts in the text.
+Correct callout usage in figures.
+Figure and Table layout issues.
+Footnote formatting and Glossaries have not been checked.
+En dash usage for negative values, en dash usage to indicate relationships, en dash usage to indicate bonds (especially in chemistry).
+The English Editing Department is not responsible for correct italic usage for genes, proteins and technical terminology.
+This responsibility belongs to the authors.
+The following are also not checked: spacing between numbers and units of measurement, ratios, en dashes for ranges, date and time formats, punctuation in equation lines, and less than/more than spacing (< >).
+Finally, capitalization and layout of titles/headings must be properly checked as well as ensuring 'Eq.' and 'Fig.' are properly spelled out, as these are layout issues.
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Building Suitable Datasets for Soft Computing and Machine Learning Techniques from Meteorological Data Integration: A Case Study for Predicting Significant Wave Height and Energy Flux}
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0001-8849-6036} % Add \orcidA{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0001-9773-6783} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0002-2657-776X} % Add \orcidC{} behind the author's name
-\newcommand{\orcidauthorD}{0000-0003-4564-1816} % Add \orcidD{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0002-1929-2408} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-8849-6036} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0001-9773-6783} % Add \orcidC{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0002-2657-776X} % Add \orcidD{} behind the author's name
+\newcommand{\orcidauthorE}{0000-0003-4564-1816}
 
@@
 %\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
-\Author{Antonio Manuel G\'omez-Orellana $^{1}$*, Juan Carlos Fern\'andez $^{1}$*\orcidA{}, Manuel Dorado-Moreno $^{1}$*\orcidB{}, Pedro Antonio Guti\'errez $^{1}$*\orcidC{} and C\'esar Herv\'as-Mart\'inez $^{1}$*\orcidD{}}
+\Author{Antonio Manuel G\'omez-Orellana\orcidA{}*, Juan Carlos Fern\'andez *\orcidB{}, Manuel Dorado-Moreno *\orcidC{}, Pedro~Antonio~Guti\'errez *\orcidD{} and C\'esar Herv\'as-Mart\'inez *\orcidE{}}
 
@@
 
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{G\'omez-Orellana, A.M.; Fern\'andez, J.C.; Dorado-Moreno, M.; Guti\'errez, P.A.; Herv\'as-Mart\'inez, C.}
+%If this is Chicago journal, please change author citation format as: Lastname, Firstname, Firstname Lastname, and First-name Last-name.
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
@@
 \address{%
-$^{1}$ \quad Department of Computer Science and Numerical Analysis, University of Cordoba, 14071, C\'ordoba, Spain.}
+Department of Computer Science and Numerical Analysis, University of Cordoba, 14071 C\'ordoba, Spain}
 
@@
 %\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
-\corres{Correspondence: am.gomez@uco.es (A.M.G.-O); jfcaballero@uco.es (J.C.F.); manuel.dorado@uco.es (M.D.-M.); pagutierrez@uco.es (P.A.G.); chervas@uco.es (C.H.-M.) }
+\corres{\hangafter=1 \hangindent=1.05em \hspace{-0.82em} Correspondence: am.gomez@uco.es (A.M.G.-O.); jfcaballero@uco.es (J.C.F.); manuel.dorado@uco.es (M.D.-M.); pagutierrez@uco.es (P.A.G.); chervas@uco.es (C.H.-M.) }
 
@@
 
-% Abstract (Do not insert blank lines, i.e. \\) 
+% Abstract (Do not insert blank lines, i.e., \\) 
 \abstract{Meteorological data are extensively used to perform environmental learning.
@@
 Such datasets are not always available in an appropriate format and its preparation and pre-processing implies a lot of time and effort by researchers.
-This paper presents a novel software tool with an user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two data sources: the \textit{National Data Buoy Center} and the \textit{National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project}.
+This paper presents a novel software tool with a user-friendly GUI to create datasets by means of management and data integration of meteorological observations from two data sources: the National Data Buoy Center and the National Centers for Environmental Prediction and for Atmospheric Research Reanalysis Project.
 Such datasets can be created using buoys and reanalysis data through customisable procedures, in terms of temporal resolution, predictive and objective variables, and can be used by SC and ML methodologies for prediction tasks (classification or regression).
 The objective is providing the research community with an automated and versatile system for the casuistry that entails well-formed and quality data integration, potentially leading to better prediction models.
-The software tool can be used as a supporting tool for coastal and ocean engineering applications, sustainable energy production or environmental modelling; as well as for decision making in the design and building of coastal protection structures, marine transport, ocean energy converters and well-planned running of offshore and coastal engineering activities.
+The software tool can be used as a supporting tool for coastal and ocean engineering applications, sustainable energy production, or environmental modelling; as well as for decision-making in the design and building of coastal protection structures, marine transport, ocean energy converters, and well-planned running of offshore and coastal engineering activities.
 Finally, to illustrate the applicability of the proposed tool, a case study to classify waves depending on their significant height and to predict energy flux in the Gulf of Alaska is presented.}
@@
 % Keywords
-\keyword{Environmental Prediction; Renewable Energy Resource Evaluation; Meteorological Data; Reanalysis Data; Marine Energy; Soft Computing}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
+\keyword{environmental prediction; renewable energy resource evaluation; meteorological data; reanalysis data; marine energy; soft computing}  % List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.
 
@@
 		A better understanding of the environment is of vital importance for science, contributing not only to more efficient exploitation of natural resources but also to the development of new strategies aimed at its protection.
-In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g.
-energy, climate change, agriculture, etc.).
-Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: global solar radiation estimation \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}, wind power ramp events prediction \cite{DORADOMORENO2017428}, sea surface temperature prediction \cite{He2020}, study of the responses exhibited by plankton to fluid motions \cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
+In that sense, meteorological observations provide an essential and valuable source of information which is widely used by researchers to address environmental learning, comprehension, prediction and conservation in numerous oceanic and atmospheric studies of a wide variety of areas (e.g., energy, climate change, agriculture, etc.).
+Some specific examples of the diversity of fields in which meteorological data can be used in are, among others: global solar radiation estimation \cite{SHAHRUKHANIS2019179}, directional analysis of sea storms \cite{LAFACE201545}, estimation of hybrid energy systems taking into account economic and environmental objectives\cite{Kumar2020}, wind power ramp events prediction \cite{DORADOMORENO2017428}, sea surface temperature prediction \cite{He2020}, study of the responses exhibited by plankton to fluid motions~\cite{FUCHS2016109}, trends in solar radiation \cite{SILVA20101852} or simulation of extreme near shore sea conditions \cite{GOULDBY201415}.
 All these studies require a prior data collection and its adaptation to a specific format that allows the interpretation of them.
@@
 The choice of the appropriate model, either in engineering problems or in any other problem, is also an important factor in addition to data \cite{Alizadeh2019, Alizadeh2020}.
-\textcolor{blue}{In this sense, public and research organisms are increasingly promoting the use of open and robust datasets to boost policies coherent with the environmental exploitation, protection and conservation, as well as modelling tools available to the scientific community and decision-makers \cite{Manfren2020}}.
-Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modelling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depend mainly on the the quality of the information used, so that if a researcher does not have quality data there will be no quality models.
+{In this sense, public and research organisms are increasingly promoting the use of open and robust datasets to boost policies coherent with the environmental exploitation, protection and conservation, as well as modelling tools available to the scientific community and decision-makers \cite{Manfren2020}}.
+Although Soft Computing (SC) and Machine Learning (ML) techniques have the ability to handle uncertainty in data and are extensively used for modelling purposes, the real challenge in modelling studies is due to the inadequacy of data, since the adequacy of the models depends mainly on the the quality of the information used, so that, if a researcher does not have quality data, there will be no quality models.
 
@@
 A framework for integrating information from offshore wind farms is implemented in \cite{NGUYEN2013150} in order to ease data interchange and enhance operation and maintenance practices.
-		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool`` is presented for comparing PCM - phase change materials- storage systems with conventional systems that do not involve energy storage, being beneficial for supporting decision-making on energy concepts for buildings.
-		A risk assessment tool to improve safety standards and emergency management in onshore wind farms, is presented in \cite{ASTIASOGARCIA201648}.
+		In \cite{Roberta2020}, a new software tool named ``Storage LCA Tool'' is presented for comparing PCM---phase change materials- storage systems with conventional systems that do not involve energy storage, being beneficial for supporting decision-making on energy concepts for buildings.
+		A risk assessment tool to improve safety standards and emergency management in onshore wind farms is presented in~\cite{ASTIASOGARCIA201648}.
 Raabe et al.
-\cite{RAABE2010213} developed two software tools, \textit{Model of Equilibrium of Bay Beaches} (MEPBAY) and \textit{Coastal Modelling System} (SMC), for supporting distinct operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
-\cite{MOTAHHIR20199} developed an open hardware/software test bench for solar tracker.
-
-		Marine energy prediction is currently a hot topic where meteorological data is used in.
+\cite{RAABE2010213} developed two software tools, Model of Equilibrium of Bay Beaches (MEPBAY) and Coastal Modelling System (SMC), for supporting distinct operational levels of headland-bay beach in coastal engineering projects, and Motahhir et al.
+\cite{MOTAHHIR20199} developed an open hardware/software test bench for a solar tracker.
+
+		Marine energy prediction is currently a hot topic where meteorological data are used.
 Marine Renewable Energy (MRE) is one of the most important renewable and sustainable energy sources available in our environment \cite{en12050787}, and it includes ocean thermal energy, marine tidal current energy and wave energy, among others.
-Its benefits and great potential \cite{ZEYRINGER20181281} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint society \cite{en12091657, BREDE201344, Alizadeh2020a}.
+Its benefits and great \mbox{potential~\cite{ZEYRINGER20181281}} make it one of the most relevant natural resources, playing a crucial role not only in the reduction of the emission of greenhouse gases but also in all other aspects involved in the difficult challenge of the transition to a low carbon footprint \mbox{society \cite{en12091657, BREDE201344, Alizadeh2020a}}.
 Wave energy exhibits a more stable power supply than wind energy and even solar energy.
-In recent years \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed, \textcolor{blue}{even arranged in array form \cite{Amini2020}}, to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing offshore oil and gas platforms \cite{OLIVEIRAPINTO2019556} or seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
+In recent years, \textit{Wave Energy Converters} (WECs) \cite{FALCAO2010899} have been developed and widely installed, {even arranged in array form \cite{Amini2020}}, to transform this wave energy into electricity, which can be injected into the electric network or supplied to existing offshore oil and gas \mbox{platforms \cite{OLIVEIRAPINTO2019556}} or seawater desalination plants \cite{FERNANDEZPRIETO2019546}, among others.
 WECs are mechanical devices that convert kinetic energy into electrical energy through either the vertical oscillation or the linear motion of waves.
-Nevertheless, waves are difficult to be characterised due to their stochastic nature, because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
+Nevertheless, waves are difficult to be characterised due to their stochastic nature because of the influence of a large number of environmental factors that exert on them \cite{ochi1998}.
 As a consequence of this complexity, many aspects of WEC design, deployment and operation \cite{CROWLEY2018159, Abdelkhalik2016, 6898109} need a proper prediction of waves \cite{en11010011, Kaloop2020}, in order to maximise the wave energy extraction \cite{en80910370}.
-For this purpose WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most relevant wave parameters related to this aspect: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
-
-		Currently, and as a support to traditional study procedures, SC and ML techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264} are being widely used in numerous research fields related to classification, regression and optimisation tasks, obtaining significant improvements in the performance of the results, either in engineering \cite{Alizadeh2020}, energy or environmental problems \cite{Fi-John2019,Mosavi2019,GUO201816}.
+For this purpose, WECs use wave \textit{flux of energy} ($F_e$) which can be calculated from the two most relevant wave parameters related to this aspect: \textit{significant wave height} ($H_s$) and \textit{wave energy period} ($T_e$).
+
+		\textls[-20]{Currently, and as a support to traditional study procedures, SC and ML \mbox{techniques \cite{Sang-Yong2014,Bishop:2006:PRM:1162264}}} are being widely used in numerous research fields related to classification, regression, and optimisation tasks, obtaining significant improvements in the performance of the results, either in engineering \cite{Alizadeh2020}, energy, or environmental problems \cite{Fi-John2019,Mosavi2019,GUO201816}.
 SC and ML methodologies can be used not only by experienced computer scientists but also by other researchers.
@@
 ML techniques have been already applied to tackle wave characterisation, accurately estimating $H_{s}$ and $T_{e}$ parameters \cite{DURANROSAL2017268, KUMAR2017605}, given that robustness of ML methods can tackle the previously explained difficulties in wave energy prediction.
-In \cite{Ali2020}, a reliable ML model based on multiple linear regression and covariant-weighted least square estimation for $H_s$ modelling is presented in order to predict significant wave height 30 minutes in advance.
+In \cite{Ali2020}, a reliable ML model based on multiple linear regression and covariant-weighted least square estimation for $H_s$ modelling is presented in order to predict significant wave height 30 min in advance.
 In \cite{Cornejo-Bueno2016}, an approach for feature selection problems is developed and applied for $H_s$ and $F_e$ prediction in oceanic buoys, obtaining very good results.
@@
 
-		The information to create these datasets related with MRE can be obtained from meteorological observations, but such information may be available in a inappropriate format and even contain missing values or measurements.
-Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection or data normalisation, among others.
+		The information to create these datasets related with MRE can be obtained from meteorological observations, but such information may be available in an inappropriate format and even contain missing values or measurements.
+Consequently, it is usually required to perform pre-processing tasks for improving the quality of the data, such as the replacement of missing values, outlier detection, or data normalisation, among others.
 Furthermore, if more than one source of information is used to achieve a better characterisation of the problem under study \cite{JOHANSSON2015143, FERNANDEZ201544, Adams2010}, then a data integration process, denominated as the matching process in this document, has to be carried out by researchers to manually create the datasets with the needed information.
@@
 Moreover, depending on the subject and the SC and ML technique to be applied, or even if the researcher considers other factors in order to enhance the performance obtained or have more in-depth conclusions, the datasets would have to be updated afterwards.
-In summary, many important details and different intermediate steps have to be considered when creating suitable datasets, specially when data integration is required, resulting in an extremely tedious task.
+In summary, many important details and different intermediate steps have to be considered when creating suitable datasets, especially when data integration is required, resulting in an extremely tedious task.
 
 		The main purpose of this paper is to present a new open source tool for the creation of datasets integrated by meteorological variables from two sources of information.
-Given that the tool provides an user-friendly graphical interface, no knowledge in programming languages is needed.
-It also prevents researchers from performing the mentioned tedious work and greatly simplify all the steps involved in it, avoiding possible errors in the intermediate steps, at least as a preliminary study in certain areas where some kind of environmental prediction is needed.
+Given that the tool provides a user-friendly graphical interface, no knowledge in programming languages is needed.
+It also prevents researchers from performing the mentioned tedious work and greatly simplifes all the steps involved in it, avoiding possible errors in the intermediate steps, at least as a preliminary study in certain areas where some kind of environmental prediction is needed.
 The meteorological data used by the tool come from two well-known sources of information: the \textit{National Oceanic and Atmospheric Administration} (NOAA) \textit{National Data Buoy Center} (NDBC) \cite{NOAA} and the \textit{National Centers for Environmental Prediction} (NCEP)/\textit{National Center for Atmospheric Research} (NCAR) \textit{Reanalysis Project} (NNRP or R1) \cite{Kalnay1996, Kistler2001}.
-The open source software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets).
-As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows researchers focus on the study of the meteorological aspects of the observations.
+The open source software tool presented in this work is named SPAMDA (Software for Pre-processing and Analysis of Meteorological DAta to build datasets), and it is available at \url{https://github.com/ayrna}.
+As SPAMDA performs all this data processing, it reduces the time involving these tasks and allows researchers to focus on the study of the meteorological aspects of the observations.
 The datasets obtained are ready to be used as input for SC and ML techniques in prediction tasks (classification or regression), although researchers can use them for other purposes.
@@
 The format of the generated datasets will be \textit{Attribute-Relation File Format} (ARFF) \cite{WEKA_ARFF}, which is the one used by WEKA.
-Besides, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling researchers to use others tools.
+In addition, the datasets can also be generated in \textit{Comma-Separated Values} (CSV) format, enabling researchers to use other tools.
 
 		In order to address the problem previously discussed, meteorological data integration from NDBC and NNRP and the casuistry that it entails, SPAMDA offers to researchers novelties and functionalities that will be detailed in Section \ref{sec:SPAMDA}, although some of them are briefly summarised below:
-
 		\begin{itemize}
@@
 
-			\item It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures, that imply a great deal of time and endeavour in early stages of the research.
+			\item \textls[-10]{It makes the researcher focus on environmental modelling, without having to worry about the development of scripts or mechanical tasks, avoiding laborious pre-processing procedures that imply a great deal of time and endeavour in early stages of} \mbox{the~research.}
 
@@
 			\item Estimation of the amount of energy flux that can be produced at different prediction horizons: short-term, mid-term or long-term.
-Although this work does not focus on models performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.
+Although this work does not focus on model performance, it should be taken into account that models tend to generalise worse with greater prediction horizons.
 
@@
 
-			\item Possibility of selecting one or more reanalysis nodes near the localisation under study, which could provide a better description of the problem to achieve more accurate models.
-
-			\item Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoys measurements, such as normalisation and missing data recovery.
+			\item Possibility of selecting one or more reanalysis nodes near the localisation under study, which could provide a better description of the problem to achieve more \mbox{accurate models.}
+
+			\item Although pre-processing is not the main objective of SPAMDA, the tool also provides some basic pre-processing filters on buoy measurements, such as normalisation and missing data recovery.
 
@@
 
-			\item SPAMDA is distributed as open source tool, its modular design allows the implementation of new modules for managing meteorological data from others sources, benefiting future renewable energy and environmental research.
+			\item SPAMDA is distributed as an open source tool, its modular design allows the implementation of new modules for managing meteorological data from other sources, benefiting future renewable energy and environmental research.
 
@@
 
-	\section{Meteorological data sources}\label{sec:DataSources}
+	\section{Meteorological Data Sources}\label{sec:DataSources}
 
 		The data provided by the above-mentioned sources of information of SPAMDA is described below:
-
 		\begin{itemize}
@@
 One of them corresponds to standard meteorological information of the historical data collected by each buoy, which can be downloaded as annual text files and whose format was adopted by NDBC since January 2007 \cite {NOAA_2}.
-These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC (Universal Time Coordinated) and from $23$:$50$ 31th December of the previous desired year to $22$:$50$ 31th December of the desired year.
+These files contain hourly measurements per day from $00$:$50$ to $23$:$50$ UTC (Universal Time Coordinated) and from $23$:$50$ 31 December of the previous desired year to $22$:$50$ 31 December of the desired year.
 In Table \ref{tab:measurementsDescription}, a comprehensive measurement description and the corresponding units are provided as a summary for the reader.
-A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Fig.
-\ref{fig:fragmentAnnualTexFile}.
+A fragment of one of these files, which contains the measurements collected during year $2017$ by the buoy identified as \textit{Station 46001} in NDBC, is shown in Figure \ref{fig:fragmentAnnualTexFile}.
 Each column corresponds to a meteorological variable or attribute, and each row or instance corresponds to the values of the measurements collected by the buoy for each attribute at a specific date and time.
 
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[scale=0.56]{figures/FigureFragmentAnnualTextFile.png}
+			\begin{figure}[H]
+%				\centering
+				\includegraphics[scale=0.5]{figures/FigureFragmentAnnualTextFile.png}
 				\caption{A fragment of an annual text file of the \textit{Station 46001}.}
@@
 			\end{figure}
-
-			\begin{table}[!ht]
-
-				\caption{Measurements descriptions and units of each meteorological variable or attribute collected by the buoys (a detailed description can be found in NDBC website \cite{NOAA_4})}.
-				\label{tab:measurementsDescription}
-				\footnotesize
-				\centering
-
-				\begin{tabular}{ccm{7.0cm}@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
-
-					\cline{1-4}
-
-					\textbf{Attribute}&\textbf{Units}&\textbf{Description}&\\[0.30cm]
-
-					\cline{1-4}
-
-					WDIR & degT & The direction the wind is coming from true North. \\
-
-					\cellcolor{gray090}WSPD & \cellcolor{gray090} m/s & \cellcolor{gray090} The speed of the wind. \\
-
-					GST &  m/s & Peak of gust speed. \\
-
-					\cellcolor{gray090} WVHT & \cellcolor{gray090} m & \cellcolor{gray090} Significant wave height. \\
-
-					DPD & sec & Dominant wave period (maximum wave energy). \\
-
-					\cellcolor{gray090}APD & \cellcolor{gray090} sec & \cellcolor{gray090} Average wave period of all waves. \\
-
-					MWD & degT & The direction from which the waves at the dominant period are coming. \\
-
-					\cellcolor{gray090}PRES & \cellcolor{gray090} hPa & \cellcolor{gray090} Sea level pressure. \\
-
-					ATMP & degC & Air temperature. &\\[0.10cm]
-
-					\cellcolor{gray090}WTMP & \cellcolor{gray090} degC & \cellcolor{gray090} Sea surface temperature.\\
-
-					DEWP & degC & Dewpoint temperature. \\
-
-					\cellcolor{gray090}VIS & \cellcolor{gray090} nmi & \cellcolor{gray090} Visibility of the station. \\
-
-					TIDE & ft & The water level. \\
-
-					\cline{1-4}
-
-				\end{tabular}
-
-			\end{table}
-
 			Note that the data collected by the network of buoys may be incomplete due to diverse circumstances such as the weather conditions in which the buoys have to operate, failures or malfunctioning elements of the buoys, among others.
@@
 
-			\item NNRP provides three-dimensional global reanalysis of numerous meteorological observations (e.g.
-components  \textit{Zonal} and \textit{Meridional} of the velocity of the wind, relative humidity, pressure, etc.), which is available monthly, daily and every $6$ hours at $00$ Z (Zulu time), $06$ Z, $12$ Z and $18$ Z from $1948$ on a global $2.5\degree$ x $2.5\degree$ grid.
-Weather observations are from different sources, such as ships, satellites and radar, among others.
-Reanalysis data is created assimilating such observations employing the same climate model along the whole period of reanalysis in order to decrease the impact of modelling changes on climate statistics.
-Such information has become a substantial support of the needs of the research community, even more in locations where instrumental (real time) data is not available.
-
-			The reanalysis data is available in the NNRP website \cite{NNRP}, which it is accessible through different sections.
-Such data can be fully (a global $2.5\degree$ x $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data, which provides a description of the file contents and also includes the spatial and temporal properties of the data.
+			\item NNRP provides three-dimensional global reanalysis of numerous meteorological observations (e.g., components  \textit{Zonal} and \textit{Meridional} of the velocity of the wind, relative humidity, pressure, etc.), which is available monthly, daily, and every six hours at $00$ Z (Zulu time), $06$ Z, $12$ Z, and $18$ Z from $1948$ on a global $2.5\degree$ $\times$ $2.5\degree$ grid.
+Weather observations are from different sources, such as ships, satellites, and radar, among others.
+Reanalysis data are created assimilating such observations employing the same climate model along the whole period of reanalysis in order to decrease the impact of modelling changes on climate statistics.
+Such information has become a substantial support of the needs of the research community, even more in locations where instrumental (real time) data are not available.
+
+ \begin{specialtable}[H]
+
+				\caption{Measurements descriptions and units of each meteorological variable or attribute collected by the buoys (a detailed description can be found in NDBC website \cite{NOAA_4}).}
+				\label{tab:measurementsDescription}
+%				\footnotesize
+%				\centering
+
+				\setlength{\cellWidtha}{\columnwidth/3-2\tabcolsep-1.2in}
+\setlength{\cellWidthb}{\columnwidth/3-2\tabcolsep-1.2in}
+\setlength{\cellWidthc}{\columnwidth/3-2\tabcolsep+2.4in}
+\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\centering}m{\cellWidtha}>{\PreserveBackslash\centering}m{\cellWidthb}>{\PreserveBackslash\raggedright}m{\cellWidthc}}
+\toprule
+
+					\textbf{Attribute}&\textbf{Units}&\textbf{Description}\\
+
+					\midrule
+
+					WDIR & degT & The direction the wind is coming from true North. \\ \midrule
+
+					WSPD &  m/s & The speed of the wind. \\ \midrule
+
+					GST &  m/s & Peak of gust speed. \\ \midrule
+
+					WVHT &  m &  Significant wave height. \\ \midrule
+
+					DPD & sec & Dominant wave period (maximum wave energy). \\ \midrule
+
+					APD &  sec & Average wave period of all waves. \\ \midrule
+
+					MWD & degT & The direction from which the waves at the dominant period are coming. \\ \midrule
+
+					PRES &  hPa &  Sea level pressure. \\ \midrule
+
+					ATMP & degC & Air temperature. \\ \midrule
+
+					WTMP &  degC & Sea surface temperature.\\ \midrule
+
+					DEWP & degC & Dewpoint temperature. \\ \midrule
+
+					VIS &  nmi &  Visibility of the station. \\ \midrule
+
+					TIDE & ft & The water level. \\
+
+					\bottomrule
+
+				\end{tabularx}}
+
+			    \end{specialtable}
+
+			\textls[-20]{The reanalysis data are available in the NNRP website \cite{NNRP}, which is accessible through different sections.
+Such data can be fully (a global $2.5\degree$ $\times$ $2.5\degree$ grid) or partially (only the desired reanalysis nodes or sub-grid) downloaded as \textit{Network Common Data Form} (NetCDF) files \cite{NetCDF}, a special binary format for representing scientific data, which provides a description of the file contents and also includes the spatial and temporal properties of the data.
 Each reanalysis file contains the values of a meteorological variable estimated by a mathematical model for each reanalysis node.
-For the sake of clarity, in Fig.
-\ref{fig:subGrid}, an example to approximately illustrate a sub-grid containing six nodes of reanalysis surrounding the geographic localisation of a buoy (obtained from NDBC) is shown.
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[scale=0.55]{figures/FigureSubGrid.jpg}
+For the sake of clarity, in \mbox{Figure \ref{fig:subGrid}}, an example to approximately illustrate a sub-grid containing six~nodes of reanalysis surrounding the geographic localisation of a buoy (obtained from NDBC) is shown.}
+
+			\begin{figure}[H]
+%				\centering
+				\includegraphics[scale=0.56]{figures/FigureSubGrid.jpg}
 				\caption{Sub-grid representation of six nodes of reanalysis surrounding the \textit{Station 46001}.}
@@
 				\item \textit{Final datasets}: Created by merging an intermediate or pre-processed dataset (which contain the information from NDBC) with the reanalysis data from NNRP.
-This procedure is referenced in SPAMDA as matching process and will be carried out according to the study to be performed (classification or regression).
-			\end{itemize}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.33]{figures/FigureSPAMDA.eps}
-			\caption{Brief outline of the functionality provided by SPAMDA.}
-			\label{fig:SPAMDA}
-		\end{figure}
-		SPAMDA consists of three main functional modules, whose main features, represented in Fig. \ref{fig:SPAMDA}, are the following:
+This procedure is referenced in SPAMDA as a matching process and will be carried out according to the study to be performed (classification or regression).
+			\end{itemize}
+
+		SPAMDA consists of three main functional modules, whose main features, represented in Figure \ref{fig:SPAMDA}, are the following:
 		\begin{itemize}
@@
 				\item Matching process to merge the information from NDBC and NNRP.
-				\item Creation of the final datasets accordingly to the ML technique to use (classification or regression).
+				\item Creation of the final datasets according to the ML technique to use (classification or regression).
 			\end{enumerate}
@@
 In this way, researchers can keep the reanalysis data files updated for their studies.
-Such files will be used, depending on researchers needs, in the matching process when obtaining the final datasets.
+Such files will be used, depending on researchers' needs, in the matching process when obtaining the final datasets.
 
@@
 
-		In the following subsections each integrated functional module is described in detail.
+		In the following subsections, each integrated functional module is described in detail.
+
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+\widefigure
+			\includegraphics[scale=0.38]{figures/FigureSPAMDA.eps}
+			\caption{Brief outline of the functionality provided by SPAMDA.}
+			\label{fig:SPAMDA}
+		\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+\vspace{-12pt}
 
@@
 
-				When a new buoy is included in SPAMDA the following information, which can be obtained from NDBC, is requested:
+				When a new buoy is included in SPAMDA, the following information, which can be obtained from NDBC, is requested:
 				\begin{itemize}
@@
 
-				For clarification, an example is presented in Fig. \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.
-
-				\begin{figure}[ht!]
-					\centering
+				For clarification, an example is presented in Figure \ref{fig:buoys}, where the buoy ID1 has three annual text files and the buoy ID2 has two annual text files.
+
+				\begin{figure}[H]
+%					\centering
 					\includegraphics[scale=0.4]{figures/FigureBuoys.eps}
@@
 				Once a buoy has been included as described in Section \ref{sec:Buoys}, it is possible to create datasets with one or more annual text files, which are referenced in SPAMDA as intermediate datasets.
-In this module, researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary ones.
+In this module, researchers can manage intermediate datasets of each buoy, which are the baseline for their studies, by creating new ones or deleting the unnecessary~ones.
 
 				When an intermediate dataset is created, it is associated with its corresponding buoy.
-Besides, a summary of its content is also created, providing relevant information such as the number of instances, the dates of the first and last measurements, the annual text files included, and the missing and duplicated dates.
-
-				An example where three intermediate datasets have been created is presented in Fig.
-\ref{fig:datasets}.
+In addition, a summary of its content is also created, providing relevant information such as the number of instances, the dates of the first and last measurements, the annual text files included and the missing and duplicated dates.
+
+				An example where three intermediate datasets have been created is presented in \mbox{Figure \ref{fig:datasets}}.
 The two intermediate datasets of the buoy ID1 contain meteorological data of different years, and the intermediate dataset of the buoy ID2 contains meteorological data of two years.
@@
 
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.40]{figures/FigureDatasets.eps}
+				\begin{figure}[H]
+%					\centering
+					\includegraphics[scale=0.4]{figures/FigureDatasets.eps}
 					\caption{Example of the creation of the intermediate datasets.}
@@
 
-			\subsection{Pre-process} \label{sec:Preprocess}
+			\subsection{Pre-Process} \label{sec:Preprocess}
 
@@
 
-				SPAMDA provides several filters grouped in three categories, \textit{Attribute}, \textit{Instance} and \textit{Recover missing data}, including the configuration of their parameters and a short description of them:
-
-				\begin{itemize}
-
-				 \item \textit{Attribute}: All theses filters can be applied to the attributes (variables of the buoy from NDBC) of the intermediate dataset.
+				SPAMDA provides several filters grouped in three categories, \textit{Attribute}, \textit{Instance}, and \textit{Recover missing data}, including the configuration of their parameters and a short description of them:
+				\begin{itemize}
+
+				 \item \textit{Attribute}: All of these filters can be applied to the attributes (variables of the buoy from NDBC) of the intermediate dataset.
 
@@
 						\item \textit{ReplaceMissingValues}: For each attribute, all the missing values will be replaced by the average value of the attribute.
-						\item \textit{ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with an user-supplied constant value.
-					\end{itemize}
-
-				 \item \textit{Instance}: All theses filters can be applied to the instances (hourly measurements of the buoy from NDBC) of the intermediate dataset.
+						\item \textit{ReplaceMissingWithUserConstant}: This filter replaces all the missing values of the attributes with a user-supplied constant value.
+					\end{itemize}
+
+				 \item \textit{Instance}: All these filters can be applied to the instances (hourly measurements of the buoy from NDBC) of the intermediate dataset.
 					\begin{itemize}
@@
 						\item \textit{RemoveWithValues}: This filter removes all the instances that match the attribute and the value supplied by the user.
-						\item \textit{SubsetByExpression}: It removes all the instances which do not match a user-specified expression.
-					\end{itemize}
-
-				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the intermediate dataset.
+						\item \textit{SubsetByExpression}: It removes all the instances that do not match a user-specified expression.
+					\end{itemize}
+
+				 \item \textit{Recover missing data}: All these filters can be applied to the instances of the intermediate~dataset.
 					\begin{itemize}
@@
 				SPAMDA allows researchers to undo the last filter applied or to restore the initial content of the intermediate dataset.
-Besides, the content and relevant statistical information of the intermediate and the pre-processed datasets can be visualised in this module, for example: minimum and maximum values, mean, standard deviation or even the number of instances with missing values.
-
-				Fig.
-\ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one.
+In addition, the content and relevant statistical information of the intermediate and the pre-processed datasets can be visualised in this module, for example: minimum and maximum values, mean, standard deviation, or even the number of instances with missing values.
+
+				Figure \ref{fig:preprocess} shows an example where the intermediate datasets 1 and 2 of the buoy ID1 have been pre-processed, obtaining as a result the pre-processed dataset 1 of each one.
 The intermediate dataset 1 of the buoy ID2 has been also pre-processed.
-\textit{Pre-processed dataset n} represents that researchers can create as many pre-processed datasets as they consider opportune.
-
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.40]{figures/FigurePreprocess.eps}
+\textit{Pre-processed dataset n} represents that researchers can create as many pre-processed datasets as they consider~opportune.
+
+				\begin{figure}[H]
+%					\centering
+					\includegraphics[scale=0.45]{figures/FigurePreprocess.eps}
 					\caption{Example of the creation of pre-processed datasets.}
@@
 
-			\subsection{Matching configuration}
+			\subsection{Matching Configuration}
 			\label{sec:matching_conf}
@@
 
-				Such process merges the information of both sources that match on time, but, given that the reanalysis data is available with a minimum time horizon of $6$ hours at $00$ Z, $06$ Z, $12$ Z and $18$ Z, and the measurements of the buoys are recorded at hourly intervals, from $00$:$50$ to $23$:$50$ UTC, the matching can only be carried every $6$ hours (discarding the rest of measurements from the buoy data).
-Besides, and since there is still a difference of 10 minutes, the matching with the reanalysis data will be performed with the nearest buoy measurement (before or after) within a maximum of 60 minutes of difference.
+				Such process merges the information of both sources that match on time, but, given that the reanalysis data are available with a minimum time horizon of $6$ h at $00$ Z, \mbox{$06$ Z}, \mbox{$12$ Z} and $18$ Z, and the measurements of the buoys are recorded at hourly intervals, from $00$:$50$ to $23$:$50$ UTC, the
+%Please confirm if the time format is correct 				
+matching can only be carried every six hours (discarding the rest of measurements from the buoy data).
+In addition, and since there is still a difference of 10 min, the matching with the reanalysis data will be performed with the nearest buoy measurement (before or after) within a maximum of 60 min of difference.
 Finally, the matched instances of both sources will form the final datasets.
 
-				Fig.
-\ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year.
-In this way, only the instances from both sources that are linked with arrows (highlighted in green colour) will be used in the creation of the final datasets.
+				Figure \ref{fig:matchingProcess} presents an example of matching with the measurements collected during $2017$ by \textit{Station 46001} (NDBC) and the reanalysis data (NNRP) of the variable \textit{pressure} for reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W in the same year.
+In this way, only the instances from both sources that are linked with arrows (highlighted in green) will be used in the creation of the final datasets.
 Although the reanalysis dates have been presented in a human readable format, note that reanalysis dates are stored in hours from $01$-$01$-$1800$, and they have to be transformed for comparison taking into account the time zone.
-Such transformation is automatically done by SPAMDA when matching the instances.
-
-				The reader can check in Appendix \ref{app:AppendixA} an example with a more complex case of the procedure.
-				\begin{figure}[ht!]
-					\centering
-					\includegraphics[scale=0.43]{figures/FigureMatchingProcess.png}
-					\caption{An example of matching the data from NDBC (left) and NNRP (right).}
-					\label{fig:matchingProcess}
-				\end{figure}
+Such transformation is automatically done by SPAMDA when matching the~instances.
+
+				The reader can check in Appendix \ref{app:AppendixA} for an example with a more complex case of the procedure.
 
@@
 				Prediction tasks are based on the estimation of the output attribute using the information provided by the input attributes.
-Depending on the task, the datasets must be prepared and configured differently:
-				\begin{itemize}
-					\item \textit{Classification}: The final datasets will be ready to use as input for ML classifiers, requiring a nominal output attribute, whose specific preparation is detailed in Section \ref{sec:FinalDatasets}.
+Depending on the task, the datasets must be prepared and \mbox{configured~differently}:
+				\begin{itemize}
+					\item \textit{Classification}: The final datasets will be ready to use as an input for ML classifiers, requiring a nominal output attribute, whose specific preparation is detailed in \mbox{Section \ref{sec:FinalDatasets}}.
 					\item \textit{Regression}: The final datasets will be ready to use as input for regression methods, requiring a real output attribute, whose preparation is also explained in Section \ref{sec:FinalDatasets}.
-					\item \textit{Direct matching}: In this case the inputs attributes have a direct correspondence with the output attribute, and it is not necessary to perform any additional preparation.
+					\item \textit{Direct matching}: In this case, the inputs' attributes have a direct correspondence with the output attribute, and it is not necessary to perform any additional preparation.
 Both input and target attributes are synchronised in time, in such a way that the final dataset is not intended for prediction purposes.
-For example, the final datasets may be used in lost data recovering tasks, in correlation studies, in descriptive analyses, etc.
+For example, the final datasets may be used in lost data recovering tasks, in correlation studies, in descriptive \mbox{analyses, etc}.
 				\end{itemize}
@@
 					\end{equation}
-
 where $H_s$ is measured in meters and $T_e$ in seconds.
-$F_e$ is referred to as flux of energy but it is defined as an average energy flux because $H_s$ is an average wave height (see descriptions of the measurements on the NDBC website).
-
-					\item \textit{Attribute to predict}: Instead of using $F_e$, researchers can select any of the attributes collected by the buoys as output (e.g.
-significant wave height, WVHT, wind direction, WDIR, sea level pressure, PRES, etc.).
+$F_e$ is referred to as flux of energy, but it is defined as an average energy flux because $H_s$ is an average wave height (see descriptions of the measurements on the NDBC website).
+
+					\item \textit{Attribute to predict}: Instead of using $F_e$, researchers can select any of the attributes collected by the buoys as output (e.g., significant wave height, WVHT, wind direction, WDIR, sea level pressure, PRES, etc.).
 Therefore, they can conduct different studies by selecting one attribute or other.
@@
 					\item \textit{Reanalysis data files}: In order to have a possible better description of the problem under study, more than one reanalysis variable can be considered as input.
-Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Fig.
-\ref{fig:subGrid}) for each variable of reanalysis.
+Remember that these files have to be previously downloaded from the NNRP website \cite{NNRP}, which should set the range of dates (temporal properties) and the desired sub-grid (spatial properties, see Figure \ref{fig:subGrid}) for each variable of reanalysis.
 
 					In that sense, the reanalysis data files must have the same spatial and temporal properties but related to different variables.
-SPAMDA simplifies this task by showing the reanalysis data files that are compatibles each other, and checking that the selection made by the researches meets that condition.
-
-					\item \textit{Buoys attributes}: In addition to the reanalysis variables, the final datasets will also include the selected attributes as inputs (of the intermediate or pre-processed dataset used), providing a possible better characterisation of the problem under study, although it will depend on how correlated the attributes are.
+SPAMDA simplifies this task by showing the reanalysis data files that are compatible with each other, and checking that the selection made by the research meets that condition.
+
+					\item \textit{Buoy attributes}: In addition to the reanalysis variables, the final datasets will also include the selected attributes as inputs (of the intermediate or pre-processed dataset used), providing a possible better characterisation of the problem under study, although it will depend on how correlated the attributes are.
 
@@
 As a consequence, the matching of instances between both sources of information may not be possible (missing dates).
-In that situation, researchers can consider two options: 1) discard the instances affected or 2) include them.
-In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft\textit{?}\guillemotright.
-
-					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Fig.
-\ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
+In that situation, researchers can consider two options: (1) discard the instances affected or (2) include them.
+In the latter case, the final datasets will contain the affected instances, but the measurements of the buoy will be stored as missing values in WEKA format, denoted as \guillemotleft?\guillemotright.
+
+					\item \textit{Nearest reanalysis nodes to consider}: As already shown in Figure \ref{fig:subGrid} (which represents six reanalysis nodes), the reanalysis data files may contain information of several reanalysis nodes.
 In this way, researchers can:
@@
 To do that, SPAMDA uses the \textit{Haversine} equation \cite{Haversine_2009} (or the great-circle distance) to calculate the distance from the location of the buoy to each node of reanalysis and obtain the closest ones.
-Haversine equation performs calculation from main point to destination point with a trigonometric function:
+The Haversine equation performs calculation from main point to destination point with a trigonometric function:
 								\begin{eqnarray}
@@
 
-							\item \textit{One (using weighted mean of the $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of the reanalysis data used (one per selected reanalysis data file).
-This weighted mean is obtained by SPAMDA and uses Eq.
-\ref{eq:Haversine} to calculate the distance from the geographical position of the buoy to each node of reanalysis.
+							\item \textit{One (using weighted mean of the $N$ nearest reanalysis nodes)}: Only one final dataset will be created, which will contain the attributes (the selected one as output and the selected ones as inputs) of the intermediate or pre-processed dataset used, along with a weighted mean of each variable of the reanalysis data used (one~per selected reanalysis data file).
+This weighted mean is obtained by SPAMDA and uses Equation (\ref{eq:Haversine}) to calculate the distance from the geographical position of the buoy to each node of reanalysis.
 Once the distances have been computed, they are normalised and inverted as shown in the following equation:
@@
 
-							Considering as example the two nearest reanalysis nodes represented in Fig.
-\ref{fig:subGrid} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes $57.5$ N $\times$ $147.5$ W and $55.0$ N $\times$ $147.5$ W.
-
-							\item \textit{'N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by researcher will be created.
+							\textls[-20]{Considering as an example the two nearest reanalysis nodes represented in \mbox{Figure \ref{fig:subGrid}} and the reanalysis variables air temperature and pressure, the weighted mean of each reanalysis variable will be calculated using the reanalysis nodes \mbox{$57.5$ N $\times$ $147.5$ W} and $55.0$ N $\times$ $147.5$ W.}
+
+							\item \textit{`N' (one per each reanalysis node)}: As many final datasets as the number of nearest $N$ reanalysis nodes configured by researcher will be created.
 Therefore, each final dataset will contain the value of each reanalysis variable used of the nearest corresponding reanalysis node, along with the selected attributes of the intermediate or pre-processed dataset used.
@@
 
-							In this case, and considering as example the four closest reanalysis nodes (see Fig.
-\ref{fig:subGrid}) and the reanalysis variables air temperature and pressure, four final datasets will be created, containing each one the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, $55.0$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.
-
-						\end{itemize}
-
-				\end{itemize}
-
-				Once the matching parameters have been described, for a better understanding of them,  Fig.
-\ref{fig:directMatching} presents an example of the data integration considering the data shown in Fig.
-\ref{fig:matchingProcess} and using the following configuration\footnote{Note that the date is shown just for a better understanding, but it will not be included in the final dataset.}:
-					\begin{itemize}
-						\item Attribute to predict: variable WVHT (Fig. \ref{fig:directMatchingWave}) / flux of energy (Fig. \ref{fig:directMatchingFlux}).
+							In this case, and considering as example the four closest reanalysis nodes (see \mbox{Figure  \ref{fig:subGrid}}) and the reanalysis variables air temperature and pressure, four final datasets will be created, each one containing the information of both reanalysis variables of the corresponding reanalysis node: $57.5$ N $\times$ $147.5$ W, \mbox{$55.0$ N $\times$ $147.5$ W}, $57.5$ N $\times$ $150.0$ W and $55.0$ N $\times$ $150.0$ W, along with the selected attributes of the intermediate or pre-processed dataset used.
+
+						\end{itemize}
+
+				\end{itemize}
+
+				Once the matching parameters have been described, for a better understanding of them, Figure \ref{fig:directMatching} presents an example of the data integration considering the data shown in Figure \ref{fig:matchingProcess} and using the following configuration (Note that the date is shown just for a better understanding, but it will not be included in the final dataset):
+					\begin{itemize}
+						\item Attribute to predict: variable WVHT (Figure  \ref{fig:directMatching}a/flux of energy (Figure  \ref{fig:directMatching}b).
 						\item Variable Pres as reanalysis input attribute.
@@
 
-				\begin{figure}[ht!]
-					\centering
-					\subfloat[attribute to predict: WVHT\label{fig:directMatchingWave}]{\includegraphics[width=0.4\textwidth]{figures/FigureDirectMatching.png}}\hfill
-					\subfloat[attribute to predict: flux of energy\label{fig:directMatchingFlux}]{\includegraphics[width=0.47\textwidth]{figures/FigureDirectMatching_EF.png}}\hfill
+				% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+	\begin{figure}[H]
+\widefigure
+					\includegraphics[scale=0.42]{figures/FigureMatchingProcess.png}
+					\caption{An example of matching the data from NDBC (\textbf{left}) and NNRP (\textbf{right}).}
+					\label{fig:matchingProcess}
+				\end{figure}
+\vspace{-6pt}
+
+				\begin{figure}[H]
+\widefigure
+					\subfloat[attribute to predict: WVHT\label{fig:directMatchingWave}]{\includegraphics[width=0.35\textwidth]{figures/FigureDirectMatching.png}}\hspace{24pt}
+					\subfloat[attribute to predict: flux of energy\label{fig:directMatchingFlux}]{\includegraphics[width=0.42\textwidth]{figures/FigureDirectMatching_EF.png}}
 					\caption{Example of data integration for \textit{Direct matching}.} \label{fig:directMatching}
 				\end{figure}
-
-			\subsection{Final datasets} \label{sec:FinalDatasets}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			\vspace{-12pt}	
+			\subsection{Final Datasets} \label{sec:FinalDatasets}
 
@@
 
-				SPAMDA allows researchers to make such preparation by means of the following options:
+				SPAMDA allows researchers to make such preparation by means of the\linebreak following~options:
 
@@
 						\item \textit{Prediction horizon} (Classification and Regression): This option indicates the time gap for moving backward the attribute to predict (output attribute).
-In this way, the input attributes (variables of the buoy and reanalysis data) will be used to predict the output attribute in a specific future time (e.g.
-+6h, +12h, +18h, +1 day, etc.).
-
-						The minimum interval for increasing and decreasing the prediction horizon is $6$h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out.
+In this way, the input attributes (variables of the buoy and reanalysis data) will be used to predict the output attribute in a specific future time (e.g., +6 h, +12 h, +18 h, +1 day, etc.).
+
+						The minimum interval for increasing and decreasing the prediction horizon is $6$ h (due to reanalysis data temporal resolution) \cite{DORADOMORENO2017428}, the same interval used when the matching process is carried out.
 Therefore, for each increment of the prediction horizon, an instance of the dataset is lost (as this future information is not available).
-As the minimum prediction horizon is $6$h, at least one instance will be lost.
+As the minimum prediction horizon is $6$ h, at least one instance will be lost.
 The relation between the inputs and the output (attribute to predict) is defined as follows:
@@
 						where $t$ is the time instant to study, $\Delta t$ is the prediction horizon, $o$ is the attribute to be predicted, $\mathbf{b}_t$ represents the vector that contains the selected NDBC variables, and, finally, $\mathbf{r}_t$ represents the vector that contains the selected reanalysis variables.
-In this way and considering the matched information shown in Fig.
-\ref{fig:directMatchingWave}, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD and the vector $\mathbf{r}$ contains Pres.
+In this way and considering the matched information shown in Figure \ref{fig:directMatching}a, WVHT is $o$, the vector $\mathbf{b}$ contains the variable WSPD, and the vector $\mathbf{r}$ contains Pres.
 
@@
 
- 						For the sake of clarity, considering the matched information shown in Fig.
-\ref{fig:directMatchingWave}, an example of building a dataset for a \textit{Regression} task is shown in Fig.
-\ref{fig:regressionNoSync}.
+ 						For the sake of clarity, considering the matched information shown in Figure \ref{fig:directMatching}a, an example of building a dataset for a \textit{Regression} task is shown in Figure  \ref{fig:regression}a.
 As mentioned earlier, this prediction task requires a real output variable (in this case, WVHT, the last one).
-The options considered for the preparation of each final dataset are the following:
- 							\begin{itemize}
- 								\item Do not synchronise the reanalysis data (see Eq. \ref{eq:noSynchronisingRD} for the relation between the inputs and the output).
- 								\item A prediction horizon of $6$h.
- 							\end{itemize}
-
- 						Note that, due to prediction horizon is $6$h, the values of WVHT attribute are moved backward one instance (up).
-As a consequence, the last instance ($2017$/$12$/$31$ $18$:$00$) is lost and is not included in the final dataset.
-Besides, and because the reanalysis data has not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
-\ref{eq:noSynchronisingRD}).
-
- 						Moreover, considering again the matched information shown in Fig.
-\ref{fig:directMatchingWave}, an example of the creation of the same dataset but applying synchronisation (see Eq.
-\ref{eq:synchronisingRD}) is shown in Fig.
-\ref{fig:regressionSync}.
-
-							\begin{figure}[ht!]
-								\centering
-								\subfloat[without synchronisation\label{fig:regressionNoSync}]{\includegraphics[width=0.39\textwidth]{figures/FigureRegressionNoSync.png}}\hfill
-								\subfloat[with synchronisation\label{fig:regressionSync}]{\includegraphics[width=0.39\textwidth]{figures/FigureRegressionSync.png}}\hfill
-								\caption{Example of the creation of a \textit{Regression} dataset with a prediction horizon of $6$h.} \label{fig:regression}
-							\end{figure}
-
-						Again, and due to the prediction horizon selected ($6$h), the values of the WVHT attribute are moved backward one instance (up) and the last instance ($2017$/$12$/$31$ $18$:$00$) is not included in the final dataset.
-But now, the values of the Pres variable are also moved backward one instance (due to the synchronisation).
-Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Eq.
-\ref{eq:synchronisingRD}).
+The options considered for the preparation of each final dataset are \mbox{the following}:
+ 							\begin{itemize}
+ 								\item Do not synchronise the reanalysis data (see Equation (\ref{eq:noSynchronisingRD}) for the relation between the inputs and the output).
+ 								\item A prediction horizon of $6$ h.
+ 							\end{itemize}
+
+ 						Note that, due to prediction horizon is $6$ h, the values of WVHT attribute are moved backward one instance (up).
+As a consequence, the last instance (31 December 2017 $18$:$00$) is lost and is not included in the final dataset.
+In addition, and because the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation (\ref{eq:noSynchronisingRD})).
+
+ 						Moreover, considering again the matched information shown in Figure \ref{fig:directMatching}a, an example of the creation of the same dataset but applying synchronisation (see Equation (\ref{eq:synchronisingRD})) is shown in Figure  \ref{fig:regression}b.
+
+						Again, and due to the prediction horizon selected ($6$ h), the values of the WVHT attribute are moved backward one instance (up) and the last instance (31 December 2017 $18$:$00$) is not included in the final dataset.
+However, now, the values of the Pres variable are also moved backward one instance (due to the synchronisation).
+Therefore, in this case, Pres is at the same time instant as the attribute to predict (${t+\Delta t}$ in Equation (\ref{eq:synchronisingRD})).
 
@@
 
-						Considering again the matched information shown in Fig.
-\ref{fig:directMatchingWave}, an example of the creation of a \textit{Classification} dataset is shown in Fig.
-\ref{fig:prediction}.
+						Considering again the matched information shown in Figure \ref{fig:directMatching}a, an example of the creation of a \textit{Classification} dataset is shown in Figure \ref{fig:prediction}.
 The options considered for the preparation of the final dataset are the following:
+%						\newpage
 							\begin{itemize}
 								\item Do not synchronise the reanalysis data.
-								\item A prediction horizon of $6$h.
+								\item A prediction horizon of $6$ h.
 								\item The thresholds shown in Table \ref{tab:thresholds}.
 							\end{itemize}
-
-						\begin{table}[!h]
-
-							\caption{Thresholds for the classification example represented in Fig. \ref{fig:prediction}}
-							\label{tab:thresholds}
-							\footnotesize
-							\centering
-
-							\begin{tabular}{cm{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
-
-								\cline{1-5}
-
-								\textbf{Class}&\textbf{Description}&\textbf{Lower end [}&\textbf{Upper end )}&\\[0.20cm]
-
-								\cline{1-5}
-
-								Low & Low wave height & $0.36$ & $1.5$&\\[0.15cm]
-
-								\cellcolor{gray090}Average & \cellcolor{gray090}Average wave height & \cellcolor{gray090}$1.5$ & \cellcolor{gray090}$2.5$&\\[0.15cm]
-
-								Big & Big wave height & $2.5$ & $4.0$&\\[0.15cm]
-
-								\cellcolor{gray090}Huge & \cellcolor{gray090}Huge wave height & \cellcolor{gray090}$4.0$ & \cellcolor{gray090}$9.9$&\\[0.15cm]
-
-								\cline{1-5}
-
-							\end{tabular}
-
-						\end{table}
-
-						\begin{figure}[ht!]
-							\centering
-							\includegraphics[scale=0.6]{figures/FigureClassification.png}
-							\caption{An example of the creation a \textit{Classification} dataset, with a prediction horizon of $6$h and without synchronisation.}
-							\label{fig:prediction}
-						\end{figure}
-
-						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable, because its values have been discretised according to the thresholds (usually defined by an expert).
-Besides, and due to the $6$h prediction horizon, the last instance is lost ($2017$/$12$/$31$ $18$:$00$) and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up).
-As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Eq.
-\ref{eq:noSynchronisingRD}).
+%							\vspace{-3pt}
+
+						Note that the attribute to be predicted has been renamed to \textit{Class\_WVHT} to show that it is now a nominal variable because its values have been discretised according to the thresholds (usually defined by an expert).
+In addition, and due to the $6$ h prediction horizon, the last instance is lost (31 December 2017 $18$:$00$), and the values of the attribute \textit{Class\_WVHT} are moved backward one instance (up).
+As the reanalysis data have not been synchronised, the values of the Pres and WSPD variables are at the same time instant ($t$ in Equation (\ref{eq:noSynchronisingRD})).
 
@@
 
-				A text file that summarises the configuration used in matching process and in the preparation of the matched data is also generated.
+				A text file that summarises the configuration used in matching process and in the preparation of the matched data are also generated.
 It can be saved and loaded, enabling researchers to resume their studies at any other time.
 
-			\subsection{Manage reanalysis data}
+    \begin{specialtable}[H]
+
+							\caption{Thresholds for the classification example represented in Figure \ref{fig:prediction}.}
+							\label{tab:thresholds}
+%							\footnotesize
+%							\centering
+\setlength{\cellWidtha}{\columnwidth/4-2\tabcolsep+0.0in}
+\setlength{\cellWidthb}{\columnwidth/4-2\tabcolsep+0.0in}
+\setlength{\cellWidthc}{\columnwidth/4-2\tabcolsep+0.0in}
+\setlength{\cellWidthd}{\columnwidth/4-2\tabcolsep+0.0in}
+\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\centering}m{\cellWidtha}>{\PreserveBackslash\centering}m{\cellWidthb}>{\PreserveBackslash\centering}m{\cellWidthc}>{\PreserveBackslash\centering}m{\cellWidthd}}
+\toprule
+
+%							\begin{tabular}{cm{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
+
+								\textbf{Class}&\textbf{Description}&\textbf{Lower End [}&\textbf{Upper End )}\\
+
+									\midrule
+
+								Low & Low wave height & $0.36$ & $1.5$\\ \midrule
+
+								Average & Average wave height &$1.5$ & $2.5$\\ \midrule
+
+								Big & Big wave height & $2.5$ & $4.0$\\ \midrule
+
+								Huge & Huge wave height & $4.0$ &$9.9$\\
+
+								\bottomrule
+
+							\end{tabularx}}
+
+						\end{specialtable}
+
+	\vspace{-10pt}
+
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+								\widefigure
+								\subfloat[without synchronisation\label{fig:regressionNoSync}]{\includegraphics[width=0.36\textwidth]{figures/FigureRegressionNoSync.png}}\hspace{24pt}
+								\subfloat[with synchronisation\label{fig:regressionSync}]{\includegraphics[width=0.36\textwidth]{figures/FigureRegressionSync.png}}\hfill
+								\caption{Example of the creation of a \textit{Regression} dataset with a prediction horizon of $6$ h.} \label{fig:regression}
+							\end{figure}	
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+	\vspace{-6pt}
+	\begin{figure}[H]
+%							\centering
+							\includegraphics[scale=0.6]{figures/FigureClassification.png}
+							\caption{An example of the creation a \textit{Classification} dataset, with a prediction horizon of $6$ h and without synchronisation.}
+							\label{fig:prediction}
+						\end{figure}
+
+			\subsection{Manage Reanalysis Data}
 
@@
 
-				In this module (see Fig.
-\ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones.
-Besides, useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final time, latitude and longitude.
-All these fields summarise the temporal and spatial properties of the data.
-Thus, researcher can quickly and easily identify each reanalysis file entered in SPAMDA.
-
-				An example where two reanalysis data files have been entered in SPAMDA is shown in Fig. \ref{fig:manageReanalisys}.
-
-				\begin{figure}[ht!]
-					\centering
+				In this module (see Figure \ref{fig:SPAMDA}), SPAMDA includes features for entering new files and deleting the unnecessary ones.
+In addition, useful information about the content of each reanalysis file can be consulted such as name of the file and the reanalysis variable, number of instances and reanalysis nodes, initial and final time, latitude and longitude.
+All of these fields summarise the temporal and spatial properties of the data.
+Thus, researchers can quickly and easily identify each reanalysis file entered in SPAMDA.
+
+				An example where two reanalysis data files have been entered in SPAMDA is shown in Figure \ref{fig:manageReanalisys}.
+
+				\begin{figure}[H]
+%					\centering
 					\includegraphics[scale=0.45]{figures/FigureManageReanalisys.eps}
@@
 
-	\section{A case study applied to Gulf of Alaska}\label{sec:CaseStudy}
+	\section{A Case Study Applied to Gulf of Alaska}\label{sec:CaseStudy}
 
@@
 		On the one hand, wave classification is addressed as a multi-class approach, given that a continuous attribute can be discretised, using different thresholds, in distinct classes.
-Such waves modelling can be applied with different purposes, such as missing buoy data reconstruction, extreme significant wave heights detection or decision-making and risk assessment about operational works in the sea.
+Such wave modelling can be applied with different purposes, such as missing buoy data reconstruction, extreme significant wave heights detection or decision-making and risk assessment about operational works in the sea.
 
 		On the other hand, the prediction of the energy flux is addressed as a regression problem.
-Energy flux prediction is related to marine energy and it is useful to characterise the wave energy production from WECs facilities, which could be injected into the electric network or supplied to existing marine platforms.
-
-		\subsection{Gathering the information and introducing it in SPAMDA}\label{sec:ObtainingFinalDataset}
+Energy flux prediction is related to marine energy, and it is useful to characterise the wave energy production from WEC facilities, which could be injected into the electric network or supplied to existing marine platforms.
+
+		\subsection{Gathering the Information and Introducing it in SPAMDA}\label{sec:ObtainingFinalDataset}
 
@@
 			\item The measurements obtained from 2013 to 2017 by the buoy with ID 46001, placed in the Gulf of Alaska, which are provided by NDBC as annual text files.
-This data is publicly available at the NDBC website.
-			\item Complementary information collected from reanalysis data containing air temperature (air), pressure (pres) and two components of wind speed measurements, South-North (vwind) and West-East (uwind).
+This data are publicly available at the NDBC website.
+			\item Complementary information collected from reanalysis data containing air temperature (air), pressure (pres) and two components of wind speed measurements, South--North (vwind) and West--East (uwind).
 This information can be downloaded from the NNRP website in NetCDF format for the four closest nodes of reanalysis surrounding the position of the buoy.
-Concretely, the closest reanalysis nodes downloaded are $57.5$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150$ W, $55$ N $\times$ $147.5$ W and $55$ N $\times$ $150$ W.
+Concretely, the closest reanalysis nodes downloaded are $57.5$ N $\times$ $147.5$ W, $57.5$ N $\times$ $150$ W, $55$ N $\times$ $147.5$ W, and $55$ N $\times$ $150$ W.
 However, as will be seen later, only the information from the nearest node will be used in the data integration process.
@@
 			After gathering the information described above, researchers can open SPAMDA.
-In Fig.
-\ref{fig:main_view}, the main view is shown.
+In Figure \ref{fig:main_view}, the main view is shown.
 In order to input the reanalysis data which will be used in further steps for creating the final dataset, researchers has to select the option \textbf{\textit{Manage reanalysis data}}.
 
-			\begin{figure}[ht]
-				\centering
-				\includegraphics[width=0.5\textwidth]{figures/FigureMain_view.png}
+			\begin{figure}[H]
+%				\centering
+				\includegraphics[width=0.55\textwidth]{figures/FigureMain_view.png}
 				\caption{SPAMDA main view.}\label{fig:main_view}
@@
 
-			Then, the view of Fig.
-\ref{fig:reanalysis} is shown.
-Here, using the buttons located at the bottom, it is possible to add, delete or consult any data from the different reanalysis files.
+			Then, the view of Figure  \ref{fig:reanalysis} is shown.
+Here, using the buttons located at the bottom, it is possible to add, delete, or consult any data from the different reanalysis files.
 Once the information has been introduced in the application, this view can be closed and the user can go back to the main view to continue entering the information related to the buoy under study.
-
-			\begin{figure}[ht!]
-				\centering
+\newpage			
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.89\textwidth]{figures/FigureManage_reanalysis_data.png}
@@
 			\end{figure}
-
-			\begin{figure}[ht!]
-				\centering
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			After that, the researcher has to select \textbf{\textit{Manage buoys data}} to open the view shown in Figure \ref{fig:manage_buoys}, where several tabs are available.
+In the \textbf{\textit{Buoys}} tab, the researcher can consult, modify, add, or delete different data related to the buoy.
+
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.89\textwidth]{figures/FigureManage_buoys.png}
@@
 			\end{figure}
-
-			After that, the researcher has to select \textbf{\textit{Manage buoys data}} to open the view shown in Fig.
-\ref{fig:manage_buoys}, where several tabs are available.
-In \textbf{\textit{Buoys}} tab, the researcher can consult, modify, add or delete different data related to the buoy.
-
-			In order to enter such data, click on the \textbf{\textit{New}} button, and then the view shown in Fig. \ref{fig:new_buoy} pops up.
-
-			\begin{figure}[ht!]
-				\centering
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			In order to enter such data, click on the \textbf{\textit{New}} button, and then the view shown in Figure \ref{fig:new_buoy} pops up.
+
+			\begin{figure}[H]
+%				\centering
 				\includegraphics[width=0.65\textwidth]{figures/FigureNew_buoy.png}
@@
 
-			Here the information about the buoy has to be included: the \textit{Station ID}, its description, geographical localisation and the corresponding annual text files.
+			Here, the information about the buoy has to be included: the \textit{Station ID}, its description, geographical localisation, and the corresponding annual text files.
 In this case, the files containing the data from year $2013$ to $2017$ are inserted by clicking on the \textbf{\textit{Add file}} button.
-Once the data has been introduced, it is necessary to click on the \textbf{\textit{Save}} button to insert the buoy in SPAMDA database.
+Once the data have been introduced, it is necessary to click on the \textbf{\textit{Save}} button to insert the buoy in SPAMDA database.
 After that, the view can be closed.
 
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.83\textwidth]{figures/FigureDatasets.png}
-				\caption{\textit{Datasets} tab: intermediate datasets of the buoy ID 46001.}\label{fig:show_datasets}
-			\end{figure}
-
-			To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textbf{\textit{Datasets}} tab (see Fig.
-\ref{fig:manage_buoys}) to switch to the corresponding view (see Fig.
-\ref{fig:show_datasets}).
+			To create the intermediate dataset, the researcher has to double-click on the buoy under study or click on the \textbf{\textit{Datasets}} tab (see Figure \ref{fig:manage_buoys}) to switch to the corresponding view (see Figure \ref{fig:show_datasets}).
 In this view, the researcher can delete or consult a summary of each intermediate or pre-processed dataset by selecting it from the corresponding list.
 It can also create new ones.
-To proceed with the creation of the intermediate dataset, the user clicks on the \textbf{\textit{New}} button, and the view shown in Fig.
-\ref{fig:intermediate} appears.
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.67\textwidth]{figures/FigureNew_intermediate_dataset.png}
-				\caption{\textit{New intermediate dataset} view: creating the intermediate dataset with $5$ annual text files.}\label{fig:intermediate}
-			\end{figure}
-
-			Here the researcher can select the annual text files to be included in the intermediate dataset, by clicking on the \textbf{\textit{->}} and \textbf{\textit{<-}} buttons.
+To proceed with the creation of the intermediate dataset, the user clicks on the \textbf{\textit{New}} button, and the view shown in Figure \ref{fig:intermediate} appears.
+
+			Here, the researcher can select the annual text files to be included in the intermediate dataset, by clicking on the \textbf{\textit{->}} and \textbf{\textit{<-}} buttons.
 In this case, all the files introduced before, which correspond to the buoy under study, are selected.
-When the file selection is finished, \textbf{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and then, with the \textbf{\textit{Save}} button, the creation process starts, showing the status of the process during it.
-After that, in order to prepare the intermediate dataset, the dataset is selected (see Fig.
-\ref{fig:show_datasets}), and then the button \textbf{\textit{Open}} is clicked to jump to the tab \textbf{\textit{Pre-process}} (shown in Fig.
-\ref{fig:preprocess_data}).
-
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.84\textwidth]{figures/FigurePreprocess.png}
-				\caption{\textit{Pre-process} tab: pre-processing the created intermediate dataset.}\label{fig:preprocess_data}
-			\end{figure}
+When the file selection is finished, \textbf{\textit{Create}} button has to be clicked in order to introduce the description and the file name of the current intermediate dataset, and, then, with the \textbf{\textit{Save}} button, the creation process starts, showing the status of the process during it.
+After that, in order to prepare the intermediate dataset, the dataset is selected (see Figure \ref{fig:show_datasets}), and then the button \textbf{\textit{Open}} is clicked to jump to the tab \textbf{\textit{Pre-process}} (shown in Figure \ref{fig:preprocess_data}).
 
 			In \textbf{\textit{Pre-process}} tab, relevant statistical information about the selected dataset is shown, and also the content of the dataset can be consulted, providing the researcher the capacity to evaluate the pre-processing being performed.
-Here the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter.
-As mentioned earlier, this case study is focused on classifying waves considering their height, so any missing data from wave height ($376$ values) and the remaining attributes are recovered, using the filter \textit{Replace missing values with symmetric $3$ hours mean}.
+Here, the researcher can apply (and configure) the necessary filters (explained in Section \ref{sec:Preprocess}) to the selected dataset, and, in the bottom part, the main statistics of the dataset are displayed, which can be used to observe the changes produced when applying a filter.
+As mentioned earlier, this case study is focused on classifying waves considering their height, so any missing data from wave height ($376$~values) and the remaining attributes are recovered, using the filter \textit{Replace missing values with symmetric $3$ h mean}.
 Furthermore, the attributes MWD, DEWP, VIS and TIDE are removed from the dataset by applying the filter \textit{RemoveByName}, since the first two had more than $92$\% of missing data and the last two $100$\%.
@@
 			At this point, the researcher has registered the buoy in SPAMDA, then entered its raw data and selected the required data for the problem (intermediate dataset).
-Finally, the data has been pre-processed in order to be ready for its future use in ML algorithms.
+Finally, the data have been pre-processed in order to be ready for its future use in ML algorithms.
 Then, a data integration process can be carried out to merge the processed data from NDBC with the reanalysis data (also included previously) from NNRP.
 
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.84\textwidth]{figures/FigureMatching_configuration.png}
-				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (waves classification).}\label{fig:matching_conf_wc}
-			\end{figure}
-
-			The next step is to customise (or load) the parameters of the matching process according to the problem being studied and to select the prediction task (described in Section \ref{sec:matching_conf}) that the final dataset will be used for, in this case, waves classification or energy flux prediction.
-
-			\subsection{Waves classification}
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
+				\includegraphics[width=0.85\textwidth]{figures/FigureDatasets.png}
+				\caption{\textit{Datasets} tab: intermediate datasets of the buoy ID 46001.}\label{fig:show_datasets}
+			\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn	
+\vspace{-6pt}
+
+			\begin{figure}[H]
+%				\centering
+				\includegraphics[width=0.67\textwidth]{figures/FigureNew_intermediate_dataset.png}
+				\caption{\textit{New intermediate dataset} view: creating the intermediate dataset with five annual text files.}\label{fig:intermediate}
+			\end{figure}
+\vspace{-6pt}
+
+		\newpage	
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
+				\includegraphics[width=0.90\textwidth]{figures/FigurePreprocess.png}
+				\caption{\textit{Pre-process} tab: pre-processing the created intermediate dataset.}\label{fig:preprocess_data}
+			\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			The next step is to customise (or load) the parameters of the matching process according to the problem being studied and to select the prediction task (described in \mbox{Section \ref{sec:matching_conf}}) that the final dataset will be used for, in this case, waves classification or energy flux prediction.
+
+			\subsection{Waves Classification}
 
 			As mentioned above, the objective of the final dataset is to be used with SC and ML algorithms to classify waves depending on their significant height.
-Following sections describe the procedure of performing the data integration provided by SPAMDA, modelling wave height by using classification algorithms available in WEKA.
-
-			\subsubsection{Obtaining the final dataset}
-
-			By clicking on the \textbf{\textit{Matching configuration}} tab, the view shown in Fig.
-\ref{fig:matching_conf_wc} will be opened.
+The following sections describe the procedure of performing the data integration provided by SPAMDA, modelling wave height by using classification algorithms available in WEKA.
+
+			\subsubsection{Obtaining the Final Dataset}
+
+			By clicking on the \textbf{\textit{Matching configuration}} tab, the view shown in Figure \ref{fig:matching_conf_wc} will be opened.
 In this view, the researcher can configure the parameters of the data integration process.
 For this problem, the following parameters were selected:
-
 			\begin{itemize}
@@
 				\item Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).
-				\item Number of final datasets: In this example that option is disabled, because only one reanalysis node is considered.
+				\item Number of final datasets: In this example, that option is disabled because only one reanalysis node is considered.
 				\item Prediction task: Classification.
@@
 
-			\begin{figure}[ht!]
-				\centering
-				\includegraphics[width=0.90\textwidth]{figures/FigureFinal_datasets.png}
-				\caption{\textit{Final datasets} tab: content of the final dataset created after data integration and discretisation of the output variable in $4$ classes.}\label{fig:final_dataset}
-			\end{figure}
-
-			After configuring the matching process, the researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Fig.
-\ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
-Given that, in the previous view (Fig.
-\ref{fig:matching_conf_wc}), \textit{Classification} was selected, the researcher can now add, modify or delete the thresholds (usually defined by an expert) for discretising the output variable (top left of Fig.
-\ref{fig:final_dataset}).
-After this, the next step is to set the time horizon desired (6 hours by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output (top right of Fig.
-\ref{fig:final_dataset}), as explained in Section {\ref{sec:FinalDatasets}}.
-Then the researcher can click on the \textbf{\textit{Update final dataset}} button to see the content shown in the bottom left corner (NDBC observations, NNRP variables, missing values, dates).
-Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textbf{\textit{Create final datasets}} button (bottom right of Fig.
-\ref{fig:final_dataset}).
+			After configuring the matching process, the researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset} and proceed to define the final dataset structure according to the selected prediction task.
+Given that, in the previous view (Figure \ref{fig:matching_conf_wc}), \textit{Classification} was selected, the researcher can now add, modify, or delete the thresholds (usually defined by an expert) for discretising the output variable (top left of Figure \ref{fig:final_dataset}).
+After this, the next step is to set the time horizon desired (6 h by default) and also to activate (if desired) the synchronisation (in time) of reanalysis variables with the output (top right of Figure \ref{fig:final_dataset}), as explained in Section {\ref{sec:FinalDatasets}}.
+Then, the researcher can click on the \textbf{\textit{Update final dataset}} button to see the content shown in the bottom left corner (NDBC observations, NNRP variables, missing values, dates).
+Finally, after checking that everything is correct, the last step would be to select the name and path of the dataset file, and its output format (CSV or ARFF) and click on the \textbf{\textit{Create final datasets}} button (bottom right of Figure  \ref{fig:final_dataset}).
 For this example, the following configuration was applied:
@@
 				\item Thresholds: see Table \ref{tab:thresholds}.
-				\item Prediction horizon: 6 hours.
+				\item Prediction horizon: 6 h.
 				\item Synchronisation: Disabled.
@@
 			At this point, the final dataset would be created according to the tailored configuration and stored in the computer of the researcher, which already can apply the ML techniques to address the problem of wave classification.
-Concretely, the final dataset consists of $7302$ instances and whose distribution is represented in Table \ref{tab:datasetDistribution}.
+Concretely, the final dataset consists of \mbox{$7302$ instances} and whose distribution is represented in Table \ref{tab:datasetDistribution}.
+
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
+				\includegraphics[width=0.90\textwidth]{figures/FigureMatching_configuration.png}
+				\caption{\textit{Matching configuration} tab: parameters for the data integration of the intermediate dataset and the reanalysis files (waves classification).}\label{fig:matching_conf_wc}
+			\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+\vspace{-6pt}
+
 			%
-				\begin{table}[!h]%
-					\caption{Distribution of instances of the final dataset}
+		    \begin{specialtable}[H]
+					\caption{Distribution of instances of the final dataset.}
 					\label{tab:datasetDistribution}
-					\centering
-					\begin{tabular}{cc}
-						\cline{1-2}
-						\textbf{Year}&\textbf{Number of instances}\\
-						\cline{1-2}
+					\setlength{\cellWidtha}{\columnwidth/2-2\tabcolsep+0.0in}
+\setlength{\cellWidthb}{\columnwidth/2-2\tabcolsep+0.0in}
+\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\centering}m{\cellWidtha}>{\PreserveBackslash\centering}m{\cellWidthb}}
+\toprule
+						\textbf{Year}&\textbf{Number of Instances}\\
+						\midrule
 						$2013$&$1460$\\
@@
 						$2017$&$1458$\\
-						\cline{1-2}
+						\midrule
 						&$7302$\\
-						\cline{1-2}
-					\end{tabular}
-				\end{table}
+						\bottomrule
+					\end{tabularx}}
+				\end{specialtable}
+			\vspace{-6pt}
+
+				\newpage	
+				% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\centering
+				\includegraphics[width=0.90\textwidth]{figures/FigureFinal_datasets.png}
+				\caption{\textit{Final datasets} tab: content of the final dataset created after data integration and discretisation of the output variable in four~classes.}\label{fig:final_dataset}
+			\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+\vspace{-12pt}
+
 		%
-		\subsubsection{Obtaining classification models with ML algorithms}
+		\subsubsection{Obtaining Classification Models with ML Algorithms}
 
 			Now, the process to obtain wave classification models is described using the final dataset previously created with SPAMDA.
-The modelling will be performed using WEKA as SC and ML tool, which can be opened through SPAMDA, as shown in Fig.
-\ref{fig:openigFinalDatasetWeka}.
-Nevertheless, as mentioned above, the researcher can create the final dataset in CSV format in order to use any other ML tools, such as KEEL, Python or R, among others.
-
-			\begin{figure}[ht!]
-				\centering
+The modelling will be performed using WEKA as SC and ML tool, which can be opened through SPAMDA, as shown in Figure \ref{fig:openigFinalDatasetWeka}.
+Nevertheless, as mentioned above, the researcher can create the final dataset in CSV format in order to use any other ML tools, such as KEEL, Python, or R, among others.
+
+			Since the final dataset is a time series of meteorological data (collected from $2013$ to $2017$), a hold-out scheme (60\% train/40\% test) will be used.
+In this way, years from $2013$ to $2015$ will be used for the training phase ($4380$ instances), whereas $2016$ and $2017$ years will be used for the test phase ($2922$ instances).
+Previous to the learning phase, the attributes are normalised to avoid some attributes dominating others because of a larger scale.
+
+			The classification algorithms that will be considered for wave modelling are Logistic Regression \cite{hosmer2013applied}, C4.5 \cite{quinlan2014c4}, Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}, which will be applied with the default values of the parameters provided by WEKA.
+Given that Logistic Regression and C4.5 algorithms are deterministic, only one run will be considered for each one.
+However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so, in this case, 30 executions for each one will be carried out.
+Table \ref{tab:results} shows the results of \mbox{this experimentation}.
+
+			As can be seen, Random Forest and Multilayer Perceptron algorithms have achieved similar accuracy, but the performance of the latter is slightly better.
+Although this is an illustrative classification example using datasets built with SPAMDA, both models have obtained good performance, despite the fact that the problem being tackled is difficult (prediction is approached six hours in advance).
+
+			% start a new page without indent 4.6cm
+\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka.png}
@@
 			\end{figure}
-
-			Since the final dataset is a time series of meteorological data (collected from $2013$ to $2017$), a hold-out scheme (60\% train / 40\% test) will be used.
-In this way, years from $2013$ to $2015$ will be used for the training phase ($4380$ instances) whereas $2016$ and $2017$ years will be used for the test phase ($2922$ instances).
-Previous to the learning phase, the attributes are normalised to avoid that some attributes dominate others because of a larger scale.
-
-			The classification algorithms that will be considered for wave modelling are Logistic Regression \cite{hosmer2013applied}, C4.5 \cite{quinlan2014c4}, Random Forest \cite{breiman2001random}, Support Vector Machine \cite{cortes1995support} and Multilayer Perceptron \cite{haykin1994neural}, which will be applied with the default values of the parameters provided by WEKA.
-Given that Logistic Regression and C4.5 algorithms are deterministic, only one run will be considered for each one.
-However, Random Forest, Support Vector Machine and Multilayer Perceptron algorithms have a stochastic component, so, in this case, 30 executions for each one will be carried out.
-Table \ref{tab:results} shows the results of this experimentation.
-
-				\begin{table}[!h]
-
-					\caption{Results (mean$\pm$SD) obtained by the algorithms}
+			\vspace{-6pt}
+
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+   \begin{specialtable}[H]
+
+					\caption{Results (mean $\pm$ SD) obtained by the algorithms.}
 					\label{tab:results}
-					\footnotesize
-					\centering
-
-					\begin{tabular}{m{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
-
-						\cline{1-4}
-
-						\textbf{Algorithm}&\textbf{Accuracy (CCR)}&\textbf{Kappa}&\\[0.20cm]
-
-						\cline{1-4}
-
-						Logistic Regression & $59.0691$ & $0.44447$&\\[0.15cm]
-
-						\cellcolor{gray090}C4.5 & \cellcolor{gray090}$61.7385$ & \cellcolor{gray090}$0.47852$&\\[0.15cm]
-
-						Random Forest & $\textit{68.6516} \pm \textit{0.3083}$ & $\textit{0.57040} \pm \textit{0.0042}$&\\[0.15cm]
-
-						\cellcolor{gray090}Support Vector Machine &\cellcolor{gray090} $61.0016 \pm 0.0522$ &\cellcolor{gray090} $0.46770 \pm 0.0007$&\\[0.15cm]
-
-						Multilayer Perceptron & $\mathbf{69.7045} \pm \mathbf{1.3033}$ & $\mathbf{0.58576} \pm \mathbf{0.0178}$&\\[0.15cm]
-
-						\cline{1-4}
-
-					\end{tabular}
-
-				\end{table}
-
-			As can be seen, Random Forest and Multilayer Perceptron algorithms have achieved similar accuracy, but the performance of the latter is slightly better.
-Although this is an illustrative classification example using datasets built with SPAMDA, both models have obtained good performance, despite the problem tackled is difficult (prediction is approached six hours in advance).
-
-			\subsection{Energy flux prediction}
+					\setlength{\cellWidtha}{\columnwidth/3-2\tabcolsep+0.0in}
+\setlength{\cellWidthb}{\columnwidth/3-2\tabcolsep+0.0in}
+\setlength{\cellWidthc}{\columnwidth/3-2\tabcolsep+0.0in}
+\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\centering}m{\cellWidtha}>{\PreserveBackslash\centering}m{\cellWidthb}>{\PreserveBackslash\centering}m{\cellWidthc}}
+\toprule
+
+						\textbf{Algorithm}&\textbf{Accuracy (CCR)}&\textbf{Kappa}\\
+
+						\midrule
+
+						Logistic Regression & $59.0691$ & $0.44447$\\ 		\midrule
+
+						C4.5 & $61.7385$ & $0.47852$\\ 		\midrule
+
+						Random Forest & $68.6516 \pm 0.3083$ & $0.57040 \pm 0.0042$\\ 		\midrule
+
+						Support Vector Machine & $61.0016 \pm 0.0522$ & $0.46770 \pm 0.0007$\\ 		\midrule
+
+						Multilayer Perceptron & $69.7045 \pm 1.3033$  & $0.58576 \pm 0.0178$\\
+
+						\bottomrule
+
+					\end{tabularx}}
+
+				\end{specialtable}			
+			\subsection{Energy Flux Prediction}
 
 			As mentioned above, the final dataset of this example is also used with SC and ML algorithms to predict flux of energy.
-Following sections explain the process of performing the data integration provided by SPAMDA to build the final dataset, modelling the flux of energy by using regression algorithms available in WEKA.
-
-			\subsubsection{Obtaining the final dataset}
+The following sections explain the process of performing the data integration provided by SPAMDA to build the final dataset, modelling the flux of energy by using regression algorithms available in WEKA.
+
+			\subsubsection{Obtaining the Final Dataset}
 
 			The researcher can configure the parameters of the data integration by clicking on the \textbf{\textit{Matching configuration}} tab.
-For this problem, as shown in Fig.
-\ref{fig:matching_conf_ef}, the following parameters were selected:
-
+For this problem, as shown in Figure \ref{fig:matching_conf_ef}, the following parameters were selected:
 			\begin{itemize}
@@
 				\item Reanalysis nodes to consider: $1$ (only the closest reanalysis node will be used).
-				\item Number of final datasets: In this example, this option is disabled, because only one reanalysis node is considered.
+				\item Number of final datasets: In this example, this option is disabled because only one reanalysis node is considered.
 				\item Prediction task: Regression.
@@
 
-			\begin{figure}[ht!]
-				\centering
+			After configuring the parameters of the matching process, the next step is to define the final dataset structure according to the selected prediction task.
+Researchers can click on the \textbf{\textit{Run}} button to jump to the view shown in Figure \ref{fig:final_dataset_EF}.
+Note that the thresholds for discretising the output variable (top left of  Figure \ref{fig:final_dataset_EF}) are disabled due to, in this case, energy flux prediction being a regression problem.
+
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.85\textwidth]{figures/FigureMatching_configuration_EF.png}
@@
 			\end{figure}
-
-			\begin{figure}[ht!]
-				\centering
+\vspace{-6pt}
+
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.85\textwidth]{figures/FigureFinal_datasets_EF.png}
@@
 			\end{figure}
-
-			After configuring the parameters of the matching process, the next step is to define the final dataset structure according to the selected prediction task.
-Researcher can click on the \textbf{\textit{Run}} button to jump to the view shown in Fig.
-\ref{fig:final_dataset_EF}.
-Note that, the thresholds for discretising the output variable (top left of Fig.
-\ref{fig:final_dataset_EF}) are disabled due to, in this case, energy flux prediction is a regression problem.
-
-			By default, the time horizon is set to $6$ hours, that is, the energy flux prediction will be performed $6$ hours in advance (top right of Fig.
-\ref{fig:final_dataset_EF}), but researchers can increase such time horizon depending on their needs.
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			By default, the time horizon is set to six hours, that is, the energy flux prediction will be performed six hours in advance (top right of Figure \ref{fig:final_dataset_EF}), but researchers can increase such time horizon depending on their needs.
 The synchronisation (in time) of reanalysis variables with the output (explained in Section {\ref{sec:FinalDatasets}}) can be set in this view.
-By clicking on the \textbf{\textit{Update final dataset}} button, researchers can preview the content of the final dataset (bottom left corner of Fig.
-\ref{fig:final_dataset_EF}).
-Finally, the last step would be to set the name, path and output format (CSV or ARFF) of the dataset file, and then the user should click on the \textbf{\textit{Create final datasets}} button (bottom right of Fig.
-\ref{fig:final_dataset_EF}).
+By clicking on the \textbf{\textit{Update final dataset}} button, researchers can preview the content of the final dataset (bottom left corner of  Figure  \ref{fig:final_dataset_EF}).
+Finally, the last step would be to set the name, path and output format (CSV or ARFF) of the dataset file, and then the user should click on the \textbf{\textit{Create final datasets}} button (bottom right of  Figure  \ref{fig:final_dataset_EF}).
 For this example, the following configuration was applied:
-
-			\begin{itemize}
-				\item Prediction horizon: 6 hours.
+			\begin{itemize}
+				\item Prediction horizon: 6 h.
 				\item Synchronisation: Disabled.
@@
 			After that, the final dataset would be created and stored in the computer of the researcher according to the introduced configuration, ready to be used as input for SC and ML techniques to tackle the problem of energy flux prediction.
-The number of instances ($7302$) and the distribution of the final dataset (Table \ref{tab:datasetDistribution}) are the same as in the previous example (waves classification) since the data used to create the final dataset and the time horizon selected (6h) are the same.
-
-		\subsubsection{Obtaining prediction models with ML algorithms}
-
-			In this example, WEKA is used as SC and ML tool to obtain energy flux prediction models, as shown in Fig.
-\ref{fig:openigFinalDatasetWeka_EF}.
+The number of instances ($7302$) and the distribution of the final dataset (Table \ref{tab:datasetDistribution}) are the same as in the previous example (waves classification) since the data used to create the final dataset and the time horizon selected (6 h) are the same.
+
+		\subsubsection{Obtaining Prediction Models with ML Algorithms}
+
+			In this example, WEKA is used as SC and ML tool to obtain energy flux prediction models, as shown in Figure \ref{fig:openigFinalDatasetWeka_EF}.
 Nonetheless, the final dataset can be created in CSV format so that the researcher can use any other SC and ML tool.
 
-			\begin{figure}[ht!]
-				\centering
+			% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+				\widefigure
 				\includegraphics[width=0.95\textwidth]{figures/FigureOpeningFinalDatasetWeka_EF.png}
@@
 			\end{figure}
-
-			For this problem, the same partitioning scheme used in the wave classification problem is considered (60\% train / 40\% test), that is, years from $2013$ to $2015$ for the training phase ($4380$ instances) and years $2016$ and $2017$ for the test phase ($2922$ instances).
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+			For this problem, the same partitioning scheme used in the wave classification problem is considered (60\% train/40\% test), that is, from $2013$ to $2015$ for the training phase ($4380$ instances) and $2016$ and $2017$ for the test phase ($2922$ instances).
 Again, the attributes are normalised prior to the learning phase.
@@
 
-				\begin{table}[!h]
-
-					\caption{Results (mean$\pm$SD) obtained by the algorithms}
+				 \begin{specialtable}[H]
+
+					\caption{Results (mean $\pm$ SD) obtained by the algorithms.}
 					\label{tab:results_EF}
-					\footnotesize
-					\centering
-
-					\begin{tabular}{m{3.20cm}cc@{\setlength{\tabcolsep}{0pt}}m{0.0cm}}
-
-						\cline{1-4}
-
-						\textbf{Algorithm}&\textbf{Root mean squared error}&\textbf{Correlation coefficient}&\\[0.20cm]
-
-						\cline{1-4}
-
-						Linear Regression & $29.6368$ & $0.7296$&\\[0.15cm]
-
-						\cellcolor{gray090}Random Forest &\cellcolor{gray090} $\mathbf{23.4353} \pm \mathbf{0.1313}$ &\cellcolor{gray090} $\textit{0.8408} \pm \textit{0.0021}$&\\[0.15cm]
-
-						Support Vector Machine & $31.3008 \pm 0.1197$ & $0.7275 \pm 0.0015$&\\[0.15cm]
-
-						\cellcolor{gray090}Multilayer Perceptron &\cellcolor{gray090} $\textit{27.1151} \pm \textit{7.5536}$ &\cellcolor{gray090} $\mathbf{0.8444} \pm \mathbf{0.0193}$&\\[0.15cm]
-
-						\cline{1-4}
-
-					\end{tabular}
-
-				\end{table}
+%					\footnotesize
+%					\centering
+
+					\setlength{\cellWidtha}{\columnwidth/3-2\tabcolsep+0.0in}
+\setlength{\cellWidthb}{\columnwidth/3-2\tabcolsep+0.0in}
+\setlength{\cellWidthc}{\columnwidth/3-2\tabcolsep+0.0in}
+\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\centering}m{\cellWidtha}>{\PreserveBackslash\centering}m{\cellWidthb}>{\PreserveBackslash\centering}m{\cellWidthc}}
+\toprule
+
+						\textbf{Algorithm}&\textbf{Root Mean Squared Error}&\textbf{Correlation Coefficient}\\
+
+						\midrule
+
+						Linear Regression & $29.6368$ & $0.7296$\\ 		\midrule
+
+						Random Forest & $23.4353 \pm 0.1313$ & $0.8408 \pm 0.0021$\\ 		\midrule
+
+						Support Vector Machine & $31.3008 \pm 0.1197$ & $0.7275 \pm 0.0015$\\ 		\midrule
+
+						Multilayer Perceptron & $27.1151 \pm 7.5536$ & $0.8444 \pm 0.0193$\\
+
+						\bottomrule
+
+					\end{tabularx}}
+
+				\end{specialtable}
 
 			As can be checked, Random Forest has achieved the best performance for the Root mean squared error.
-The standard deviation of the results obtained by Multilayer Perceptron indicates that this algorithm may has been slightly affected by its stochastic component.
+The standard deviation of the results obtained by Multilayer Perceptron indicates that this algorithm may have been slightly affected by its stochastic component.
 However, both Multilayer Perceptron and Random Forest have obtained an excellent Correlation coefficient.
@@
 An exhaustive comparison of regression algorithms is not the purpose of this work.
-However, note that Multilayer Perceptron and Random Forest algorithms have achieved very good results despite the fact that the energy flux prediction has been performed with a time horizon of 6h.
-
-		\subsection{Important remarks}
+However, note that Multilayer Perceptron and Random Forest algorithms have achieved very good results despite the fact that the energy flux prediction has been performed with a time horizon of 6 h.
+
+		\subsection{Important Remarks}
 
 			In this section, it has been described how to use SPAMDA to create final datasets with the aim of classifying waves and predicting flux of energy.
-However, using the same data described in Section \ref{sec:ObtainingFinalDataset} the researcher can quickly address other objectives or different studies by merely tailoring the matching configuration of the data integration process.
+However, using the same data described in Section \ref{sec:ObtainingFinalDataset}, the researcher can quickly address other objectives or different studies by merely tailoring the matching configuration of the data integration process.
 For example, longer-term wave or energy flux prediction can be addressed by changing the time horizon, waves modelling can be approached from another perspective by creating the final dataset for regression, or environmental modelling can be focused in diverse fields by changing the output meteorological variable.
@@
 			As SPAMDA performs all data processing and management to create the datasets, it not only prevents researchers from performing repetitive tasks but also prevents them from making possible errors.
-In this way, researchers can focus on the studies they are carrying out.
+In this way, researchers can focus on the studies they are carrying~out.
 
@@
 
-		%A new open source tool named SPAMDA with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
+		%A new open source tool named SPAMDA with a user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP has been presented in this work.
 The aim of the tool presented in this work is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.
 
-		%Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
+		%Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc) on data using custom-made implementations or scripts developed in some programming language; but they do not allow for build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
 These datasets can be easily obtained with SPAMDA by means of the selection of different input parameters, such as predictive and objective variables, output discretisation or prediction horizon.
@@
 
-		\textcolor{blue}{Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc.) on data using custom-made implementations or scripts developed in some programming language; but they do not allow to build datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
-In this sense, a new open source tool named SPAMDA has been presented in this work, with an user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP.
+{Studies on marine energy using ML and SC methodologies apply specific algorithms (extreme learning machine, metaheuristics, Bayesian networks, neural networks, etc.) on data using custom-made implementations or scripts developed in some programming language; but they do not allow for building datasets in an automated way ready to be used as input for prediction tasks (classification or regression).
+In this sense, a new open source tool named SPAMDA has been presented in this work, with a user-friendly GUI for creating datasets using meteorological data from NDBC and NNRP.
 The aim of the tool is to provide the research community with an automated, customisable and robust integration for NDBC and NNRP data, serving as a tool for analysis and decision support in marine energy and engineering applications, among others.}
@@
 		The case study described in Section \ref{sec:CaseStudy} illustrates how SPAMDA can be used by researchers in a practical approach for environmental modelling, concretely, to classify waves in the Gulf of Alaska depending on their height.
-The case study also covers an example of energy flux prediction, to predict the wave energy that could be exploited by WECs facilities six hours in advance, although such time horizon is customisable.
+The case study also covers an example of energy flux prediction, to predict the wave energy that could be exploited by WEC facilities six hours in advance, although such time horizon is customisable.
 Given that this work does not focus on models performance, a more extensive validation or comparison study of the results obtained in both examples has not been carried out.
-\textcolor{blue}{The final datasets obtained with SPAMDA can be replicated by researchers using the same meteorological data from NDBC and NNRP (publicly available) and applying the same parameters for the pre-processing tasks and the data integration process.
+{The final datasets obtained with SPAMDA can be replicated by researchers using the same meteorological data from NDBC and NNRP (publicly available) and applying the same parameters for the pre-processing tasks and the data integration process.
 After that, the models and results obtained, using such final datasets, will depend on the SC or ML tool used.}
@@
 
-    \section*{Additional material}
-
-		The source code and the software tool are available at \url{https://github.com/ayrna}.
+%    \section*{Additional material}
+%	
+%		The source code and the software tool are available at \url{https://github.com/ayrna}.
 
@@
 %% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
+\supplementary{The source code and the software tool are available at \url{https://github.com/ayrna}.}
 
@@
 
-\authorcontributions{Conceptualization, Formal analysis and Investigation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Funding acquisition, Project administration, Resources and Supervision, Pedro Antonio Gutiérrez and César Hervás-Martínez; Methodology, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Software, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno; Validation, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández, Manuel Dorado-Moreno, Pedro Antonio Gutiérrez and César Hervás-Martínez; Writing – original draft, Antonio Manuel Gómez-Orellana, Juan Carlos Fernández and Manuel Dorado-Moreno.}
+\authorcontributions{Conceptualization, Formal analysis and Investigation, A.M.G.-O., J.C.F., M.D.-M., P.A.G.
+and C.H.-M.; Funding acquisition, Project administration, Resources and Supervision, P.A.G.
+and C.H.-M.; Methodology, A.M.G.-O., J.C.F., M.D.-M., P.A.G.
+and C.H.-M.; Software, \mbox{A.M.G.-O.}, J.C.F.
+and M.D.-M.; Validation, A.M.G.-O., J.C.F., M.D.-M., P.A.G.
+and C.H.-M.; Writing---original draft, A.M.G.-O., J.C.F.
+and M.D.-M.
+All authors have read and agreed to the published version of the manuscript.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+%\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER Grant No.
+XXX.'' and  and ``The APC was funded by XXX''.
 Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
@@
 \funding{This work has been partially subsidised by the projects with references TIN2017-85887-C2-1-P of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), UCO-1261651 of the ``Consejer\'ia de Econom\'ia, Conocimiento, Empresas y Universidad'' of the ``Junta de Andaluc\'ia'' (Spain) and FEDER funds of the European Union.}
+\institutionalreview{Not applicable.}%MDPI: In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals.}
+
+\informedconsent{Not applicable.}%MDPI: Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.}
+
+\dataavailability{Publicly available datasets were analyzed in this study.
+This data can be found in the National Data Buoy Center (NDBC) and the NOAA Physical Sciences Laboratory (PSL).}%MDPI: Please refer to suggested Data Availability Statements in section “MDPI Research Data Policies” at \href{https://www.mdpi.com/ethics}{https://www.mdpi.com/ethics}}.
 
@@
 
-\acknowledgments{The authors also thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to University of Waikato for the WEKA (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
+\acknowledgments{The authors also thank NOAA/OAR/ESRL PSD, Boulder, CO, USA for the NCEP Reanalysis data provided from their Web site at \url{https://www.esrl.noaa.gov/psd/}, to NOAA/NDBC by its data that were collected and made freely available, to the University of Waikato for the WEKA (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.}
 
@@
 \conflictsofinterest{The authors declare no conflict of interest.} 
-
+%\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \appendixtitles{yes}
-
+\appendixstart
 \appendix
-\section{Managing the casuistry of incomplete data}\label{app:AppendixA}
+\section{Managing the Casuistry of Incomplete Data}\label{app:AppendixA}
 
@@
 		The measurements collected by the buoys may be incomplete or recorded at a different time than the expected one, due to the weather conditions in which the buoys have to operate.
-To illustrate this casuistry, the following examples are shown in Fig.
-\ref{fig:measurements}:
-		\begin{itemize}
-			\item In the instance marked with a), the measurement of $17$:$50$ was collected at $17$:$45$, $5$ minutes earlier.
-			\item In the instance marked with b), the measurement of $23$:$50$ was collected at $23$:$30$, $20$ minutes earlier.
-			\item In the instance marked with c), the measurement of $05$:$50$ is duplicated.
-			\item In the instance marked with d), the measurement of $11$:$50$ is missing (missing date or instance).
-			\item In the instance marked with e), the measurement of $17$:$50$ and $18$:$50$ are missing (missing dates or instances).
-			\item Missing values highlighted in red colour.
-		\end{itemize}
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.47]{figures/FigureMeasurements.png}
+To illustrate this casuistry, the following examples are shown in Figure \ref{fig:measurements}:
+		\begin{itemize}
+			\item In the instance marked with (a), the measurement of $17$:$50$ was collected at $17$:$45$, \mbox{$5$ min} earlier.
+			\item In the instance marked with (b), the measurement of $23$:$50$ was collected at $23$:$30$, \mbox{$20$ min} earlier.
+			\item In the instance marked with (c), the measurement of $05$:$50$ is duplicated.
+			\item In the instance marked with (d), the measurement of $11$:$50$ is missing (missing date or instance).
+			\item In the instance marked with (e), the measurement of $17$:$50$ and $18$:$50$ are missing (missing dates or instances).
+			\item Missing values highlighted in red.
+		\end{itemize}
+
+		SPAMDA has been designed to tackle these situations, and it informs researchers of any incidence found while reading the annual text files for creating the intermediate datasets.
+For the case of measurements that were recorded at a different time than expected, a time gap of 6 min ($10\%$ of an hour) has been established.
+Therefore, if the time difference exceeds such value, the date will be considered as an unexpected.
+
+		Figure \ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Figure \ref{fig:measurements}.
+Note that the instance marked with a) has not been informed by SPAMDA as an unexpected date because its time difference is less than six minutes.
+Depending on the affected attribute, NDBC uses a specific value \cite{NOAA_3} to indicate the presence of lost data (e.g., $99$ for VIS and TIDE attributes, $999$ for DEWP, MWD and WDIR, etc.).
+SPAMDA interprets these specific values and, after creating the intermediate dataset, researchers can check if it contains missing values by visualising its statistical information or content.
+Remember that SPAMDA provides several filters for recovering missing data, which were described in Section \ref{sec:Datasets}.
+
+		SPAMDA takes into account this casuistry when carrying out the matching process.
+An example is given in Figure \ref{fig:matchingMeasurements}.
+As mentioned above, the matching process is performed with the nearest measurement (previous or next) within a maximum of 60 min of difference.
+However, in the instance marked with $e)$, given that the measurements dates 1 May 2017 $17$:$50$ and 1 May 2017 $18$:$50$ are missing, the reanalysis date 1 May 2017 $18$:$00$ cannot be matched with buoy data (this date is highlighted in mauve in Figure \ref{fig:matchingMeasurements}).
+Depending on the selection made by researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.
+
+	% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\appendix
+\begin{figure}[H]
+	\widefigure
+			\includegraphics[scale=0.55]{figures/FigureMeasurements.png}
 			\caption{A fragment of an annual text file with different missing value examples.}
+			%Figures A1 and A3, 1.
+Please fill in the brackets parentheses of Numbering subfigure.
+2.
+Please add explanation of subfigure (a)--(e).
+3.
+Please add explanation of different colors.
+AUTHORS: the explanation is in the maintext
 			\label{fig:measurements}
 		\end{figure}	
-
-		SPAMDA has been designed to tackle these situations, and it informs researchers of any incidence found while reading the annual text files for creating the intermediate datasets.
-For the case of measurements that were recorded at a different time than expected, it has been established a time gap of 6 minutes ($10\%$ of an hour).
-Therefore, if the time difference exceeds such value the date will be considered as an unexpected.
-
-		Fig.
-\ref{fig:creatingDataset} shows the status of the creation of an intermediate dataset with the information of Fig \ref{fig:measurements}.
-Note that the instance marked with a) has not been informed by SPAMDA as an unexpected date because its time difference is less than $6$ minutes.
-Depending on the affected attribute, NDBC uses a specific value \cite{NOAA_3} to indicate the presence of lost data (e.g.
-$99$ for VIS and TIDE attributes, $999$ for DEWP, MWD and WDIR, etc.).
-SPAMDA interprets these specific values and, after creating the intermediate dataset, researchers can check if it contains missing values by visualising its statistical information or content.
-Remember that SPAMDA provides several filters for recovering missing data, which were described in subsection \ref{sec:Datasets}.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.47]{figures/FigureCreatingDataset.png}
-			\caption{Status of the creation of the intermediate dataset for the example of Fig \ref{fig:measurements}.}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+		\vspace{-6pt}
+
+		\begin{figure}[H]
+%			\centering
+			\includegraphics[scale=0.6]{figures/FigureCreatingDataset.png}
+			\caption{Status of the creation of the intermediate dataset for the example of Figure \ref{fig:measurements}.}
 			\label{fig:creatingDataset}
 		\end{figure}
-
-		SPAMDA takes into account this casuistry when carrying out the matching process.
-An example is given in Fig.
-\ref{fig:matchingMeasurements}.
-As above-mentioned, the matching process is performed with the nearest measurement (previous or next) within a maximum of 60 minutes of difference.
-However, in the instance marked with $e)$, given that the measurements dates $01$/$05$/$2017$ $17$:$50$ and $01$/$05$/$2017$ $18$:$50$ are missing, the reanalysis date $01$/$05$/$2017$ $18$:$00$ cannot be matched with buoy data (this date is highlighted in mauve colour in the Figure).
-Depending on the selection made by researchers in the parameter \textit{Include missing dates}, this instance will be included in the final dataset (with missing values for buoy variables) or not.
-
-		\begin{figure}[ht!]
-			\centering
-			\includegraphics[scale=0.4]{figures/FigureMatchingMeasurements.png}
-			\caption{Matching the measurements (left) and the reanalysis data (right).}
+		\vspace{-6pt}
+
+		% start a new page without indent 4.6cm
+\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{figure}[H]
+			\widefigure
+			\includegraphics[scale=0.45]{figures/FigureMatchingMeasurements.png}
+			\caption{Matching the measurements (\textbf{left}) and the reanalysis data (\textbf{right}).}
 			\label{fig:matchingMeasurements}
 		\end{figure}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{paracol}
 \reftitle{References}
 
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Please provide either the correct journal abbreviation (e.g., according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
 % Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
@@
 %\bibliography{your_external_BibTeX_file}
-\externalbibliography{yes}
-\bibliography{bibfile}
+%\externalbibliography{yes}
+%\bibliography{bibfile}
+\begin{thebibliography}{999}
+
+\bibitem[Anis \em{et~al.}(2019)Anis, Jamil, Ansari, and
+  Bellos]{SHAHRUKHANIS2019179}
+Anis, M.S.; Jamil, B.; Ansari, M.A.; Bellos, E.
+\newblock {Generalized models for estimation of global solar radiation based on
+  sunshine duration and detailed comparison with the existing: A case study for
+  India}.
+\newblock {\em Sustain. Energy Technol. Assess.} {\bf 2019},
+  {\em 31},~179--198, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.seta.2018.12.009}{\detokenize{10.1016/j.seta.2018.12.009}}}.
+
+\bibitem[Laface \em{et~al.}(2015)Laface, Arena, and Soares]{LAFACE201545}
+Laface, V.; Arena, F.; Soares, C.G.
+\newblock {\textls[-15]{Directional analysis of sea storms}}.
+\newblock {\em Ocean Eng.} {\bf 2015}, {\em 107},~45--53, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.oceaneng.2015.07.027}{\detokenize{10.1016/j.oceaneng.2015.07.027}}}.
+
+\bibitem[Shivam \em{et~al.}(2020)Shivam, Tzou, and Wu]{Kumar2020}
+Shivam, K.; Tzou, J.C.; Wu, S.C.
+\newblock Multi-Objective Sizing Optimization of a Grid-Connected Solar–Wind
+  Hybrid System Using Climate Classification: A Case Study of Four Locations in
+  Southern Taiwan.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~2505, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en13102505}{\detokenize{10.3390/en13102505}}}.
+
+\bibitem[Dorado-Moreno \em{et~al.}(2017)Dorado-Moreno, Cornejo-Bueno,
+  Guti{\'e}rrez, Prieto, Herv{\'a}s-Mart{\'i}nez, and
+  Salcedo-Sanz]{DORADOMORENO2017428}
+Dorado-Moreno, M.; Cornejo-Bueno, L.; Guti{\'e}rrez, P.A.; Prieto, L.;
+  Herv{\'a}s-Mart{\'i}nez, C.; Salcedo-Sanz, S.
+\newblock {Robust estimation of wind power ramp events with reservoir
+  computing}.
+\newblock {\em Renew. Energy} {\bf 2017}, {\em 111},~428--437, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.renene.2017.04.016}{\detokenize{10.1016/j.renene.2017.04.016}}}.
+
+\bibitem[He \em{et~al.}(2020)He, Zha, Song, Hao, Du, Liotta, and Perra]{He2020}
+He, Q.; Zha, C.; Song, W.; Hao, Z.; Du, Y.; Liotta, A.; Perra, C.
+\newblock Improved Particle Swarm Optimization for Sea Surface Temperature
+  Prediction.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~1369, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en13061369}{\detokenize{10.3390/en13061369}}}.
+
+\bibitem[Fuchs and Gerbi(2016)]{FUCHS2016109}
+Fuchs, H.L.; Gerbi, G.P.
+\newblock {Seascape-level variation in turbulence- and wave-generated
+  hydrodynamic signals experienced by plankton}.
+\newblock {\em Prog. Oceanogr.} {\bf 2016}, {\em 141},~109--129, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.pocean.2015.12.010}{\detokenize{10.1016/j.pocean.2015.12.010}}}.
+
+\bibitem[da~Silva \em{et~al.}(2010)da~Silva, Ara{\'u}jo~e Silva, Cavalcanti,
+  Braga~Campos, Vieira~de Azevedo, Singh, and Rodrigues~Pereira]{SILVA20101852}
+da~Silva, V.D.P.R.; Ara{\'u}jo~e Silva, R.; Cavalcanti, E.P.; Braga~Campos, C.;
+  Vieira~de Azevedo, P.; Singh, V.P.; Rodrigues~Pereira, E.R.
+\newblock {Trends in solar radiation in NCEP/NCAR database and measurements in
+  northeastern Brazil}.
+\newblock {\em Sol. Energy} {\bf 2010}, {\em 84},~1852--1862, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.solener.2010.07.011}{\detokenize{10.1016/j.solener.2010.07.011}}}.
+
+\bibitem[Gouldby \em{et~al.}(2014)Gouldby, M{\'e}ndez, Guanche, Rueda, and
+  M{\'i}nguez]{GOULDBY201415}
+Gouldby, B.; M{\'e}ndez, F.J.; Guanche, Y.; Rueda, A.; M{\'i}nguez, R.
+\newblock {A methodology for deriving extreme nearshore sea conditions for
+  structural design and flood risk analysis}.
+\newblock {\em Coast. Eng.} {\bf 2014}, {\em 88},~15--26, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.coastaleng.2014.01.012}{\detokenize{10.1016/j.coastaleng.2014.01.012}}}.
+
+\bibitem[Alizadeh \em{et~al.}(2019)Alizadeh, Jia, Nellippallil, Wang, Hao,
+  Allen, and Mistree]{Alizadeh2019}
+Alizadeh, R.; Jia, L.; Nellippallil, A.B.; Wang, G.; Hao, J.; Allen, J.K.;
+  Mistree, F.
+\newblock Ensemble of surrogates and cross-validation for rapid and accurate
+  predictions using small data sets.
+\newblock {\em Artif. Intell. Eng. Des. Anal. Manuf.} {\bf 2019}, {\em 33},~484–501, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1017/S089006041900026X}{\detokenize{10.1017/S089006041900026X}}}.
+
+\bibitem[Alizadeh \em{et~al.}(2020)Alizadeh, Allen, and Mistree]{Alizadeh2020}
+Alizadeh, R.; Allen, J.K.; Mistree, F.
+\newblock Managing computational complexity using surrogate models: A critical
+  review.
+\newblock {\em Res. Eng. Des.} {\bf 2020}, {\em 31},~275–298,  
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1007/s00163-020-00336-7}{\detokenize{10.1007/s00163-020-00336-7}}}.
+
+\bibitem[Manfren \em{et~al.}(2020)Manfren, Groppi, and
+  Astiaso~Garcia]{Manfren2020}
+Manfren, M.; Groppi, D.; Astiaso~Garcia, D.
+\newblock Open data and energy analytics---An analysis of essential information
+  for energy system planning, design and operation.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~2334, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.energy.2020.118803}{\detokenize{10.1016/j.energy.2020.118803}}}.
+
+\bibitem[Dhanraj~Bokde \em{et~al.}(2020)Dhanraj~Bokde, Mundher~Yaseen, and
+  Bruun~Andersen]{Neeraj2020}
+Dhanraj~Bokde, N.; Mundher~Yaseen, Z.; Bruun~Andersen, G.
+\newblock ForecastTB---An R Package as a Test-Bench for Time Series
+  Forecasting---Application of Wind Speed and Solar Radiation Modeling.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~2578, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en13102578}{\detokenize{10.3390/en13102578}}}.
+
+\bibitem[Lo \em{et~al.}(2015)Lo, Lim, and Rahman]{LO2015293}
+Lo, C.K.; Lim, Y.S.; Rahman, F.A.
+\newblock {New integrated simulation tool for the optimum design of bifacial
+  solar panel with reflectors on a specific site}.
+\newblock {\em Renew. Energy} {\bf 2015}, {\em 81},~293--307, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.renene.2015.03.047}{\detokenize{10.1016/j.renene.2015.03.047}}}.
+
+\bibitem[Nguyen \em{et~al.}(2013)Nguyen, Prinz, Friis{\o}, Nossum, and
+  Tyapin]{NGUYEN2013150}
+Nguyen, T.H.; Prinz, A.; Friis{\o}, T.; Nossum, R.; Tyapin, I.
+\newblock {A framework for data integration of offshore wind farms}.
+\newblock {\em Renew. Energy} {\bf 2013}, {\em 60},~150--161, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.renene.2013.05.002}{\detokenize{10.1016/j.renene.2013.05.002}}}.
+
+\bibitem[Di~Bari \em{et~al.}(2020)Di~Bari, Horn, Nienborg, Klinker,
+  Kieseritzky, and Pawelz]{Roberta2020}
+Di~Bari, R.; Horn, R.; Nienborg, B.; Klinker, F.; Kieseritzky, E.; Pawelz, F.
+\newblock The Environmental Potential of Phase Change Materials in Building
+  Applications. A Multiple Case Investigation Based on Life Cycle Assessment
+  and Building Simulation.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~3045, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en13123045}{\detokenize{10.3390/en13123045}}}.
+
+\bibitem[Astiaso~Garcia and Bruschi(2016)]{ASTIASOGARCIA201648}
+Astiaso~Garcia, D.; Bruschi, D.
+\newblock {A risk assessment tool for improving safety standards and emergency
+  management in Italian onshore wind farms}.
+\newblock {\em Sustain. Energy Technol. Assess.} {\bf 2016},
+  {\em 18},~48--58, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.seta.2016.09.009}{\detokenize{10.1016/j.seta.2016.09.009}}}.
+
+\bibitem[Raabe \em{et~al.}(2010)Raabe, Klein, Gonz{\'a}lez, and
+  Medina]{RAABE2010213}
+Raabe, A.L.A.; Klein, A.H.d.F.; Gonz{\'a}lez, M.; Medina, R.
+\newblock {MEPBAY and SMC: Software tools to support different operational
+  levels of headland-bay beach in coastal engineering projects}.
+\newblock {\em Coast. Eng.} {\bf 2010}, {\em 57},~213--226, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.coastaleng.2009.10.008}{\detokenize{10.1016/j.coastaleng.2009.10.008}}}.
+
+\bibitem[Motahhir \em{et~al.}(2019)Motahhir, EL~Hammoumi, EL~Ghzizal, and
+  Derouich]{MOTAHHIR20199}
+Motahhir, S.; EL~Hammoumi, A.; EL~Ghzizal, A.; Derouich, A.
+\newblock {Open hardware/software test bench for solar tracker with virtual
+  instrumentation}.
+\newblock {\em Sustain. Energy Technol. Assess.} {\bf 2019},
+  {\em 31},~9--16, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.seta.2018.11.003}{\detokenize{10.1016/j.seta.2018.11.003}}}.
+
+\bibitem[Cascajo \em{et~al.}(2019)Cascajo, Garc{\'i}a, Quiles, Correcher, and
+  Morant]{en12050787}
+Cascajo, R.; Garc{\'i}a, E.; Quiles, E.; Correcher, A.; Morant, F.
+\newblock {Integration of Marine Wave Energy Converters into Seaports: A Case
+  Study in the Port of Valencia}.
+\newblock {\em Energies} {\bf 2019}, {\em 12},~787, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en12050787}{\detokenize{10.3390/en12050787}}}.
+
+\bibitem[Zeyringer \em{et~al.}(2018)Zeyringer, Fais, Keppo, and
+  Price]{ZEYRINGER20181281}
+Zeyringer, M.; Fais, B.; Keppo, I.; Price, J.
+\newblock {The potential of marine energy technologies in the UK---Evaluation
+  from a systems perspective}.
+\newblock {\em Renew. Energy} {\bf 2018}, {\em 115},~1281--1293, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.renene.2017.07.092}{\detokenize{10.1016/j.renene.2017.07.092}}}.
+
+\bibitem[{De Jong} \em{et~al.}(2019){De Jong}, Hoppe, and Noori]{en12091657}
+{De Jong}, M.; Hoppe, T.; Noori, N.
+\newblock {City Branding, Sustainable Urban Development and the Rentier State.
+  How do Qatar, Abu Dhabi and Dubai present Themselves in the Age of Post Oil
+  and Global Warming?}
+\newblock {\em Energies} {\bf 2019}, {\em 12},~1657, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en12091657}{\detokenize{10.3390/en12091657}}}.
+
+\bibitem[Brede and de~Vries(2013)]{BREDE201344}
+Brede, M.; de~Vries, B.J.M.
+\newblock {The energy transition in a climate-constrained world: Regional vs.
+  global optimization}.
+\newblock {\em Environ. Model. Softw.} {\bf 2013}, {\em
+  44},~44--61, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.envsoft.2012.07.011}{\detokenize{10.1016/j.envsoft.2012.07.011}}}.
+
+\bibitem[Alizadeh \em{et~al.}(2020)Alizadeh, Lund, and
+  Soltanisehat]{Alizadeh2020a}
+Alizadeh, R.; Lund, P.D.; Soltanisehat, L.
+\newblock Outlook on biofuels in future studies: A systematic literature
+  review.
+\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2020}, {\em
+  134},~110326, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/doi.org/10.1016/j.rser.2020.110326}{\detokenize{10.1016/j.rser.2020.110326}}}.
+
+\bibitem[Falc{\~a}o(2010)]{FALCAO2010899}
+Falc{\~a}o, A.F.D.O.
+\newblock {Wave energy utilization: A review of the technologies}.
+\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2010}, {\em
+  14},~899--918, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.rser.2009.11.003}{\detokenize{10.1016/j.rser.2009.11.003}}}.
+
+\bibitem[Amini \em{et~al.}(2020)Amini, Golbaz, Amini, Majidi~Nezhad, Neshat,
+  and Astiaso~Garcia]{Amini2020}
+Amini, E.; Golbaz, D.; Amini, F.; Majidi~Nezhad, M.; Neshat, M.;
+  Astiaso~Garcia, D.
+\newblock A Parametric Study of Wave Energy Converter Layouts in Real Wave
+  Models.
+\newblock {\em Energies} {\bf 2020}, {\em 13},~6095, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en13226095}{\detokenize{10.3390/en13226095}}}.
+
+\bibitem[Oliveira-Pinto \em{et~al.}(2019)Oliveira-Pinto, Rosa-Santos, and
+  Taveira-Pinto]{OLIVEIRAPINTO2019556}
+Oliveira-Pinto, S.; Rosa-Santos, P.; Taveira-Pinto, F.
+\newblock {Electricity supply to offshore oil and gas platforms from renewable
+  ocean wave energy: Overview and case study analysis}.
+\newblock {\em Energy Convers. Manag.} {\bf 2019}, {\em
+  186},~556--569, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.enconman.2019.02.050}{\detokenize{10.1016/j.enconman.2019.02.050}}}.
+
+\bibitem[Fern{\'a}ndez~Prieto \em{et~al.}(2019)Fern{\'a}ndez~Prieto,
+  Rodr{\'i}guez~Rodr{\'i}guez, and
+  Schallenberg~Rodr{\'i}guez]{FERNANDEZPRIETO2019546}
+Fern{\'a}ndez~Prieto, L.; Rodr{\'i}guez~Rodr{\'i}guez, G.;
+  Schallenberg~Rodr{\'i}guez, J.
+\newblock {Wave energy to power a desalination plant in the north of Gran
+  Canaria Island: Wave resource, socioeconomic and environmental assessment}.
+\newblock {\em J. Environ. Manag.} {\bf 2019}, {\em
+  231},~546--551, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.jenvman.2018.10.071}{\detokenize{10.1016/j.jenvman.2018.10.071}}}.
+
+\bibitem[Ochi(1998)]{ochi1998}
+Ochi, M.K.
+\newblock {\em {Ocean Waves: The Stochastic Approach}}; {Cambridge Ocean
+  Technology Series}; Cambridge University Press:  Cambridge, United Kingdom, 1998.
+
+\bibitem[Crowley \em{et~al.}(2018)Crowley, Porter, Taunton, and
+  Wilson]{CROWLEY2018159}
+Crowley, S.; Porter, R.; Taunton, D.J.; Wilson, P.A.
+\newblock {Modelling of the WITT wave energy converter}.
+\newblock {\em Renew. Energy} {\bf 2018}, {\em 115},~159--174, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.renene.2017.08.004}{\detokenize{10.1016/j.renene.2017.08.004}}}.
+
+\bibitem[Abdelkhalik \em{et~al.}(2016)Abdelkhalik, Robinett, Zou, Bacelli, Coe,
+  Bull, Wilson, and Korde]{Abdelkhalik2016}
+Abdelkhalik, O.; Robinett, R.; Zou, S.; Bacelli, G.; Coe, R.; Bull, D.; Wilson,
+  D.; Korde, U.
+\newblock {On the control design of wave energy converters with wave
+  prediction}.
+\newblock {\em J. Ocean. Eng. Mar. Energy} {\bf 2016}, {\em
+  2},~473--483, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1007/s40722-016-0048-4}{\detokenize{10.1007/s40722-016-0048-4}}}.
+
+\bibitem[Ringwood \em{et~al.}(2014)Ringwood, Bacelli, and Fusco]{6898109}
+Ringwood, J.V.; Bacelli, G.; Fusco, F.
+\newblock {Energy-Maximizing Control of Wave-Energy Converters: The Development
+  of Control System Technology to Optimize Their Operation}.
+\newblock {\em IEEE Control Syst.} {\bf 2014}, {\em 34},~30--55, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1109/MCS.2014.2333253}{\detokenize{10.1109/MCS.2014.2333253}}}.
+
+\bibitem[Wei(2018)]{en11010011}
+Wei, C.C.
+\newblock {Nearshore Wave Predictions Using Data Mining Techniques during
+  Typhoons: A Case Study near Taiwan{\rq}s Northeastern Coast}.
+\newblock {\em Energies} {\bf 2018}, {\em 11},~11, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en11010011}{\detokenize{10.3390/en11010011}}}.
+
+\bibitem[Kaloop \em{et~al.}(2020)Kaloop, Kumar, Zarzoura, Roy, and
+  Hu]{Kaloop2020}
+Kaloop, M.R.; Kumar, D.; Zarzoura, F.; Roy, B.; Hu, J.W.
+\newblock A wavelet---Particle swarm optimization---Extreme learning machine
+  hybrid modeling for significant wave height prediction.
+\newblock {\em Ocean Eng.} {\bf 2020}, {\em 213},~107777, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/doi.org/10.1016/j.oceaneng.2020.107777}{\detokenize{10.1016/j.oceaneng.2020.107777}}}.
+
+\bibitem[Rusu(2015)]{en80910370}
+Rusu, L.
+\newblock {Assessment of the Wave Energy in the Black Sea Based on a 15-Year
+  Hindcast with Data Assimilation}.
+\newblock {\em Energies} {\bf 2015}, {\em 8},~10370--10388, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en80910370}{\detokenize{10.3390/en80910370}}}.
+
+\bibitem[Rhee \em{et~al.}(2014)Rhee, Park, and Inoue]{Sang-Yong2014}
+Rhee, S.Y.; Park, J.; Inoue, A. (Eds.)
+\newblock {\em Soft Computing in Machine Learning}; Springer: Berlin, Germany, 2014.
+
+\bibitem[Bishop(2006)]{Bishop:2006:PRM:1162264}
+Bishop, C.M.
+\newblock {\em {Pattern Recognition and Machine Learning}}; Springer: New York, NY, USA,  2006.
+
+\bibitem[Chang \em{et~al.}(2019)Chang, Hsu, and Chang]{Fi-John2019}
+Chang, F.J.; Hsu, K.; Chang, L.C., Eds.
+\newblock {\em Flood Forecasting Using Machine Learning Methods}; MPDI: Basel, Switzerland, 2019.
+
+\bibitem[Dineva \em{et~al.}(2019)Dineva, Mosavi, Faizollahzadeh~Ardabili,
+  Vajda, Shamshirband, Rabczuk, and Chau]{Mosavi2019}
+Dineva, A.; Mosavi, A.; Faizollahzadeh~Ardabili, S.; Vajda, I.; Shamshirband,
+  S.; Rabczuk, T.; Chau, K.W.
+\newblock Review of Soft Computing Models in Design and Control of Rotating
+  Electrical Machines.
+\newblock {\em Energies} {\bf 2019}, {\em 12},~1049, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/en12061049}{\detokenize{10.3390/en12061049}}}.
+
+\bibitem[Guo \em{et~al.}(2018)Guo, Wang, Chen, Li, Liu, Xu, Huang, and
+  Huang]{GUO201816}
+Guo, Y.; Wang, J.; Chen, H.; Li, G.; Liu, J.; Xu, C.; Huang, R.; Huang, Y.
+\newblock {Machine learning-based thermal response time ahead energy demand
+  prediction for building heating systems}.
+\newblock {\em Appl. Energy} {\bf 2018}, {\em 221},~16--27, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.apenergy.2018.03.125}{\detokenize{10.1016/j.apenergy.2018.03.125}}}.
+
+\bibitem[Frank \em{et~al.}(2016)Frank, Hall, and Witten]{WEKA}
+Frank, E.; Hall, M.A.; Witten, I.H.
+\newblock {\emph{The WEKA Workbench. Online Appendix for Data Mining: Practical
+  Machine Learning Tools and Techniques}};  Morgan Kaufmann: Cambridge, MA, USA, 2016.
+
+\bibitem[Dur{\'a}n-Rosal \em{et~al.}(2017)Dur{\'a}n-Rosal, Fern{\'a}ndez,
+  Guti{\'e}rrez, and Herv{\'a}s-Mart{\'i}nez]{DURANROSAL2017268}
+Dur{\'a}n-Rosal, A.M.; Fern{\'a}ndez, J.C.; Guti{\'e}rrez, P.A.;
+  Herv{\'a}s-Mart{\'i}nez, C.
+\newblock {Detection and prediction of segments containing extreme significant
+  wave heights}.
+\newblock {\em Ocean Eng.} {\bf 2017}, {\em 142},~268--279, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.oceaneng.2017.07.009}{\detokenize{10.1016/j.oceaneng.2017.07.009}}}.
+
+\bibitem[Kumar \em{et~al.}(2017)Kumar, Savitha, and Al~Mamun]{KUMAR2017605}
+Kumar, N.K.; Savitha, R.; Al~Mamun, A.
+\newblock {Regional ocean wave height prediction using sequential learning
+  neural networks}.
+\newblock {\em Ocean Eng.} {\bf 2017}, {\em 129},~605--612, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.oceaneng.2016.10.033}{\detokenize{10.1016/j.oceaneng.2016.10.033}}}.
+
+\bibitem[Ali \em{et~al.}(2020)Ali, Prasad, Xiang, and Deo]{Ali2020}
+Ali, M.; Prasad, R.; Xiang, Y.; Deo, R.C.
+\newblock Near real-time significant wave height forecasting with hybridized
+  multiple linear regression algorithms.
+\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2020}, {\em
+  132},~110003, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/doi.org/10.1016/j.rser.2020.110003}{\detokenize{10.1016/j.rser.2020.110003}}}.
+
+\bibitem[Cornejo-Bueno \em{et~al.}(2016)Cornejo-Bueno, Nieto-Borge,
+  García-Díaz, Rodríguez, and Salcedo-Sanz]{Cornejo-Bueno2016}
+Cornejo-Bueno, L.; Nieto-Borge, J.; García-Díaz, P.; Rodríguez, G.;
+  Salcedo-Sanz, S.
+\newblock Significant wave height and energy flux prediction for marine energy
+  applications: A grouping genetic algorithm---Extreme Learning Machine
+  approach.
+\newblock {\em Renew. Energy} {\bf 2016}, {\em 97},~380--389, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/doi.org/10.1016/j.renene.2016.05.094}{\detokenize{10.1016/j.renene.2016.05.094}}}.
+
+\bibitem[Emmanouil \em{et~al.}(2020)Emmanouil, Aguilar, Nane, and
+  Schouten]{Emmanouil2020}
+Emmanouil, S.; Aguilar, S.G.; Nane, G.F.; Schouten, J.J.
+\newblock Statistical models for improving significant wave height predictions
+  in offshore operations.
+\newblock {\em Ocean Eng.} {\bf 2020}, {\em 206},~107249, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/doi.org/10.1016/j.oceaneng.2020.107249}{\detokenize{10.1016/j.oceaneng.2020.107249}}}.
+
+\bibitem[Shamshirband \em{et~al.}(2020)Shamshirband, Mosavi, Rabczuk, Nabipour,
+  and wing Chau]{Shamshirband2020}
+Shamshirband, S.; Mosavi, A.; Rabczuk, T.; Nabipour, N.; Wing Chau, K.
+\newblock Prediction of significant wave height; comparison between nested grid
+  numerical model, and machine learning models of artificial neural networks,
+  extreme learning and support vector machines.
+\newblock {\em Eng. Appl. Comput. Fluid Mech.} {\bf
+  2020}, {\em 14},~805--817, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1080/19942060.2020.1773932}{\detokenize{10.1080/19942060.2020.1773932}}}.
+
+\bibitem[Johansson \em{et~al.}(2015)Johansson, Epitropou, Karatzas, Karppinen,
+  Wanner, Vrochidis, Bassoukos, Kukkonen, and Kompatsiaris]{JOHANSSON2015143}
+Johansson, L.; Epitropou, V.; Karatzas, K.; Karppinen, A.; Wanner, L.;
+  Vrochidis, S.; Bassoukos, A.; Kukkonen, J.; Kompatsiaris, I.
+\newblock {Fusion of meteorological and air quality data extracted from the web
+  for personalized environmental information services}.
+\newblock {\em Environ. Model. Softw.} {\bf 2015}, {\em
+  64},~143--155, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.envsoft.2014.11.021}{\detokenize{10.1016/j.envsoft.2014.11.021}}}.
+
+\bibitem[Fern{\'a}ndez \em{et~al.}(2015)Fern{\'a}ndez, Salcedo-Sanz,
+  Guti{\'e}rrez, Alexandre, and Herv{\'a}s-Mart{\'i}nez]{FERNANDEZ201544}
+Fern{\'a}ndez, J.C.; Salcedo-Sanz, S.; Guti{\'e}rrez, P.A.; Alexandre, E.;
+  Herv{\'a}s-Mart{\'i}nez, C.
+\newblock {Significant wave height and energy flux range forecast with machine
+  learning classifiers}.
+\newblock {\em Eng. Appl. Artif. Intell.} {\bf 2015},
+  {\em 43},~44--53, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.engappai.2015.03.012}{\detokenize{10.1016/j.engappai.2015.03.012}}}.
+
+\bibitem[Adams and Flora(2010)]{Adams2010}
+Adams, J.; Flora, S.
+\newblock {Correlating seabird movements with ocean winds: linking satellite
+  telemetry with ocean scatterometry}.
+\newblock {\em Mar. Biol.} {\bf 2010}, {\em 157},~915--929, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1007/s00227-009-1367-y}{\detokenize{10.1007/s00227-009-1367-y}}}.
+
+\bibitem[{National Data Buoy Center}()]{NOAA}
+{National Data Buoy Center}.
+\newblock {National Oceanic and Atmospheric Administration of the USA (NOAA)}.
+\newblock  Available online: \url{http://www.ndbc.noaa.gov/}
+\newblock {(accessed on 10 December 2020)}.
+
+\bibitem[Kalnay \em{et~al.}(1996)Kalnay, Kanamitsu, Kistler, Collins, Deaven,
+  Gandin, Iredell, Saha, White, Woollen, Zhu, Leetmaa, Reynolds, Chelliah,
+  Ebisuzaki, Higgins, Janowiak, Mo, Ropelewski, Wang, Jenne, and
+  Joseph]{Kalnay1996}
+Kalnay, E.; Kanamitsu, M.; Kistler, R.; Collins, W.; Deaven, D.; Gandin, L.;
+  Iredell, M.; Saha, S.; White, G.; Woollen, J.; et al.
+\newblock {The NCEP/NCAR 40-Year Reanalysis Project}.
+\newblock {\em Bull. Am. Meteorol. Soc.} {\bf 1996},
+  {\em 77},~437--471, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1175/1520-0477(1996)077<0437:TNYRP>2.0.CO;2}{\detokenize{10.1175/1520-0477(1996)077<0437:TNYRP>2.0.CO;2}}}.
+
+\bibitem[Kistler \em{et~al.}(2001)Kistler, Collins, Saha, White, Woollen,
+  Kalnay, Chelliah, Ebisuzaki, Kanamitsu, Kousky, van~den Dool, Jenne, and
+  Fiorino]{Kistler2001}
+Kistler, R.; Collins, W.; Saha, S.; White, G.; Woollen, J.; Kalnay, E.;
+  Chelliah, M.; Ebisuzaki, W.; Kanamitsu, M.; Kousky, V.; et al.
+\newblock {The NCEP--NCAR 50--Year Reanalysis: Monthly Means CD--ROM and
+  Documentation}.
+\newblock {\em Bull. Am. Meteorol. Soc.} {\bf 2001},
+  {\em 82},~247--267, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1175/1520-0477(2001)082<0247:TNNYRM>2.3.CO;2}{\detokenize{10.1175/1520-0477(2001)082<0247:TNNYRM>2.3.CO;2}}}.
+
+\bibitem[WEK()]{WEKA_ARFF}
+{The WEKA Data Mining Software: {Attribute-Relation File Format (ARFF)}}.
+\newblock Available online: \url{https://www.cs.waikato.ac.nz/ml/weka/arff.html}
+\newblock {(accessed  on 10 December 2020)}.
+
+\bibitem[Ali and Prasad(2019)]{Ali2019}
+Ali, M.; Prasad, R.
+\newblock Significant wave height forecasting via an extreme learning machine
+  model integrated with improved complete ensemble empirical mode
+  decomposition.
+\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2019}, {\em
+  104},~281--295, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.rser.2019.01.014}{\detokenize{10.1016/j.rser.2019.01.014}}}.
+
+\bibitem[Chatziioannou \em{et~al.}(2017)Chatziioannou, Katsardi, Koukouselis,
+  and Mistakidis]{CHATZIIOANNOU2017126}
+Chatziioannou, K.; Katsardi, V.; Koukouselis, A.; Mistakidis, E.
+\newblock {The effect of nonlinear wave-structure and soil-structure
+  interactions in the design of an offshore structure}.
+\newblock {\em Mar. Struct.} {\bf 2017}, {\em 52},~126--152, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.marstruc.2016.11.003}{\detokenize{10.1016/j.marstruc.2016.11.003}}}.
+
+\bibitem[Dalgic \em{et~al.}(2015)Dalgic, Lazakis, Dinwoodie, McMillan, and
+  Revie]{DALGIC2015211}
+Dalgic, Y.; Lazakis, I.; Dinwoodie, I.; McMillan, D.; Revie, M.
+\newblock {Advanced logistics planning for offshore wind farm operation and
+  maintenance activities}.
+\newblock {\em Ocean Eng.} {\bf 2015}, {\em 101},~211--226, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.oceaneng.2015.04.040}{\detokenize{10.1016/j.oceaneng.2015.04.040}}}.
+
+\bibitem[Spaulding \em{et~al.}(2016)Spaulding, Grilli, Damon, Crean, Fugate,
+  Oakley, and Stempel]{Spaulding2020}
+Spaulding, M.L.; Grilli, A.; Damon, C.; Crean, T.; Fugate, G.; Oakley, B.A.;
+  Stempel, P.
+\newblock STORMTOOLS: Coastal Environmental Risk Index (CERI).
+\newblock {\em J. Mar. Sci. Eng.} {\bf 2016}, {\em
+  4},~54, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/jmse4030054}{\detokenize{10.3390/jmse4030054}}}.
+
+\bibitem[{National Data Buoy Center}({\natexlab{a}})]{NOAA_1}
+{National Data Buoy Center}.
+\newblock {NDBC---Historical NDBC Data}.
+\newblock Available online: \url{http://www.ndbc.noaa.gov/historical\_data.shtml}
+\newblock {(accessed on 10 December 2020)}.
+
+\bibitem[{National Data Buoy Center}({\natexlab{b}})]{NOAA_2}
+{National Data Buoy Center}.
+\newblock {NDBC---Important NDBC Web Site Changes}.
+\newblock Available online:  \url{http://www.ndbc.noaa.gov/mods.shtml}
+\newblock {(accessed on 10 December 2020)}.
+
+\bibitem[{National Data Buoy Center}({\natexlab{c}})]{NOAA_4}
+{National Data Buoy Center}.
+\newblock {Measurement Descriptions and Units}.
+\newblock Available online: \url{https://www.ndbc.noaa.gov/measdes.shtml#stdmet}
+\newblock (accessed on 10 December 2020).
+
+\bibitem[{NOAA/OAR/ESRL PSD}()]{NNRP}
+{NOAA/OAR/ESRL PSD}.
+\newblock {ESRL : PSD : NCEP/NCAR Reanalysis 1}.
+\newblock
+  Available online: \url{https://www.esrl.noaa.gov/psd/data/gridded/data.ncep.reanalysis.html}
+\newblock {(accessed on 15 January 2019)}.
+
+\bibitem[{Unidata}(2017)]{NetCDF}
+{Unidata}.
+\newblock {\emph{Network Common Data Form (NetCDF) Version 4.6.10 [Software]};
+  UCAR/Unidata: Boulder, CO, USA},  2017;
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.5065/D6H70CW6}{\detokenize{10.5065/D6H70CW6}}}.
+
+\bibitem[de~Smith \em{et~al.}(2009)de~Smith, Goodchild, and
+  Longley]{Haversine_2009}
+de~Smith, M.J.; Goodchild, M.F.; Longley, P.A.
+\newblock {\em {Geospatial Analysis: A Comprehensive Guide to Principles,
+  Techniques and Software Tools}}, 3rd ed.; Matador:  Leicester, United Kingdom, 2009.
+
+\bibitem[{Hosmer Jr} \em{et~al.}(2013){Hosmer Jr}, Lemeshow, and
+  Sturdivant]{hosmer2013applied}
+{Hosmer}, D.W., Jr.; Lemeshow, S.; Sturdivant, R.X.
+\newblock {\em {Applied Logistic Regression}}, 3rd ed.; John Wiley \& Sons: New Jersey, USA, 2013.
+
+\bibitem[Quinlan(1992)]{quinlan2014c4}
+Quinlan, J.R.
+\newblock {\em {C4. 5: Programs for Machine Learning}}; Morgan Kaufmann:  Burlington, Massachusetts, USA, 1992.
+
+\bibitem[Breiman(2001)]{breiman2001random}
+Breiman, L.
+\newblock {Random forests}.
+\newblock {\em Mach. Learn.} {\bf 2001}, {\em 45},~5--32, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1023/A:1010933404324}{\detokenize{10.1023/A:1010933404324}}}.
+
+\bibitem[Cortes and Vapnik(1995)]{cortes1995support}
+Cortes, C.; Vapnik, V.
+\newblock {Support-vector networks}.
+\newblock {\em Mach. Learn.} {\bf 1995}, {\em 20},~273--297, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1007/BF00994018}{\detokenize{10.1007/BF00994018}}}.
+
+\bibitem[Haykin(1994)]{haykin1994neural}
+Haykin, S.
+\newblock {\em {Neural Networks: A Comprehensive Foundation}}; Prentice Hall
+  PTR:  New Jersey, USA, 1994.
+
+\bibitem[{National Data Buoy Center}()]{NOAA_3}
+{National Data Buoy Center}.
+\newblock {NDBC---Measurement Descriptions and Units}.
+\newblock  Available online: \url{https://www.ndbc.noaa.gov/measdes.shtml}
+\newblock {(accessed on 10 December 2020)}.
+
+\bibitem[Dur{\'a}n-Rosal \em{et~al.}(2016)Dur{\'a}n-Rosal,
+  Herv{\'a}s-Mart{\'i}nez, Tall{\'o}n-Ballesteros, Mart{\'i}nez-Estudillo, and
+  Salcedo-Sanz]{DuranRosal2016}
+Dur{\'a}n-Rosal, A.M.; Herv{\'a}s-Mart{\'i}nez, C.; Tall{\'o}n-Ballesteros,
+  A.J.; Mart{\'i}nez-Estudillo, A.C.; Salcedo-Sanz, S.
+\newblock {{M}assive missing data reconstruction in ocean buoys with
+  evolutionary product unit neural networks}.
+\newblock {\em Ocean Eng.} {\bf 2016}, {\em 117},~292--301, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1016/j.oceaneng.2016.03.053}{\detokenize{10.1016/j.oceaneng.2016.03.053}}}.
+
+\bibitem[Alcal{\'a}-Fdez \em{et~al.}(2009)Alcal{\'a}-Fdez, S{\'a}nchez,
+  Garc{\'i}a, del Jes{\'u}s, Ventura, Garrell, Otero, Romero, Bacardit, Rivas,
+  Fern{\'a}ndez, and Herrera]{AlcalFdez2009KEELAS}
+Alcal{\'a}-Fdez, J.; S{\'a}nchez, L.; Garc{\'i}a, S.; del Jes{\'u}s, M.J.;
+  Ventura, S.; Garrell, J.M.; Otero, J.; Romero, C.; Bacardit, J.; Rivas, V.M.;
+  et al.
+\newblock {KEEL: A software tool to assess evolutionary algorithms for data
+  mining problems}.
+\newblock {\em Soft Comput.} {\bf 2009}, {\em 13},~307--318, 
+\newblock
+  doi:{\changeurlcolor{black}\href{https://doi.org/10.1007/s00500-008-0323-y}{\detokenize{10.1007/s00500-008-0323-y}}}.
+
+\end{thebibliography}
 
